% SPDX-FileCopyrightText: 2026 Will Cook
% SPDX-License-Identifier: CC-BY-4.0
\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,booktabs,seqsplit}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,linkcolor=black,urlcolor=RoyalBlue,citecolor=black]{hyperref}
\newcommand{\commit}{0e85f7ebadb0f81a26791bc167ce224bd2534f36}
\newcommand{\repobase}{https://github.com/wcook04/plectis-lean-erdos249-257/blob/\commit}
\newcommand{\PK}{Erdos249257}
\input{module-aliases.tex}
\newcommand{\idn}[1]{\texttt{\expandafter\seqsplit\expandafter{\detokenize{#1}}}}
\newcommand{\lref}[3]{\href{\repobase/\PK/#1\#L#2}{\texttt{\modulesigil{#1}}}}
\emergencystretch=2em
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{remark}\newtheorem*{remark}{Remark}
\newcommand{\Z}{\mathbb Z}
\newcommand{\ph}{\varphi}
\title{Transport and Curvature Certificates for the Totient Constant\\
{\large a Lean~4 companion to the Erd\H{o}s~\#249 formalisation}}
\author{Will Cook}
\date{July 2026}

\begin{document}
\sloppy
\maketitle
\begin{abstract}
Let $S=\sum_{n\geq1}\ph(n)/2^n$ and
$R_N=\sum_{j\geq1}\ph(N+j)/2^j$.  The irrationality of $S$ is
Erd\H{o}s Problem~249 and remains open.  We formalise exact secondary
reductions built from finite differences of the tails $R_N$: curvature,
exponent-transport and prime-jump commutators, and an anchored four-vertex
operator at $H,3H,5H,15H$.  Each object splits into an integer dyadic window
and a rigorously bounded tail, giving a decidable modular certificate for
non-integrality.  We prove that old M\"obius divisor channels are affine and
are annihilated by the relevant moments, isolate the new-channel migration at
a prime jump, and show that every fixed-precision valuation--unit word admits
a centred completion.  The finite non-integrality criteria and no-go theorem are
unconditional; every irrationality conclusion remains conditional on an
explicitly named unbounded supply.  No such supply, and hence no solution of
Problem~249, is claimed.
\end{abstract}

\section{Scope and logical boundary}\label{sec:scope}
The main exposition~\cite{mainpaper} develops the direct certificate reduction
for $S$, proves an unconditional denominator exclusion, and records the open
boundary.  This companion studies finite-difference operators on the same
tail sequence.  They annihilate affine contributions from divisor channels
already present at an lcm height, but the remaining new channels must still be
shown to produce non-integral values at unboundedly many heights.  Thus every
endpoint below has the form
\[
 \text{unbounded supply of finite certificates}\Longrightarrow S\notin\mathbb Q,
\]
with the antecedent open.  A certificate at one height is finite evidence,
not evidence for the cofinal quantifier.  The fixed-precision no-go theorem is
a boundary on one local strategy, not a global impossibility result.

The Lambert-series background is classical.  The open problem is recorded in
the Erd\H{o}s--Graham collection and maintained catalogue
\cite{erdosgraham1980,erdosproblems}.  We make no priority claim for finite
differences or affine moment cancellation as general techniques.  The
contribution here is their precise integration into checked identities,
finite non-integrality criteria, conditional hypotheses, and an obstruction
theorem.

Small blue source marks link the displayed claims to their formal proofs and
may be ignored on a first reading.

\section{Dyadic windows}\label{sec:windows}
For a finite integer combination $X(H)=\sum_i c_iR_{m_iH}$, truncate each
tail at depth $L$.  Multiplication by $2^L$ gives
\[
 2^LX(H)=A(H,L)+E(H,L),
\]
where $A(H,L)\in\Z$ is computable.  If $|E(H,L)|\le B(H,L)$, then
$X(H)\notin\Z$ whenever
\[
 B(H,L)<A(H,L)\bmod2^L<2^L-B(H,L).
\]
The reusable soundness theorem is
\idn{notMem_int_of_window_remainder_bound}
\lref{PrimeJumpWindow.lean}{53}{notMem_int_of_window_remainder_bound}.
Lean checks the exact decomposition, radius, and concrete modular arithmetic;
an external analytic or arithmetic argument must provide unbounded instances.

\section{Curvature on the lcm cone}\label{sec:curvature}
Set
\[
 Q_H=R_{4H}-3R_{2H}+2R_H.
\]
Its depth-$L$ window $C(H,L)$ satisfies
\[
 2^LQ_H=C(H,L)+E_C(H,L),\qquad |E_C(H,L)|\le6H+3L+3.
\]
The exact split and bound are
\idn{two_pow_mul_curvature_eq_window_add_shifted}
\lref{CurvatureCarry.lean}{119}{two_pow_mul_curvature_eq_window_add_shifted}
and \idn{abs_curvatureShiftedTail_le}
\lref{CurvatureCarry.lean}{137}{abs_curvatureShiftedTail_le}.
The central residue criterion therefore proves $Q_H\notin\Z$ by
\idn{curvature_notMem_int_of_sharpCurvatureCert}
\lref{CurvatureCarry.lean}{159}{curvature_notMem_int_of_sharpCurvatureCert}.
An unbounded supply implies irrationality through
\idn{irrational_totientSeries_of_sharpCurvatureSupply}
\lref{CurvatureCarry.lean}{206}{irrational_totientSeries_of_sharpCurvatureSupply},
but that supply is not proved.

The window evolves by
$C(H,L+1)=2C(H,L)+d(H,L+1)$, checked in
\idn{curvatureWindow_succ}
\lref{CurvatureCarry.lean}{259}{curvatureWindow_succ}.  Exact phase-certificate
and common-carrier formulas expose possible routes to a supply theorem without
asserting one.

\section{Exponent transport and affine cancellation}\label{sec:transport}
If $p\mid H$, multiplication by $p$ adds no squarefree divisor support:
\[
 d\mid pH\Longleftrightarrow d\mid H\qquad(d\text{ squarefree}).
\]
This is \idn{squarefree_dvd_mul_iff_of_dvd}
\lref{ExponentOnlyTransport.lean}{36}{squarefree_dvd_mul_iff_of_dvd}.
The exact residue kernel is affine, not merely periodic:
\[
 K_d(N)=N\,a_d(N)+b_d(N),
\]
where both $a_d$ and $b_d$ are periodic modulo $d$.  The decomposition is
\idn{transportResidueKernel_eq_affineState}
\lref{ExponentOnlyTransport.lean}{141}{transportResidueKernel_eq_affineState}.
Keeping both blocks prevents an intercept-only matrix calculation from being
misread as global non-cancellation.  The exact $H\mapsto3H$ two-block transfer
is \idn{residueTransportDefect_three_eq_affineState}
\lref{ExponentOnlyTransport.lean}{176}{residueTransportDefect_three_eq_affineState}.

Every old channel $d\mid H$ is killed by the first-order defect
\idn{residueTransportDefect_eq_zero_of_dvd}
\lref{ExponentOnlyTransport.lean}{112}{residueTransportDefect_eq_zero_of_dvd}.
The endpoint
\idn{irrational_totient_series_of_exponentOnlyThreeTransportSupply}
\lref{ExponentOnlyTransport.lean}{205}{irrational_totient_series_of_exponentOnlyThreeTransportSupply}
is conditional on a named fixed-three tower supply.

More generally, coefficients satisfying $\sum_i c_i=\sum_i c_im_i=0$
annihilate every old affine channel.  This is
\idn{oldChannel_affine_moment_annihilation}
\lref{JointExponentTransport.lean}{36}{oldChannel_affine_moment_annihilation}.
For $q(X,Y)=XY-3X-2Y+4$, the vanishing $q(1,1)=q(3,5)=0$ gives
\[
 K_d(15H)-3K_d(3H)-2K_d(5H)+4K_d(H)=0\quad(d\mid H),
\]
checked by \idn{joint35_oldChannel_zero}
\lref{JointExponentTransport.lean}{71}{joint35_oldChannel_zero}.
The associated window radius is $19H+5L+5$; the resulting finite
non-integrality criterion is
\idn{joint35Cone_notMem_int_of_cert}
\lref{JointExponentTransport.lean}{172}{joint35Cone_notMem_int_of_cert}.
The conditional endpoint is
\idn{irrational_totient_series_of_anchoredJoint35ConeSupply}
\lref{JointExponentTransport.lean}{217}{irrational_totient_series_of_anchoredJoint35ConeSupply}.

\section{Prime jumps and three-transport boundaries}\label{sec:primejump}
When $p\nmid H$, a jump from $H$ to $pH$ creates new squarefree support.  The
residue Hecke defect splits into a post-jump channel and an explicit migration
correction in \idn{residueHeckeDefect_eq_postJump_add_migration}
\lref{PrimeJumpMigration.lean}{38}{residueHeckeDefect_eq_postJump_add_migration}.
For $d\mid H$ the total old-channel defect is zero
\lref{PrimeJumpMigration.lean}{46}{residueHeckeDefect_eq_zero_of_dvd}; the
$(d,H,p)=(5,12,5)$ calculation records the two cancelling terms.

Set
\[
 J(H,p)=(R_{2pH}-R_{pH})-p(R_{2H}-R_H).
\]
Its exact remainder radius is $3pH+(p+1)(L+2)$, proved with the window split by
\idn{primeJumpTailCommutator_eq_window_add_remainder}
\lref{PrimeJumpWindow.lean}{91}{primeJumpTailCommutator_eq_window_add_remainder}.
The finite non-integrality criterion is
\idn{primeJumpTailCommutator_notMem_int_of_sharpKill}
\lref{PrimeJumpWindow.lean}{178}{primeJumpTailCommutator_notMem_int_of_sharpKill}.
Lean verifies $(H,p,L)=(12,5,15)$ in
\idn{primeJumpSharpKill_twelve_five}
\lref{PrimeJumpWindow.lean}{186}{primeJumpSharpKill_twelve_five}; this is one
finite deposit, not an unbounded sequence.

The balanced vertices $H,2H,3H,6H$ give a related radius $9H+4L+4$.
The corresponding criterion is \idn{threeTransport_notMem_int_of_cert}
\lref{ThreeTransportBoundary.lean}{126}{threeTransport_notMem_int_of_cert};
the irrationality theorem
\idn{irrational_totientSeries_of_roughnessBoundaryLatticeSupply}
\lref{ThreeTransportBoundary.lean}{177}{irrational_totientSeries_of_roughnessBoundaryLatticeSupply}
is conditional.  The concrete check $(H,L)=(60,12)$ is
\idn{sharpThreeTransportCert_60_12}
\lref{ThreeTransportBoundary.lean}{203}{sharpThreeTransportCert_60_12}.

\section{A first-harmonic criterion}\label{sec:harmonic}
For the window discrepancy $A(h,N,L)$ define
\[
 W_1(h,N,L)=\cos\left(2\pi\frac{A(h,N,L)\bmod2^L}{2^L}\right).
\]
Assume $N\in[X,2X)$ and $16(2X+h+L+2)\le2^L$.  If every $N$ failed the
central-residue certificate, each phase would lie within $\pi/8$ of an
integer turn, so each cosine would exceed $9/10$.  Hence
\[
 \sum_{X\le N<2X}W_1(h,N,L)\le\frac9{10}X
\]
forces a certificate in the block.  This implication is
\idn{exists_certifiedKill_of_first_harmonic_gap}
\lref{FirstHarmonicGap.lean}{128}{exists_certifiedKill_of_first_harmonic_gap}.
More generally, if $T$ is any nonempty finite set of indices below $2X$ and
\[
 \sum_{N\in T}W_1(h,N,L)\le\frac9{10}|T|,
\]
then a certificate occurs at some $N\in T$.  This subset form is
\idn{exists_certifiedKill_of_first_harmonic_gap_subset}
\lref{FirstHarmonicGap.lean}{104}{exists_certifiedKill_of_first_harmonic_gap_subset}.
By certificate completeness, the witness produced here is an exact
non-integral tail difference, not a heuristic phase test.  Hence a
first-harmonic saving on one finite block produces exactly the kind of witness
required by the tail-difference criterion.  What is still missing is a
first-harmonic estimate that supplies such blocks at unbounded parameters; no
such estimate is proved here.

\section{Adjacent-pair phase separation}\label{sec:adjacentphase}
At a scale satisfying the same room inequality, with $N+1<2X$, suppose that
the first harmonic phases at two adjacent indices obey
\[
 \left\lVert
 \operatorname{windowFirstExp}(h,N+1,L)-\operatorname{windowFirstExp}(h,N,L)
 \right\rVert^2\ge\frac{19}{25}.
\]
Then a certificate occurs at one of $N$ and $N+1$.  This finite implication is
\idn{exists_certifiedKill_of_phasePairSeparation}
\lref{AdjacentPhaseSeparation.lean}{258}{exists_certifiedKill_of_phasePairSeparation}.
Consequently, a cofinal supply of adjacent pairs satisfying this inequality
implies irrationality of the totient constant, by
\idn{irrational_totient_series_of_adjacentPhaseSeparation}
\lref{AdjacentPhaseSeparation.lean}{295}{irrational_totient_series_of_adjacentPhaseSeparation}.
No cofinal adjacent-pair separation is proved here.

\section{Pivot anti-reconstruction and window-pair geometry}\label{sec:pivot}
The first-harmonic phase of a finite window is the exact additive character of
its dyadic discrepancy residue.  It differs from the corresponding infinite
tail-orbit phase by the explicitly bounded omitted tail.  On a finite supplier
fibre, the anchored defect is the squared distance of the window phases from
the constant phase $1$; its centred form is the usual pairwise-energy identity.
Thus a counted family of separated phase pairs of sufficient total squared
chord mass produces a finite certificate.

The formal proof records several equivalent sufficient cofinal hypotheses.  In
particular, a cofinal pivot
anti-reconstruction bound, a cofinal centred reconstruction-variance bound, or
a cofinal counted-separated-pairs bound is sufficient for irrationality of
$S$.  The direct finite criterion is
\idn{exists_certifiedKill_of_pivotFiberSeparatedPairs}
\lref{PivotAntiReconstruction.lean}{1105}{exists_certifiedKill_of_pivotFiberSeparatedPairs},
and the resulting irrationality theorem is
\idn{irrational_totient_series_of_pivotAntiReconstruction}
\lref{PivotAntiReconstruction.lean}{1502}{irrational_totient_series_of_pivotAntiReconstruction}.

There is also an exact fibre-free formulation: the cofinal window-pair
predicate is equivalent to irrationality of $S$, not merely sufficient for it.
The reverse implication uses the expansive doubling orbit to obtain adjacent
infinite-tail separation and then chooses a finite window depth.  This is
formalised by
\idn{dtwWindowSeparatedPairs_iff_irrational_totient_series}
\lref{PivotAntiReconstruction.lean}{1744}{dtwWindowSeparatedPairs_iff_irrational_totient_series}.
The development does not prove any of these cofinal arithmetic hypotheses, so it does
not settle Erd\H{o}s~\#249.

\section{Fixed local precision cannot finish the argument}\label{sec:nogo}
A finite valuation--unit symbol records a bounded amount of local transport
data.  For every positive precision $u$, every finite compatible word admits
a prefix-locked centred completion.  Thus no finite word over this fixed local
alphabet excludes all centred endpoints.  The exact statement is
\idn{fixedPrecisionTropicalNoGo}
\lref{TropicalCurvatureCarry.lean}{137}{fixedPrecisionTropicalNoGo}.
This does not say curvature certificates are rare.  A proof may use precision
growing with scale, global correlations, or a harmonic gap.  It rules out only
the inference that a bounded local signature alone forces escape.

\section{LCM factor ideals and whole-ray anchors}\label{sec:factorideal}
For every $t\geq 3$, there is a nonzero synthetic affine carry whose forcing
letters lie in the principal ideal generated by
$\varphi(\operatorname{lcm}(1,\ldots,t))$, satisfy the exact whole-ray anchor
values, and remain inside the relevant diagonal bounds.  Nevertheless its
state is a strict survivor.  The construction is formalised by
\idn{lcm_factorIdeal_sparseAnchor_not_sufficient}
\lref{LcmFactorIdealPulseObstruction.lean}{866}{lcm_factorIdeal_sparseAnchor_not_sufficient}.

The same example remains a dyadic coboundary after every finite integer shift
polynomial.  In particular, the finite-rank LCM shift algebra alone cannot
remove its terminal compensation; the corresponding formal statement is
\idn{lcm_factorIdeal_finiteRank_shiftAlgebra_not_sufficient}
\lref{LcmFactorIdealPulseObstruction.lean}{798}{lcm_factorIdeal_finiteRank_shiftAlgebra_not_sufficient}.
This is a synthetic countermodel: it does not assert that these forcing letters
are actual totient differences, and it does not rule out a nonlinear argument
using the missing fresh divisor channels.

\section{Statement matrix and open boundary}\label{sec:index}
\begin{center}\small
\begin{tabular}{@{}p{0.24\textwidth}p{0.46\textwidth}p{0.2\textwidth}@{}}
\toprule Layer & Checked statement & Status \\
\midrule
Curvature & exact split, radius, recurrence, criterion & unconditional \\
Exponent transport & stable support, affine state, cancellation & unconditional \\
Joint $(3,5)$ transport & moment annihilation and criterion & unconditional \\
Prime jump & migration, sharp window, $(12,5,15)$ & finite \\
Three transport & sharp window and $(60,12)$ & finite \\
First harmonic & cosine gap forces a certificate & conditional input \\
Adjacent phases & squared chord separation forces a certificate & conditional input \\
Pivot anti-reconstruction & pairwise phase energy and exact window-pair equivalence & conditional / equivalence \\
Tropical signature & fixed precision has a completion & no-go \\
LCM factor ideals & factor/anchor and finite shift algebra have a synthetic survivor & no-go \\
Irrationality of $S$ & follows from a named unbounded supply & supply open \\
\bottomrule
\end{tabular}
\end{center}
The development proves no cofinal supply of curvature, prime-jump, joint, or
three-transport certificates, and no cofinal first-harmonic gap.  It therefore
does not prove Erd\H{o}s~\#249.

\begin{thebibliography}{9}
\bibitem{erdosgraham1980} P.~Erd\H{o}s and R.~L. Graham,
\emph{Old and New Problems and Results in Combinatorial Number Theory},
Monographies de l'Enseignement Math\'ematique 28 (1980), 61--62,
\href{https://mathweb.ucsd.edu/~ronspubs/80_11_number_theory.pdf}{scan}.
\bibitem{erdosproblems} T.~Bloom (ed.), \emph{Erd\H{o}s Problems}, entry 249,
\href{https://www.erdosproblems.com/249}{\texttt{erdosproblems.com/249}},
accessed July 2026.
\bibitem{mainpaper} W.~Cook, \emph{Erd\H{o}s Problems 249 and 257 in Lean~4:
formalised results, certificate reductions, and open questions}, companion
exposition in this release.
\end{thebibliography}
\end{document}
