% SPDX-FileCopyrightText: 2026 Will Cook
% SPDX-License-Identifier: CC-BY-4.0
%
% A mathematician-facing exposition of the Erdos249257 Lean
% development. Every formalised statement links to the exact Lean source line
% at the pinned formal-source checkpoint. Build with `tectonic erdos249-257-exposition.tex` or `make`.
\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,linkcolor=black,urlcolor=RoyalBlue,citecolor=black]{hyperref}

% ---------------------------------------------------------------------------
% Drill-down machinery: each formal anchor links to the GitHub blob at the
% pinned formal-source commit. \commit names that checkpoint; scripts/check_release.py in
% the repository verifies that it matches docs/claims.json and that every
% \lref/\lrefx/\lloc target resolves to the named declaration at that line.
% ---------------------------------------------------------------------------
\newcommand{\commit}{81150c1c6fdd4cd51e76d3aa0de344341e0e9e7c}
\newcommand{\repobase}{https://github.com/wcook04/plectis-lean-erdos249-257/blob/\commit}
\newcommand{\sourceurl}{https://github.com/wcook04/plectis-lean-erdos249-257/tree/\commit}
\newcommand{\rootbase}{https://github.com/wcook04/plectis-lean-erdos249-257/blob/main}
\newcommand{\PK}{Erdos249257}
\input{module-aliases.tex}
% Formal source coordinates remain in the hyperlink targets and verifier
% manifest.  The printed link names the declaration instead of using an
% indistinguishable source glyph.
\newcommand{\lref}[3]{\href{\repobase/\PK/#1\#L#2}{\nolinkurl{#3}}}
\newcommand{\lrefx}[3]{\href{\repobase/\PK/#1\#L#2}{\nolinkurl{#3}}}
\newcommand{\lloc}[2]{\href{\repobase/\PK/#1\#L#2}{Lean source at line #2}}
% \rref{rootfile}: link a repository-root navigation file at the public main branch.
\newcommand{\rref}[1]{\href{\rootbase/\detokenize{#1}}{repository record}}
\emergencystretch=2em

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{problem}[theorem]{Problem}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\ph}{\varphi}
\newcommand{\lcm}{\operatorname{lcm}}
\newcommand{\Irr}{\text{ irrational}}
\newcommand{\Qzero}{Q_0}

\title{\vspace{-4em}Tail Certificates and Achievement-Set Geometry\\
{\large for Erd\H{o}s Problems 249 and 257}}
\author{Will Cook \qquad July 2026}
\date{}

\begin{document}
\sloppy
\maketitle

\begin{abstract}
\noindent
For the totient constant
$S=\sum_{n\ge1}\ph(n)/2^n$, we prove that every rational representation
$S=p/q$ has $q>\Qzero\approx7.96\times10^{34}$.  We also prove that
irrationality of $S$ is equivalent to non-integrality of every positive binary
tail difference, with a complete finite certificate at each fixed pair of
parameters.  The missing input is an unbounded supply of these certificates;
the checks through $t=64$ are finite evidence only.  For Erd\H{o}s~\#257, we
formalise Erd\H{o}s's classical full-support theorem in every integer base and
several structured-support cases.  At base $2$, the Mersenne achievement set is
compact, perfect, nowhere dense, and of measure one.  Membership of $1/2$,
which would refute the universal statement, is equivalent to infinitely many
greedy skips.  The upper last producer and the middle coordinate $-3$ are
impossible; the remaining sufficient condition is a future-tail inequality
for the coordinates $-2$ and $-1$.  Both problems reduce exact finite
arithmetic to one unproved cofinal producer.  Neither is settled.
\end{abstract}

\section{Introduction}\label{sec:intro}

The two questions are whether
$S=\sum_{n\ge1}\ph(n)/2^n$ is irrational (\#249), and whether
$\sum_{n\in A}(2^n-1)^{-1}$ is irrational for every infinite
$A\subseteq\N$ (\#257).  The formulations occur in the Erd\H{o}s--Graham
collection and Erd\H{o}s's later survey~\cite{erdosgraham1980,erdos1988};
the numbering follows the maintained catalogue~\cite{erdosproblems}.

For \#257, proving irrationality for every infinite support is the universal
direction; constructing an infinite support of sum $1/2$ would refute it.
The corpus contains unconditional positive cases and an exact conditional
counterexample route, but reaches neither endpoint.  Across both problems,
tail and carry identities separate exact finite arithmetic from one analytic
or combinatorial cofinality statement.  The status classes are:

\begin{center}
\scriptsize
\begin{tabular}{@{}p{2.3cm}p{5.65cm}p{5.65cm}@{}}
\toprule
status & Erd\H{o}s \#249 & Erd\H{o}s \#257\\
\midrule
\textbf{Classical formalisation} &
Lambert-series and convolution identities used to represent $S$. &
Full support in every base $b\ge2$; pairwise-coprime and periodic support
families.\\[2pt]
\textbf{Unconditional or exact here} &
$q>\Qzero$ for every rational $S=p/q$; irrationality is equivalent to
non-integrality of every positive tail difference and to pointwise finite
certificates. &
Achievement-set geometry; exact greedy classification of $1/2$; impossibility
of the upper last producer and of the middle coordinate $-3$.\\[2pt]
\textbf{Conditional / finite} &
First-harmonic and transport consumers; 28 diagonal certificates through
$t=64$. &
Cofinal seam or carry producers imply an infinite support of sum $1/2$;
local bands, finite reset rows, and scoped no-go theorems.\\[2pt]
\textbf{Open producer} &
An unbounded certificate supply, or a cofinal estimate that supplies it. &
Universal direction: all infinite supports.  Counterexample direction:
cofinal false terminals; concretely, the unresolved middle-tail estimate.\\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Reading map.}
Section~\ref{sec:spines} states the two exact proof spines and their unresolved
producers.  Section~\ref{sec:ladder} gives the common Mersenne--Lambert
transform.  Sections~\ref{sec:eb} and~\ref{sec:249} supply the theorem-level
detail for \#257 and \#249.
Sections~\ref{sec:architecture} and~\ref{sec:verify} describe the formalisation
and its verification; the appendices contain auxiliary carry criteria and a
formal-source guide.  Blue declaration links may be ignored on a
first reading.  The separate transport-and-curvature companion develops the
technical affine and first-harmonic material without changing the open status
of \#249.

\medskip
\noindent\textbf{Keywords.} irrationality; Erd\H{o}s--Borwein constant;
Euler totient; Lambert series; achievement sets; binary carry systems; Lean~4.
\qquad
\textbf{MSC 2020.} 11J72 (primary); 11A25, 68V20 (secondary).

\section{The two exact spines}\label{sec:spines}

\subsection{Erd\H{o}s \#249: an exact certificate normal form}
\label{res:249story}

Write
\[
 R_N=\sum_{m\geq1}\frac{\ph(N+m)}{2^m}.
\]
The formal source proves the exact chain
\[
\begin{split}
 S\notin\Q
 &\quad\Longleftrightarrow\quad
   R_{N+h}-R_N\notin\Z
   \quad\text{for every }h>0\text{ and every }N\\
 &\quad\Longleftrightarrow\quad
   \text{for every }h>0,N\text{ there is an }L
   \text{ with }\textsf{certifiedKill}(h,N,L).
\end{split}
\]
Here \(\textsf{certifiedKill}(h,N,L)\) is a finite residue computation with a
rigorous tail allowance.  The second equivalence is completeness of the
certificate: at fixed \(h,N\), some finite depth succeeds exactly when the
tail difference is non-integral.  Conversely, integrality of one
positive-shift difference forces \(S\) to be rational.

This normal form is accompanied by an unconditional theorem: if \(S=p/q\),
then
\[
 q>\Qzero
 =79\,639\,646\,646\,701\,375\,323\,355\,774\,875\,831\,053.
\]
The exact open producer is an unbounded supply of non-integral tail
differences, equivalently of finite certificates.  The first-harmonic
criterion gives one sufficient finite mechanism, but no cofinal harmonic
saving is proved.  Formalised:
\lrefx{LcmConeFlatness.lean}{386}{irrational_totient_series_iff_all_tail_diffs_nonintegral},
\lrefx{LcmConeFlatness.lean}{316}{exists_certifiedKill_iff_tail_diff_notMem_int},
\lrefx{LcmConeFlatness.lean}{399}{irrational_totient_series_iff_pointwise_certificates},
\lloc{CertificateKernel.lean}{18055}, and
\lrefx{FirstHarmonicGap.lean}{128}{exists_certifiedKill_of_first_harmonic_gap}.

\subsection{Erd\H{o}s \#257: the exact half-value counterexample spine}
\label{res:halfstory}

Let
\[
  \mathcal A=\left\{\sum_{n\in A}\frac1{2^n-1}:A\subseteq\N,\ 0\notin A\right\}.
\]
The weights are strictly superincreasing, so the normalized support of a value
is unique.  For \(1/2\), the following are equivalent:
\[
\begin{aligned}
\frac12\in\mathcal A
&\quad\Longleftrightarrow\quad
\text{the greedy skipped support is infinite}\\
&\quad\Longleftrightarrow\quad
\text{false successor terminals occur beyond every bound}.
\end{aligned}
\]
A finite normalized support cannot have value \(1/2\).  Thus membership would
produce an infinite support of rational sum and refute the universal statement
of Erd\H{o}s~\#257.  The required counterexample producer is therefore exact:
cofinally many false terminals, equivalently no last greedy skip.

The strongest current last-skip contradiction is conditional.  At the seam
following a putative last skip, the upper producer is impossible, and among
the middle coordinates \(-3,-2,-1\), the coordinate \(-3\) is impossible.
For \(-2\) and \(-1\), it is sufficient that the producer carry dominate the
complete future divisor-incidence tail.  That inequality is not proved for
the actual greedy orbit.  Formalised:
\lrefx{GreedyAchievementSet.lean}{2514}{half_mem_mersenneAchievementSet_iff_greedySkippedSupport_infinite},
\lrefx{HalfCylinderHalfMembershipClassification.lean}{126}{half_mem_mersenneAchievementSet_iff_unboundedTerminalFalse},
\lrefx{HalfCylinderHalfMembershipClassification.lean}{235}{half_mem_mersenneAchievementSet_iff_no_lastHalfGreedySkip},
\lrefx{HalfCarryReachability.lean}{589}{finite_boolSupport_ne_half},
\lrefx{HalfCylinderFinalMiddleCellEscape.lean}{587}{finalMiddleCell_neg_three_not_last},
and
\lrefx{HalfCylinderLastProducerContradiction.lean}{387}{half_mem_mersenneAchievementSet_of_middleProducerTailEscapeExceptNegThree}.

\paragraph{Prior-art boundary.}
The full-support theorem is Erd\H{o}s's, and Borwein proved a broader
irrationality theorem containing the same Mersenne--Lambert
specialisation~\cite{erdos1948,borwein1992}.  Pairwise-coprime supports are
classical~\cite{erdos1968}, and Luca--Tachiya treat periodic rational
coefficients by another method~\cite{lucatachiya}.  Kova\v{c}--Tao supply the
strict-tail achievement-set context~\cite{kovactao}; Wang--Grau Ribas supply
the closest recorded integral carry recurrence~\cite{wanggrauribas}.
Crandall and Campbell concern adjacent binary-expansion questions
~\cite{crandall,campbell}.  The explicit denominator bound, certificate
equivalence, and last-producer package have no exact antecedent in the
maintained comparison record; that absence is not itself a novelty claim.
The theorem-specific comparison record is maintained with the release.

\section{The Mersenne--Lambert ladder}\label{sec:ladder}

Define the Lambert-type transform of an arithmetic function $f$ by
\[
L(f)\;=\;\sum_{n\ge1}\frac{f(n)}{2^n-1}\;=\;\sum_{m\ge1}\frac{(f*\mathbf 1)(m)}{2^m},
\qquad (f*\mathbf 1)(m)=\sum_{d\mid m} f(d),
\]
the second equality being the standard Lambert rearrangement (one Dirichlet
convolution by the constant $\mathbf 1=\zeta$); Apostol supplies the classical
arithmetic-function and Dirichlet-convolution background~\cite{apostol}; Merca
introduced a modern Lambert-series factorisation theorem~\cite{merca2017}, and
Merca--Schmidt give a synthesis/unification treatment~\cite{mercaschmidt}.
Walking the ladder by convolution gives five values with
rational, irrational, open, or transcendental status.

\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
input $f$ & value & mathematical status & treatment here\\
\midrule
$\mu$ & $L(\mu)=1/2$ & rational & formalised\\
$\ph$ & $L(\ph)=2$ & rational & formalised\\
$\mathbf 1$ & $L(\mathbf 1)=E$ & irrational & formalised\\
$A=\ph*\mu$ & $L(A)=S$ & open (\#249) & identities and bounds formalised\\
$\mathrm{Id}$ & $L(\mathrm{Id})=\sum\sigma(m)/2^m$ & transcendental & cited~\cite{nesterenko}\\
\bottomrule
\end{tabular}
\end{center}

The weight $A=\ph*\mu$ is the primitive-conductor count: $A(p)=p-2$ on primes,
$A\ge0$, and $A*\zeta=\ph$. One convolution by $\zeta$ separates the open
constant $S$ from the rational $2$, and one more separates rational from
transcendental ($\ph*\zeta=\mathrm{Id}$, $\mathrm{Id}*\zeta=\sigma$). So $S$ is
the rung one convolution from the machine-checked irrational $E$: the same series
$\sum(\cdot)/(2^d-1)$, with the constant weight $1$ replaced by the nonnegative
weight $A$.

\begin{proposition}[the rational rungs]
$\sum_{d\ge1}\mu(d)/(2^d-1)=1/2$ and $\sum_{d\ge1}\ph(d)/(2^d-1)=2$, exactly.
\end{proposition}
\noindent Formalised: \lref{MersenneLambertLadder.lean}{587}{tsum_moebius_div_two_pow_sub_one_eq_half}
and \lref{MersenneLambertLadder.lean}{575}{tsum_totient_div_two_pow_sub_one_eq_two}; restated on the
\#249 series shape at \lrefx{CertificateKernel.lean}{18106}{tsum_moebius_div_two_pow_sub_one_eq_half}
and \lrefx{CertificateKernel.lean}{18101}{tsum_totient_div_two_pow_sub_one_eq_two}.

\begin{proposition}[the positive lift, $L(A)=S$]\label{res:lift}
With the primitive-conductor weight $A=\ph*\mu$,
\[
\sum_{d\ge1}\frac{A(d)}{2^d-1}\;=\;\sum_{n\ge1}\frac{\ph(n)}{2^n}\;=\;S.
\]
\end{proposition}
\noindent Formalised: \lref{CertificateKernel.lean}{18117}{tsum_primWeight_div_two_pow_sub_one_eq_totient_series},
with the weight $A$ defined at \lrefx{MersenneLambertLadder.lean}{252}{primWeight}. This
places $S$ in the same $\sum(\cdot)/(2^d-1)$ family as $E$, with a nonnegative weight.

\paragraph{Periodic weights.}\label{sec:periodicweights}
For \(L_b(\gamma)=\sum_{a\ge1}\gamma(a)/(b^a-1)\), the checked endpoints are:
\begin{theorem}[eventually-periodic nonnegative weights]
\label{res:periodicnonnegative}
An eventually-periodic nonnegative rational weight with a positive periodic
value gives an irrational series.
\end{theorem}
\lrefx{CertificateKernel.lean}{12397}{irrational_weightedErdosSeries_eventuallyPeriodic}
This is contained in a broader theorem of Luca--Tachiya~\cite{lucatachiya}.
\begin{theorem}[periodic signed-weight dichotomy]\label{res:periodicsigned}
For a periodic integer weight, the series is irrational or terminating in
base \(b\).
\end{theorem}
\lrefx{CertificateKernel.lean}{13847}{irrational_or_bpow_mul_eq_intCast_intWeightedErdosSeries_periodic}
\begin{proposition}[residue-blind obstruction]\label{res:periodfourobstruction}
For the period-four weight \(1,0,-1,0\), the divisor coefficient vanishes on
\(n\equiv3\pmod4\).
\end{proposition}
\lref{CertificateKernel.lean}{13898}{intWeightedCoeff_periodFourSignWeight_eq_zero_of_mod_four_eq_three}

\paragraph{Finite algebraic boundaries.}\label{sec:primitiveindex}
Put \(a_n=(\ph*\mu)(n)/n\).  If an integer \(D\) clears these coordinates
through \(N\), the exact obstruction is:
\begin{proposition}[primitive-coordinate finite-jet obstruction]
\label{res:primitiveindex}
\(D\) is divisible by an explicit two-tier primorial \(T_N\); hence no fixed
\(D\) clears every \(a_n\).
\end{proposition}
\lref{PrimitiveDeterminantLift.lean}{148}{twoTierPrimorial_dvd_of_integral_jet}
Finite Mersenne-atom determinants likewise recover only their explicit
denominator product, with no additional height surplus.
\lrefx{MersenneTailAtoms.lean}{181}{one_le_abs_weightedMersenneAtom_cleared_det}
These are scoped finite-denominator obstructions, not irrationality criteria
for \(S\).

\begin{proposition}[the M\"obius-square lens]\label{res:lens}
\[
S \;=\; \tfrac12 \;+\; \sum_{d\ge1}\frac{\mu(d)}{(2^d-1)^2}.
\]
\end{proposition}
\noindent Formalised: \lref{CertificateKernel.lean}{18126}{totient_series_eq_half_add_moebius_mersenne_square}.
Writing $T=\sum_{d\ge1}\mu(d)/(2^d-1)^2$, irrationality of $S$ is equivalent to
irrationality of $T$, since $S=\tfrac12+T$. The factor $1/(2^d-1)^2$ is the
probability that $d$ divides two independent fair-coin waiting times
(Section~\ref{sec:geom}). The signed constant, the coprime-pair mass, and the
tail differences of Section~\ref{sec:reduction} are three readings of the same
quantity.

\paragraph{Squared Lambert calculus.}\label{sec:squared}
For \(L_2(f)=\sum_{d\ge1}f(d)/(2^d-1)^2\), the denominator is the probability
that \(d\) divides two independent geometric waiting times.  The formal
transfer theorem therefore gives a divisor calculus for their gcd.
\lref{GcdMomentCalculus.lean}{105}{tsum_lambert_linear_weight_sq_pure}
\begin{proposition}[squared Lambert ladder]\label{res:gcdmoment}
At \(r=1/2\), the three rows \(f=\mu,\mathbf 1,\ph\) yield respectively
\[
 L_2(\mu)=S-\tfrac12,\qquad
 L_2(\mathbf 1)=\sum_{n\ge1}\frac{\sigma(n)-\tau(n)}{2^n},\qquad
 L_2(\ph)=\sum_{n\ge1}\frac{P(n)-n}{2^n},
\]
where \(P\) is Pillai's gcd-sum function.
\end{proposition}
\lrefx{GcdMomentCalculus.lean}{235}{tsum_totient_div_mersenne_sq_eq_gcd_moment_series}
Postelmans--Van Assche imply irrationality of the middle row
~\cite{postelmansvanassche}; this does not transfer to the M\"obius row, which
is exactly the open constant \(S-\tfrac12\).  The further coprime-pair and
Stern--Brocot identities are retained in the living reference rather than
given equal narrative weight here.

\section{Erd\H{o}s--Borwein-type irrationality (the \#257 direction)}\label{sec:eb}

\subsection{Full support: the Erd\H{o}s--Borwein constant, every base}

\begin{theorem}[full-support irrationality]\label{res:full}
For every integer $b\ge2$, the series $\displaystyle\sum_{n\ge1}\frac{1}{b^n-1}$ is
irrational. In particular the Erd\H{o}s--Borwein constant $E=\sum_{n\ge1}1/(2^n-1)$
is irrational.
\end{theorem}
\noindent Formalised: \lref{CertificateKernel.lean}{8000}{irrational_erdosSum_full_support}
(all bases) and \lref{CertificateKernel.lean}{8007}{irrational_erdosBorwein_series}
(the base-$2$ corollary). This is Erd\H{o}s's 1948 theorem~\cite{erdos1948}. In the ladder of
Section~\ref{sec:ladder} it is the rung $L(\mathbf 1)$, and it is the fixed
machine-checked point one convolution away from the open $S$.

The proof begins with the Lambert identity
\[
 \sum_{n\geq 1}\frac{1}{b^n-1}
   =\sum_{m\geq 1}\frac{\tau(m)}{b^m},
\]
where $\tau(m)$ is the number of positive divisors of $m$
(\lrefx{CertificateKernel.lean}{5865}{erdosSum_full_support_eq_tsum_divisor_count}).
If the right-hand side were rational with reduced denominator $q$, then a
non-integral integer multiple could not lie at distance less than $1/q$ from an
integer.  It is therefore enough, for every $q$, to construct $N$ such that
$b^N\sum_m\tau(m)b^{-m}$ is non-integral but lies within $1/q$ of an integer.
This elementary near-integer criterion is formalised at
\lrefx{CertificateKernel.lean}{5792}{irrational_of_int_mul_near_int}.

The construction of $N$ is divided into three finite parts.  First, a bounded
Bertrand argument supplies disjoint prime blocks, and the Chinese remainder
theorem chooses an arithmetic progression on which
\[
        b^r\mid \tau(N+r) \qquad (1\leq r\leq K).
\]
These divisibilities make the first $K$ terms of the shifted tail integral.
The multiplicativity step that converts prescribed prime valuations into the
displayed divisibility is
\lrefx{CertificateKernel.lean}{6177}{bpow_dvd_card_divisors_of_exact_prime_block}.
Second, divisor pairing bounds the average of the weighted middle window
\[
   \sum_{K<r\leq L}\tau(N+r)b^{L-r}
\]
along that progression.  A pigeonhole choice then selects one translate with a
small middle contribution; see
\lrefx{CertificateKernel.lean}{7246}{exists_first_block_progression_with_selected_weighted_middle}.
Third, the elementary estimate $\tau(n)\leq n$ controls the infinite tail after
$L$.  The parameters are chosen so that the middle and far-tail bounds together
are smaller than $1/q$; the shifted tail is strictly positive because every
divisor count is positive, so the resulting dilation is not an integer.

Thus the analytic series is consumed only at the Lambert identity and the
geometric tail estimate.  The remaining proof is a finite CRT construction,
divisor counting, and explicit parameter arithmetic.  The final certificate
existence theorem holds for every $b\geq2$ and every precision $q$
(\lrefx{CertificateKernel.lean}{7859}{exists_weighted_block_certificates}); its
composition with the near-integer criterion is the theorem cited above.  This
is a formalisation of the classical full-support theorem, not a new
irrationality result.

\subsection{Named infinite-support cases}

Beyond full support, the development proves irrationality for a family of infinite
supports $A$, at every base $b\ge2$. Each is a formalised case of the \#257
statement family. None is the universal statement, and none is claimed as new
mathematics.

\begin{theorem}[support-class family]\label{res:support}
For every integer $b\ge2$, the series $\sum_{n\in A} 1/(b^n-1)$ is irrational for
each of the following supports $A$:
\begin{enumerate}[label=\textup{(\alph*)},leftmargin=2em]
\item factorial support $A=\{n! : n\ge1\}$, giving $\sum 1/(b^{n!}-1)$;
\item power-of-two support $A=\{2^n : n\ge0\}$, giving $\sum 1/(b^{2^n}-1)$;
\item multiples $A=d\N$;
\item pairwise-coprime supports with summable reciprocals (Erd\H{o}s's
condition~\cite{erdos1968});
\item eventually-periodic supports, residue classes, and the odd numbers.
\end{enumerate}
\end{theorem}
\noindent These sort into three groups. The classical case is (d): infinite
pairwise-coprime $A$ with $\sum_{a\in A}1/a<\infty$, formalised from Erd\H{o}s's
condition. The lcm-gap named cases are (a) and (b), whose support gaps outgrow the
running $\lcm$; the structured families are (c) and (e).\footnote{Luca--Tachiya
prove the broader eventual-periodic rational-coefficient theorem; its indicator
specialisation directly contains case (e)~\cite{lucatachiya}. Their Chowla--Erd\H{o}s/
large-modulus-prime proof is distinct from the periodic-divisor certificate route
formalised here, and the full mixed-sign coefficient theorem is not claimed here.}
Formalised: (a)
\lrefx{CertificateKernel.lean}{5707}{irrational_erdosSum_factorial_support};
(b) \lrefx{CertificateKernel.lean}{5731}{irrational_erdosSum_two_pow_support};
(c) \lrefx{CertificateKernel.lean}{8775}{irrational_erdosSupportSeries_multiples};
(d) \lrefx{CertificateKernel.lean}{10448}{irrational_erdosSupportSeries_pairwise_coprime};
(e) \lrefx{CertificateKernel.lean}{11276}{irrational_erdosSupportSeries_eventuallyPeriodic},
\lrefx{CertificateKernel.lean}{11344}{irrational_erdosSupportSeries_residueClass},
\lrefx{CertificateKernel.lean}{11358}{irrational_erdosSupportSeries_odd}. The two
base-$2$ instances that name the problem directly are
\lrefx{CertificateKernel.lean}{5754}{erdos257_family_factorial_instance} and
\lrefx{CertificateKernel.lean}{5762}{erdos257_family_two_pow_instance}.

The prime support is now known unconditionally: Tao and Ter\"av\"ainen proved
that
\[
  \sum_p\frac1{2^p-1}=\sum_{n\ge1}\frac{\omega(n)}{2^n}
\]
is irrational~\cite{taoteravainen}. Their proof uses quantitative two-point
correlations of multiplicative functions and is not formalised here. This
settles an important special case, while the universal support statement
remains open.

The factorial and power-of-two cases are worth a word because they are not
reachable by a Liouville shortcut: the power-of-two gap $2^k-\lcm$ stays within a
bounded ratio of the $\lcm$, so the terms do not decay fast enough for a
transcendence-measure argument, and a uniform criterion is needed. The mechanism
is \emph{period noncollapse}: a
prime-valuation-deficit witness forces the multiplicative order of the base modulo
the relevant modulus not to collapse, which rules out the long repeated digit
blocks a rational value would require.

\subsection{The base-$2$ achievement set}

Let
\[
  \mathcal A=\left\{\sum_{n\in A}\frac1{2^n-1}:A\subseteq\N,\ 0\notin A\right\}.
\]
The formal development verifies that $\mathcal A$ is compact, perfect,
totally disconnected, nowhere dense, and has Lebesgue measure one.  Its
strictly superincreasing weights make normalized support coding injective,
and membership is equivalent to exact greedy survival at every level.
Formalised:
\lrefx{GreedyAchievementSet.lean}{996}{volume_mersenneAchievementSet},
\lrefx{GreedyAchievementSet.lean}{1445}{mem_mersenneAchievementSet_iff_greedy_survival},
\lrefx{GreedyAchievementSet.lean}{1527}{positiveMersenneSupportValue_injective_normalized},
\lrefx{GreedyAchievementSet.lean}{1620}{perfect_mersenneAchievementSet},
\lrefx{GreedyAchievementSet.lean}{1636}{isTotallyDisconnected_mersenneAchievementSet},
and
\lref{GreedyAchievementSet.lean}{1645}{isNowhereDense_mersenneAchievementSet}.
Kova\v{c}--Tao supply the strict-tail Cantor-set context~\cite{kovactao}; no
novelty claim is made for that geometry.  The exact rational execution and
finite non-membership certificates used for computation are described in
Appendix~\ref{sec:carry}.

The exact half-value spine and its open producer are stated in
Section~\ref{res:halfstory}.  The rest of this section develops sufficient
criteria and finite case analyses around the remaining middle-tail estimate;
it may be skipped on a first reading.

\subsection{A band formula for the final skip between two takes}
\label{res:twothirdsband}

One local configuration admits a simple exact description in full generality.
Suppose a take at rank $b$ is followed by the next take at $c>b+1$.  Put
$q=2^b-1$, $m=2^{c-1}$, and write the residual immediately before the first
take as $1/R$.  When $0<R<q$, the final intervening skip is dyadically unsafe
if and only if
\[
  \frac{q(m-1)}{q+m-1}<R<\frac{qm}{q+m}.
\]
The interval has exact width
\[
  \frac{q^2}{(q+m)(q+m-1)}.
\]
Formalised:
\lrefx{HalfGreedyTwoThirdsBand.lean}{88}{postTakeUnsafeAt_iff_band}
and
\lref{HalfGreedyTwoThirdsBand.lean}{104}{band_width_general}.

For a single intervening skip, $c=b+2$ and $m=2q+2$.  The band becomes
\[
  \frac{q(2q+1)}{3q+1}<R<\frac{2q(q+1)}{3q+2},
\]
with width $q^2/((3q+1)(3q+2))<1/9$, entirely between $2q/3$ and
$2q/3+2/9$.  No integral reciprocal lies in this window.  In the
corresponding positive odd integer formulation, with residual $p/(2D)$,
membership in the cleared band forces $p\ge7$.  Formalised:
\lrefx{HalfGreedyTwoThirdsBand.lean}{127}{singletonUnsafe_iff_twoThirdsBand},
\lrefx{HalfGreedyTwoThirdsBand.lean}{154}{band_width},
\lrefx{HalfGreedyTwoThirdsBand.lean}{163}{band_width_lt_ninth},
\lrefx{HalfGreedyTwoThirdsBand.lean}{185}{not_twoThirdsBand_of_int},
and
\lref{HalfGreedyTwoThirdsBand.lean}{231}{seven_le_of_intBand_odd}.

These are localizers, not orbit-level eliminations.  They do not show that
the greedy orbit of $1/2$ avoids any such band, and they do not identify the
two surviving last-producer coordinates $-2,-1$ with dyadically unsafe skips.

\subsection{Conditional support and reachability producers}

\subsubsection{A conditional orthogonal-petal criterion}\label{res:sunflower}

Let $A$ be a finite-core orthogonal-petal bouquet: outside a finite frame, its
elements have the form $c_i p_i$, where the cores $c_i$ divide one fixed
integer and the nontrivial petals $p_i$ are pairwise coprime to that frame and
to one another, with summable reciprocal mass.  The formal construction also
contains an explicit infinite example with alternating cores $2$ and $3$ and
squared prime petals; every member of this support has exactly three prime
factors with multiplicity.

For such a bouquet, the remaining input is an exact tail-selection predicate:
at every positive binary scale it chooses a carry slot with a normalized tail
at most $16$.  Lean proves that this predicate supplies the needed forced-carry
certificates and hence irrationality of the base-$2$ support series.  The
result is
\lref{SupportSunflowerDichotomy.lean}{540}{irrational_erdosSupportSeries_of_orthogonalPetalBouquet},
with the carry bridge at
\lrefx{SupportSunflowerDichotomy.lean}{531}{sunflowerForcedCarrySupply_of_orthogonalPetalBouquet}.
The tail-selection predicate itself is not proved here.  Thus this is a
conditional reduction for one structured support family, not a further
unconditional case of Erd\H{o}s~\#257.

\subsubsection{Composite dilation defects}\label{res:compositedefect}

For a support divisor count $f_A(x)=\#\{d\in A:d\mid x\}$, multiplication by
a composite support element $a$ has an exact correction term.  Besides the
distinguished divisor $a$, the new divisors are counted by a finite defect
which excludes divisors already present at $x$.  The formal identity is
\lref{CompositeDilationDefect.lean}{30}{supportCoeff_mul_eq_add_defect}.
For a finite-core orthogonal-petal bouquet, the defect at a ray is bounded by
the finite exceptional frame plus the divisor count of the petal support:
\lrefx{CompositeDilationDefect.lean}{151}{compositeDilationDefect_le_exceptional_add_petalCoeff}.
This records the local correction needed when dilation is not prime.  It does
not prove a correlation estimate, the sunflower tail selector, or an
irrationality statement.

\subsubsection{Mixed prime-power layers}\label{res:mixedlayer}

The one-prime-power layer has a two-signature extension.  Exact $p$-adic and
$q$-adic layer operators commute, and, for distinct primes with the residual
argument coprime to $pq$, their composite extracts the iterated exact-valuation
pullback of the support coefficient.  This is formalised by
\lref{MaximalOmegaLayer.lean}{39}{mixedPrimePowerLayerTwo_supportCoeffInt}.
The singleton support $\{12\}$ gives a checked $(2^2,3)$ example with value
one at residual argument $1$:
\lrefx{MaximalOmegaLayer.lean}{65}{mixedPrimePowerLayerTwo_twelve_fixture}.
These are finite coefficient identities.  They do not provide a decimation
transport, a bounded-$\Omega$ conclusion, or an irrationality theorem.

\subsubsection{Half-trapping return carries}\label{res:halftrapping}

If rational-linear channels vanish on every relation killed by one scalar
evaluation, their evaluation matrix has rank at most one; every nontrivial
square minor therefore vanishes.  The all-rank determinant statement is
\lref{HalfTrappingReturnCarry.lean}{42}{relationInvariantLinearChannels_det_eq_zero}.
For two reverse carries with a common coefficient word, a $1/0$ seam, and a
common following overlap, the terminal difference is a power of two times an
odd integer.  A shared nonnegative Archimedean bound then gives the matching
dyadic spacing inequality, formalised at
\lrefx{HalfTrappingReturnCarry.lean}{191}{overlappingReverseCarryWords_twoPow_le_realBound}.
The argument does not produce reverse-carry presentations or an earlier
overlapping return; it supplies only the algebraic core of that conditional
criterion.

\subsubsection{Half-carry reachability}\label{res:halfcarryreachability}

There is also a finite-prefix criterion for the target $1/2$.  At an unresolved
even seam, the scalar transition reaches every integer terminal carry except
the single value $2\delta-c$; this is the exact equivalence
\lref{HalfCarryReachability.lean}{235}{evenSeamReachable_iff}.  A canonical
description of such a seam at every sufficiently large even depth is an
unproved hypothesis.  Under that hypothesis, the finite admissible words are
cofinal, by
\lrefx{HalfCarryReachability.lean}{999}{cofinalAdmissibility_of_canonicalEvenSeamSupply},
and compactness then produces an infinite support with sum exactly $1/2$:
\lrefx{HalfCarryReachability.lean}{1027}{exists_infinite_support_half_of_canonicalEvenSeamSupply}.
Thus the formal result is a conditional reduction.  It does not establish the
canonical seam supply and does not prove Erd\H{o}s~\#257.

\paragraph{Cofinal strip returns.}  The all-level strip hypothesis can be
weakened for the canonical greedy half carry.  Its uncentred, binary-scaled
form is antitone.  Consequently, if it returns to the square-root strip
cofinally, then one sufficiently late return bounds every later scaled carry;
the strip envelope divided by $2^N$ tends to zero.  The resulting tempered
centred carry gives an infinite greedy support with support-series value
$1/2$.

\begin{proposition}[cofinal strip-return closure]\label{res:cofinalstripreturn}
If the canonical greedy half carry returns cofinally to the stated
square-root strip, then $1/2$ belongs to the Mersenne achievement set.
\end{proposition}
\noindent Formalised:
\lref{CofinalStripReturn.lean}{84}{greedy_integerHalfCarry_scaled_tendsto_zero_of_cofinalStripReturn}
and
\lref{CofinalStripReturn.lean}{157}{half_mem_mersenneAchievementSet_of_cofinalStripReturn}.
The cofinal return condition itself is not established.  It is weaker than a
uniform strip bound and still does not give an unconditional membership
theorem.

\paragraph{Terminal-only cofinal approximation.}  Coherence can be removed
entirely from a further version of this reduction.  At depth $M$, take only a
normalised finite word whose terminal half carry lies in the square-root
strip.  Its finite support then has support-series value within
\[
 \frac{4M+12}{2^M}
\]
of $1/2$.  If such terminal witnesses occur cofinally, these finite values
converge to $1/2$; closedness of the Mersenne achievement set supplies the
limit.  No control of the earlier rows, and no compatibility between the
witnesses, is used.

\begin{proposition}[terminal-only cofinal closure]\label{res:terminalonlycofinal}
Cofinal normalised terminal witnesses in the square-root strip imply that
$1/2$ belongs to the Mersenne achievement set, and hence that an infinite
support has support-series value $1/2$.
\end{proposition}
\noindent Formalised:
\lref{TerminalOnlyCofinal.lean}{68}{dist_half_erdosSupportSeries_wordSupport_le_of_terminalOnlyWitness},
\lref{TerminalOnlyCofinal.lean}{134}{half_mem_mersenneAchievementSet_of_cofinalTerminalOnlyStrip},
and
\lref{TerminalOnlyCofinal.lean}{192}{exists_infinite_support_half_of_cofinalTerminalOnlyStrip}.
The cofinal terminal-witness supply is not proved.  In particular, the
proposition does not establish membership of $1/2$ unconditionally.

\paragraph{Scaled terminal vanishing.}  The square-root strip is stronger
than the analytic condition used by the terminal-only argument.  Let the
depths of normalised finite words tend to infinity.  If their terminal carries,
after division by the corresponding power of $2$, tend to zero, then the
universal coefficient-tail error also tends to zero.  The finite support
values therefore converge to $1/2$, and closedness of the achievement set
again supplies the limit.  The earlier terminal-only strip condition implies
this weaker sequential hypothesis.

\begin{proposition}[scaled terminal-vanishing closure]\label{res:terminalscaledvanishing}
A normalised terminal-word sequence with depths tending to infinity and
vanishing binary-scaled terminal carries yields an infinite support with
support-series value $1/2$.
\end{proposition}
\noindent Formalised:
\lref{TerminalOnlyScaledVanishing.lean}{24}{dist_half_erdosSupportSeries_wordSupport_le_carry_add_tail},
\lref{TerminalOnlyScaledVanishing.lean}{165}{half_mem_mersenneAchievementSet_of_terminalScaledVanishing},
and
\lref{TerminalOnlyScaledVanishing.lean}{221}{exists_infinite_support_half_of_terminalScaledVanishing}.
No such scaled-vanishing sequence is constructed here; the result is a
conditional reduction, not an unconditional membership theorem.

\subsection{Window, seam, rewind, and compactness sockets}

\subsubsection{Finite shadows of the half cylinder}\label{res:halfcylindershadow}

For an admissible finite word, the residual from $1/2$ is given exactly by its
terminal half carry minus the future coefficient tail, scaled by a power of
two; see
\lref{HalfCylinderFiniteShadow.lean}{55}{halfStripAdmissible_residual_eq}.
At a first seam, avoiding the integer strip is equivalent to full scalar
reachability of that strip:
\lrefx{HalfCylinderFiniteShadow.lean}{807}{evenSeam_fullIntegerStrip_iff_escape}.
The rank-three greedy example is raw dyadically safe while its seam has an
in-strip hole, so raw safety alone does not imply this escape.  Finally, an
explicit frozen-margin producer implies $1/2$ belongs to the Mersenne
achievement set, by
\lrefx{HalfCylinderFiniteShadow.lean}{1256}{half_mem_mersenneAchievementSet_of_governedFrozenMarginProducer}.
That producer is an unproved hypothesis; this is a conditional reduction, not
a proof of Erd\H{o}s~\#257.

\subsubsection{Fixed-coefficient rewind}\label{res:fixedcoeffrewind}

For a finite list of common coefficients, the canonical inverse-parent map has
an exact closed form with a power-of-two denominator:
\lref{HalfCarryCeilParentContraction.lean}{49}{rewind_eq_closedForm}.
An interval of width at most that denominator contracts either to one ancestor
or to two adjacent ancestors.  The singleton alternative is equivalent to the
explicit dyadic phase inequality, while the other alternative is a seam pair;
see
\lrefx{HalfCarryCeilParentContraction.lean}{180}{rewind_singleton_or_isSeamPair}.
Thus width alone does not justify a singleton conclusion.  This is a finite
rewind lemma, and supplies neither a seam exclusion nor an Erd\H{o}s~\#257
conclusion.

\subsubsection{Selected half-carry windows}\label{res:selectedhalfwindow}

The selected-window construction keeps an explicit admissible word for each
terminal carry in a finite protected interval.  A supply of such windows at
all sufficiently large levels gives cofinal admissibility and hence an
infinite support with sum $1/2$:
\lref{HalfCarrySelectedWindow.lean}{831}{cofinalAdmissibility_of_selectedHalfWindows}
and
\lrefx{HalfCarrySelectedWindow.lean}{864}{exists_infinite_support_half_of_selectedHalfWindows}.
The unresolved input is the coherent history at each step: either a common
next-row divisor cylinder, or the canonical adjacent seam.  The module makes
this input explicit; it does not construct the window supply and does not
prove Erd\H{o}s~\#257.

\subsubsection{Realized rewind phase}\label{res:rewindphase}

For a realized fixed-coefficient history, the dyadic rewind phase is the
complementary binary suffix numeral:
\lref{HalfCarryRewindPhase.lean}{39}{phase_eq_denom_sub_one_sub_suffixNumeral}.
Consequently the singleton criterion is exactly the inequality that this
suffix numeral covers the rest of the terminal interval,
formalised by
\lrefx{HalfCarryRewindPhase.lean}{65}{rewind_interval_singleton_iff_suffixNumeral_covers}.
This converts the phase condition into support-history coordinates, but does
not prove the required coverage inequality globally.

\subsubsection{Localized protected seams}\label{res:protectedseam}

The compactness argument needs only one admissible word at cofinally many
depths.  At a localized one-hole seam, one of the two protected carries $3$
and $4$ survives; therefore a cofinal supply of realized protected seams
implies cofinal admissibility,
\lref{HalfCarryProtectedSeamConsumer.lean}{31}{cofinalAdmissibility_of_cofinalProtectedEvenSeamRealization},
and hence an infinite support with sum $1/2$,
\lrefx{HalfCarryProtectedSeamConsumer.lean}{53}{exists_infinite_support_half_of_cofinalProtectedEvenSeamRealization}.
The cofinal realization supply is not proved.

\subsubsection{Combining the two cofinal half-carry criteria}\label{res:cofinalwindowseam}

At each requested scale, the public mixed producer permits either a selected
protected window or a localized realized seam.  Each branch supplies the one
admissible word needed for compactness, so this cofinal disjunction implies
an infinite support with sum $1/2$:
\lref{HalfCarryCofinalWindowOrSeamConsumer.lean}{32}{cofinalAdmissibility_of_cofinalSelectedWindowOrProtectedSeam}
and
\lrefx{HalfCarryCofinalWindowOrSeamConsumer.lean}{58}{exists_infinite_support_half_of_cofinalSelectedWindowOrProtectedSeam}.
The mixed cofinal producer remains an unproved hypothesis.

\subsubsection{Integer-greedy first-wrap reduction}\label{res:integergreedy}

The first-wrap seam has exact truncated Mersenne weights with a quantitative
gap-dominance property.  For these weights, a small positive subset-sum defect
exists exactly when the descending integer-greedy remainder is small and
positive; this is formalised by
\lref{HalfCylinderIntegerGreedy.lean}{654}{seam_smallPositiveDefect_iff_integerGreedyRemainder}.
The remaining arithmetic input is an unbounded lower bound for that
deterministic remainder sequence.  No such lower bound is proved here.

\subsubsection{Concrete seam adapter}\label{res:concreteseam}

The finite seam-word adapter realizes the abstract adjacent cut by concrete
Boolean words and identifies its next word with the integer-greedy choice:
\lref{HalfCylinderConcreteSeamAdapter.lean}{667}{seamCutNextWord_eq_greedyWord_succ}.
It implements a finite recurrence, not the unbounded remainder estimate.

\subsection{Fatal-gap and last-producer spine}

\subsubsection{Fatal gaps and eventual right tails}\label{res:fatalright}

For the concrete integer seam words, eventual right extension is equivalent
to nonmembership of $1/2$ in the Mersenne achievement set:
\lref{HalfCylinderFatalGapRightTail.lean}{787}{seamGreedyEventuallyRight_iff_half_not_mem}.
The same condition is equivalent to the existence of a finite fatal half gap,
by
\lrefx{HalfCylinderFatalGapRightTail.lean}{781}{seamGreedyEventuallyRight_iff_existsFatalHalfGap},
and also to a final skipped exponent in the real greedy orbit,
by
\lrefx{HalfCylinderFatalGapRightTail.lean}{802}{seamGreedyEventuallyRight_iff_exists_isLastHalfGreedySkip}.
Thus this equivalence gives an exact classification of the nonmembership branch.
It neither decides whether $1/2$ belongs to the achievement set nor proves
the universal Erd\H{o}s~\#257 statement.

This is the nonmembership half of the argument summarized earlier: a last
greedy skip is equivalent to eventual right extension.  The following
paragraph records the first exceptional middle case that can be ruled out.

\paragraph{First final-middle cell.}  The first exceptional middle cell has a
separate obstruction: if the orbit extends right after the last false rank,
then the cell with producer carry $-3$ makes the lazy cofinite tail strictly
smaller than $1/2$ while its centred carry becomes negative.  This is
impossible by the analytic nonnegativity identity,
\lrefx{HalfCylinderFinalMiddleCellEscape.lean}{587}{finalMiddleCell_neg_three_not_last}.
It removes that one conditional terminal configuration; it does not decide
the right-tail branch or the remaining two middle cells.

The remaining global reduction is now precise.  If every genuine middle
producer other than that discharged cell has producer carry larger than its
complete future divisor-incidence tail, then an eventual right seam is
impossible and $1/2$ belongs to the Mersenne achievement set:
\lrefx{HalfCylinderLastProducerContradiction.lean}{387}{half_mem_mersenneAchievementSet_of_middleProducerTailEscapeExceptNegThree}.
The stronger square-root lower-bound hypothesis is also sufficient,
\lrefx{HalfCylinderLastProducerContradiction.lean}{409}{half_mem_mersenneAchievementSet_of_middleProducerSqrtEscape}.
Neither hypothesis is supplied here.

For finite seam prefixes the remaining tail bound sharpens further: the
future divisor-incidence tail is bounded by the cardinality of the prefix.
Thus the explicit row-scale inequality
\(|c|+\mathrm{belowPulse}+5<4\,\mathrm{remainder}\) is sufficient for
the same half-membership conclusion,
\lrefx{HalfCylinderMiddleCarryLowerBound.lean}{1503}{half_mem_mersenneAchievementSet_of_middleProducerCardEscape};
it is enough in turn that every late middle remainder is at least its row,
\lrefx{HalfCylinderMiddleCarryLowerBound.lean}{1510}{half_mem_mersenneAchievementSet_of_middleProducerRowEscape}.
These are exact conditional reductions, not proofs of the row bound.

\paragraph{Finite upper-reset band certificates.}  The companion upper
branch has a kernel-checked finite certificate: for every upper transition
from rows $13$ through $30$, every subsequent dyadic danger band is avoided.
The certificate shares each exact successor remainder across all its band
indices,
\lrefx{HalfCylinderUpperResetBandCertificates.lean}{78}{seamUpperResetDyadicBandEscape_through_thirty}.
It is a bounded producer check and does not supply the cofinal band-escape
hypothesis.

\subsubsection{Positive half-membership classification}\label{res:halfmembership}

The complementary branch has an equally exact seam formulation.  Membership
of $1/2$ in the Mersenne achievement set is equivalent to false successor
terminal bits beyond every bound,
\lref{HalfCylinderHalfMembershipClassification.lean}{126}{half_mem_mersenneAchievementSet_iff_unboundedTerminalFalse};
equivalently, it is equivalent to a cofinal sequence of such false terminals,
\lrefx{HalfCylinderHalfMembershipClassification.lean}{203}{half_mem_mersenneAchievementSet_iff_cofinalTerminalFalse}.
It is also equivalent to a cofinal integer-seam sequence with skipped ranks
tending to infinity,
\lrefx{HalfCylinderHalfMembershipClassification.lean}{213}{half_mem_mersenneAchievementSet_iff_exists_unboundedSkippedRanksAlong}.
These equivalences isolate the positive arithmetic producer; they do not
establish that the producer exists.

\subsubsection{Largest skipped ranks}\label{res:largestskip}

At a named largest skipped rank $d$, the lower seam word and its adjacent
upper competitor have an exact late-range gap.  In division-free form, the
identity is
\[
  3L+\bigl(3\cdot2^{s+1}+2\cdot4^{s-d}+4\bigr)=3U
\]
whenever $2s<3d$; this is formalised at
\lref{HalfCylinderLargestSkipGap.lean}{259}{three_mul_largestSkipLowerWeight_add_exactLateGap_eq_upperWeight}.
The upper and middle successor branches create a terminal largest skipped
rank, while the right branch preserves an existing one:
\lrefx{HalfCylinderLargestSkipGap.lean}{328}{seamGreedyWord_succ_isLargestFalseRank_terminal_of_upperOrMiddle}
and
\lrefx{HalfCylinderLargestSkipGap.lean}{339}{IsLargestFalseRank.seamGreedyWord_succ_of_rightBranch}.
The first-crossing branch producer required to use this identity globally is
not proved.

\subsubsection{Boundary-pulse normalization}\label{res:boundarypulse}

Above a late largest skipped rank, every filled suffix rank has zero row
pulse, so the word pulse is carried entirely by the lower prefix:
\lref{HalfCylinderBoundaryPulse.lean}{85}{exists_lowerPrefix_wordPulse_eq_of_largestFalse_late}.
At the first crossing of the two-thirds boundary there are exactly two cases,
$3d=2s+1$ and $3d=2s+2$, and the skipped rank has pulse respectively two or
one:
\lrefx{HalfCylinderBoundaryPulse.lean}{150}{rowPulse_boundary_of_late_firstCrossing}.
This supplies the incidence normalization for a producer estimate; it does
not establish that estimate.

\subsubsection{Largest-skip induction}\label{res:largestskipinduction}

Assume that at the first crossing a late largest skipped rank either
remains late at the next row or forces an upper-or-middle successor branch.
This hypothesis propagates a late largest skipped rank from row
$14$ onward,
\lref{HalfCylinderLargestSkipInduction.lean}{90}{LargestSkipLateStepSocket.largestSkipLateAt_succ},
and hence yields cofinal skipped ranks and membership of $1/2$ in the Mersenne
achievement set,
\lrefx{HalfCylinderLargestSkipInduction.lean}{167}{half_mem_mersenneAchievementSet_of_largestSkipLateStepSocket}.
The first-crossing hypothesis itself remains unproved.

\subsubsection{The fixed-tail survival criterion}\label{res:fixedtailsocket}

A skipped rank is final exactly when its post-decision remainder lies strictly
above the complete remaining Mersenne tail:
\lref{HalfCylinderFixedTailSocket.lean}{22}{isLastHalfGreedySkip_iff_skip_and_fatal}.
Consequently membership of $1/2$ is equivalent to the pointwise assertion that
every actual skipped rank has remainder at most its full future tail,
\lrefx{HalfCylinderFixedTailSocket.lean}{73}{half_mem_iff_every_actual_skip_survives}.
This is an exact local reformulation; the tail-survival inequality is not
proved.

\subsubsection{The producer-carry identity}\label{res:producercarry}

For a producer-aligned finite support, the residual from $1/2$ is exactly the
integer producer carry minus the complete future divisor-incidence tail,
scaled by $2^{-(2d+2)}$:
\lref{HalfCylinderProducerCarrySocket.lean}{27}{producerCarry_residual_identity}.
Thus the carry exceeds that tail exactly when the support series lies below
$1/2$, and a negative carry places it above $1/2$;
\lrefx{HalfCylinderProducerCarrySocket.lean}{48}{producerCarry_gt_tail_iff_supportSeries_lt_half}
and
\lrefx{HalfCylinderProducerCarrySocket.lean}{145}{producerCarry_neg_forces_half_lt_supportSeries}.
An upper carry at most $-8$ even gives a full Mersenne-gap margin,
\lrefx{HalfCylinderProducerCarrySocket.lean}{190}{producerCarry_le_neg_eight_forces_gap_margin}.
These are local criteria, not a proof that the required carries occur.

\subsubsection{Quarter-band endpoint cells}\label{res:quarterband}

For the adjacent seam recurrence, failure of the desired next-row upper bound
on the middle branch is exactly one multiple-of-four pulse cell,
\lref{HalfCylinderQuarterBandEndpoints.lean}{153}{seamAdjacentCut_middleNextUpperFailure_iff_pulseCell}.
On the terminal-reset branch, the same failure is exactly a constrained
$3\cdot2^{s-1}+k$ pulse cell,
\lrefx{HalfCylinderQuarterBandEndpoints.lean}{175}{seamAdjacentCut_rightNextUpperFailure_iff_pulseCell}.
These are arithmetic normal forms for the exceptional cells; no avoidance
statement for the concrete seam orbit is proved.

\subsubsection{Reset-deficit escape}\label{res:resetdeficit}

At a late largest skipped rank, a right successor branch pins the integer
greedy remainder between an explicit quarter threshold and carry threshold:
\lref{HalfCylinderResetDeficitEscape.lean}{200}{rightBranch_remainder_window}.
Along such a branch the remainder satisfies an exact affine recurrence,
\lrefx{HalfCylinderResetDeficitEscape.lean}{218}{rightBranch_remainder_succ_eq}.
Moreover, a doubled-scale deficit forces an upper-or-middle branch within the
corresponding number of rows,
\lrefx{HalfCylinderResetDeficitEscape.lean}{261}{upperOrMiddle_within_of_two_pow_deficit}.
The remaining input is the deterministic anti-concentration needed to exclude
the pinned right-branch window globally.

\subsubsection{Seam-to-producer carry alignment}\label{res:seamproducer}

At a first-wrap seam, the producer-aligned carry of the left boundary support
is four times the shifted seam hole, less the explicit paired incidence pulse
at the next two rows:
\lref{HalfCylinderSeamProducerAlignment.lean}{24}{producerCarry_insert_eq_four_mul_shiftedHole_sub_pulse}.
This is an exact coordinate change between the seam and producer pictures; it
does not supply a sign or size estimate for that carry.

\subsection{Finite deposits, identities, and counterexamples}

\subsubsection{A computed selected-window base}\label{res:selectedwindowbase}

The selected half-carry construction has a checked depth-$18$ base: twelve
Boolean words cover every terminal carry in the strip $[1,12]$, with all
intermediate carries in the corresponding half strip.  The reflected table is
formalised at
\lref{HalfCarrySelectedWindowBase.lean}{79}{depth18TableCheck_eq_true};
the resulting selected window is defined at
\lrefx{HalfCarrySelectedWindowBase.lean}{129}{depth18SelectedHalfWindow}.
All twelve words share the same restriction at depth $13$, giving the first
next-row divisor-agreement certificate,
\lrefx{HalfCarrySelectedWindowBase.lean}{172}{depth18SelectedHalfWindow_nextRowDivisorAgreement}.
This finite base does not provide selected windows at unbounded depths.

\subsubsection{Operational rewind-seam bridge}\label{res:rewindseambridge}

For a selected window with a rewind seam, a two-value coefficient profile on
the adjacent base ancestors induces the live next-row profile,
\lref{RewindSeamOperationalBridge.lean}{140}{nextRow_profile_of_rewindSeam}.
Under the stated strip and buffer inequalities, that profile realizes a
protected one-hole seam at the next row,
\lrefx{RewindSeamOperationalBridge.lean}{217}{protectedEvenSeamRealizedAt_succ_of_profile}.
The base-ancestor coefficient drop is an explicit input, not a theorem here.

\subsubsection{Selected suffix cylinders}\label{res:selectsuffix}

The depth-$18$ selected window has one common terminally indexed suffix
cylinder after its depth-$13$ prefix is removed,
\lref{SelectedSuffixCylinder.lean}{88}{depth18SelectedHalfWindow_hasSuffixCylinderAt_thirteen}.
Under the canonical selected-window successor, the endpoint evolves exactly by
$E' = 2E-C$ and still covers the new carry interval,
\lrefx{SelectedSuffixCylinder.lean}{136}{SelectedHalfWindow.hasSuffixCylinderAt_stepExplicit}.
This preserves a finite cylinder invariant; it does not produce the required
coefficient profile at every later step.  For a rewind seam, the next prefix
bit has an exact exhaustive outcome: the cylinder promotes by one bit, or the
two base ancestors form the literal half-divisor boundary pair and hence
produce a protected seam,
\lrefx{SuffixCylinderPromotion.lean}{322}{promotedSuffixCylinder_or_protectedEvenSeamRealizedAt}.
At an exact half-divisor row, the crossing-cylinder profile gives the more
concrete advance-or-seam alternative,
\lrefx{SuffixCylinderGlobalProducer.lean}{174}{CylinderStage.feedbackStep_or_protectedSeam}.
The below-strip and above-strip hole cases have explicit finite word and
suffix-cylinder sheet witnesses,
\lrefx{SuffixCylinderBelow.lean}{117}{SelectedHalfWindow.word_stepTailExplicit}
and
\lrefx{SuffixCylinderTwoSheet.lean}{26}{hasSuffixCylinderAt_take}; combining
them leaves only a realized seam whose exact hole is inside the next strip,
\lrefx{SuffixCylinderFeedbackFanIn.lean}{62}{CylinderStage.feedbackAdvance_or_protectedSeam_inStrip}.
The strengthened in-strip alternative retains its two endpoint cylinders,
complementary literal prefixes, every admissible word off the hole, and the
canonical parent-failure coordinate required by the finite-gap construction,
\lrefx{SuffixCylinderInStrip.lean}{665}{CylinderStage.feedbackAdvance_or_inStripTwoSheet}.
The one-hole state embeds as a singleton finite gap; provided a common
next-row coefficient and the stated child-gap and buffer inequalities, this
two-sheet package advances one row while preserving its complete data,
\lrefx{SuffixCylinderFiniteGap.lean}{256}{InStripGapStage.advanceOfCommonNextCoeff}.
After cutoff promotion the two live sheets may have different coefficients.
The profiled-gap interface retains their independent literal prefixes,
derives the ready next-row profile, and advances under its explicit gap and
buffer hypotheses,
\lrefx{SuffixCylinderProfiledGap.lean}{124}{ProfiledGapStage.advanceOfNextCoeffProfile}.
At the next cutoff boundary, the lower and upper sheets respectively gain
their forced false and true prefix bits; the upper endpoint is adjusted by
the corresponding head weight,
\lrefx{SuffixCylinderProfiledPromotion.lean}{107}{selectedHalfIntervalPromoteHigh}.
The two retained prefixes remain adjacent, so their next-row coefficients
differ by at most one in the directed form needed for the profiled update,
\lrefx{SuffixCylinderProfiledAdjacency.lean}{232}{ProfiledGapStage.lowerNextCoeff_le_upperNextCoeff_add_one}.
The same adjacency invariant survives the false/true cutoff promotion,
\lrefx{SuffixCylinderProfiledPromotionAdjacency.lean}{65}{ProfiledGapStage.promoteCutoff_hasAdjacentPrefixes}.
Given its explicit next-row arithmetic cell, the promoted stage then advances
with its coefficient profile and buffer synthesized internally, preserving
adjacency through the whole feedback operation,
\lrefx{SuffixCylinderProfiledAutoStep.lean}{110}{ProfiledGapStage.promoteThenAdvance_hasAdjacentPrefixes}.
If the resulting macro-step swallows the next strip, its coefficient-pulse
bound forces the preceding gap into the exact quarter-band lower bound;
this is a quantified residual branch, not an impossibility claim,
\lrefx{SuffixCylinderProfiledSwallow.lean}{254}{nextBound_le_four_gapWidth_add_two_of_macro_swallow}.
An explicit full-cylinder run reaches row $51$ with endpoint $51{,}327{,}745$,
which exceeds the certified feedback head threshold and promotes the next
cutoff,
\lrefx{SuffixCylinderStage51.lean}{536}{fullCylinderStage51_feedback_headThreshold_margin}.
Any cofinal family of full suffix-cylinder stages would yield the terminal-only
cofinal producer consumed by the final support construction,
\lrefx{SuffixCylinderTerminalOnlyBridge.lean}{125}{cofinalTerminalOnlyStrip_of_cofinalCylinderStages}.
This is explicitly conditional: the public development does not establish
such a cofinal family.
This is a conditional local feedback classification, not a cofinal supply
theorem: the required crossing coefficient profile is still an input.

\subsubsection{Pre-feedback selected-cylinder run}\label{res:prefeedback}

Before the first half-divisor feedback row, Lean constructs a full selected
window at every depth from $18$ through $27$,
\lref{SelectedSuffixCylinderPreFeedback.lean}{118}{nonempty_stage_of_eighteen_le_of_le_twentySeven}.
At depth $27$ the common suffix cylinder remains and its endpoint covers the
whole protected strip,
\lrefx{SelectedSuffixCylinderPreFeedback.lean}{141}{depth27SelectedHalfWindow_hasSuffixCylinderAt_thirteen}
and
\lrefx{SelectedSuffixCylinderPreFeedback.lean}{147}{depth27SelectedHalfWindow_endpoint_covers_strip}.
The shared depth-$13$ prefix gives an explicit checked coefficient recurrence,
including the resulting depth-$27$ endpoint,
\lrefx{SuffixCylinderFirstFeedback.lean}{144}{stage27_endpoint_eq}.
This is a finite run, not an unbounded induction.
The threshold-stage continuation extends this explicit window through depth
$29$: its endpoint still covers the strip and its common suffix cylinder is
certified at every cutoff from $14$ through $25$, while cutoff $26$ is
explicitly not common,
\lrefx{SuffixCylinderThreshold.lean}{700}{depth29SelectedHalfWindow_endpoint_covers_strip}
and
\lrefx{SuffixCylinderThreshold.lean}{806}{depth29SelectedHalfWindow_not_commonRestrictionAt_twentySix}.

\subsubsection{Half-divisor unit drop}\label{res:halfdivisordrop}

At the doubled feedback row, changing only the terminal half-divisor bit from
false to true increases the support coefficient by exactly one:
\lref{HalfDivisorUnitDrop.lean}{20}{supportCoeff_extend_true_eq_false_add_one_at_double}.
The same holds for any explicitly identified adjacent boundary pair,
\lrefx{HalfDivisorUnitDrop.lean}{35}{supportCoeff_boundaryPair_unitDrop_at_double}.
The rewind adapter transfers such a pair to the next protected seam while
retaining the exact unit-drop witness,
\lrefx{RewindHalfDivisorAdapter.lean}{86}{rewindBaseUnitDropAt_of_halfDivisorBoundaryPair}
and
\lrefx{RewindHalfDivisorAdapter.lean}{147}{protectedEvenSeamRealizedAt_succ_of_fullStrip_rewindHalfDivisorBoundaryPair}.
This is the local coefficient input for the rewind-seam profile, not a global
construction of such boundary pairs.

\subsubsection{A finite rewind-boundary obstruction}\label{res:rewindboundarycounterexample}

The scalar one-row rewind data alone still does not supply that missing
boundary predicate.  Lean gives a concrete depth-$26$ selected two-word
window whose depth-$27$ explicit step has rewind history $[1]$, is a width-two
seam pair, and satisfies the doubled-row identity, but has no depth-$13$
half-divisor boundary-pair witness:
\lrefx{RewindBoundaryPairCounterexample.lean}{169}{depth27_actualRewindSeam_not_boundaryPair}.
Thus the protected-window buffer in the cofinal consumer is a genuine
hypothesis, not a consequence of the scalar seam facts alone.

\subsubsection{Concrete producer-carry coordinates}\label{res:producercarrycoordinates}

For every adjacent cut at rank $s\ge5$, Lean identifies the producer carry of
the terminal-augmented below support with the concrete middle coordinate:
\[
  4\,\mathrm{remainder}-\mathrm{belowPulse}-4,
\]
and identifies the terminal-augmented above support with the negative upper
coordinate
\[
  -\bigl(4\,\mathrm{overshoot}+\mathrm{abovePulse}+4\bigr).
\]
The generic row-word identity is
\lref{HalfCylinderProducerLowerBound.lean}{105}{producerCarry_insert_seamWordSupport_eq},
with the below and above specializations at
\lrefx{HalfCylinderProducerLowerBound.lean}{150}{producerCarry_insert_seamBelowSupport_eq_middleCoordinate}
and
\lrefx{HalfCylinderProducerLowerBound.lean}{183}{producerCarry_insert_seamAboveSupport_eq_neg_upperCoordinate}.
These identities make the remaining quantitative task exact: they do not
prove exponential lower bounds, sign persistence, or a universal escape
mechanism.

\subsubsection{Strategy boundary: Campbell shift synchronization}
\label{res:campbellboundary}

A proposed route combines a finite Mersenne shift from a putative last greedy
skip with Campbell's quarter-exponent prime progression.  The exact parameter
requirements cannot be synchronized: if the combined progression modulus is
$d\ge2$, the phase factor is at most $d$, the fatal-window prime cap is at
most four times that phase, and Campbell's applicability requires
$d^4$ below the cap, then the inequalities are inconsistent.  Formalised:
\lrefx{CampbellShiftSynchronization.lean}{299}{campbellQuarterExponent_periodFreeze_noSync}
and
\lref{CampbellShiftSynchronization.lean}{320}{campbellPhaseFreeze_parameters_impossible}.

The associated shifted-zero interface is not an independent producer.  It is
exactly equivalent both to infinitely many actual greedy skips and to
$1/2\in\mathcal A$:
\lrefx{CampbellShiftSynchronization.lean}{471}{shiftWindowZero_iff_greedyMersenneSkippedSupport_infinite}
and
\lref{CampbellShiftSynchronization.lean}{488}{shiftWindowZero_iff_half_mem_mersenneAchievementSet}.
This closes one synchronization strategy and one reformulation; it does not
advance either endpoint beyond the already stated half-achievement problem.

\subsection{What stays open}

The universal Erd\H{o}s \#257 statement, irrationality of $\sum_{n\in A}1/(2^n-1)$
for \emph{every} infinite $A$, is not proved. The proved cases are a family, not
the quantified theorem.

\section{The totient constant $S$ (Erd\H{o}s \#249)}\label{sec:249}

Whether $S=\sum_{n\ge1}\ph(n)/2^n$ is irrational is open. The development gives two
things and claims neither as a solution: an unconditional exclusion of
small-denominator rationals, and a conditional reduction to finite certificates.

The exact certificate spine and its open producer are stated in
Section~\ref{res:249story}.  This section gives the denominator theorem, the
geometry of $S$, and the finite arithmetic behind that summary.

\subsection{Unconditional: no small denominator}\label{sec:uncond}

\begin{theorem}[denominator exclusion]\label{res:farey}
Set
\[
\Qzero:=79\,639\,646\,646\,701\,375\,323\,355\,774\,875\,831\,053
       \approx7.96\times10^{34}.
\]
For every integer $p$ and every $q$ with $1\le q\le\Qzero$,
\[
S \;\ne\; \frac{p}{q}.
\]
Equivalently, if $S$ is rational then its reduced denominator exceeds that bound.
\end{theorem}
\noindent Formalised: \lloc{CertificateKernel.lean}{18055}.
The bound is reached by a ladder of finite, elementary exclusions,
\[
4838 \;\to\; 4\,194\,304 = 2^{22} \;\to\; 2.49\times10^{17}\ (\text{Farey } K{=}120)
\;\to\; 7.96\times10^{34}\ (\text{Farey } K{=}240),
\]
each a mediant/Farey argument that traps any candidate $m/q$ inside a forbidden
gap. The mediant framework is historically associated with Farey's 1816 note
on vulgar fractions~\cite{farey1816}; the finite exclusions themselves are
Lean-checked arguments in the formal-source checkpoint. See
\lrefx{CertificateKernel.lean}{15347}{tsum_totient_div_pow_two_ne_ratCast_of_den_le_4838}
for the first rung and \lrefx{GapFareyBound.lean}{88}{gap_check_window_1_120_le_248672326362367909},
\lrefx{GapFareyBound.lean}{176}{gap_check_window_1_240_le_79639646646701375323355774875831053}
for the two Farey windows.

This is an explicit lower bound on the denominator of any rational
representation, not a proof of irrationality. A rational with a larger
denominator is not excluded by it.

\begin{remark}[one bound, four representations]
Because the ladder identities of Section~\ref{sec:ladder} are equalities, the same
finite record transfers verbatim to three other representations of the constant:
to the positive lift $\sum_d A(d)/(2^d-1)$
(\lloc{CertificateKernel.lean}{18144}),
to the signed M\"obius-square constant $T=\sum_d\mu(d)/(2^d-1)^2$ at half the bound
(\lloc{CertificateKernel.lean}{18158}),
and to the coprimality probability of Section~\ref{sec:geom} at half the bound.
Half appears because $S=\tfrac12+T$ turns a denominator $d$ for $T$ into $2d$ for $S$.
\end{remark}

\subsection{The geometric picture: $S$ as coprime-pair mass}\label{sec:geom}

Let $X,Y$ be independent fair-coin waiting times, $\Pr(X=n)=2^{-n}$ for $n\ge1$.
Then $\Pr(d\mid X)=1/(2^d-1)$, so the Lambert transform is a divisor calculus of
$X$: $L(f)=\mathbb E[(f*\zeta)(X)]$. The rungs read as $L(\mu)=\Pr(X=1)=1/2$,
$L(\ph)=\mathbb E[X]=2$, $L(\mathbf 1)=\mathbb E[\tau(X)]=E$, and $L(A)=\mathbb E[\ph(X)]=S$.

Counting visible lattice points $(a,b)$ with $a+b=n$, $a>0$, $\gcd(a,b)=1$ gives
exactly $\ph(n)$, uniformly in $n$, so $S$ is itself a coprime-pair mass.

\begin{proposition}[coprimality probability]\label{res:coprime}
\[
S \;=\; \tfrac12 \;+\; \Pr\!\big(\gcd(X,Y)=1\big)
\;=\; \tfrac12 \;+\!\!\sum_{\substack{a,b\ge1\\ \gcd(a,b)=1}}\!\! 2^{-(a+b)}.
\]
\end{proposition}
\noindent Formalised: \lref{CertificateKernel.lean}{18229}{totient_series_eq_half_add_visible_coprime_pairs},
with the underlying lattice identity at
\lrefx{CertificateKernel.lean}{18216}{tsum_visible_coprime_pairs_eq_totient_series}.
Base $2$ is the one point where $\Pr(X=a)\Pr(Y=b)=2^{-(a+b)}$ needs no normalising
constant, so the totient sum and the coprimality probability line up with no
boundary correction. In this reading, Erd\H{o}s \#249 asks whether two independent
fair-coin waiting times are coprime with irrational probability.

\subsection{Conditional: reduction to finite certificates}\label{sec:reduction}

The reduction uses only the elementary bound $\ph(n)\le n$, geometric tail
estimates, and finite integer arithmetic.
For $t\ge1$, write $M_t=\lcm(1,\ldots,t)$.

\paragraph{Step 1: rationality forces an integer tail-period law.} Write the
fractional layer of $2^N S$ as the tail
\[
R_N \;=\; \sum_{m\ge1}\frac{\ph(N+m)}{2^m}.
\]
If $S$ is rational, its binary expansion is eventually periodic, which forces, for
a suitable period $h$ and all large $N$, the integrality
$R_{N+h}-R_N\in\Z$. To refute integrality at a finite depth $L$, form the window
discrepancy
\[
\Delta_{h,N,L}=\sum_{j=0}^{L-1}\big(\ph(N+h+1+j)-\ph(N+1+j)\big)\,2^{\,L-1-j}\in\Z,
\]
the first $L$ binary digits of $R_{N+h}-R_N$ up to a controlled deep tail (bounded
using only $\ph(n)\le n$), and call the pair \emph{certified} when the residue of
$\Delta_{h,N,L}$ modulo $2^L$ misses a neighbourhood of $0$:
\[
\textsf{certifiedKill}(h,N,L):\quad
N+h+L+2 \;<\; \big(\Delta_{h,N,L}\bmod 2^L\big) \;<\; 2^L-(N+h+L+2).
\]
This is a finite integer computation, decidable by the kernel. It is the natural
finite object here because a rational $S$ would pin the tail difference to an
integer, and a residue that stays a fixed distance from $0$ is a finite proof that
it did not.

\begin{corollary}[exact tail-period normal form]\label{res:red21}
$S$ is irrational if and only if, for every period $h\ge1$ and every threshold
$N_0$, there exist $N\ge N_0$ and $L$ with $\textsf{certifiedKill}(h,N,L)$.
\end{corollary}
\noindent Formalised: \lref{LcmConeFlatness.lean}{412}{irrational_totient_series_iff_certificate_supply};
the tail $R_N$ and the predicate are
\lrefx{TotientTailPeriodKiller.lean}{57}{totientTail} and
\lrefx{TotientTailPeriodKiller.lean}{72}{certifiedKill}.

\paragraph{Step 2: collapse the two parameters to one.} Every multiple of a period
is a period, so $\lcm(1,\dots,t)$ is the universal-period envelope: any eventual
period of a rational $S$ divides it once $t$ is large. Standing on that one
diagonal ray covers all possible periods at once, and the position parameter
drops out too.

\begin{corollary}[diagonal reduction]\label{res:diag}
Suppose that for every threshold $t_0$ there are parameters $t\ge t_0$ and
$L$ for which $\textsf{certifiedKill}(M_t,M_t,L)$ holds. Then $S$ is
irrational.
\end{corollary}
\noindent Formalised: \lref{LcmDiagonalReduction.lean}{112}{irrational_totient_series_of_lcm_diagonal_certificate_supply}.

\paragraph{Step 3: the cone, flatness, and completeness.} On the cone
$\{kM_t:k\ge1\}$, rationality forces a single fractional constant across the
whole cone.

\begin{proposition}[cone flatness]\label{res:flat}
If $S$ is rational, then for every sufficiently large $t$ there is
$\theta_t\in[0,1)$ such that the fractional part of the tail $R_{kM_t}$ equals
$\theta_t$ for every $k\ge1$.
\end{proposition}
\noindent Formalised: \lref{LcmConeFlatness.lean}{90}{rational_totient_series_forces_lcm_cone_flatness}.

The certificates are exact: a certificate exists exactly
when the object it certifies is non-integral. Thus the finite computations are
machine-checked witnesses rather than numerical experiments.

\begin{proposition}[certificate completeness]\label{res:complete}
For all $h,N$,
\[
\big(\exists L,\ \textsf{certifiedKill}(h,N,L)\big)
\quad\Longleftrightarrow\quad
R_{N+h}-R_N\notin\Z.
\]
\end{proposition}
\noindent Formalised: \lref{LcmConeFlatness.lean}{316}{exists_certifiedKill_iff_tail_diff_notMem_int}.
Conversely, integrality of one positive-shift tail difference forces $S$ to be
rational.  Thus $S$ is irrational if and only if every positive-shift tail
difference is non-integral, equivalently if every such pair admits a finite
certificate.  These normal forms are formalised at
\lrefx{LcmConeFlatness.lean}{357}{not_irrational_of_tail_diff_mem_int},
\lrefx{LcmConeFlatness.lean}{386}{irrational_totient_series_iff_all_tail_diffs_nonintegral},
and \lrefx{LcmConeFlatness.lean}{399}{irrational_totient_series_iff_pointwise_certificates}.
Finite certificates witness non-integrality. A joint refuter
can interrogate a menu of cone vertices at once and is sharper than any pairwise
test: it refutes the single-fractional-part model that rationality predicts on a
finite sample of cone vertices
(\lref{LcmConeNonflat.lean}{126}{exists_nonintegral_pair_of_coneNonflatCert}).

\paragraph{A reduced-denominator finite predicate.}  Separately, the formal
development records a stricter finite condition for an odd reduced denominator
$u$.  Its predicate $\textsf{ReducedDenominatorUnitGapCert}(u,N,K)$ asks that
every candidate integer in the associated dyadic tail interval be non-coprime
to $u$.  At a prime power $u=p^e$, the condition is equivalent to the exact
alternative that the interval has no candidate, or that it has one candidate
and $p$ divides that candidate.  Thus such an interval has at most one
candidate.

\begin{proposition}[prime-power unit-gap ceiling]\label{res:unitgap}
For every prime power $p^e$ with $e>0$, a
$\textsf{ReducedDenominatorUnitGapCert}(p^e,N,K)$ has candidate count at most
one.  The row $(u,N,K)=(3,3,5)$ satisfies this predicate, although its
corresponding ordinary empty-gap inequality fails.
\end{proposition}
\noindent Formalised:
\lref{PrimitiveWeightCertificate.lean}{54}{reducedDenominatorUnitGapCert_prime_pow_iff},
\lrefx{PrimitiveWeightCertificate.lean}{111}{reducedDenominatorCandidateCount_prime_pow_le_one},
\lrefx{PrimitiveWeightCertificate.lean}{151}{reducedDenominatorUnitGapCert_three_window_3_5}, and
\lrefx{PrimitiveWeightCertificate.lean}{157}{not_ordinary_gap_certificate_three_window_3_5}.
This is a finite certificate-level strengthening.  We do not establish that
rationality of $S$ supplies these predicates at unbounded parameters.

\paragraph{The certificates are non-empty.} The initial segment of the required
supply is machine-checked, so the reduction is not vacuous. By completeness, each
of these is a kernel-checked proof of a specific non-integral tail difference,
not a numerical experiment.

\begin{example}[verified finite instances]\label{res:deposits}
The following are checked by the Lean kernel:
\begin{itemize}[leftmargin=1.4em]
\item $\textsf{certifiedKill}(h,12,16)$ for every period $h\in[1,8]$;
\item certified kills for every period $h\in[1,16]$ at a fixed window;
\item diagonal certificates $\textsf{certifiedKill}(M_t,M_t,L)$ at the 28 scales
  $t\in\{1,2,3,4,5,7,8,9,11,13,16,17,19,23,25,27,29,31,32,37,41,43,
  47,49,53,59,61,64\}$;
\item a tabulated joint cone-vertex certificate.
\end{itemize}
\end{example}
\noindent Formalised: \lrefx{TotientTailPeriodKiller.lean}{404}{certifiedKill_all_small},
\lrefx{CarrySurvivorExtinction.lean}{574}{certifiedKill_all_upto_sixteen},
\lrefx{LcmDiagonalReduction.lean}{237}{certifiedKill_periodLcm_diagonal_upto_six}
(with \lrefx{LcmDiagonalReduction.lean}{242}{certifiedKill_periodLcm_diagonal_at_seven},
\lrefx{LcmDiagonalReduction.lean}{246}{certifiedKill_periodLcm_diagonal_at_eight}),
\lrefx{DiagonalPincerCertificatesT64.lean}{1967}{certifiedKill_diagonal_all_imported_through_t64},
and \lrefx{LcmConeNonflat.lean}{506}{coneNonflatCert_cells}. These instances show
that each layer of the reduction is non-vacuous and, by completeness, exact. They
are not evidence that the unbounded supply exists.

The first row of this finite table can be read without any search procedure.
Take $(h,N,L)=(1,12,16)$.  The integer in the definition is
\[
 \Delta_{1,12,16}
 =\sum_{j=0}^{15}\bigl(\ph(14+j)-\ph(13+j)\bigr)2^{15-j}
 =-143140.
\]
Its least nonnegative residue modulo $2^{16}$ is $53468$.  The tail allowance
is $N+h+L+2=31$, and hence
\[
       31<53468<65536-31=65505.
\]
This is precisely $\textsf{certifiedKill}(1,12,16)$.  The soundness theorem
(\lrefx{TotientTailPeriodKiller.lean}{262}{tail_diff_notMem_int_of_certifiedKill})
therefore gives $R_{13}-R_{12}\notin\Z$.  It is one exact non-integral tail
difference, not an assertion that such differences occur at unbounded scales.

\paragraph{An exact target interval.} The diagonal calculation also admits a
more local normal form.  Put $H_t=M_t$ and
\[
 D_t=R_{2H_t}-R_{H_t}.
\]
The formal development separates an exact rational M\"obius shadow
$Q_t=a_t/d_t$ in lowest terms and defines the remaining \emph{foreign defect}
by $F_t=D_t-Q_t$.  The full target is the residue class of $-a_t$ modulo $d_t$:
\[
 \textsf{FullTargetHit}(t)
 \quad\Longleftrightarrow\quad
 \exists k\in\Z,\qquad d_tF_t=-a_t+d_tk.
\]
Since $D_t=a_t/d_t+F_t$, this is exactly the condition $D_t\in\Z$; no
complementary term is discarded or assigned a conjectural sign.

\begin{proposition}[diagonal pincer decomposition]\label{res:pincer}
For every $t$, diagonal integrality is equivalent to the foreign diagonal
defect meeting the full target.  Consequently, full-target avoidance at
unbounded scales implies irrationality of $S$.
\end{proposition}
\noindent Formalised:
\lref{DiagonalPincerDecomposition.lean}{215}{diagonal_int_iff_foreignDiagonalDefect_hits_fullTarget}
and
\lref{DiagonalPincerDecomposition.lean}{290}{irrational_totientSeries_of_full_target_avoidance_supply}.
The proposition is an exact reduction; the unbounded avoidance hypothesis is
still open.

Two algebraic inputs make this interval test explicit.  First, truncate the
M\"obius-square identity of Proposition~\ref{res:lens} after $D$ terms.  The
resulting rational quantity gives a Lambert-projected centre $C_{H,D}$ for the
scaled diagonal difference.  The error is exactly a signed squared-Mersenne
tail:
\[
 \textsf{scaledDiagonal}(H)-C_{H,D}
   =c_H\sum_{d>D}\frac{\mu(d)}{(2^d-1)^2},
\]
where $c_H$ is the explicit diagonal coefficient in the formal source.  The
absolute value of the sum is bounded geometrically by
$4/[3(2^{D+1}-1)^2]$.  Thus a rational separation of the centre from every
integer by more than the scaled bound proves target avoidance; the argument
does not estimate the sign of the tail.

\begin{proposition}[squared-Mersenne enclosure]\label{res:squareenclosure}
The scaled diagonal tail difference differs from its Lambert-projected centre
by the squared-Mersenne tail.  Separation of the centre from the full target by
more than the proved tail bound forces target avoidance.
\end{proposition}
\noindent Formalised:
\lref{SquaredMersenneDiagonalEnclosure.lean}{116}{scaleDiagonalTailDifference_sub_lambertProjectedDiagonal}
and
\lref{SquaredMersenneDiagonalEnclosure.lean}{237}{scaleFullTarget_miss_of_lambert_projected_separation}.

\paragraph{Finite foreign-residue projection.}  There is a second finite
presentation of the same obstruction.  For a cutoff $D$, split the finite
M\"obius residue diagonal into the channels $d\leq D$ which divide $H$ and
the remaining channels.  The divisor part is exactly the truncated M\"obius
shadow, while every retained nondivisor channel is explicit.  Above the
stable cutoff $2H$, a finite window of omitted channels is bounded by
\[
 c_H\left(\frac{2}{2^D}+\frac{4}{3\,4^D}\right),
\]
where $c_H$ is the diagonal coefficient.  Thus, if the actual foreign defect
agrees with this projection within the displayed bound, and the projected
quantity is farther than that bound from every integer, then the full target
is missed.

\begin{proposition}[finite foreign-residue projection]\label{res:foreignresidue}
The finite residue diagonal splits exactly into projected foreign and divisor
channels.  A controlled projected foreign defect with the stated integer
separation cannot meet the full target.
\end{proposition}
\noindent Formalised:
\lref{ActualForeignResidueProjection.lean}{307}{finiteResidueDiagonal_eq_projectedForeign_add_divisor},
\lref{ActualForeignResidueProjection.lean}{275}{abs_foreignTailWindow_le_foreignComplementBound},
and
\lref{ActualForeignResidueProjection.lean}{348}{scaleFullTarget_miss_of_projected_separation}.
The passage from finite windows to the analytic foreign complement, including
the corresponding kernel identity, is not supplied here.  Nor is an
unbounded family of separated projections; these are parts of the open
certificate-supply problem.

Second, the rational M\"obius shadow has an exact reduced denominator.  If
$r_t$ is the square-free kernel of $H_t$, $s_t=H_t/r_t$, and $J(r_t)$ is the
odd Jordan scalar defined in the algebraic shadow chain, then
\[
 d_t=
 \frac{2^{r_t}-1}
 {\gcd\!\bigl(2^{r_t}-1,\,s_t\lvert J(r_t)\rvert\bigr)}.
\]
For $t\geq5$, the product of the Mersenne factors $2^p-1$ over primes
$t/2<p\leq t$ divides $d_t$; in particular $d_t\geq2^{t/2}$.  This is a
denominator theorem for $Q_t$ only.  Cancellation by $F_t$ is exactly what the
full-target condition retains.

\begin{proposition}[Mersenne-shadow denominator]\label{res:shadowden}
At lcm height $t$, the reduced denominator of the scaled M\"obius shadow is
given exactly by the formal denominator formula.  The upper-half Mersenne
product divides it and yields the stated lower bounds.
\end{proposition}
\noindent Formalised:
\lref{MersenneShadowDenominatorGrowth.lean}{99}{exists_upperHalf_mersenne_channel_dvd_den_with_bounds}
and
\lref{MersenneShadowDenominatorGrowth.lean}{147}{lcmHeight_scaledMobiusShadow_den_exact}.

Finally, the remaining fresh-loss term can be represented by a finite integer
projection, and then by binary suffix data.  At depth $L$, define
\[
 W_{t,L}=P_L(2H_t)-P_L(H_t),\qquad
 \rho_{t,L}=W_{t,L}\bmod 2^L,
\]
where $P_L(M)$ is the length-$L$ window numerator from Step~1.  The finite
residue certificate is the asymmetric safe interval
\[
 H_t+L+2<\rho_{t,L}<2^L-(2H_t+L+2).
\]
The two unequal margins are the tail bounds at the two cone vertices $H_t$ and
$2H_t$; replacing them by a symmetric slogan would lose part of the checked
statement.

\begin{corollary}[fresh-loss projection reduction]\label{res:freshloss}
The fresh-loss residue certificate is equivalent to its projection
inequality, and the projection forces full-target avoidance.  Hence an
unbounded projection supply, or either stronger adjacent-suffix supply stated
in the formal development, implies irrationality of $S$.
\end{corollary}
\noindent Formalised:
\lref{DiagonalFreshLossBridge.lean}{61}{diagonalFreshLossResidueCert_iff_projectionCert}
and
\lref{DiagonalFreshLossBridge.lean}{191}{irrational_totientSeries_of_diagonalFreshLossProjectionSupply}.
The development proves finite fixtures for this bridge but does not assert an
unbounded projection or suffix supply.

\paragraph{T64 certificate closure.}\label{res:t64closure}
The generated Lucas-certificate closure supplies the primality dependencies for
the finite diagonal-pincer certificate family through $t=64$; for example,
\lref{DiagonalPincerPrimeCertificates/ClosureT64.lean}{51759}{prime_lucas_2364533768205644535022723273}.
It is finite certificate infrastructure, not an unbounded certificate supply
or an irrationality proof.

\paragraph{T64 endpoint certificate.}\label{res:t64endpoint}
Using that closure, Lean checks the individual diagonal-pincer certified kill
at $t=64$,
\lref{DiagonalPincerCertificateT64Endpoint.lean}{1928}{certifiedKill_diagonal_t64}.
This is one finite endpoint certificate, not an unbounded supply.

\paragraph{Flexible odd three-rank window.}\label{res:flexibleoddwindow}
At an analytically admissible odd power-of-two rank, centrality at any one of
the canonical odd rank or its next two base-four successors yields the landed
adjacent-suffix gap interface,
\lref{DiagonalFlexibleOddWindowSupply.lean}{93}{diagonalAdjacentSuffixGapSupply_of_powerTwoOddGuardThreeRankBand}.
Thus a cofinal three-rank band supply implies irrationality,
\lrefx{DiagonalFlexibleOddWindowSupply.lean}{136}{irrational_totientSeries_of_powerTwoOddGuardThreeRankBandSupply};
that supply is not proved.

\paragraph{Affine frontier for that window.}\label{res:flexibleoddaffine}
With the stated short-correction bounds, the three central-band predicates are
equivalent to one three-scale affine escape condition,
\lref{DiagonalFlexibleOddWindowAffine.lean}{236}{threeRankBandAt_iff_threeScaleAffineEscapeAt_of_shortCorrections}.
A zero-start affine orbit shows that a uniform correction envelope alone does
not force escape at any finite number of ranks,
\lrefx{DiagonalFlexibleOddWindowAffine.lean}{225}{zeroAffineOrbit_stays_in_scaled_edge}.

\paragraph{Fresh-prime deficit decomposition.}  The literal foreign channel
also has an endpoint form suited to curvature tests.  Retain only the prime
support already present at $H$ in the two endpoints $H+s$ and $2H+s$, and
write the resulting discrepancy from the actual totient as an endpoint
deficit.  Each such deficit is nonnegative.  The actual diagonal increment
then equals the old-prime increment plus the lower deficit minus the upper
deficit.  The second and five-point curvature operators preserve this exact
split; the latter has coefficient row $[-8,4,2,1,1]$.  Consequently, the old
curvature margin loses at most the stated adverse weighted deficit.

\begin{proposition}[fresh-prime deficit decomposition]\label{res:freshprimedeficit}
The endpoint fresh deficits are nonnegative, give an exact decomposition of
the foreign channel, and bound the loss from the old-prime curvature margins.
\end{proposition}
\noindent Formalised:
\lref{FreshPrimeDeficitDecomposition.lean}{91}{endpointFreshDeficit_nonneg},
\lref{FreshPrimeDeficitDecomposition.lean}{184}{finiteForeignChannelIncrement_eq_endpointFreshDeficit_sub},
and
\lref{FreshPrimeDeficitDecomposition.lean}{269}{oldFivePoint_sub_secondBranchAdverseDeficit_le_actual}.
This supplies the exact correction term for the finite probes.  It supplies
neither a cofinal deficit bound nor an irrationality criterion.

\paragraph{One-bit lifting.}  The power-of-two signed-margin socket admits a
separate exact arithmetic reduction.  If a signed difference is already
divisible by $2^k$, lifting its congruence to $2^{k+1}$ is equivalent to
evenness of the quotient after division by $2^k$.  Failure of the lift gives
the unique translated residue class.  In particular, a displayed difference
$P-N=16c$ agrees modulo $32$ exactly when $c$ is even; otherwise the residue
is the offset class by $16$.

\begin{proposition}[one-bit signed-margin lift]\label{res:powertwobitlift}
The fifth-bit question after fourth-bit agreement is exactly one cofactor
parity test, with the complementary offset class characterising failure.
\end{proposition}
\noindent Formalised:
\lref{PowerTwoBitLift.lean}{34}{two_pow_succ_dvd_iff_quotient_even},
\lref{PowerTwoBitLift.lean}{111}{modEq_thirtyTwo_iff_even_difference_cofactor},
and
\lref{PowerTwoBitLift.lean}{129}{not_modEq_thirtyTwo_iff_offset_sixteen}.
The centered-state formulation identifies the signed-margin socket with
failure of this next lift, subject to its stated correction and margin
budget:
\lrefx{PowerTwoCenteredBitLift.lean}{121}{actualPenultimateSignedMargin_centeredLift_iff_not_twice_dvd}
and
\lrefx{PowerTwoCenteredBitLift.lean}{159}{actualPenultimateSignedMargin_centeredThirtyTwo_iff_not_thirtyTwo_dvd}.
This is a finite arithmetic interface.  It does not supply the required
cofactors at unbounded scales and gives no irrationality conclusion.

\paragraph{Parity is not a margin certificate.}  The preceding lift does not
settle the reduced signed-margin socket.  For every positive even modulus and
every radius in the stated nontrivial range, each parity class contains both a
socket hit and a socket miss.  The same statement applies to a positive
power-of-two modulus.  Thus even a fifth-bit computation, by itself, cannot
decide the required margin condition.

\begin{proposition}[parity obstruction for the reduced margin socket]
\label{res:paritysocket}
Under the explicit radius inequalities, parity does not decide membership in
the reduced margin socket, including at a positive power-of-two modulus.
\end{proposition}
\noindent Formalised:
\lref{Totient827ParitySocketCeiling.lean}{73}{reducedMarginSocket_not_decided_by_parity}
and
\lref{Totient827ParitySocketCeiling.lean}{110}{powerTwo_reducedMarginSocket_not_decided_by_parity}.
The module contains no $t=827$ numeral and proves no signed-margin
certificate; it excludes parity as a sufficient standalone criterion.

\paragraph{A closed prime-adjunction route.} One tempting way to seek an
unbounded obstruction is to compare the full target at $H$, $pH$, $qH$, and
$pqH$.  The formal development rules out this route in its unprojected form.
For every positive $k$, the actual diagonal satisfies an affine transport law
\[
 D_{kH}=Q_k(H)D_H+Z_k(H),
 \qquad Q_k(H)\in\N,\quad Z_k(H)\in\Z.
\]
Hence integrality at $H$ propagates to every multiple, and the four target hits
at the prime-adjunction diamond are equivalent to the single hit at its root
(\lref{FullTargetPrimeAdjunctionNoGo.lean}{196}{fullTarget_primeAdjunction_diamond_iff_root}).
The corresponding transport of the complete foreign defect has zero diamond
curvature
(\lrefx{FullTargetPrimeAdjunctionNoGo.lean}{247}{foreignTransportCorrection_primeAdjunction_flat}).
Thus explicit-shadow fibres alone cannot create a holonomy obstruction.  A
future positive route would need a separately defined projection together with
a theorem controlling what that projection discards.  This is a proved
obstruction to one strategy, not progress on the unbounded-supply proposition.

\begin{proposition}[scalar-localisation obstruction]\label{res:adelic}
Let $x\in\Q$ and $c\in\Z$.  If $H\mid\operatorname{den}(x)$ and the reduced denominator
of $cx$ divides $H$, then
\[
 \frac{\operatorname{den}(x)}{H}\mid |c|.
\]
\end{proposition}
\noindent Formalised:
\lrefx{AdelicHeightObstruction.lean}{23}{scalarLocalization_complement_dvd}
and \lrefx{AdelicHeightObstruction.lean}{56}{scalarLocalization_integer_eq_mul_num}.
Thus a scalar localisation can move a complementary denominator factor into
its coefficient, but cannot erase it.  This rules out a scalar-denominator
shortcut of that form.  It neither proves full-target avoidance nor supplies
any certificates, so it gives no irrationality criterion for $S$.

\begin{proposition}[square-CRT correction suppression]\label{res:squarecrt}
Let $E$ be a finite family indexed by distinct primes $p_i$, with prescribed
anchors $A_i$ and residues $a_i$.  There is a bounded common base $n$ such
that, for each $i\in E$ and each shift $h$ with $p_i\nmid a_i+h$, one has
\[
 n=A_i+p_i(a_i+p_i t)\quad\text{and}\quad
 \ph(n+p_i h-A_i)=(p_i-1)\ph(a_i+p_i t+h)
\]
for a suitable $t\in\N$.  In particular, if every $p_i$ exceeds a fixed
horizon $J$, the choice $a_i=0$ works simultaneously for all $1\le h\le J$.
\end{proposition}
\noindent Formalised:
\lrefx{SquareCRTCube.lean}{198}{exists_squareCRT_clean_totient_family}
and \lrefx{SquareCRTCube.lean}{228}{exists_squareCRT_clean_horizon}.
The finite mechanism does not force a cube coefficient to be nonzero: there
is a checked clean two-step cube with both displayed coefficients zero
(\lrefx{SquareCRTCube.lean}{285}{squareCRT_clean_block_can_vanish}) and a
separate checked clean cube with second coefficient $-4$
(\lrefx{SquareCRTCube.lean}{303}{squareCRT_clean_block_can_be_nonzero}).
Thus the result gives neither an unbounded certificate supply nor an
irrationality criterion for $S$.

\begin{proposition}[signed dyadic-moment algebra]\label{res:signedmoment}
For matrices $M:\iota\times\kappa\to R$ and
$N:\kappa\times\iota\to R$ over a commutative ring, the determinant of
$MN$ is the sum over all maps $p:\iota\to\kappa$ of the determinant of the
corresponding selected-column matrix times the diagonal product
$\prod_i N_{p(i),i}$.  Separately, if a selected integer numerator is odd and
its dyadic exponent is strictly larger than every other selected exponent,
then the common numerator after dyadic denominator clearing is odd and hence
nonzero.
\end{proposition}
\noindent Formalised:
\lrefx{SignedQMomentObstruction.lean}{28}{det_mul_rectangular},
\lrefx{SignedQMomentObstruction.lean}{77}{scaled_dyadic_sum_odd}, and
\lrefx{SignedQMomentObstruction.lean}{95}{scaled_dyadic_sum_ne_zero}.
These are finite algebraic inputs to a signed-moment route.  They do not show
that any actual totient Hankel determinant is nonzero, and they give no
irrationality criterion for $S$.

\begin{proposition}[dyadic totient certificate interface]\label{res:dyadictotient}
For every fixed level $e$, a separated nonzero evaluation minor of the
canonical dyadic totient-kernel family gives its span dimension
$2^e+1$.  Every even residue channel is exactly a scalar multiple of its
odd-core channel.  Moreover, a compressed-adjoint certificate whose nonzero
multiple is strictly smaller in absolute value than its positive modulus is
inconsistent.
\end{proposition}
\noindent Formalised:
\lrefx{TotientMahlerDefect.lean}{98}{finrank_canonicalTotientKernel_eq_of_certificate},
\lrefx{TotientMahlerDefect.lean}{156}{totientKernel_even_residue_reduce}, and
\lrefx{TotientMahlerDefect.lean}{197}{false_of_compressedAdjointCertificate}.
The unformalised producer is the all-level construction of the separated
minors (by a CRT--Dirichlet argument); the interface does not provide those
certificates for every $e$, and it gives no irrationality criterion for $S$.

\begin{proposition}[residual-gauge obstruction for monomial minors]\label{res:residualgauge}
Let a monomial matrix have a common nonzero residual weight in each column.
The residual weights act by a right diagonal gauge, and therefore preserve
determinant nonvanishing.  If one exponent is $1$, the inverse-phase locked
gauge makes that row identically $1$ while preserving a nonzero determinant
whenever the coefficient-side determinant is nonzero.  If the residual is
allowed to depend on both row and column, it reconstructs a consecutive-power
matrix whose zeroth row is identically $1$.
\end{proposition}
\noindent Formalised:
\lrefx{ResidualGaugeObstruction.lean}{55}{det_residualMonomialMatrix_ne_zero_iff},
\lrefx{ResidualGaugeObstruction.lean}{94}{locked_reconstruction_preserves_nonzero_minor}, and
\lrefx{ResidualGaugeObstruction.lean}{145}{rowReconstruction_zero_row_one}.
Thus a residual-blind rank, determinant, or conditioning criterion alone cannot
exclude the target.  The proposition does not exclude a criterion that also
uses an arithmetic coupling identity, and it gives no irrationality criterion
for $S$.

\paragraph{What is open.} The reduction converts \#249 into the statement that the
certificate supply is \emph{unbounded}: certificates exist for arbitrarily large
parameters. By Proposition~\ref{res:complete} this is equivalent to the
tail differences being non-integral infinitely often, and equivalently to the
Farey gap denominators of Section~\ref{sec:uncond} being unbounded: the same
obstruction seen as a rational denominator, an eventual binary period, and a flat
tail on the $\lcm$ cone. That unboundedness is the remaining analytic question,
and it is left open. The reduction is real; the closure is not asserted.

\section{Formalisation architecture and mathematical lessons}\label{sec:architecture}

The formal development repeatedly separates a finite, kernel-checkable core
from the analytic statement that would have to hold uniformly. For the
full-support Erd\H{o}s--Borwein theorem, the core is a bounded certificate
engine assembled from a Bertrand/CRT frame, divisor-pair averaging, and an
explicit parameter closure. For \#249, the same design principle appears more
sharply: eventual binary periodicity predicts an integral tail difference;
finite modular arithmetic can refute that prediction at any fixed pair of
parameters; and completeness proves that this computation loses no
information. Lean therefore makes the open boundary visible: the missing
ingredient is not another finite verification but a theorem producing
witnesses at unbounded parameters.

The representation of eventual periodicity was itself consequential. Rather
than reason about an informal repeating digit string, the development uses the
scaled tails $R_N$ and proves that rationality forces their period differences
to be integers. Certificate completeness then separates two logically distinct
tasks: finite evaluation decides non-integrality at fixed parameters, while an
analytic theorem would still be needed to produce such parameters without
bound. This separation is what prevents verified computation from being
mistaken for the open theorem.

Three changes of coordinates serve different mathematical purposes. The
Mersenne--Lambert transform places $S$ next to the Erd\H{o}s--Borwein constant
and exposes the nonnegative primitive-conductor weight. The M\"obius-square
identity turns the same number into a signed rapidly convergent series. The
coprime-pair identity turns it into a probability. Formalising all three and
proving that the denominator exclusion transfers between them prevents an
informal change of representation from silently changing the claim.

The auxiliary carry modules ask what rationality would force rather than
claiming a new irrationality theorem. They give an exact tempered-orbit
criterion, Boolean--M\"obius coordinates for support series, constraints from
residue wraps and reciprocal mass, and a greedy description of the value set.
The sublogarithmic coverage theorem adds a further lesson: the elementary
fixed-power estimate $\tau(n)^k\le(k^{2^k})^k n$ becomes useful only after it is
transported through the binary tail and composed with the exact carry
recurrence. Rationality then forbids long zero windows in $f_A$, uniformly in
the support. In the reciprocal-mass argument, divergent series had to be kept
as an explicit alternative rather than passed through a default value for a
divergent formal sum.
Together they rule out several tempting but false proof strategies---for
example, rationality need not give a bounded carry alphabet---and identify the
kind of infinite Boolean carry path that a future proof would have to exclude.
The precise statements and source links are retained in
Appendix~\ref{sec:carry}; their supporting role is why they are not a second
main narrative here.

This architecture is reusable beyond the two named problems: choose a
representation in which rationality imposes a rigid tail law, isolate a finite
predicate that soundly refutes that law, prove the predicate complete at fixed
parameters, and state the remaining uniformity theorem separately. The last
step matters as much as the first three. It prevents a large verified finite
range from being mistaken for the infinite theorem.

The repository separately records claim status and editorial change rules;
the mathematical content remains the statements and proofs above.

\section{Artefact availability and verification}\label{sec:verify}

The Lean~4 proof assistant~\cite{lean} and Mathlib library~\cite{mathlib} check
the development. The source, exact informal-to-formal links, and build
instructions are maintained in the
\href{https://github.com/wcook04/plectis-lean-erdos249-257}{public repository}.
This paper is tied to an
\href{\sourceurl}{archived formal-source revision}; every blue declaration name
links directly to the corresponding checked theorem there. The repository
pins the theorem prover and library versions needed to reproduce the build.

Every declaration is elaborated and every proof term is checked by the Lean
kernel. Finite computations use kernel-checked decision procedures rather than
an external evaluator, and the development contains no proof placeholders or
added axioms. Continuous integration repeats the complete project build.
Appendix~\ref{app:index} explains how a reader can move from the mathematics
to the exact formal sources without exposing implementation names in the
printed paper.

\section{Complements and further questions}\label{sec:further}

The formalisation establishes checked results in two directions and leaves two
exact frontiers. It
verifies the all-integer-base Erd\H{o}s--Borwein theorem and several structured
support cases toward \#257. For \#249 it proves that a rational value of $S$
would have denominator exceeding $\Qzero$, and it reduces irrationality to an
unbounded supply of finite certificates whose fixed-parameter semantics are
complete. The unresolved \#249 step is precisely that unboundedness; the
unresolved \#257 step is the passage from the formalised support classes to
every infinite support. Neither frontier is softened by the size of the
verified artefact.

The remaining questions can be stated without reference to the size of the
formal development.

\subsection{Two exact-arithmetic reductions}

Two registered reductions put parts of the remaining obligations into scalar
exact arithmetic.  They do not change the status of either problem.

For \#249, let $m_t$ be the canonical adjacent-suffix depth and let $d_t$ be
the corresponding residue in $[0,2^{m_t})$.  Define
\[
  \sigma_t=\min\bigl(d_t-2^{m_t-5},
    (2^{m_t}-2^{m_t-5})-d_t\bigr)\in\Z.
\]

\begin{proposition}[canonical strict-jump slack]\label{res:slackfrontier}
The canonical residue lies in the central band
$[2^{m_t-5},2^{m_t}-2^{m_t-5}]$ if and only if $\sigma_t\ge0$.
If nonnegative slack occurs at arbitrarily large strict jumps
$M_t<M_{t+1}$, then $S$ is irrational.
\end{proposition}
\noindent Formalised:
\lrefx{DiagonalFreshLossBridge.lean}{2469}{canonicalAdjacentSuffixCentral_iff_slack_nonneg},
\lrefx{DiagonalFreshLossBridge.lean}{2559}{canonicalAdjacentSuffixJumpCentralSupply_iff_slackSupply},
and
\lref{DiagonalFreshLossBridge.lean}{2731}{irrational_totientSeries_of_canonicalAdjacentSuffixJumpSlackSupply}.
The proposition is an equivalence and a conditional consumer.  It does not
prove that $\sigma_t$ is nonnegative cofinally.

For \#257, write $r_n$ for the exact greedy residual of the target $1/2$ and
put
\[
  u_n=4^n(2r_n-2^{-n}).
\]
The second-channel condition is the rational inequality
\[
  \frac16+\frac{37}{56}2^{-n}\le \left|u_n-\frac13\right|.
\]

\begin{proposition}[rational second-channel reduction]\label{res:secondfrontier}
The second-channel condition is decidable over $\Q$ and holds for
$1\le n\le6$.  If it holds for every $n\ge7$, then
$1/2$ belongs to the Mersenne achievement set $\mathcal A$.
\end{proposition}
\noindent Formalised:
\lrefx{GreedyAchievementSet.lean}{2695}{HalfSecondChannelSeparatedRat},
\lrefx{GreedyAchievementSet.lean}{2723}{halfSecondChannelSeparatedRat_of_pos_le_six},
and
\lref{GreedyAchievementSet.lean}{2901}{half_mem_mersenneAchievementSet_of_secondChannelSeparationRat_from_seven}.
The membership conclusion is itself exact:
\[
  1/2\in\mathcal A
  \quad\Longleftrightarrow\quad
  \text{the canonical greedy orbit omits infinitely many exponents}.
\]
This is formalised at
\lref{GreedyAchievementSet.lean}{2514}{half_mem_mersenneAchievementSet_iff_greedySkippedSupport_infinite}.
The forward implication uses the already formalised irrationality of the full
Erd\H{o}s--Borwein series; the reverse implication uses the absorbing fatal-state
criterion.  Thus the second-channel hypothesis is one sufficient route to the
infinite-skip condition.  It does not prove that condition, and membership alone
is not the universal \#257 theorem.

\begin{problem}[Erd\H{o}s \#249]
Prove that
\[
  \sum_{n\ge1}\frac{\ph(n)}{2^n}
\]
is irrational, without imposing a bound on the denominator of a possible
rational representation.
\end{problem}

\begin{problem}[unbounded certificate supply]
Produce certified non-integrality witnesses at unbounded parameters.  A
sufficient one-parameter form is to produce them along the lcm diagonal, as
in Corollary~\ref{res:diag}; the verified scales in
Example~\ref{res:deposits} form a finite initial set and do not establish the
unbounded supply.
\end{problem}

\begin{problem}[universal Erd\H{o}s \#257]
Prove that
\[
  \sum_{n\in A}\frac{1}{2^n-1}
\]
is irrational for every infinite set $A\subseteq\N$, rather than only for the
support families in Theorem~\ref{res:support}.
\end{problem}

On the \#257 side, rationality would impose two complementary forms of
rigidity: divisor-coverage zero windows have length $o(\log N)$, whereas an
infinite support with rational series value forces the associated positive
carry states to be unbounded.
These statements constrain different objects and do not yet exclude a rational
infinite-support value, but together they narrow the structure that any such
value would have to exhibit.

The main lesson from formalisation is the alignment of representations and
proof obligations. The Lambert, M\"obius, probabilistic, periodic-tail, and
carry descriptions are not competing stories: each makes a different
rationality obstruction explicit. A future advance must exploit one of those
obstructions uniformly, not merely extend the finite range already checked.
The public package therefore records exact implications, finite results, and
negative results without treating corpus volume as mathematical evidence or
claiming that an unchecked composition closes either open problem.

\subsection*{Statements and declarations}

\paragraph{Artefact and data availability.} The
\href{\sourceurl}{archived formal-source revision} contains the Lean sources,
pinned toolchain, library manifest, and generated certificate data used in the
verification. Repository metadata and this manuscript provide navigation
rather than proof authority; Section~\ref{sec:verify} gives the verification
route.

\paragraph{Use of AI assistance.} Large-language-model assistants were used
during development for proof drafting, refactoring, and prose editing. Every
registered formal claim is linked to a declaration in the archived source
revision. Lean checks each proof term against the pinned library; the project
contains no proof placeholders or project-defined axioms.

\paragraph{Funding and competing interests.} This work received no external
funding. The author declares no competing interests.

\paragraph{Acknowledgements.} The problem numbering and status follow the
Erd\H{o}s Problems catalogue maintained by Thomas Bloom~\cite{erdosproblems}.

\appendix
\section{Auxiliary rationality criteria from binary carry systems}\label{sec:carry}

Five further modules provide additional criteria. They prove no new
irrationality. Instead they pin down, in machine-checked form, what a rational
value of the series in question would have to look like: as an integer carry
orbit, as a M\"obius-inverted support coordinate, as a residue-wrap cycle
modulo the odd part of the denominator, and as a point of a greedy subset-sum
geometry. Everything in this section is an equivalence or an unconditional
constraint on the rational side; nothing excludes a rational value of either
open series, and each module states that boundary in its header. The prior-art
boundary is theorem-family-specific. The integral carry recurrence has a
direct antecedent in Wang--Grau Ribas~\cite{wanggrauribas}; the strict-tail
geometry is recorded by Kova\v{c}--Tao~\cite{kovactao}; M\"obius inversion,
repetend algebra, and divisor averaging are classical (see, for example,
Apostol~\cite{apostol}). We make no claim of mathematical priority for the
converse/rigidity, certificate-normal-form, or coupled reciprocal-mass/collision
statements collected here.

\subsection{Tail-orbit rigidity: a rationality criterion}

For a coefficient sequence $c:\N\to\N$ with $c(n)\le n$, write
\[
X_c=\sum_{n\ge1}\frac{c(n)}{2^n},
\qquad
T_c(N)=\sum_{j\ge1}\frac{c(N+j)}{2^j},
\]
the series and its scaled tail after $N$ binary digits. Call an integer
sequence $u$ a \emph{tempered orbit} for $c$ with multiplier $v\ge1$ if
\[
u(N+1)=2\,u(N)-v\,c(N+1)\ \text{ for all } N,
\qquad\text{and}\qquad
u(N)/2^N\to0 .
\]
The recurrence is exact binary long division against the coefficient stream;
the boundary condition forbids the homogeneous solution. Wang and Grau Ribas
use the integral carry recurrence forced by rationality for the weighted binary
special case $c(n)=n d_n$~\cite{wanggrauribas}. The module abstracts that
forward mechanism to arbitrary nonnegative $c(n)\le n$; its explicit converse
and subexponential-orbit rigidity are the local additions, with no priority
claim attached. Wang--Grau Ribas's positive-density theorem and its Erd\H{o}s
\#260 corollary are not used or formalised here.

\begin{theorem}[tail-orbit rigidity]\label{res:rigidity}
Let $c(n)\le n$ for all $n$. Every tempered orbit is the scaled tail,
$u(N)=v\,T_c(N)$ for all $N$; and $X_c$ is rational if and only if a tempered
orbit exists for some multiplier $v\ge1$.
\end{theorem}
\noindent Formalised:
\lref{GenericTailOrbitRigidity.lean}{338}{temperedBinaryOrbit_eq_scaledTail}
(rigidity) and
\lref{GenericTailOrbitRigidity.lean}{425}{binaryCoeffSeries_rational_iff_exists_temperedBinaryOrbit}
(the criterion), restated in Mathlib's irrationality vocabulary at
\lrefx{GenericTailOrbitRigidity.lean}{434}{not_irrational_binaryCoeffSeries_iff_exists_temperedBinaryOrbit}.
The engine is one observation: a real orbit with $d(N+1)=2\,d(N)$ and
$d(N)=o(2^N)$ vanishes identically
(\lrefx{GenericTailOrbitRigidity.lean}{314}{doublingOrbit_eq_zero_of_tempered}).

The tempered boundary is essential: the
homogeneous parasite $N\mapsto2^N$ can be added to any orbit without
disturbing the recurrence, so positivity of an orbit certifies nothing by
itself, and positivity is nowhere advertised as an equivalent criterion. The
linear-growth hypothesis covers both constants introduced in
Section~\ref{sec:intro}:
$\ph(n)\le n$ for the \#249 coefficients, and $f_A(n)\le\tau(n)\le n$ for the
\#257 support coefficients $f_A(n)=\#\{a\in A:a\mid n\}$. The modules below
instantiate the criterion on supports; the totient instantiation is not taken
here.

\begin{proposition}[finite-level carry anti-compression]\label{res:carryrank}
Suppose a tempered integral carry orbit for the totient sequence has positive
multiplier.  If the canonical level-$e$ totient kernel family is linearly
independent, then the span of its canonical carry sections has dimension at
least $2^e-1$.  Consequently, an all-level independence producer together
with rationality of the totient coefficient series would force one tempered
carry orbit with unbounded finite-level section ranks.
\end{proposition}
\noindent Formalised:
\lref{TotientCarryKernelRigidity.lean}{211}{finrank_canonicalCarryKernel_ge_of_linearIndependent}
and
\lref{TotientCarryKernelRigidity.lean}{261}{not_irrational_totientSeries_implies_unbounded_carryRank}.
The all-level independence producer is not supplied here.  Thus the
proposition records an exact conditional anti-compression consequence, not an
irrationality criterion for $S$.

\subsection{Boolean--M\"obius carry coordinates (the \#257 direction)}

At base $2$ the support series of Section~\ref{sec:eb} is itself a binary
coefficient series, $\sum_{n\in A}1/(2^n-1)=X_{f_A}$
(\lrefx{BooleanMobiusCarry.lean}{376}{erdosSupportSeries_two_eq_binaryCoeffSeries}).
Two exact coordinate changes then turn a hypothetical rational value into a
support-free object. The divisor transform and M\"obius inversion below are
classical Lambert-series infrastructure~\cite{apostol,merca2017,mercaschmidt}; the
contribution here is their Lean-checked composition with the tempered carry
interface.

The signed identity $L(\mu)=1/2$ does not by itself provide a Boolean support
for the target value.  If the negative M\"obius indices are selected and the
positive indices (apart from $1$) are discarded, the selected support sum is
exactly $1/2$ plus the positive M\"obius tail, hence exceeds $1/2$ by at least
$1/63$ from the index $6$.  This closes one sign-truncation route only; it does
not exclude another infinite Boolean support with value $1/2$.
\noindent Formalised:
\lref{MobiusSignSupportNoGo.lean}{111}{tsum_negativeMobius_eq_half_add_positiveMobiusTail}
and
\lref{MobiusSignSupportNoGo.lean}{150}{half_add_one_div_sixty_three_le_tsum_negativeMobius}.

\begin{proposition}[Boolean M\"obius inversion, both directions]\label{res:boolmob}
On positive integers, $f_A=\mathbf 1_A*\zeta$ and $\mu*f_A=\mathbf 1_A$, where
$\mathbf 1_A$ is the indicator of $A$. Conversely, every integer-valued
arithmetic function $f$ whose M\"obius transform is Boolean,
$(\mu*f)(n)\in\{0,1\}$ for all $n\ge1$, satisfies $f=f_B$ for the support
$B=\{n\ge1:(\mu*f)(n)=1\}$ selected by that transform.
\end{proposition}
\noindent Formalised:
\lref{BooleanMobiusCarry.lean}{76}{positiveSupportBitAF_mul_zeta},
\lref{BooleanMobiusCarry.lean}{94}{moebius_mul_supportCoeffAF}, and, with no
search or finiteness hypothesis, the converse
\lref{BooleanMobiusCarry.lean}{180}{eq_supportCoeffAF_booleanMobiusSupport_of_boolean};
the selected support is
\lrefx{BooleanMobiusCarry.lean}{139}{booleanMobiusSupport}.

On the carry side, Theorem~\ref{res:rigidity} applies at $c=f_A$: the support
series is rational exactly when a tempered carry orbit for $f_A$ exists
(\lrefx{BooleanMobiusCarry.lean}{383}{erdosSupportSeries_rational_iff_exists_temperedCarry});
a displayed fraction $p/q$ corresponds to an orbit $U$ with multiplier $q$ and
initial value $U(0)=p$
(\lrefx{BooleanMobiusCarry.lean}{466}{support_fraction_iff_exists_temperedCarry});
and exact division recovers the coefficient from the orbit,
$\bigl(2U(N)-U(N+1)\bigr)/q=f_A(N+1)$
(\lrefx{BooleanMobiusCarry.lean}{680}{temperedCarry_quotient_eq_supportCoeff}).
The elementary divisor-pair bound $\tau(n)\le2\lfloor\sqrt n\rfloor$
(\lrefx{BooleanMobiusCarry.lean}{208}{card_divisors_le_two_mul_sqrt}) confines
the tails to a square-root strip, $T_{f_A}(N)\le2\sqrt N+4$
(\lrefx{BooleanMobiusCarry.lean}{289}{binaryCoeffTail_supportCoeff_le_two_sqrt_add_four}),
hence every orbit state to $0<U(N)\le q\,(2\sqrt N+4)$
(\lrefx{BooleanMobiusCarry.lean}{690}{temperedCarry_pos_of_exists_pos_mem},
\lrefx{BooleanMobiusCarry.lean}{705}{temperedCarry_le_denominator_mul_two_sqrt_add_four});
and the strip is strong enough to re-derive the tempered boundary, so the
analytic side condition can be traded for one inequality. The two coordinate
changes compose.

\begin{theorem}[certificate-level equivalence]\label{res:carrycert}
Fix an integer $p$ and an integer $q\ge1$. There is a nonempty support $A$ of
positive integers with $\sum_{n\in A}1/(2^n-1)=p/q$ if and only if there is an
integer sequence $U$ with $U(0)=p$ such that: every $U(N)$ is positive;
$U(N)\le q\,(2\sqrt N+4)$; $q$ divides $2U(N)-U(N+1)$ for every $N$; and the
M\"obius transform of the quotient sequence $\bigl(2U(N)-U(N+1)\bigr)/q$ is
Boolean. The support is not guessed: it is reconstructed from the certificate
by Proposition~\ref{res:boolmob}.
\end{theorem}
\noindent Formalised:
\lref{BooleanMobiusCarry.lean}{948}{exists_normalized_support_fraction_iff_exists_booleanMobiusCarry},
with the certificate packaged as
\lrefx{BooleanMobiusCarry.lean}{775}{BooleanMobiusCarryCertificate} and the
quotient normalized at index zero
(\lrefx{BooleanMobiusCarry.lean}{766}{carryQuotient}).

\begin{example}[the support $\{2,3\}$]
The support $\{2,3\}$ has value
$\tfrac13+\tfrac17=\tfrac{10}{21}$
(\lrefx{BooleanMobiusCarry.lean}{1054}{erdosSupportSeries_support23_eq_ten_div_twenty_one}),
its orbit is the pure period-six cycle $10,20,19,17,13,26$ with multiplier
$21$ (\lrefx{BooleanMobiusCarry.lean}{1103}{carryOrbit23_isTempered}), and the
M\"obius transform of the carry quotient recovers exactly $\{2,3\}$
(\lrefx{BooleanMobiusCarry.lean}{1140}{mobius_carryOrbit23_recovers_support}).
\end{example}

These are coordinates, not by themselves a strategy: nothing here excludes
infinite Boolean carry paths, and no finite search is promoted to a
completeness argument.

\subsection{Sublogarithmic divisor coverage}

The coefficient $f_A(n)$ counts the elements of $A$ dividing $n$. Say that a
zero window of length $h$ starts after $c+N$ when
$f_A(c+N+j+1)=0$ for every $0\le j<h$. Thus no integer in that interval is
divisible by an element of the support.

\begin{theorem}[sublogarithmic zero windows]\label{res:sublog}
Let $A$ contain a positive integer and suppose
\[
  \sum_{a\in A}\frac1{2^a-1}=\frac{p}{2^c v},
  \qquad p\in\Z,\quad c\in\N,\quad v\ge1.
\]
For every $\varepsilon>0$ there is a constant $B\ge0$, depending only on
$\varepsilon,c,$ and $v$, such that every zero window after $c+N$ satisfies
\[
  h\le \varepsilon\log_2(N+1)+B,
  \qquad \log_2 x:=\frac{\log x}{\log 2}.
\]
The bound is uniform in $A,p,N,$ and $h$.
\end{theorem}
\noindent Formalised:
\lref{SublogDivisorCoverage.lean}{392}{supportCoeffZeroWindow_length_le_eps_logb_add}.

The proof first establishes, for each integer $k\ge2$, the explicit divisor
estimate
\[
  \tau(n)^k\le \bigl(k^{2^k}\bigr)^k n.
\]
It transports the resulting $n^{1/k}$ bound through the binary tail and then
uses the exact carry recurrence across a zero window. Raising the resulting
inequality to the $k$th power absorbs the occurrence of $h$ on the right and
leaves a coefficient $1/(k-1)$ in front of the binary logarithm. Choosing $k$
after $\varepsilon$ gives the theorem. This controls gaps in divisor coverage;
it does not assert a subpower bound for the carry state itself.

\subsection{Residue wraps, reciprocal mass, and unbounded carries}

Suppose now the support series equals $p/(2^c v)$ with $v$ odd and $A$
containing a positive element. Rationality makes the scaled tails integers:
$u(N)=v\,T_{f_A}(c+N)$ is a positive integer, satisfies the carry recurrence
$u(N+1)+v\,f_A(c+N+1)=2\,u(N)$, and reduces modulo $v$ to the doubling
residues $r_N=p\,2^N\bmod v$
(\lrefx{RationalSupportCarrySkeleton.lean}{668}{exists_shifted_odd_tail_nat_states_of_support_fraction}).
Doubling a least residue either stays under $v$ or wraps past it exactly once,
so each step emits a wrap digit $k_N=\lfloor2r_N/v\rfloor\in\{0,1\}$, and over
any cycle length $h$ with $2^h\equiv1\pmod v$ the repetend identity holds:
\[
\sum_{N<h}r_N \;=\; v\cdot w,
\qquad
w=\sum_{N<h}k_N,
\]
with $w\ge1$ when $p$ is coprime to $v>1$. Formalised:
\lref{RationalSupportCarrySkeleton.lean}{171}{sum_doublingResidue_eq_mul_wrapCount}
and \lrefx{RationalSupportCarrySkeleton.lean}{181}{doublingWrapCount_pos};
residues, digits, and counts are
\lrefx{RationalSupportCarrySkeleton.lean}{35}{doublingResidue},
\lrefx{RationalSupportCarrySkeleton.lean}{39}{doublingWrapDigit},
\lrefx{RationalSupportCarrySkeleton.lean}{43}{doublingWrapCount}.

\begin{proposition}[one-wrap classification]\label{res:onewrap}
Let $v>1$ be odd, $p$ coprime to $v$, and $h\ge1$ with $2^h\equiv1\pmod v$.
If the cycle of $p$ wraps exactly once, then $v=2^h-1$ and the starting
residue is a power of two: $r_0=2^a$ for some $a<h$.
\end{proposition}
\noindent The one-wrap cycles are thus exactly the Mersenne repetends
$2^a/(2^h-1)$. Formalised:
\lref{RationalSupportCarrySkeleton.lean}{415}{one_doublingWrap_classification},
from the generic closed-cycle form
\lrefx{RationalSupportCarrySkeleton.lean}{353}{one_wrap_cycle_classification},
and instantiated at the true order $h=\operatorname{ord}_v(2)$
(\lrefx{RationalSupportCarrySkeleton.lean}{478}{one_oddOrder_doublingWrap_classification},
with \lrefx{RationalSupportCarrySkeleton.lean}{436}{oddDoublingOrder}). The
classification is algebraic; a finite check of $446$ coprime starting residues
across twelve modulus/order rows is retained as kernel-checked validation, not
as the proof
(\lrefx{RationalSupportCarrySkeleton.lean}{2450}{orderWrapValidatedStartCount_eq_446},
\lrefx{RationalSupportCarrySkeleton.lean}{2454}{orderWrapValidation446_passes},
\lrefx{RationalSupportCarrySkeleton.lean}{2460}{orderWrap_minima_table_passes}).

The analytic bridge is Ces\`aro averaging. Write $\rho(A)=\sum_{a\in A}1/a$
for the reciprocal mass
(\lrefx{RationalSupportCarrySkeleton.lean}{790}{reciprocalMass}). When the
reciprocal family is summable, the mean of $f_A(1),\dots,f_A(N)$ converges to
$\rho(A)$ --- the density of the multiples of $a$, summed over the support ---
and the mean of the scaled tails has the same limit
(\lref{RationalSupportCarrySkeleton.lean}{945}{tendsto_supportCoeff_mean_reciprocalMass},
\lrefx{RationalSupportCarrySkeleton.lean}{1036}{tendsto_binaryCoeffTail_supportCoeff_mean_reciprocalMass}).
The divergent case is carried as an explicit disjunction, not through a
default value assigned to a divergent formal sum
(\lrefx{RationalSupportCarrySkeleton.lean}{796}{ReciprocalMassDivergesOrAtLeast}).
Each tail state splits exactly as $u(N)=v\,e_N+r_N$ with integral excess
$e_N=\lfloor u(N)/v\rfloor\ge0$
(\lrefx{RationalSupportCarrySkeleton.lean}{501}{oddTailExcess}).

\begin{proposition}[order-wrap lower bound, with exact excess mean]\label{res:orderwrap}
Let $A$ and $p/(2^c v)$ be as above with $v>1$ odd, and let $w$ be the wrap
count of $p$ over one cycle of length $h=\operatorname{ord}_v(2)$. Then
$\sum_{a\in A}1/a$ diverges or $\rho(A)\ge w/h$; if moreover $p$ is coprime to
$v$, then $w\ge1$, so $\rho(A)\ge1/\operatorname{ord}_v(2)$. In the summable
case the bound is the periodic part of an exact identity: the Ces\`aro mean of
the excess $e_N$ converges automatically, and
\[
\rho(A)\;=\;\frac{w}{h}\;+\;\lim_{N\to\infty}\frac1N\sum_{n<N}e_n .
\]
\end{proposition}
\noindent Formalised:
\lref{RationalSupportCarrySkeleton.lean}{1454}{support_fraction_oddOrder_wrapRatio_le_reciprocalMass},
\lref{RationalSupportCarrySkeleton.lean}{1478}{one_div_oddOrder_le_reciprocalMass_of_support_fraction},
the form retaining the divergent alternative
\lloc{RationalSupportCarrySkeleton.lean}{1582},
and the excess mean
\lref{RationalSupportCarrySkeleton.lean}{1334}{tendsto_oddTailExcess_mean}
(fraction-facing shifted form
\lrefx{RationalSupportCarrySkeleton.lean}{1577}{exists_shifted_oddTailExcess_mean_of_support_fraction},
named-limit form
\lrefx{RationalSupportCarrySkeleton.lean}{1353}{reciprocalMass_eq_wrapRatio_add_oddTailExcessMean}).
A rational value thus ties the odd part of its denominator to the sparsity of
its support: a summable support with small $\rho(A)$ can only display
denominators whose odd part has small multiplicative order of $2$.

\begin{corollary}[collision strengthening at common multiples]\label{res:collision}
In the summable case, let $F\subseteq A$ be finite and let $L$ be a common
multiple of $F$. Then
\[
\rho(A)\;\ge\;\frac{w}{h}\;+\;\frac{\lceil(|F|-1)/2\rceil}{L}.
\]
In particular, an infinite support whose value is a dyadic rational $p/2^c$
has $\sum_{a\in A}1/a$ divergent or $\rho(A)>1$.
\end{corollary}
\noindent The mechanism is a collision: at every positive multiple of $L$ at
least $|F|$ support divisors land on the same index, so $f_A\ge|F|$ there,
while the exact carry equation $f_A(n)+e_n=k_{n-1}+2e_{n-1}$ with
$k_{n-1}\le1$ lets one step absorb at most one unit --- so the preceding
excess must spike to at least $\lceil(|F|-1)/2\rceil$, and the spikes recur
with density $1/L$. Formalised:
\lref{RationalSupportCarrySkeleton.lean}{2037}{booleanCollision_wrap_bound_of_common_multiple}
(denominator-shifted form
\lloc{RationalSupportCarrySkeleton.lean}{1863}),
the carry equation
\lrefx{RationalSupportCarrySkeleton.lean}{1651}{supportCoeff_add_oddTailExcess_succ_eq_wrap_add_two_excess},
and the dyadic corollary
\lref{RationalSupportCarrySkeleton.lean}{2200}{dyadic_support_fraction_reciprocalMass_diverges_or_gt_one}
(via \lloc{RationalSupportCarrySkeleton.lean}{2003}).

\begin{theorem}[carry states are unbounded over infinite supports]\label{res:unbounded}
Let $A$ be infinite with value $p/(2^c v)$, $v\ge1$. Then the positive integer
carry state $u(N)=v\,T_{f_A}(c+N)$ is unbounded: for every $B$ there is an $N$
with $u(N)>B$.
\end{theorem}
\noindent The proof picks $2B+1$ positive support elements; at a common
multiple $L$ of all of them, the local inequality $1+v\,|F|\le2\,u(L-c-1)$
pushes the state past $B$. Formalised:
\lref{RationalSupportCarrySkeleton.lean}{2377}{exists_unbounded_shifted_odd_tail_nat_state_of_support_fraction},
from
\lrefx{RationalSupportCarrySkeleton.lean}{2320}{shifted_state_unbounded_of_infinite_support}
and the local inequality
\lrefx{RationalSupportCarrySkeleton.lean}{2228}{one_add_mul_card_le_two_mul_shifted_state};
the $\{2,3\}$ fixture revalidates it at sixty-six common multiples
(\lrefx{RationalSupportCarrySkeleton.lean}{2401}{carryOrbit23_common_multiple_bound_validation66}).
So a rational value over an infinite support cannot run on a bounded carry
alphabet. That closes finite-state readings of the carry system; it does not
close the problem. Nothing in the constraints above
prevents an infinite support from satisfying every one of these constraints.
The repetend identities and divisor averages used here are classical. We make
no claim of mathematical novelty for the coupled reciprocal-mass bounds,
collision strengthening, or global unboundedness statement.

\subsection{Greedy geometry of the \#257 value set}

The fourth module studies the set of all values a base-$2$ support series can
take. For $n\ge1$ put $w_n=1/(2^n-1)$, let $T(n)$ be the mass strictly after
exponent $n$, and let
\[
\mathcal A\;=\;\Bigl\{\,\sum_{n\in A}w_n \;:\; A\subseteq\N,\ 0\notin A\Bigr\},
\]
the \emph{Mersenne achievement set}, the exponent $0$ being normalised away as
analytically invisible. Coding by supports of positive exponents is literally
the formal support series
(\lrefx{GreedyAchievementSet.lean}{554}{positiveMersenneSupportValue_eq_erdosSupportSeries};
the set is \lrefx{GreedyAchievementSet.lean}{571}{mersenneAchievementSet}). In
this language the base-$2$ \#257 statement reads: the rational members of
$\mathcal A$ are exactly the finite subset sums. Nothing below decides that.
Kova\v{c} and Tao verify, for every fixed integer $t\ge2$, that
\[
  \sum_{\ell>n}\frac1{t^\ell-1}<\frac1{t^n-1},
\]
and deduce distinct Lambert subsums and a Cantor set~\cite{kovactao}. At
$t=2$ this is the relevant Lambert-series instance of Kakeya's classical
strict-tail criterion for subsum sets~\cite{kakeya1914}. The local
contribution is the executable Lean interface, quantitative enclosure, and
finite non-membership certificates over that known geometry; it does not
attribute those refinements to the cited sources or settle \#257.

\begin{theorem}[superincreasing geometry and greedy membership]\label{res:greedy}
For every $n\ge1$ the weight dominates the whole tail after it, $T(n)<w_n$,
with the two-scale gap $w_n-T(n)=\tfrac23\,4^{-n}+O(8^{-n})$ and explicit
valid $O$-constant $3$. Consequently a real $x$ belongs to $\mathcal A$
exactly when $x\ge0$ and every greedy residual of $x$ is at most the remaining
tail; and normalised support coding is injective, so each member of
$\mathcal A$ has a unique support, which the greedy recursion recovers bit by
bit.
\end{theorem}
\noindent Formalised:
\lref{GreedyAchievementSet.lean}{180}{mersenneTail_lt_weight},
\lref{GreedyAchievementSet.lean}{473}{mersenneGap_isBigOWith} (explicit bound
\lrefx{GreedyAchievementSet.lean}{448}{mersenneGap_asymptotic_bound}),
\lref{GreedyAchievementSet.lean}{1445}{mem_mersenneAchievementSet_iff_greedy_survival},
and
\lref{GreedyAchievementSet.lean}{1527}{positiveMersenneSupportValue_injective_normalized},
transferred to the kernel series at
\lrefx{GreedyAchievementSet.lean}{1536}{erdosSupportSeries_two_injective_normalized}.
The nonnegativity guard in the membership criterion is necessary: without it a
negative target would survive every level while lying outside $\mathcal A$.

The binary coding also permits the global geometry to be checked directly.

\begin{proposition}[fat-Cantor geometry]\label{res:greedytopology}
The achievement set $\mathcal A$ is compact, perfect, totally disconnected,
and nowhere dense.  Its Lebesgue measure is exactly one.
\end{proposition}
\noindent Formalised:
\lref{GreedyAchievementSet.lean}{996}{volume_mersenneAchievementSet},
\lref{GreedyAchievementSet.lean}{1620}{perfect_mersenneAchievementSet},
\lref{GreedyAchievementSet.lean}{1636}{isTotallyDisconnected_mersenneAchievementSet},
and
\lref{GreedyAchievementSet.lean}{1645}{isNowhereDense_mersenneAchievementSet}.
Kova\v{c}--Tao supply the strict-tail Cantor-set context cited above.  The
formalisation records these exact properties without making a novelty or
priority claim for them.

\begin{proposition}[exact rational runs and finite non-membership certificates]\label{res:death}
The greedy recursion executed in exact rational arithmetic agrees step by step
with the real recursion under casting. A finite decidable certificate --- the
exact rational residual at some level is at least an exact rational upper
enclosure of the remaining tail --- soundly proves non-membership in
$\mathcal A$; the level-one instance shows $3/4\notin\mathcal A$. And a
rational number already known to lie in $\mathcal A$ has computable support
bits.
\end{proposition}
\noindent Formalised:
\lref{GreedyAchievementSet.lean}{1130}{cast_greedyMersenneRemainderRat};
\lref{GreedyAchievementSet.lean}{1726}{certifiedGreedyMersenneDeath_not_mem},
with the certificate
\lrefx{GreedyAchievementSet.lean}{1720}{CertifiedGreedyMersenneDeath} and the
enclosure \lrefx{GreedyAchievementSet.lean}{1687}{mersenneTailUpperRat};
\lref{GreedyAchievementSet.lean}{1748}{three_fourths_not_mem_mersenneAchievementSet};
and \lref{GreedyAchievementSet.lean}{1832}{rational_member_support_bit_iff}.
The certificates are one-sided: survival through any finite depth proves
nothing about membership, and no decidability of membership in $\mathcal A$ is
asserted.  Neither the global geometry nor the finite death certificates
exclude rational values for either open series.

\section{Guide to the formal sources}\label{app:index}

The paper is meant to be read as mathematics. The blue declaration names
following formalised statements are optional links to the exact checked
theorems in the \href{\sourceurl}{archived source revision}. They preserve
line-level traceability while saying which declaration each link opens.

For a first reading, the main landmarks are Theorem~\ref{res:full} for the
full-support irrationality theorem, Theorem~\ref{res:farey} for the totient
denominator exclusion, Corollary~\ref{res:red21} for the exact
tail-certificate reduction, and Proposition~\ref{res:greedytopology} for the
geometry of the Mersenne achievement set. The public repository contains a
complete searchable cross-reference for readers who need declaration-level
detail.

% Retain the exhaustive source-link manifest in the authored TeX for automated
% validation and query tooling, but keep it outside the printed paper.
\iffalse

Each row links to the declaration at the archived source revision.

\begin{center}
\footnotesize
\begin{tabular}{@{}p{7.2cm}p{7.8cm}@{}}
\toprule
informal statement & linked Lean source\\
\midrule
$\sum 1/(b^n-1)$ irrational, all $b\ge2$ & \lrefx{CertificateKernel.lean}{8000}{irrational_erdosSum_full_support}\ (\texttt{CertificateKernel})\\
Erd\H{o}s--Borwein $E$ irrational & \lrefx{CertificateKernel.lean}{8007}{irrational_erdosBorwein_series}\ (\texttt{CertificateKernel})\\
factorial support & \lrefx{CertificateKernel.lean}{5707}{irrational_erdosSum_factorial_support}\ (\texttt{CertificateKernel})\\
power-of-two support & \lrefx{CertificateKernel.lean}{5731}{irrational_erdosSum_two_pow_support}\ (\texttt{CertificateKernel})\\
pairwise-coprime supports & \lrefx{CertificateKernel.lean}{10448}{irrational_erdosSupportSeries_pairwise_coprime}\ (\texttt{CertificateKernel})\\
eventually-periodic nonnegative rational weights & \lrefx{CertificateKernel.lean}{12483}{irrational_ratWeightSeries_eventuallyPeriodic}\ (\texttt{CertificateKernel})\\
periodic signed-weight irrational-or-terminating dichotomy & \lrefx{CertificateKernel.lean}{13847}{irrational_or_bpow_mul_eq_intCast_intWeightedErdosSeries_periodic}\ (\texttt{CertificateKernel})\\
period-four residue obstruction & \lrefx{CertificateKernel.lean}{13898}{intWeightedCoeff_periodFourSignWeight_eq_zero_of_mod_four_eq_three}\ (\texttt{CertificateKernel})\\
$L(\mu)=1/2$ & \lrefx{MersenneLambertLadder.lean}{587}{tsum_moebius_div_two_pow_sub_one_eq_half}\ (\texttt{MersenneLambertLadder})\\
$L(\ph)=2$ & \lrefx{MersenneLambertLadder.lean}{575}{tsum_totient_div_two_pow_sub_one_eq_two}\ (\texttt{MersenneLambertLadder})\\
positive lift $L(A)=S$ & \lrefx{CertificateKernel.lean}{18117}{tsum_primWeight_div_two_pow_sub_one_eq_totient_series}\ (\texttt{CertificateKernel})\\
M\"obius-square lens & \lrefx{CertificateKernel.lean}{18126}{totient_series_eq_half_add_moebius_mersenne_square}\ (\texttt{CertificateKernel})\\
$L_2(\mathbf1)$ and $L_2(\ph)$ gcd moments & \lrefx{GcdMomentCalculus.lean}{216}{tsum_one_div_mersenne_sq_eq_sigma_sub_tau_series}, \lrefx{GcdMomentCalculus.lean}{235}{tsum_totient_div_mersenne_sq_eq_gcd_moment_series}\ (\texttt{GcdMomentCalculus})\\
Stern--Brocot cylinder recursion and rate & \lrefx{GcdMomentCalculus.lean}{474}{cylinderMass_split}, \lrefx{GcdMomentCalculus.lean}{525}{sternBrocotDepthMass_error}\ (\texttt{GcdMomentCalculus})\\
$S=\tfrac12+\Pr(\gcd=1)$ & \lrefx{CertificateKernel.lean}{18229}{totient_series_eq_half_add_visible_coprime_pairs}\ (\texttt{CertificateKernel})\\
$S\ne p/q$ for $q\le\Qzero$ & \lrefx{CertificateKernel.lean}{18056}{tsum_totient_div_pow_two_ne_ratCast_of_den_le_79639646646701375323355774875831053}\ (\texttt{CertificateKernel})\\
tail-period reduction & \lrefx{TotientTailPeriodKiller.lean}{394}{irrational_totient_series_of_certificate_supply}\ (\texttt{TotientTailPeriodKiller})\\
diagonal reduction & \lrefx{LcmDiagonalReduction.lean}{112}{irrational_totient_series_of_lcm_diagonal_certificate_supply}\ (\texttt{LcmDiagonalReduction})\\
cone flatness from rationality & \lrefx{LcmConeFlatness.lean}{90}{rational_totient_series_forces_lcm_cone_flatness}\ (\texttt{LcmConeFlatness})\\
certificate completeness & \lrefx{LcmConeFlatness.lean}{316}{exists_certifiedKill_iff_tail_diff_notMem_int}\ (\texttt{LcmConeFlatness})\\
joint cone refuter & \lrefx{LcmConeNonflat.lean}{126}{exists_nonintegral_pair_of_coneNonflatCert}\ (\texttt{LcmConeNonflat})\\
verified finite instances & \lrefx{TotientTailPeriodKiller.lean}{404}{certifiedKill_all_small}\ (\texttt{TotientTailPeriodKiller})\\
full-target diagonal reduction & \lrefx{DiagonalPincerDecomposition.lean}{215}{diagonal_int_iff_foreignDiagonalDefect_hits_fullTarget}\ (\texttt{DiagonalPincerDecomposition})\\
squared-Mersenne enclosure & \lrefx{SquaredMersenneDiagonalEnclosure.lean}{116}{scaleDiagonalTailDifference_sub_lambertProjectedDiagonal}\ (\texttt{SquaredMersenneDiagonalEnclosure})\\
exact Mersenne-shadow denominator & \lrefx{MersenneShadowDenominatorGrowth.lean}{147}{lcmHeight_scaledMobiusShadow_den_exact}\ (\texttt{MersenneShadowDenominatorGrowth})\\
fresh-loss projection reduction & \lrefx{DiagonalFreshLossBridge.lean}{191}{irrational_totientSeries_of_diagonalFreshLossProjectionSupply}\ (\texttt{DiagonalFreshLossBridge})\\
prime-adjunction diamond no-go & \lrefx{FullTargetPrimeAdjunctionNoGo.lean}{196}{fullTarget_primeAdjunction_diamond_iff_root}\ (\texttt{FullTargetPrimeAdjunctionNoGo})\\
scalar-localisation height obstruction & \lrefx{AdelicHeightObstruction.lean}{23}{scalarLocalization_complement_dvd}\ (\texttt{AdelicHeightObstruction})\\
square-CRT correction suppression & \lrefx{SquareCRTCube.lean}{198}{exists_squareCRT_clean_totient_family}\ (\texttt{SquareCRTCube})\\
signed dyadic-moment algebra & \lrefx{SignedQMomentObstruction.lean}{28}{det_mul_rectangular}\ (\texttt{SignedQMomentObstruction})\\
dyadic totient certificate interface & \lrefx{TotientMahlerDefect.lean}{98}{finrank_canonicalTotientKernel_eq_of_certificate}\ (\texttt{TotientMahlerDefect})\\
residual-gauge obstruction for monomial minors & \lrefx{ResidualGaugeObstruction.lean}{55}{det_residualMonomialMatrix_ne_zero_iff}\ (\texttt{ResidualGaugeObstruction})\\
tail-orbit rationality criterion & \lrefx{GenericTailOrbitRigidity.lean}{425}{binaryCoeffSeries_rational_iff_exists_temperedBinaryOrbit}\ (\texttt{GenericTailOrbitRigidity})\\
Boolean carry certificate $\Leftrightarrow$ support fraction & \lrefx{BooleanMobiusCarry.lean}{948}{exists_normalized_support_fraction_iff_exists_booleanMobiusCarry}\ (\texttt{BooleanMobiusCarry})\\
one-wrap classification & \lrefx{RationalSupportCarrySkeleton.lean}{415}{one_doublingWrap_classification}\ (\texttt{RationalSupportCarrySkeleton})\\
order-wrap mass bound & \lrefx{RationalSupportCarrySkeleton.lean}{1478}{one_div_oddOrder_le_reciprocalMass_of_support_fraction}\ (\texttt{RationalSupportCarrySkeleton})\\
unbounded carry states & \lrefx{RationalSupportCarrySkeleton.lean}{2377}{exists_unbounded_shifted_odd_tail_nat_state_of_support_fraction}\ (\texttt{RationalSupportCarrySkeleton})\\
sublogarithmic zero windows & \lrefx{SublogDivisorCoverage.lean}{392}{supportCoeffZeroWindow_length_le_eps_logb_add}\ (\texttt{SublogDivisorCoverage})\\
greedy membership criterion & \lrefx{GreedyAchievementSet.lean}{1445}{mem_mersenneAchievementSet_iff_greedy_survival}\ (\texttt{GreedyAchievementSet})\\
Mersenne achievement set: measure one and nowhere dense & \lrefx{GreedyAchievementSet.lean}{996}{volume_mersenneAchievementSet}, \lrefx{GreedyAchievementSet.lean}{1645}{isNowhereDense_mersenneAchievementSet}\ (\texttt{GreedyAchievementSet})\\
Campbell shift-synchronization boundary & \lrefx{CampbellShiftSynchronization.lean}{299}{campbellQuarterExponent_periodFreeze_noSync}, \lrefx{CampbellShiftSynchronization.lean}{488}{shiftWindowZero_iff_half_mem_mersenneAchievementSet}\ (\texttt{CampbellShiftSynchronization})\\
\bottomrule
\end{tabular}
\end{center}
\fi

\clearpage
\begin{thebibliography}{99}
\small
\setlength{\itemsep}{0pt}
\bibitem{erdos1948} P.~Erd\H{o}s, \emph{On arithmetical properties of Lambert series},
  J. Indian Math. Soc. (N.S.) \textbf{12} (1948), 63--66,
  \href{https://users.renyi.hu/~p_erdos/1948-04.pdf}{archive scan}.
\bibitem{erdos1968} P.~Erd\H{o}s, \emph{On the irrationality of certain series},
  Math. Student \textbf{36} (1968), 222--226,
  \href{https://users.renyi.hu/~p_erdos/1969-09.pdf}{archive scan}.
\bibitem{erdosgraham1980} P.~Erd\H{o}s and R.~L. Graham, \emph{Old and New Problems and
  Results in Combinatorial Number Theory}, Monographies de l'Enseignement
  Math\'ematique, vol.~28, Universit\'e de Gen\`eve, Geneva, 1980, pp.~61--62,
  \href{https://mathweb.ucsd.edu/~ronspubs/80_11_number_theory.pdf}{author-hosted scan}.
\bibitem{erdos1988} P.~Erd\H{o}s, \emph{On the irrationality of certain series: problems and
  results}, in \emph{New Advances in Transcendence Theory}, ed.~A.~Baker,
  Cambridge University Press, 1988, pp.~102--109,
  doi:\href{https://doi.org/10.1017/CBO9780511897184.009}{10.1017/CBO9780511897184.009}.
\bibitem{borwein1992} P.~B. Borwein, \emph{On the irrationality of certain series},
  Math. Proc. Cambridge Philos. Soc. \textbf{112} (1992), no.~1, 141--146,
  doi:\href{https://doi.org/10.1017/S030500410007081X}{10.1017/S030500410007081X}.
\bibitem{apostol} T.~M. Apostol, \emph{Introduction to Analytic Number Theory},
  Undergraduate Texts in Mathematics, Springer, 1976,
  doi:\href{https://doi.org/10.1007/978-1-4757-5579-4}{10.1007/978-1-4757-5579-4}.
\bibitem{merca2017} M.~Merca, \emph{The Lambert series factorization theorem},
  Ramanujan J. \textbf{44} (2017), no.~2, 417--435,
  doi:\href{https://doi.org/10.1007/s11139-016-9856-3}{10.1007/s11139-016-9856-3}.
\bibitem{mercaschmidt} M.~Merca and M.~D. Schmidt, \emph{Generating special arithmetic
  functions by Lambert series factorizations}, Contrib. Discrete Math. \textbf{14}
  (2019), no.~1, 31--45,
  doi:\href{https://doi.org/10.55016/ojs/cdm.v14i1.62425}{10.55016/ojs/cdm.v14i1.62425}.
\bibitem{kakeya1914} S.~Kakeya, \emph{On the set of partial sums of an infinite series},
  Proc. Tokyo Math.-Phys. Soc., 2nd ser. \textbf{7} (1914), 250--251,
  doi:\href{https://doi.org/10.11429/ptmps1907.7.14_250}{10.11429/ptmps1907.7.14\_250}.
\bibitem{farey1816} J.~Farey, \emph{On a curious property of vulgar fractions},
  Philos. Mag. \textbf{47} (1816), 385--386,
  doi:\href{https://doi.org/10.1080/14786441608628487}{10.1080/14786441608628487}.
\bibitem{nesterenko} Yu.~V. Nesterenko, \emph{Modular functions and transcendence questions},
  Mat. Sb. \textbf{187} (1996), no.~9, 65--96; English transl., Sb. Math. \textbf{187} (1996), no.~9, 1319--1348,
  doi:\href{https://doi.org/10.1070/SM1996v187n09ABEH000158}{10.1070/SM1996v187n09ABEH000158}.
\bibitem{postelmansvanassche} K.~Postelmans and W.~Van Assche,
  \emph{Irrationality of $\zeta_q(1)$ and $\zeta_q(2)$},
  J. Number Theory \textbf{126} (2007), no.~1, 119--154,
  doi:\href{https://doi.org/10.1016/j.jnt.2006.11.011}{10.1016/j.jnt.2006.11.011}.
\bibitem{lucatachiya} F.~Luca and Y.~Tachiya, \emph{Irrationality of Lambert series associated
  with a periodic sequence}, Int. J. Number Theory \textbf{10} (2014), no.~3, 623--636,
  doi:\href{https://doi.org/10.1142/S1793042113501121}{10.1142/S1793042113501121}.
\bibitem{kovactao} V.~Kova\v{c} and T.~Tao, \emph{On several irrationality problems for
  Ahmes series}, Acta Math. Hungar. \textbf{175} (2025), 572--608,
  doi:\href{https://doi.org/10.1007/s10474-025-01528-0}{10.1007/s10474-025-01528-0}.
\bibitem{wanggrauribas} H.~Wang and J.~M. Grau Ribas, \emph{Positive dyadic density for
  rational weighted binary expansions}, arXiv:2606.24972 (2026),
  doi:\href{https://doi.org/10.48550/arXiv.2606.24972}{10.48550/arXiv.2606.24972}.
\bibitem{crandall} R.~Crandall, \emph{The googol-th bit of the Erd\H{o}s--Borwein constant},
  Integers \textbf{12} (2012), no.~5, 811--840,
  doi:\href{https://doi.org/10.1515/integers-2012-0007}{10.1515/integers-2012-0007}.
\bibitem{campbell} J.~M. Campbell, \emph{On the binary digits of the Erd\H{o}s--Borwein
  constant}, arXiv:2605.24160 (2026),
  doi:\href{https://doi.org/10.48550/arXiv.2605.24160}{10.48550/arXiv.2605.24160}.
\bibitem{taoteravainen} T.~Tao and J.~Ter\"av\"ainen, \emph{Quantitative correlations and
  some problems on prime factors of consecutive integers}, arXiv:2512.01739v2 (2026),
  doi:\href{https://doi.org/10.48550/arXiv.2512.01739}{10.48550/arXiv.2512.01739}.
\bibitem{lean} L.~de Moura, S.~Kong, J.~Avigad, F.~van Doorn, and J.~von Raumer,
  \emph{The Lean theorem prover (system description)}, in \emph{CADE-25}, LNCS 9195,
  Springer (2015), 378--388, doi:10.1007/978-3-319-21401-6\_26.
\bibitem{mathlib} The mathlib Community, \emph{The Lean mathematical library},
  in \emph{CPP 2020}, ACM (2020), 367--381, doi:10.1145/3372885.3373824.
\bibitem{erdosproblems} T.~Bloom (ed.), \emph{Erd\H{o}s Problems}, entries 249 and 257,
  \href{https://www.erdosproblems.com}{\texttt{erdosproblems.com}}, accessed July 2026.
\end{thebibliography}

\end{document}
