% SPDX-FileCopyrightText: 2026 Will Cook
% SPDX-License-Identifier: CC-BY-4.0
%
% A mathematician-facing exposition of the Erdos249257 Lean
% development. Every formalised statement links to the exact Lean source line
% at the pinned formal-source checkpoint. Build with
% `tectonic erdos249-257-main-paper.tex` or `make`.
\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{enumitem,float}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage[colorlinks=true,linkcolor=black,urlcolor=RoyalBlue,citecolor=black]{hyperref}
\hypersetup{
  pdftitle={Tail Certificates and Achievement-Set Geometry: Lean-checked exact reductions for Erdős 249 and the half-value branch of Erdős 257},
  pdfauthor={Will Cook},
  pdfsubject={Main expository paper for the Erdős 249/257 Lean formalisation}, pdflang={en-GB}, pdfkeywords={irrationality, Euler totient, Lambert series, achievement sets, Lean 4},
}

% ---------------------------------------------------------------------------
% Drill-down machinery: each formal anchor links to the GitHub blob at the
% pinned formal-source commit. \commit names that checkpoint; scripts/check_release.py in
% the repository verifies that it matches docs/claims.json and that every
% \lref/\lrefx/\lword/\lloc target resolves to the named declaration at that line.
% ---------------------------------------------------------------------------
% Public checkpoint: tag formal-source-2026-07-18 points to this source commit.
\newcommand{\commit}{ebc98ee71f6819bd9fe344a0bb343542c604e533}
\newcommand{\commitshort}{ebc98ee71f68}
\newcommand{\repobase}{https://github.com/wcook04/plectis-lean-erdos249-257/blob/\commit}
\newcommand{\sourceurl}{https://github.com/wcook04/plectis-lean-erdos249-257/tree/\commit}
\newcommand{\rootbase}{https://github.com/wcook04/plectis-lean-erdos249-257/blob/main}
\newcommand{\PK}{Erdos249257}
\input{module-aliases.tex}
% Formal source coordinates and exact declaration names remain in the
% hyperlink targets and verifier manifest. Default links turn the declaration
% name into a spaced, lower-case label; \lword supplies a shorter authored
% label when the surrounding exposition needs one.
\ExplSyntaxOn
\cs_new_protected:Npn \paperleanlabel #1
  {
    \tl_set:Nx \l_tmpa_tl { \tl_to_str:n {#1} }
    \regex_replace_all:nnN { _+ } { \c{space} } \l_tmpa_tl
    \regex_replace_all:nnN
      { ([a-z0-9])([A-Z]) }
      { \1\c{space}\2 }
      \l_tmpa_tl
    \tl_set:Nx \l_tmpa_tl { \tl_lower_case:n { \tl_use:N \l_tmpa_tl } }
    \tl_use:N \l_tmpa_tl
  }
\ExplSyntaxOff
\newcommand{\leanlabel}[1]{\paperleanlabel{#1}}
\newcommand{\lref}[3]{\href{\repobase/\PK/#1\#L#2}{\leanlabel{#3}}}
\newcommand{\lrefx}[3]{\href{\repobase/\PK/#1\#L#2}{\leanlabel{#3}}}
\newcommand{\lword}[4]{\href{\repobase/\PK/#1\#L#2}{#4}}
\newcommand{\lloc}[2]{\href{\repobase/\PK/#1\#L#2}{Lean source}}
% \rref{rootfile}: link a repository-root navigation file at the public main branch.
\newcommand{\rref}[1]{\href{\rootbase/\detokenize{#1}}{repository record}}
\emergencystretch=2em

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{problem}[theorem]{Problem}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Npos}{\mathbb{N}_{>0}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\ph}{\varphi}
\newcommand{\lcm}{\operatorname{lcm}}
\newcommand{\Irr}{\text{ irrational}}
\newcommand{\Qzero}{Q_0}

\title{\vspace{-4em}Exact Reductions for Two Open Irrationality Problems of Erd\H{o}s\\
{\large A Lean formalisation of Problem~\#249 and the half-value branch of Problem~\#257}}
\author{Will Cook \qquad July 2026}
\date{}

\begin{document}
\sloppy
\maketitle

\begin{abstract}
\noindent
Let $S=\sum_{n\geq1}\ph(n)/2^n$, and for $A\subseteq\Npos$ let
$X_A=\sum_{n\in A}(2^n-1)^{-1}$.  Erd\H{o}s asked whether $S$ is
irrational (Problem~\#249) and whether $X_A$ is irrational for every
infinite $A$ (Problem~\#257).  Both problems remain open.  We formalise in
Lean two exact reductions that isolate the missing unbounded step in each
question.

For \#249, irrationality of $S$ is equivalent to the existence, beyond
every bound, of a finite integer calculation plus a rigorous tail estimate
showing that two shifted binary tails differ by a non-integer.  We also prove
unconditionally that any rational representation $S=p/q$ has
$q>\Qzero\approx7.96\times10^{34}$.  At the lcm scales
$M_t=\lcm(1,\ldots,t)$, Lean checks 28 such calculations through $t=64$.
These checks remain finite; they do not prove the required unbounded supply.

For \#257, we formalise the classical irrationality theorem for the sum over
all positive exponents in every integer base $b\geq2$, together with several
previously known structured choices of infinite $A$.  At base $2$, each weight exceeds the
sum of all later weights, so every attainable value has a unique support.
The value $1/2$ is attainable exactly when its greedy expansion omits weights
at arbitrarily large ranks; this would refute the universal assertion in
\#257.  Nonattainment is equivalent to a final omission.  After a putative
final omission, the next-step recurrence has two branches.  We exclude one
branch and the integer remainder $-3$ in the other; the values $-2,-1$ and a
future-tail inequality remain unresolved.

An exact change of variables places both series in one common formula, but
transfers no irrationality theorem between them.  The contribution is a
Lean-checked map from each original question to finite objects and an exact
beyond-every-cutoff statement; it identifies but does not prove the missing
unbounded mechanism.  We prove neither the irrationality of $S$ nor the
universal assertion in \#257, and we do not decide whether $1/2$ is attainable.
\end{abstract}

\section{Introduction}\label{sec:intro}

The two questions are whether
$S=\sum_{n\ge1}\ph(n)/2^n$ is irrational (\#249), and whether
$\sum_{n\in A}(2^n-1)^{-1}$ is irrational
for every infinite $A\subseteq\Npos$ (\#257).
Here \(\ph(n)\) is Euler's totient, the number
of integers from \(1\) to \(n\) that are coprime to \(n\).  Problem \#249
appears on p.~61 of Erd\H{o}s--Graham and p.~102 of Erd\H{o}s's
survey~\cite[p.~61]{erdosgraham1980}\cite[p.~102]{erdos1988}; problem
\#257 appears on pp.~62 and 105, respectively~\cite[p.~62]{erdosgraham1980}
\cite[p.~105]{erdos1988}.  Numbering follows Bloom's
catalogue~\cite{erdosproblems}.  Neither problem is settled.

Throughout, \(\N=\{0,1,2,\ldots\}\) and
\(\Npos=\{1,2,3,\ldots\}\).  A \emph{support} \(A\subseteq\Npos\)
selects terms of the \#257 series, and its sum belongs to the corresponding
\emph{achievement set} \(\mathcal A\).  Exponent \(0\) is excluded because
\(2^0-1=0\).  An infinite support with sum \(1/2\) would therefore refute
the universal assertion in \#257.

\subsection*{What is proved}

Each question is reduced to a finite Lean-checkable calculation and one
unbounded condition that remains open.  For \#249, a \emph{certificate} is a
finite integer calculation plus a rigorous bound on the omitted tail.  Put
\(M_t=\lcm(1,\ldots,t)\) and
\(R_N=\sum_{m\geq1}\ph(N+m)2^{-m}\).  For \#257, skipping rank \(n\) means
omitting the weight \((2^n-1)^{-1}\).
\noindent The two central equivalences, stated before the later proof
machinery, are:
\[
 \begin{aligned}
 \substack{\text{finite certificates for }R_{2M_t}-R_{M_t}\notin\Z\\
   \text{at arbitrarily large }t}
 &\quad\Longleftrightarrow\quad S\notin\Q,\\[4pt]
 \substack{\text{the greedy expansion of }\tfrac12\text{ skips ranks}\\
   \text{beyond every bound}}
 &\quad\Longleftrightarrow\quad \dfrac12\in\mathcal A,\\[-1pt]
 &\quad\Longrightarrow\quad \text{a counterexample to universal \#257}.
 \end{aligned}
\]
The first line is an exact reformulation of \#249.  The second concerns only
the proposed half-value counterexample to \#257, not the universal statement.
Neither line proves the required unbounded behaviour.  The current analysis
of the final possible skip reduces the second open condition to two explicit
obligations; neither obligation is proved.

The contribution is therefore a verified map of the finite calculations and
the exact beyond-every-cutoff statements still missing.  It is not evidence
that either missing theorem is close.  After the common Lambert coordinates
are exposed, the two problems remain separate research programmes.

The totient constant also has the elementary form
\[
 S=\frac12+\Pr(\gcd(X,Y)=1),
\]
where \(X,Y\in\Npos\) are independent and
\(\Pr(X=n)=\Pr(Y=n)=2^{-n}\).  Both series are values of one Lambert
transform, but this common description transfers no theorem from one problem
to the other; Section~\ref{sec:architecture} gives the exact map.

\begin{center}
\small
\begin{tabular}{@{}
  >{\raggedright\arraybackslash}p{0.27\linewidth}
  >{\raggedright\arraybackslash}p{0.20\linewidth}
  >{\raggedright\arraybackslash}p{0.45\linewidth}@{}}
\toprule
Question or result & Status & What is proved \\
\midrule
Irrationality in \#249 & Open & No proof of \(S\notin\Q\) is claimed. \\
Denominator exclusion & Proved & If \(S=p/q\), then \(q>\Qzero\). \\
Tail differences and \(M_t\)-diagonal certificates & Exact equivalences & The unbounded supply is equivalent to \(S\notin\Q\), but remains open. \\
Diagonal computation & Finite evidence & Lean checks 28 diagonal certificates through \(t=64\); this is not cofinal. \\
Universal assertion in \#257 & Open & No proof covers every infinite support. \\
Full and structured supports & Prior work/formalised & Erd\H{o}s's full-support theorem and named restricted families are formalised; the prime-support theorem is cited, not formalised. \\
Membership of \(1/2\) & Open; exact reductions & Greedy and carry laws give equivalent events without proving they occur. \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{What ``Lean-checked'' means here.}
A linked theorem statement has a proof term accepted by Lean from the pinned
source and dependencies.  An exact equivalence remains an equivalence even
when both sides are open; a finite calculation remains finite; and a theorem
with a rationality hypothesis does not prove that a rational support exists.
The manuscript, its link checker, and its generated indexes help a reader find
the proof but are not themselves proof authority.  The release section states
the additional source-pin and trust checks required for a reproducible copy.

\paragraph{Reading map.}
Section~\ref{sec:spines} states the two exact reductions, and
Section~\ref{sec:ladder} introduces their common coordinates.
Sections~\ref{sec:eb} and~\ref{sec:249} give the \#257 and \#249 results.
Section~\ref{sec:architecture} explains the shared transform and the limits
of the comparison.  Sections~\ref{sec:verify} and~\ref{sec:further} give the
verification protocol and remaining open problems.  The appendices contain the local
arguments and audit tables.  Linked phrases open the corresponding Lean
proofs.

\medskip
\noindent\textbf{Keywords.} irrationality; Erd\H{o}s--Borwein constant;
Euler totient; Lambert series; achievement sets; binary carry systems; Lean~4.
\qquad
\textbf{MSC 2020.} 11J72 (primary); 11A25, 68V20 (secondary).

\section{The two exact reductions}\label{sec:spines}

The first equivalence below restates \#249 itself.  The second restates only
the proposed half-value counterexample to \#257.  Neither equivalence proves
the unbounded behaviour that it identifies.

\subsection{Erd\H{o}s \#249: tail differences and finite certificates}
\label{res:249story}

Write
\[
 R_N=\sum_{m\geq1}\frac{\ph(N+m)}{2^m}.
\]
Each \(R_N\) is \(2^NS\) minus an integer.  Rationality of \(S\) would
force one binary period \(h\) whose tail differences \(R_{N+h}-R_N\) are
eventually all integral; the certificates below refute exactly such
candidate periods (Section~\ref{sec:reduction}).

\paragraph{Exact reduction.}
The formal source proves the exact chain
\[
\begin{split}
 S\notin\Q
 &\quad\Longleftrightarrow\quad
   R_{N+h}-R_N\notin\Z
   \quad\text{for every }h>0\text{ and every }N\\
 &\quad\Longleftrightarrow\quad
   \text{for every }h>0,N\text{ there is an }L
   \text{ with }\mathsf{Sep}(h,N,L).
\end{split}
\]
Here \(\mathsf{Sep}(h,N,L)\) is a finite residue computation with a
rigorous tail allowance.  At fixed \(h,N\), some depth succeeds exactly when
the tail difference is non-integral; integrality of one positive-shift
difference forces \(S\) to be rational.  These statements are the
\lword{LcmConeFlatness.lean}{386}{irrational_totient_series_iff_all_tail_diffs_nonintegral}{tail-difference equivalence},
\lword{LcmConeFlatness.lean}{399}{irrational_totient_series_iff_pointwise_certificates}{pointwise certificate equivalence},
and the fixed-parameter
\lword{LcmConeFlatness.lean}{316}{exists_certifiedKill_iff_tail_diff_notMem_int}{certificate-completeness theorem}.

\paragraph{Unconditional finite baseline.}
Independently of the unbounded certificate supply, if \(S=p/q\), then
\[
 q>\Qzero
 =79\,639\,646\,646\,701\,375\,323\,355\,774\,875\,831\,053.
\]
This is the \lword{CertificateKernel.lean}{18056}{tsum_totient_div_pow_two_ne_ratCast_of_den_le_79639646646701375323355774875831053}{denominator exclusion}.

\paragraph{Missing theorem.}
An unbounded certificate supply is the exact open statement: for every
positive \(h\), non-integral differences \(R_{N+h}-R_N\) occur at
arbitrarily large \(N\).  Theorem~\ref{res:diag} gives the equivalent
one-parameter lcm-diagonal form.  The 28 changing-lcm scales through
\(t=64\) are verified, but neither this finite computation nor the
denominator bound proves the unbounded supply.

\subsection{Erd\H{o}s \#257: half-value membership and final greedy skips}
\label{res:halfstory}

Let
\[
  \mathcal A=\left\{\sum_{n\in A}\frac1{2^n-1}:A\subseteq\Npos\right\}.
\]

\paragraph{Exact reduction and finite exclusion.}
The weights satisfy
\(w_n>\sum_{m>n}w_m\), so a value has at most one support
\(A\subseteq\Npos\).  No finite such support can have value \(1/2\).  For
\(1/2\), the following are equivalent:
\[
\begin{aligned}
\frac12\in\mathcal A
&\quad\Longleftrightarrow\quad
\text{the set of exponents omitted by the greedy expansion is infinite}\\
&\quad\Longleftrightarrow\quad
\text{a successor terminal bit equal to zero occurs beyond every bound}
\end{aligned}
\]
(the terminal bit is defined in Section~\ref{sec:eb}).
Concretely, the greedy expansion of \(1/2\) skips rank \(1\), takes ranks
\(2,3\), skips \(4,5\), and takes \(6,7\); membership asks whether such
skips recur beyond every bound.

\paragraph{Counterexample consequence.}
Membership would produce an infinite support of rational sum and refute the
universal statement of Erd\H{o}s~\#257.  Thus the counterexample condition is
exact: terminal zeros beyond every bound, equivalently no last greedy skip.
The formal statements are the
\lword{GreedyAchievementSet.lean}{2514}{half_mem_mersenneAchievementSet_iff_greedySkippedSupport_infinite}{infinite-skip equivalence},
\lword{HalfCylinderHalfMembershipClassification.lean}{126}{half_mem_mersenneAchievementSet_iff_unboundedTerminalFalse}{unbounded-terminal equivalence},
\lword{HalfCylinderHalfMembershipClassification.lean}{235}{half_mem_mersenneAchievementSet_iff_no_lastHalfGreedySkip}{no-last-skip equivalence},
and the
\lword{HalfCarryReachability.lean}{589}{finite_boolSupport_ne_half}{finite-support exclusion}.

\paragraph{Missing theorem.}
The missing input is a proof that the terminal bit is zero beyond every
bound.  The
current last-skip route proves that, after a putative last skip, neither the
upper transition nor middle carry \(-3\) can occur.  It remains to exclude
middle carries \(-2,-1\) and to prove a future-tail inequality for the
remaining nonnegative middle carries.  Neither statement is known for the
greedy orbit.  The exact conditional implication is the
\lword{HalfCylinderLastProducerContradiction.lean}{387}{half_mem_mersenneAchievementSet_of_middleProducerTailEscapeExceptNegThree}{remaining-tail implication}.
Proposition~\ref{res:finalskipexclusions} gives the local exclusions and
Corollary~\ref{res:taildominance} states the sufficient global hypothesis.

\paragraph{Relation to prior work.}
Prior results are cited at their first use in the body, beginning in
Section~\ref{sec:ladder}.  No priority claim is made for the reductions above.

\section{Lambert-series identities and comparison values}\label{sec:ladder}

Whenever \(\sum_{n\ge1}|f(n)|/(2^n-1)<\infty\), define the Lambert transform by
\[
L(f)\;=\;\sum_{n\ge1}\frac{f(n)}{2^n-1}\;=\;\sum_{m\ge1}\frac{(f*\mathbf 1)(m)}{2^m},
\qquad (f*\mathbf 1)(m)=\sum_{d\mid m} f(d),
\]
the second equality being the absolutely convergent Lambert rearrangement.
It follows by expanding \((2^n-1)^{-1}=\sum_{k\geq1}2^{-nk}\), interchanging
the sums, and collecting the terms with \(m=nk\).  The coefficient
\((f*\mathbf 1)(m)\) is the Dirichlet convolution of \(f\) with the constant
function \(\mathbf 1(n)=1\).  We also write \(\mathrm{Id}(n)=n\), let \(\mu\)
denote the M\"obius function, and put \(\sigma(m)=\sum_{d\mid m}d\).
Apostol gives the classical arithmetic-function and Dirichlet-convolution
background~\cite{apostol}.  Merca introduced a Lambert-series factorisation
theorem~\cite{merca2017}, and Merca--Schmidt later unified the associated
matrix formulations~\cite{mercaschmidt}.

The rearrangement separates two operations: the divisor transform
\(f\mapsto f*\mathbf 1\), followed by evaluation of the resulting coefficients
at binary scales, \(\gamma\mapsto\sum_{m\ge1}\gamma(m)2^{-m}\).
Section~\ref{sec:architecture} uses this decomposition to compare the two open
questions.
Three inputs form the convolution chain
\[
 \alpha \xrightarrow{*\mathbf 1} \ph
 \xrightarrow{*\mathbf 1} \mathrm{Id},
 \qquad \alpha=\ph*\mu,
\]
whereas \(\mu\) and \(\mathbf 1\) are comparison inputs.  Write
\(E=L(\mathbf 1)=\sum_{n\ge1}(2^n-1)^{-1}\) for the Erd\H{o}s--Borwein
constant.  The weight \(\alpha\) is nonnegative; for example,
\(\alpha(p)=p-2\) at a prime \(p\).  The corresponding values have the
following statuses.

\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
input $f$ & value & mathematical status & treatment here\\
\midrule
\multicolumn{4}{@{}l}{\emph{Convolution chain}}\\
$\alpha=\ph*\mu$ & $L(\alpha)=S$ & open (\#249) & identities and bounds formalised\\
$\ph$ & $L(\ph)=2$ & rational & formalised\\
$\mathrm{Id}$ & $L(\mathrm{Id})=\sum\sigma(m)/2^m$ & transcendental & cited~\cite{nesterenko}\\
\addlinespace
\multicolumn{4}{@{}l}{\emph{Comparison inputs}}\\
$\mu$ & $L(\mu)=1/2$ & rational & formalised\\
$\mathbf 1$ & $L(\mathbf 1)=E$ & irrational & formalised\\
\bottomrule
\end{tabular}
\end{center}

The arrows apply only to the first three rows.  The rows for \(\mu\) and
\(\mathbf 1\) are separate comparisons, and the table records arithmetic
statuses rather than implications between them.  In particular, \(S\) and
\(E\) have the same Mersenne-denominator form, with weights \(\alpha\) and
\(\mathbf 1\), but no displayed convolution connects those weights.  The
known status of one row therefore does not transfer to another.

Nesterenko proved that
\(\mathbb Q(q,E_2(q),E_4(q),E_6(q))\) has transcendence degree at least \(3\)
when \(0<|q|<1\).  Since \(q=1/2\) is algebraic, the three Eisenstein values at
\(1/2\) are algebraically independent.  In particular,
\[
 L(\mathrm{Id})=\sum_{m\ge1}\frac{\sigma(m)}{2^m}
 =\frac{1-E_2(1/2)}{24}
\]
is transcendental~\cite{nesterenko}.

\begin{proposition}[Two rational Lambert identities]
\[
 L(\mu)=\sum_{d\ge1}\frac{\mu(d)}{2^d-1}=\frac12,
 \qquad
 L(\ph)=\sum_{d\ge1}\frac{\ph(d)}{2^d-1}=2.
\]
\end{proposition}
\noindent These equalities are formalised as the
\lword{MersenneLambertLadder.lean}{587}{tsum_moebius_div_two_pow_sub_one_eq_half}{M\"obius identity}
and the
\lword{MersenneLambertLadder.lean}{575}{tsum_totient_div_two_pow_sub_one_eq_two}{totient identity}.
M\"obius inversion gives \(\mu*\mathbf 1=(1,0,0,\ldots)\), while
\(\ph*\mathbf 1=\mathrm{Id}\); the rearranged sums are therefore
\(2^{-1}\) and \(\sum_{m\geq1}m2^{-m}=2\), respectively.

\begin{proposition}[Lambert representation of \(S\)]\label{res:lift}
With the weight $\alpha=\ph*\mu$,
\[
\sum_{d\ge1}\frac{\alpha(d)}{2^d-1}\;=\;\sum_{n\ge1}\frac{\ph(n)}{2^n}\;=\;S.
\]
\end{proposition}
\noindent The first equality is formalised as the
\lword{CertificateKernel.lean}{18117}{tsum_primWeight_div_two_pow_sub_one_eq_totient_series}{Lambert representation}.
The second is the definition of \(S\).  The proposition places \(S\) in the
same \(\sum(\cdot)/(2^d-1)\) family as \(E\), with the nonnegative weight
\(\alpha\).

\paragraph{Probabilistic reading.}
If \(X,Y\) are independent fair-coin waiting times,
\(\Pr(X=n)=2^{-n}\), then
\[
 S=\tfrac12+\Pr(\gcd(X,Y)=1).
\]
Thus Erd\H{o}s~\#249 also asks whether the probability that two independent
fair-coin waiting times are coprime is irrational.  The exact identity is
Proposition~\ref{res:coprime}; its M\"obius-square form and the periodic and
finite-algebraic variants of these identities are collected in
Appendices~\ref{app:lambert-weighted}--\ref{app:lambert-algebraic}.

\paragraph{What this section proves---and does not prove.}
The formalised contribution is
\(L(\mu)=1/2\), \(L(\ph)=2\), and \(L(\alpha)=S\); the transcendental row is
cited from Nesterenko.  None of these statuses transfers between inputs.  For
\#249 the remaining issue is still the unbounded certificate supply isolated
in Section~\ref{sec:249}; for \#257 this section supplies context rather than
either the universal theorem or a rational counterexample.

\section{Erd\H{o}s--Borwein-type irrationality (the \#257 direction)}\label{sec:eb}

We begin with the classical full-support theorem and named structured
supports.  We then specialise to base \(2\), where each weight exceeds the sum
of all later weights and hence determines a unique support.  In these
coordinates we derive necessary conditions on a hypothetical rational value,
then return to the target \(1/2\) and its final-skip reduction.  None of these
results covers arbitrary infinite supports, so the universal assertion in
Erd\H{o}s~\#257 remains open.

\subsection{Full support: the Erd\H{o}s--Borwein constant, every base}

\begin{theorem}[full-support irrationality]\label{res:full}
For every integer $b\ge2$, the series $\displaystyle\sum_{n\ge1}\frac{1}{b^n-1}$ is
irrational. In particular the Erd\H{o}s--Borwein constant $E=\sum_{n\ge1}1/(2^n-1)$
is irrational.
\end{theorem}
\noindent The
\lword{CertificateKernel.lean}{8000}{irrational_erdosSum_full_support}{all-base theorem}
and its
\lword{CertificateKernel.lean}{8007}{irrational_erdosBorwein_series}{base-\(2\) corollary}
are formalised.  This is Erd\H{o}s's 1948 theorem~\cite{erdos1948}.  In
Section~\ref{sec:ladder} it is the value $L(\mathbf 1)$, and it is the fixed
machine-checked comparison point in the same Mersenne--Lambert family as the
open $S$; as Section~\ref{sec:ladder} notes, it is not a further convolution
step from $S$.  Borwein later proved a wider rational-shift Lambert-series
theorem~\cite{borwein1992}; that result is not formalised here.

The proof begins with the Lambert identity
\[
 \sum_{n\geq 1}\frac{1}{b^n-1}
   =\sum_{m\geq 1}\frac{\tau(m)}{b^m},
\]
where $\tau(m)$ is the number of positive divisors of $m$; see the
\lword{CertificateKernel.lean}{5865}{erdosSum_full_support_eq_tsum_divisor_count}{Lambert identity}.
If the right-hand side were rational with reduced denominator $q$, then a
non-integral integer multiple could not lie at distance less than $1/q$ from an
integer.  It is therefore enough, for every $q$, to construct $N$ such that
$b^N\sum_m\tau(m)b^{-m}$ is non-integral but lies within $1/q$ of an integer.
This is the formal
\lword{CertificateKernel.lean}{5792}{irrational_of_int_mul_near_int}{near-integer criterion}.

The construction of $N$ is divided into three finite parts.  First, a bounded
use of Bertrand's postulate supplies disjoint prime blocks, and the Chinese remainder
theorem chooses an arithmetic progression on which
\[
        b^r\mid \tau(N+r) \qquad (1\leq r\leq K).
\]
These divisibilities make the first $K$ terms of the shifted tail integral.
The multiplicativity step that converts prescribed prime valuations into the
displayed divisibility is the
\lword{CertificateKernel.lean}{6177}{bpow_dvd_card_divisors_of_exact_prime_block}{prime-block divisibility lemma}.
Second, divisor pairing bounds the average of the weighted middle window
\[
   \sum_{K<r\leq L}\tau(N+r)b^{L-r}
\]
along that progression.  A pigeonhole choice then selects one translate with a
small middle contribution; see the
\lword{CertificateKernel.lean}{7246}{exists_first_block_progression_with_selected_weighted_middle}{weighted-middle selection lemma}.
Third, the elementary estimate $\tau(n)\leq n$ controls the infinite tail after
$L$.  The parameters are chosen so that the middle and far-tail bounds together
are smaller than $1/q$; the shifted tail is strictly positive because every
divisor count is positive, so the dilated sum is not an integer.

Thus the analytic series is used only in the Lambert identity and the
geometric tail estimate.  The remaining proof is a finite CRT construction,
divisor counting, and explicit parameter arithmetic.  The final certificate
existence theorem holds for every $b\geq2$ and every precision $q$; see the
\lword{CertificateKernel.lean}{7859}{exists_weighted_block_certificates}{uniform certificate theorem}.
Its
composition with the near-integer criterion is the theorem cited above.  This
is a formalisation of the classical full-support theorem, not a new
irrationality result.

\subsection{Named infinite-support cases}

Beyond full support, the development proves irrationality for a family of infinite
supports $A$, at every base $b\ge2$. Each is a formalised case of the \#257
statement family. None is the universal statement, and none is claimed as new
mathematics.

\begin{theorem}[irrationality for structured supports]\label{res:support}
For every integer $b\ge2$, the series $\sum_{n\in A} 1/(b^n-1)$ is irrational for
each of the following infinite supports $A$:
\begin{enumerate}[label=\textup{(\alph*)},leftmargin=2em]
\item factorial support $A=\{n! : n\ge1\}$, giving $\sum 1/(b^{n!}-1)$;
\item power-of-two support $A=\{2^n : n\ge0\}$, giving $\sum 1/(b^{2^n}-1)$;
\item multiples $A=d\Npos=\{dk:k\ge1\}$ with $d\ge1$;
\item pairwise-coprime supports with summable reciprocals (Erd\H{o}s's
condition~\cite{erdos1968});
\item eventually-periodic supports, residue classes, and the odd numbers.
\end{enumerate}
\end{theorem}
\noindent These sort into three groups. The classical case is (d): infinite
pairwise-coprime $A$ with $\sum_{a\in A}1/a<\infty$, formalised from Erd\H{o}s's
condition. The lcm-gap named cases are (a) and (b), whose support gaps outgrow the
running $\lcm$; the structured families are (c) and (e).\footnote{Luca--Tachiya
prove the broader eventual-periodic rational-coefficient theorem; its indicator
specialisation directly contains case (e)~\cite{lucatachiya}. Their
Chowla--Erd\H{o}s-style large-modulus proof is distinct from the periodic-divisor
certificate route formalised here, and the full mixed-sign coefficient theorem is
not claimed.}
\noindent Formalised case by case by the
\lword{CertificateKernel.lean}{5707}{irrational_erdosSum_factorial_support}{factorial-support theorem},
\lword{CertificateKernel.lean}{5731}{irrational_erdosSum_two_pow_support}{power-of-two-support theorem},
\lword{CertificateKernel.lean}{8775}{irrational_erdosSupportSeries_multiples}{multiple-support theorem},
\lword{CertificateKernel.lean}{10448}{irrational_erdosSupportSeries_pairwise_coprime}{pairwise-coprime-support theorem},
and
\lword{CertificateKernel.lean}{11276}{irrational_erdosSupportSeries_eventuallyPeriodic}{eventually-periodic-support theorem}.
The residue-class and odd supports in part~\textup{(e)} are direct
specialisations of the last theorem; the first two theorems already hold for
every base \(b\geq2\).

The prime support is now known unconditionally: Tao and Ter\"av\"ainen proved
that
\[
  \sum_p\frac1{2^p-1}=\sum_{n\ge1}\frac{\omega(n)}{2^n}
\]
is irrational~\cite{taoteravainen}. Their proof uses quantitative two-point
correlations of multiplicative functions and is not formalised here. This
settles the prime-support case, while the universal support statement
remains open.

Cases~\textup{(a)} and~\textup{(b)} both use the lcm-gap criterion.  The
power-of-two case is not a Liouville argument: before \(2^k\), the prefix lcm is
\(2^{k-1}\), so the gap has only the same size as the lcm.  Period noncollapse
gives the finite input: the order of \(b\) modulo the reduced denominator of a
partial sum equals that prefix lcm.  The denominator therefore exceeds the lcm,
while the additive gap to the next exponent still tends to infinity; the tail
is then too small for a rational limit.

The results above enlarge the family of supports for which irrationality is
known, but they do not supply the universal quantifier.  We now fix the base at
\(2\).  The support sums form one achievement set, and an infinite-support
counterexample is a rational point of this set whose unique support is
infinite.  We first derive necessary conditions for such a rational point and
then specialise to \(1/2\), for which finite support is impossible.

\subsection{The base-$2$ achievement set}

For \(n\geq1\), put
\[
 w_n=\frac1{2^n-1}.
\]
For \(n\geq0\), put
\[
 T_n=\sum_{m>n}w_m,
 \qquad
 \mathcal A=\left\{\sum_{n\in A}w_n:A\subseteq\N,\ 0\notin A\right\}.
\]
For \(x\geq0\), define the greedy residuals by \(r_0(x)=x\) and
\[
 \varepsilon_n(x)=
 \begin{cases}
  1,&w_n\leq r_{n-1}(x),\\
  0,&w_n>r_{n-1}(x),
 \end{cases}
 \qquad
 r_n(x)=r_{n-1}(x)-\varepsilon_n(x)w_n.
\]
When \(\varepsilon_n(x)=1\) we say the expansion \emph{takes} rank \(n\);
otherwise it \emph{skips} rank \(n\).

\begin{theorem}[strict-tail greedy coding]\label{res:greedy}
For every \(n\geq1\),
\[
 T_n<w_n.
\]
Hence every \(x\in\mathcal A\)
has a unique support by positive exponents (the ``normalised'' support in the
formal source).
Moreover, a real number \(x\) lies in \(\mathcal A\) if and only if \(x\geq0\)
and \(r_n(x)\leq T_n\) for every \(n\geq0\).  In that case
\(x=\sum_{n\geq1}\varepsilon_n(x)w_n\).
\end{theorem}
\noindent The strict-tail inequality and greedy criterion are the
\lword{GreedyAchievementSet.lean}{180}{mersenneTail_lt_weight}{strict-tail theorem}
and
\lword{GreedyAchievementSet.lean}{1445}{mem_mersenneAchievementSet_iff_greedy_survival}{greedy membership theorem}.
Uniqueness is the
\lword{GreedyAchievementSet.lean}{1527}{positiveMersenneSupportValue_injective_normalized}{support-coding theorem}.
Kova\v{c}--Tao provide the strict-tail Cantor-set context~\cite{kovactao}.

\paragraph{Geometric and finite-arithmetic consequences.}
Proposition~\ref{res:greedytopology} proves that \(\mathcal A\) is compact,
perfect, totally disconnected, nowhere dense, and of Lebesgue measure \(1\).
Proposition~\ref{res:death} shows that the greedy recursion in exact rational
arithmetic agrees step by step with the real recursion and gives sound finite
certificates of nonmembership.
For the half-value question, Theorem~\ref{res:greedy} makes the candidate
support canonical, while compactness allows one to pass from cofinal finite
approximations to a point of \(\mathcal A\).  The measure-one statement does
not decide whether \(1/2\in\mathcal A\); a finite certificate can prove
nonmembership, but survival through any finite depth proves nothing about
membership.  These consequences use the same greedy coordinates but are not
clauses of the membership equivalence.  Their detailed statements and the
one-sided certificate boundary are in Appendix~\ref{app:carry-geometry}.  No
novelty claim is made for the strict-tail geometry.  Before specialising to
\(1/2\), we record the carry constraints forced by any rational
\(X_A\in\mathcal A\); these are necessary conditions in the universal
direction, not yet a contradiction.

\subsection{Rigidity of a hypothetical rational support}

For \(A\subseteq\Npos\), define
\[
 f_A(m)=\#\{a\in A:a\mid m\},
 \qquad
 X_A=\sum_{a\in A}\frac1{2^a-1}
     =\sum_{m\geq1}\frac{f_A(m)}{2^m},
 \qquad
 T_f(N)=\sum_{j\geq1}\frac{f(N+j)}{2^j}.
\]
Thus \(f_A(m)\) counts the selected exponents dividing \(m\), \(X_A\) is the
original support sum, and \(T_f(N)\) is the binary tail of a coefficient
sequence after position \(N\).

\begin{theorem}[rational-support carry recurrence]\label{res:rigidity-main}
Suppose that \(A\neq\varnothing\) and
\[
 X_A=\frac{p}{2^c v},
 \qquad p\in\Z,\quad c\in\N,\quad v\geq1\ \text{odd}.
\]
Then \(u_N=vT_{f_A}(c+N)\) is a positive integer for every \(N\), and
\[
 u_{N+1}+v f_A(c+N+1)=2u_N. \tag{4.1}
\]
\end{theorem}
\noindent This is formalised by
\lword{RationalSupportCarrySkeleton.lean}{668}{exists_shifted_odd_tail_nat_states_of_support_fraction}{the shifted natural-state theorem}.
The recurrence is binary long division: since
\(2T_f(M)=f(M+1)+T_f(M+1)\), doubling the scaled tail separates the next
coefficient, and rationality is what makes the states \(u_N\) integers.
Its tail-orbit normal form is Theorem~\ref{res:rigidity}.  A two-element
support, with its explicit period-six orbit and multiplier \(21\), is worked
at the end of Appendix~\ref{app:carry-boolean}.
The recurrence is a necessary consequence of rationality, not a criterion by
itself.  Wang uses the same forward mechanism for
\(\gamma(n)=n d_n\)~\cite{wang2026}; Appendix~\ref{app:carry-normal-form}
separates that antecedent from the generic converse and rigidity statements
formalised here.  Wang's density theorem and its Erd\H{o}s~\#260 corollary
are not used.

\begin{corollary}[unbounded-state consequence]
\label{res:rigidity-unbounded-main}
Under the hypotheses of Theorem~\ref{res:rigidity-main}, if \(A\) is
infinite, then \((u_N)\) is unbounded.
\end{corollary}
\noindent This is formalised by
\lword{RationalSupportCarrySkeleton.lean}{2377}{exists_unbounded_shifted_odd_tail_nat_state_of_support_fraction}{the fraction-facing unbounded-state theorem};
see also Theorem~\ref{res:unbounded}.

\begin{corollary}[sublogarithmic coverage consequence]
\label{res:rigidity-sublog-main}
Under the hypotheses of Theorem~\ref{res:rigidity-main}, for every
\(\varepsilon>0\) there is
\(B=B(\varepsilon,c,v)\) such that every interval
\(\{c+N+1,\ldots,c+N+\ell\}\) on which \(f_A\) vanishes satisfies
\[
 \ell\leq\varepsilon\log_2(N+1)+B;
\]
\end{corollary}
\noindent This is formalised by
\lword{SublogDivisorCoverage.lean}{392}{supportCoeffZeroWindow_length_le_eps_logb_add}{the sublogarithmic zero-window theorem};
see also Theorem~\ref{res:sublog}.

\begin{corollary}[odd-denominator reciprocal-mass consequence]
\label{res:rigidity-mass-main}
Under the hypotheses of Theorem~\ref{res:rigidity-main}, if
\(\rho(A)=\sum_{a\in A}a^{-1}<\infty\), \(v>1\), and
\(h=\operatorname{ord}_v(2)\), then
\[
 \rho(A)=\frac{w}{h}
   +\lim_{M\to\infty}\frac1M\sum_{N<M}e_N\geq\frac{w}{h},
\]
where \(w\) is the number of wraps in one doubling-residue cycle and
\(e_N=\lfloor u_N/v\rfloor\) is the integral excess in the split
\(u_N=ve_N+\bar p_N\), with \(\bar p_N\) the least residue of
\(p2^N\) modulo \(v\).  A step wraps precisely when doubling
\(\bar p_N\) crosses \(v\); thus \(w\) counts the resulting carry digits
over one cycle of length \(h\).  The term \(w/h\) is the cycle's mean
residue contribution, while the displayed excess mean is nonnegative.  If
\((p,v)=1\), then \(w\geq1\).
\end{corollary}
\noindent The exact excess-mean identity is
\lword{RationalSupportCarrySkeleton.lean}{1577}{exists_shifted_oddTailExcess_mean_of_support_fraction}{the shifted excess-mean theorem};
see also Proposition~\ref{res:orderwrap}.

\begin{corollary}[dyadic reciprocal-mass consequence]
\label{res:rigidity-dyadic-mass-main}
Under the hypotheses of Theorem~\ref{res:rigidity-main}, if \(v=1\) and
\(A\) is infinite, then either \(\rho(A)\) diverges or \(\rho(A)>1\).
\end{corollary}
\noindent This is formalised by
\lword{RationalSupportCarrySkeleton.lean}{2200}{dyadic_support_fraction_reciprocalMass_diverges_or_gt_one}{the dyadic reciprocal-mass alternative};
see also Corollary~\ref{res:collision}.

Read together, the four statements say that rationality would force an
unbounded integer state sequence for an infinite support, severely limit
intervals containing no support divisor, and constrain the reciprocal mass
according to whether the reduced denominator has an odd factor.  They are
independent necessary consequences of one rationality premise, not a
contradiction.
Appendices~\ref{app:carry-normal-form}--\ref{app:carry-boolean} supply the
common recurrence and Boolean certificate; Appendices~\ref{app:carry-sublog}
and~\ref{app:carry-mass} supply the independent filters.  No corollary above
appears as a hypothesis of another; their appendix proofs do share the carry
machinery, in the order displayed in Appendix~\ref{app:carry-mass}.

\begin{corollary}[Boolean--M\"obius orbit form of universal \#257]
\label{res:universal-boolean-obstruction}
The universal assertion in Erd\H{o}s~\#257 is equivalent to the following
orbit statement: for every rational number \(p/q\), every
Boolean--M\"obius carry certificate for \(p/q\) reconstructs a finite
support.  Equivalently, the universal assertion fails exactly when one such
certificate reconstructs an infinite support.
\end{corollary}
\noindent Concretely, a certificate is an integer sequence \(U\) with
\(U(0)=p\), positive states bounded by \(q(2\sqrt N+4)\), and
\(q\mid 2U(N)-U(N+1)\).  Dividing these differences by \(q\) gives an
integer coefficient sequence \(f_U\).  Its Boolean M\"obius condition is
\((\mu*f_U)(n)\in\{0,1\}\), and inversion reconstructs the support as
\(A_U=\{n\geq1:(\mu*f_U)(n)=1\}\).  These are exactly the conditions in
Theorem~\ref{res:carrycert}.  That theorem is the Lean-checked
fraction-by-fraction equivalence
\lword{BooleanMobiusCarry.lean}{948}{exists_normalized_support_fraction_iff_exists_booleanMobiusCarry}{between supports and carry certificates};
the corollary merely quantifies it over \(p/q\) and separates finite from
infinite reconstructed supports.  This is an exact reformulation, not a
finite algorithm: the orbit condition and the support it reconstructs are
infinite objects.

\subsection{The target \(1/2\)}

\paragraph{Why this target.}
Any infinite support with any rational value would refute the universal
statement.  Fixing \(1/2\) turns one possible counterexample route into the
point-membership question \(1/2\in\mathcal A\).  This target is convenient,
not forced.

The reason it is distinguished is the signed M\"obius identity
\(L(\mu)=1/2\) from Section~\ref{sec:ladder}.  A support series is different:
each term has coefficient \(0\) or \(1\).  The most immediate conversion,
selecting exactly the indices with \(\mu(d)=-1\), overshoots \(1/2\) by at
least \(1/63\) (Proposition~\ref{res:mobiussignnogo}).  Moreover, no finite
support attains \(1/2\) (Proposition~\ref{res:nofinitehalf} below), while the
strict-tail coding of Theorem~\ref{res:greedy} makes any candidate support
unique.

Let
\[
 \mathcal S(x)=\{n\geq1:\varepsilon_n(x)=0\}
\]
be the set of exponents omitted by the greedy expansion.

\paragraph{A worked start.}
The greedy expansion of \(1/2\) begins: \(w_1=1>\tfrac12\), so rank \(1\)
is skipped; \(w_2=\tfrac13\) is taken, leaving \(\tfrac16\);
\(w_3=\tfrac17\) is taken, leaving \(\tfrac1{42}\); ranks \(4,5\) are
skipped because \(w_4=\tfrac1{15}\) and \(w_5=\tfrac1{31}\) exceed
\(\tfrac1{42}\); and \(w_6=\tfrac1{63}\), \(w_7=\tfrac1{127}\) are taken,
leaving
\[
 \frac12-\Bigl(\frac13+\frac17+\frac1{63}+\frac1{127}\Bigr)
 =\frac1{16002}.
\]
The exact calculation continues with the next take at rank \(14\), and through
rank \(26\) it takes exactly
\[
 2,3,6,7,14,20,21,26.
\]
The finite seam model introduced below reproduces this prefix uniformly for
every row \(s\geq27\); this is the
\lword{HalfFinalMiddlePhaseSieve.lean}{1190}{seamGreedyBits_head2367_14_20_21_26}{kernel-checked rank-(26) prefix}
used again in Appendix~\ref{app:half-final-skip}.

This is only an initial segment.  Membership of \(1/2\) in \(\mathcal A\)
requires the complementary skips to recur beyond every bound.  Nonmembership
has a finite alternative: one skip is the \emph{last}, precisely when its
residual lies strictly above the whole remaining tail
(Appendix~\ref{res:fixedtailsocket}).  Theorem~\ref{res:fatalright} below
identifies nonmembership with the existence of such a skip.

The next definition replaces the real greedy calculation by an integer one.
A \emph{row} \(s\) keeps the ranks \(2,\ldots,s-1\), rescales every weight by
\(4^s\), rounds down, and uses a target \(2^s\) below the rescaled value
\(4^s/2\).  The resulting zero--one word records the finite take/skip
decisions; its last entry is the terminal bit used in
Theorem~\ref{res:halfmembership}.

\begin{definition}[finite greedy seam and terminal bit]\label{def:terminal-bit}
For an integer \(s\geq6\), set
\[
 q_{s,d}=\lfloor4^sw_d\rfloor
 =\left\lfloor\frac{4^s}{2^d-1}\right\rfloor
 \quad(2\leq d<s),
 \qquad
 K_s=4^s\cdot\frac12-2^s=2^{2s-1}-2^s.
\]
Starting from \(R_{s,2}=K_s\), define successively, for \(2\leq d<s\),
\[
 \beta_{s,d}=
 \begin{cases}
  1,&q_{s,d}\leq R_{s,d},\\
  0,&q_{s,d}>R_{s,d},
 \end{cases}
 \qquad
 R_{s,d+1}=R_{s,d}-\beta_{s,d}q_{s,d}.
\]
The Boolean word \((\beta_{s,2},\ldots,\beta_{s,s-1})\) is the
\emph{finite greedy seam at row \(s\)}, and
\(\tau_s=\beta_{s,s-1}\) is its terminal bit.
\end{definition}

\noindent For example, at \(s=7\) we have \(K_7=8064\), and the run takes
\(q_{7,2}=5461\)
and \(q_{7,3}=2340\), skips \(1092\) and \(528\), and takes
\(q_{7,6}=260\), leaving \(3\): the word is \(1,1,0,0,1\), matching the
real greedy decisions at ranks \(2\)--\(6\), with terminal bit
\(\tau_7=1\).  The seam
is thus an exact integer coordinate model of the real greedy process.
Appendix~\ref{app:half-final-skip} develops the equivalence with
final skips and the unresolved transition cells; the finite coordinate changes in
Appendix~\ref{app:half-evidence} do not strengthen that equivalence.

\begin{theorem}[exact half-membership classification]\label{res:halfmembership}
The following are equivalent:
\[
\begin{aligned}
\textup{(i)}\;&\quad \frac12\in\mathcal A;\\
\textup{(ii)}\;&\quad \mathcal S(1/2)\ \text{is infinite};\\
\textup{(iii)}\;&\quad \mathcal S(1/2)\ \text{has no largest element};\\
\textup{(iv)}\;&\quad
  \text{for every }N\text{ there is }s\geq\max\{N,6\}
  \text{ with }\tau_s=0.
\end{aligned}
\]
\end{theorem}
\noindent The omitted-set equivalence is the
\lword{GreedyAchievementSet.lean}{2514}{half_mem_mersenneAchievementSet_iff_greedySkippedSupport_infinite}{infinite-skip theorem};
the terminal forms are the
\lword{HalfCylinderHalfMembershipClassification.lean}{126}{half_mem_mersenneAchievementSet_iff_unboundedTerminalFalse}{unbounded-terminal theorem}
and
\lword{HalfCylinderHalfMembershipClassification.lean}{235}{half_mem_mersenneAchievementSet_iff_no_lastHalfGreedySkip}{no-final-skip theorem}.

\begin{proposition}[finite half-value exclusion]\label{res:nofinitehalf}
No finite support \(A\subseteq\Npos\) has value \(1/2\).
\end{proposition}
\noindent This is the
\lword{HalfCarryReachability.lean}{589}{finite_boolSupport_ne_half}{finite-support theorem}.

\begin{corollary}[counterexample from half-membership]\label{res:halfcounterexample}
Any one of the equivalent conditions in
Theorem~\ref{res:halfmembership} produces an infinite support
\(A\subseteq\Npos\) with \(X_A=1/2\), and therefore refutes the
universal form of Erd\H{o}s~\#257.
\end{corollary}
\noindent This combines the unique greedy support from
Theorem~\ref{res:greedy} with Proposition~\ref{res:nofinitehalf}.
A tempting signed-to-Boolean shortcut also fails: selecting precisely the
negative M\"obius indices overshoots \(1/2\) by at least \(1/63\).
Proposition~\ref{res:mobiussignnogo} gives the exact sign-separation identity.

Theorem~\ref{res:halfmembership} therefore leaves one question: does the
greedy orbit of \(1/2\) skip cofinally?  It does not answer that question.  The
next subsection characterises nonmembership by a final skip and identifies
what would be required to exclude it.

\subsection{The final-skip reduction}

Appendix~\ref{app:half-final-skip} contains the detailed derivations used in
this subsection; the statements and their logical status are recorded here.
To detect whether a skip can be final, we compare the ranks selected in
consecutive seam rows.

For each \(s\geq6\), let
\[
 G_s=\{d:2\leq d<s,\ \beta_{s,d}=1\}.
\]
Thus \(G_s\) is the set chosen by the integer greedy run at row \(s\).
Among all subsets of \(\{2,\ldots,s-1\}\) whose integer weight exceeds
\(K_s\), let \(H_s\) be the unique one with the smallest weight.  It is the
cheapest strict overshoot of the target.

The recurrence from row \(s\) to row \(s+1\) has three forms:
\[
\begin{aligned}
 \textup{upper:}&\quad G_{s+1}=H_s,\\
 \textup{middle:}&\quad G_{s+1}=G_s,\\
 \textup{right:}&\quad G_{s+1}=G_s\cup\{s\}.
\end{aligned}
\]
An upper step resets the selected ranks to \(H_s\).  A middle step keeps
\(G_s\) and skips the new rank \(s\); a right step keeps \(G_s\) and takes
that rank.  A final skipped exponent \(D\) is therefore an upper or middle
transition followed only by right transitions.  The terminal bit at row
\(13\) is false.  Consequently, if the seam is eventually right, its last
false terminal row satisfies \(D\geq13\); the lower bound comes from this
calculation rather than an omitted range of cases.  This is the
\lword{HalfCylinderLastProducerContradiction.lean}{178}{exists_last_false_terminal_of_eventuallyRight}{last-false row theorem}.

At a middle transition \(D\geq13\), the next question is whether restoring the
skipped rank \(D\) still leaves the support below \(1/2\).  Let
\[
 A_D=G_D\cup\{D\},\qquad f_A(n)=\#\{a\in A:a\mid n\},
\]
and put
\[
 P_D=2f_{G_D}(2D+1)+f_{G_D}(2D+2),
 \quad C_D=4R_{D,D}-P_D-4,
 \quad \Theta_D=\sum_{j\geq1}\frac{f_{A_D}(2D+2+j)}{2^j}.
\]
Here \(C_D\in\Z\) is the transition carry and \(\Theta_D\) is the complete
future divisor-incidence tail, not a finite truncation.  The residual identity of
Appendix~\ref{res:producercarry} compares the two quantities exactly:
\[
 \frac12-\sum_{a\in A_D}w_a
 =2^{-(2D+2)}\bigl(C_D-\Theta_D\bigr),
\]
so \(\Theta_D<C_D\) is equivalent to
\(\sum_{a\in A_D}w_a<\tfrac12\).

For a finite set \(u\subseteq\{1,\ldots,d\}\), write
\(V(u)=\sum_{n\in u}w_n\).  Once this prefix through rank \(d\) is fixed,
the next decision is whether to take \(w_{d+1}\).  If it is skipped, even
taking every later weight reaches at most \(V(u)+T_{d+1}\); if it is taken,
the resulting sum is at least \(V(u)+w_{d+1}\).  We call \((u,d)\) a
\emph{fatal half-gap} when \(1/2\) lies strictly between these two bounds:
\[
 V(u)+T_{d+1}<\frac12<V(u)+w_{d+1}.
\]
No support agreeing with \(u\) through rank \(d\) can then represent \(1/2\):
extensions omitting rank \(d+1\) fall short, while those containing it
overshoot.

\begin{theorem}[final-skip classification]
\label{res:fatalright}
Nonmembership of \(1/2\) in \(\mathcal A\) is equivalent to each of the
following:
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.2em]
\item the integer seam recurrence is eventually right;
\item a fatal half-gap exists;
\item the real greedy expansion of \(1/2\) has a final skipped exponent.
\end{enumerate}
\end{theorem}
\noindent The three equivalent nonmembership criteria are formalised in
\lword{HalfCylinderFatalGapRightTail.lean}{781}{seamGreedyEventuallyRight_iff_existsFatalHalfGap}{fatal-gap classification},
\lword{HalfCylinderFatalGapRightTail.lean}{787}{seamGreedyEventuallyRight_iff_half_not_mem}{nonmembership classification},
and
\lword{HalfCylinderFatalGapRightTail.lean}{802}{seamGreedyEventuallyRight_iff_exists_isLastHalfGreedySkip}{final-skip classification}.

\begin{proposition}[excluded final-skip branches]\label{res:finalskipexclusions}
At a final skipped exponent \(D\geq13\), the upper successor is impossible.
On the middle branch, the only carry values outside the ranges
\(C_D\leq-4\) and \(C_D\geq0\) are \(C_D\in\{-3,-2,-1\}\).  At a final
skip, the value \(C_D=-3\) is impossible; the values \(-2\) and \(-1\)
remain unresolved.
\end{proposition}
\noindent The upper and \(C_D=-3\) exclusions are the
\lword{HalfCylinderLastProducerContradiction.lean}{250}{upperProducer_not_last}{upper-branch exclusion}
and
\lword{HalfCylinderLastProducerContradiction.lean}{315}{middleProducer_neg_three_not_last}{middle-\(-3\) exclusion};
the upper case is proved directly at the linked declaration, and
Appendix~\ref{app:half-final-skip} sketches a prose argument only for the
\(-3\) cell.
The three exceptional values are identified by the
\lword{HalfCylinderMiddleCarryLowerBound.lean}{3090}{seamMiddleCoordinate_not_safe_iff_three_cells}{three-cell classification}.

\begin{corollary}[tail-dominance membership criterion]\label{res:taildominance}
If every middle transition at a row \(D\geq13\) with \(C_D\neq-3\) satisfies
\[
 \Theta_D<C_D, \tag{4.2}
\]
then no final skipped exponent exists, and hence \(1/2\in\mathcal A\).
\end{corollary}
\noindent This implication is the
\lword{HalfCylinderLastProducerContradiction.lean}{387}{half_mem_mersenneAchievementSet_of_middleProducerTailEscapeExceptNegThree}{tail-dominance theorem}.
Since \(\Theta_D\geq0\), inequality~\textup{(4.2)} cannot hold when
\(C_D=-2\) or \(-1\).  The corollary therefore has two open inputs.  First,
these two carry values must be shown not to occur at any actual middle
transition.  Second, every remaining middle transition with \(C_D\geq0\)
must have complete future tail smaller than \(C_D\).  Neither input is proved
for the greedy orbit.  The word ``every'' is essential: one final middle
transition is enough for nonmembership, so a bound that failed at even one
row would not rule it out.

\paragraph{Logical status of the final-skip argument.}
The proved levels of the argument are
\[
\begin{array}{rcl}
 \frac12\notin\mathcal A
 &\Longleftrightarrow
   \text{a final skipped exponent exists}
 &[\text{proved}],\\
 \text{final upper, or final middle with }C_D=-3
 &\Longrightarrow\bot
 &[\text{proved}].
\end{array}
\]
These lines do not imply membership.  The unproved hypothesis has two parts:
exclude all actual middle transitions with \(C_D\in\{-2,-1\}\), and
prove~\textup{(4.2)} at every remaining non-\((-3)\) middle transition.
Together they establish the global hypothesis of
Corollary~\ref{res:taildominance}; only then does
Corollary~\ref{res:halfcounterexample} produce an infinite rational-valued
support and a counterexample to the universal \#257 statement.

\long\def\HalfValueTechnicalAppendix{

\subsection{Exact band for a final skip between two takes}
\label{res:twothirdsband}
\label{app:half-localizer}

At rank \(k\), call the current residual \emph{dyadically safe} when it is at
most \(2^{-k}\), and \emph{unsafe} otherwise.  Since the actual Mersenne tail
satisfies \(T_k>2^{-k}\), safety is sufficient but not necessary for
survival; an unsafe skip need not stop the orbit.

Suppose a take at rank \(b\) is followed by the next take at \(c>b+1\).  Put
\(q=2^b-1\), \(m=2^{c-1}\), and write the residual just before rank \(b\) as
\(\rho=1/R\), so \(R\) is its reciprocal.  Taking \(w_b=1/q\) changes the
reciprocal to \(R'=Rq/(q-R)\), which stays fixed through the intervening
skips.  The last skip, at rank \(c-1\), occurs exactly when \(R'>m-1\) and is
dyadically unsafe exactly when \(R'<m\).  For \(0<R<q\), solving
\(m-1<R'<m\) for the pre-take reciprocal gives
\[
  \frac{q(m-1)}{q+m-1}<R<\frac{qm}{q+m}.
\]
The interval has exact width
\[
  \frac{q^2}{(q+m)(q+m-1)}.
\]
Formalised:
\lword{HalfGreedyTwoThirdsBand.lean}{88}{postTakeUnsafeAt_iff_band}{the general unsafe-band equivalence}
and
\lword{HalfGreedyTwoThirdsBand.lean}{104}{band_width_general}{the general band-width formula}.

For a single intervening skip, $c=b+2$ and $m=2q+2$.  The band becomes
\[
  \frac{q(2q+1)}{3q+1}<R<\frac{2q(q+1)}{3q+2},
\]
with width $q^2/((3q+1)(3q+2))<1/9$, entirely between $2q/3$ and
$2q/3+2/9$.  No integral \(R\) lies in this window.  If instead
\(\rho=p/(2D)\), then \(R=2D/p\); for positive odd \(p,D,q\), clearing
denominators in the same inequalities forces \(p\ge7\).  Formalised:
\lword{HalfGreedyTwoThirdsBand.lean}{127}{singletonUnsafe_iff_twoThirdsBand}{the one-skip band equivalence},
\lword{HalfGreedyTwoThirdsBand.lean}{154}{band_width}{the one-skip width formula},
\lword{HalfGreedyTwoThirdsBand.lean}{163}{band_width_lt_ninth}{the width bound},
\lword{HalfGreedyTwoThirdsBand.lean}{185}{not_twoThirdsBand_of_int}{the integral-reciprocal exclusion},
and
\lword{HalfGreedyTwoThirdsBand.lean}{231}{seven_le_of_intBand_odd}{the odd-coordinate lower bound}.

These results neither show that the greedy orbit of \(1/2\) enters or avoids
the band nor identify the remaining transition coordinates \(-2,-1\) with
dyadically unsafe skips.

\subsection{Conditional criteria and local inputs for structured supports}
\label{app:half-conditional-producers}

The results here have three distinct roles.  The pairwise-coprime dilation
criterion is a conditional irrationality theorem for one structured support family.  The
seam, strip, and terminal criteria are sufficient hypotheses for
\(1/2\in\mathcal A\).  The dilation, prime-power, and reverse-carry spacing
results are local algebraic inputs that prove neither an irrationality conclusion nor
half-membership without an additional supply hypothesis.  Nothing in this subsection is a new unconditional case.

\subsubsection{A conditional pairwise-coprime dilation criterion}\label{res:sunflower}

Write $A=E\cup\{c_ip_i:i\geq0\}$, where $E$ is a finite set of positive
integers and $Q,c_i,p_i$ are positive.  Suppose that $Q$ is divisible by every
$d\in E$ and every $c_i$, while $p_i>1$ is coprime to $Q$ and to $p_j$ for
$j\ne i$, and $\sum_i1/p_i<\infty$.  We call this a
\emph{finite-core pairwise-coprime dilation family}: $Q$ contains the fixed
factors and the $p_i$ are mutually coprime.  Lean includes the example $E=\varnothing$,
$Q=6$, with $c_i$ alternating between $2,3$ and
$p_i=5^2,7^2,11^2,\ldots$; each support element has three prime factors with
multiplicity.

Put $f_A(n)=\#\{d\in A:d\mid n\}$ and
$T_A(M)=\sum_{j\geq1}f_A(M+j)2^{-j}$.  For every $K\geq1$, the unproved
selector requires some $N$ for which
$2^K\mid\sum_{r=1}^K f_A(N+r)2^{K-r}$ and $T_A(N+K)\leq16$.
These are the carry divisibility and scaled-tail bounds.  Lean proves that
they imply the
\lword{SupportSunflowerDichotomy.lean}{531}{sunflowerForcedCarrySupply_of_orthogonalPetalBouquet}{forced-carry certificate supply}
and hence
\lword{SupportSunflowerDichotomy.lean}{540}{irrational_erdosSupportSeries_of_orthogonalPetalBouquet}{irrationality of the support series}.
The selector itself is not proved.  This is therefore a conditional reduction
for one structured support family, not an unconditional case of
Erd\H{o}s~\#257.

\subsubsection{Composite dilation defects}\label{res:compositedefect}

Let $\Delta_A(a,x)$ count the $d\in A$ such that $d\mid ax$, $d\nmid x$,
and $d\ne a$.  For $a\in A$ and $a,x>0$, partitioning the support divisors of
$ax$ gives
$f_A(ax)=f_A(x)+\mathbf 1_{a\nmid x}+\Delta_A(a,x)$:
\lword{CompositeDilationDefect.lean}{30}{supportCoeff_mul_eq_add_defect}{the composite-dilation identity}.
For $A=\{2,6\}$ and $(a,x)=(6,1)$, the additional divisor $2$ gives
\lword{CompositeDilationDefect.lean}{218}{compositeDilationDefect_two_six_fixture}{\(\Delta_A(6,1)=1\)}.

If $A=E\cup\{c_jp_j:j\geq0\}$ is such a family and $a=c_ip_i$, every divisor
counted by $\Delta_A(a,x)$ either lies in $E$ or equals $c_jp_j$ with
$j\ne i$ and $p_j\mid x$.  Hence, for $P=\{p_j:j\geq0\}$,
$\Delta_A(c_ip_i,x)\leq |E|+f_P(x)$:
\lword{CompositeDilationDefect.lean}{151}{compositeDilationDefect_le_exceptional_add_petalCoeff}{the dilation-family defect bound}.
These finite local identities prove neither a correlation estimate, the tail
selector, nor irrationality.

\subsubsection{Mixed prime-power layers}\label{res:mixedlayer}

For $e\geq1$ and an integer-valued coefficient $g$, put
$L_{p,e}g(n)=g(p^en)-g(p^{e-1}n)$.  This finite difference records the change
between the two adjacent $p$-power arguments.  The operators commute,
$L_{q,f}L_{p,e}=L_{p,e}L_{q,f}$:
\lword{MaximalOmegaLayer.lean}{29}{primePowerLayer_comm}{the layer-commutation identity}.
If $p\ne q$ are prime, $e,f\geq1$, and $\gcd(n,pq)=1$, then, viewing $f_A$
as integer-valued, $L_{q,f}L_{p,e}f_A(n)=f_B(n)$ for
$B=\{d:\gcd(d,pq)=1,\ p^eq^fd\in A\}$:
\lword{MaximalOmegaLayer.lean}{39}{mixedPrimePowerLayerTwo_supportCoeffInt}{the two-prime layer identity}.
For $A=\{12\}$, $(p^e,q^f,n)=(4,3,1)$, and the value is one:
\lword{MaximalOmegaLayer.lean}{65}{mixedPrimePowerLayerTwo_twelve_fixture}{the singleton-support calculation}.
These finite identities provide no decimation transport, bounded-$\Omega$
conclusion, or irrationality theorem.

\subsubsection{Rank-one obstruction and reverse-carry spacing}\label{res:halftrapping}

For rational-linear maps $\varepsilon,\lambda_j:V\to\mathbb Q$, assume
$\varepsilon(v_0)=1$ and
$\ker\varepsilon\subseteq\ker\lambda_j$ for every $j$.  Each $\lambda_j$ is a
scalar multiple of $\varepsilon$; hence every square matrix
$(\lambda_j(v_i))$ of size at least two has determinant zero:
\lword{HalfTrappingReturnCarry.lean}{42}{relationInvariantLinearChannels_det_eq_zero}{the rank-one determinant criterion}.
A reverse-carry word consists of integer sequences $c_m,b_m,u_m$ satisfying
$b_m+2u_m=c_m+u_{m+1}$.  Two words with a common $c_m$ at a seam, bits $1,0$
there, and common $c_m,b_m$ for the next $\ell$ positions have terminal carry
difference $2^\ell z$ with $z$ odd:
\lword{HalfTrappingReturnCarry.lean}{124}{overlappingReverseCarryWords_carryDifference_eq_twoPow_mul_odd}{the exact carry factorisation}.
If both terminal carries lie in $[0,B]$, this forces $2^\ell\leq B$:
\lword{HalfTrappingReturnCarry.lean}{191}{overlappingReverseCarryWords_twoPow_le_realBound}{the dyadic spacing inequality}.
Neither theorem constructs the two words or an earlier overlap.

\subsubsection{Half-carry reachability}\label{res:halfcarryreachability}

For a Boolean word $a$, set
$f_a(n)=\#\{d:d\mid n,\ a_d=1\}$, $K_1=1$,
$K_n=2K_{n-1}-f_a(n)$, and $H_n=2\lfloor\sqrt n\rfloor+4$.
Admissibility through $N$ means $a_0=a_1=0$ and
$1\leq K_n\leq H_n$ for $1\leq n\leq N$; $K_N$ is its terminal carry.
At an unresolved half-divisor step to depth $2d$, let $h$ be the parent
carry, $\delta$ the last parent that takes the divisor, $c$ the remaining
coefficient, and $b\in\{0,1\}$ the new bit.  The child is $2h-c-1-b$ when
$h\leq\delta$ and $2h-c-b$ otherwise.  Hence $h,b$ realise exactly
$k\neq2\delta-c$: one missing carry and no others
(\lword{HalfCarryReachability.lean}{235}{evenSeamReachable_iff}{the exact one-hole equivalence}).

A \emph{canonical even seam} is the corresponding adjacent pair of admissible
parent words at depth $2d$ that realises this one-hole pattern.  The unproved
supply hypothesis gives one at every $2d$,
$d\geq10$.  Choose $k=1$, except $k=2$ when the hole is $1$; this gives words
cofinally
(\lword{HalfCarryReachability.lean}{999}{cofinalAdmissibility_of_canonicalEvenSeamSupply}{the cofinal-admissibility implication}).
Restriction fills earlier depths, and K\"onig compactness gives one infinite
support of value $1/2$
(\lword{HalfCarryReachability.lean}{1027}{exists_infinite_support_half_of_canonicalEvenSeamSupply}{the half-value conclusion}).
The seam supply is unproved, so Erd\H{o}s~\#257 remains undecided.

\paragraph{Cofinal strip returns.}  Let $A_g$ be the canonical greedy support
of $1/2$, and set $G_N=K_{N+1}$, $S_N=G_N/2^N$, and
$B_N=H_{N+1}/2^N$.  The recurrence yields
$S_{N+1}=S_N-f_{A_g}(N+2)/2^{N+1}\leq S_N$, so $S_N$ is non-increasing
(\lword{CofinalStripReturn.lean}{23}{scaled_integerHalfCarry_antitone}{the scaled-carry monotonicity}),
and the proved nonnegativity of $G_N-1$ gives $S_N\geq0$; meanwhile $B_N\to0$
(\lword{CofinalStripReturn.lean}{42}{tendsto_halfStripBound_succ_div_pow_zero}{the strip-envelope limit}).
Cofinal returns give arbitrarily late $M$ with $S_M\leq B_M$.  For
$\varepsilon>0$, choose one with $B_M<\varepsilon$; then
$0\leq S_N\leq S_M\leq B_M<\varepsilon$ for
$N\geq M$, so $S_N\to0$.  Hence the centred carry $C_N=G_N-1$ satisfies
$C_N/2^N\to0$, and the tempered-carry theorem gives an infinite greedy support
of value $1/2$.

\begin{proposition}[cofinal strip-return criterion]\label{res:cofinalstripreturn}
Let \(G_M\) be the uncentred integer carry of the canonical greedy expansion
of \(1/2\).  If
\[
 \forall N\in\N\;\exists M\geq N,\qquad
 G_M\leq 2\lfloor\sqrt{M+1}\rfloor+4,
\]
then \(1/2\in\mathcal A\), and its canonical greedy support is infinite.
\end{proposition}
\noindent Formalised:
\lword{CofinalStripReturn.lean}{84}{greedy_integerHalfCarry_scaled_tendsto_zero_of_cofinalStripReturn}{scaled-carry decay}
and
\lword{CofinalStripReturn.lean}{157}{half_mem_mersenneAchievementSet_of_cofinalStripReturn}{the strip-return membership criterion}.
The cofinal return condition itself is not established.  It is weaker than a
uniform strip bound and still does not give an unconditional membership
theorem.

\paragraph{Terminal-only cofinal approximation.}  A \emph{terminal-only
witness} at depth $M$ is a Boolean word $a$ with $a_0=a_1=0$; its earlier
carries are unrestricted, while its terminal carry $K_M$ satisfies
$|K_M|\leq H_M$.  Put $A=\{d:a_d=1\}$,
$X_A=\sum_{d\in A}(2^d-1)^{-1}$, and
$T_M=\sum_{j\geq1}f_A(M+j)2^{-j}$.  The exact residual identity is
$K_M=2^M(1/2-X_A)+T_M$.  Since both $|K_M|$ and $T_M$ are at most
$2\sqrt M+4$,
\[
 \left|X_A-\frac12\right|
 \leq\frac{4\sqrt M+8}{2^M}
 \leq\frac{4M+12}{2^M}.
\]
For witnesses at cofinally large depths, the last bound tends to zero.  Each
$X_A$ belongs to the closed Mersenne achievement set, so their limit $1/2$
also belongs to it.  The words at different depths need not agree, and the
finite-support obstruction forces the support representing the limit to be
infinite.

\begin{proposition}[terminal-only cofinal criterion]\label{res:terminalonlycofinal}
Cofinal normalised terminal witnesses in the square-root strip imply that
$1/2$ belongs to the Mersenne achievement set, and hence that an infinite
support has support-series value $1/2$.
\end{proposition}
\noindent Formalised:
\lword{TerminalOnlyCofinal.lean}{68}{dist_half_erdosSupportSeries_wordSupport_le_of_terminalOnlyWitness}{the terminal approximation bound},
\lword{TerminalOnlyCofinal.lean}{134}{half_mem_mersenneAchievementSet_of_cofinalTerminalOnlyStrip}{the terminal-strip compactness criterion},
and
\lword{TerminalOnlyCofinal.lean}{192}{exists_infinite_support_half_of_cofinalTerminalOnlyStrip}{the infinite-support conclusion}.
The cofinal terminal-witness supply is not proved.  In particular, the
proposition does not establish membership of $1/2$ unconditionally.

\paragraph{Scaled terminal vanishing.}  Let $M_n\to\infty$, and let $a^{(n)}$
be a normalised word of depth $M_n$.  Write $A_n$ for its finite support,
$X_n$ for its support-series value, and $K_n$ for its terminal carry.  The
carry--tail identity gives
\[
 \left|X_n-\frac12\right|
 \leq \frac{|K_n|}{2^{M_n}}
   +\frac{2\sqrt{M_n}+4}{2^{M_n}}.
\]
The second term tends to zero for every such depth sequence.  Thus the sole
additional hypothesis $|K_n|/2^{M_n}\to0$ gives $X_n\to1/2$; closedness and
the finite-support obstruction give an infinite support of value $1/2$.
The preceding square-root-strip condition is stronger, since
$|K_n|\leq H_{M_n}$ and $H_{M_n}/2^{M_n}\to0$.

\begin{proposition}[scaled terminal-vanishing criterion]\label{res:terminalscaledvanishing}
A normalised terminal-word sequence with depths tending to infinity and
vanishing binary-scaled terminal carries yields an infinite support with
support-series value $1/2$.
\end{proposition}
\noindent Formalised:
\lword{TerminalOnlyScaledVanishing.lean}{24}{dist_half_erdosSupportSeries_wordSupport_le_carry_add_tail}{the scaled terminal error bound},
\lword{TerminalOnlyScaledVanishing.lean}{165}{half_mem_mersenneAchievementSet_of_terminalScaledVanishing}{the scaled-vanishing compactness criterion},
and
\lword{TerminalOnlyScaledVanishing.lean}{221}{exists_infinite_support_half_of_terminalScaledVanishing}{the scaled-vanishing support conclusion}.
No such scaled-vanishing sequence is constructed here; the result is a
conditional reduction, not an unconditional membership theorem.

\subsection{Finite-to-cofinal criteria: windows, seams, and compactness}
\label{app:half-finite-cofinal}

Every half-value route below has the form
``finite witness at depth $N$'' plus ``a cofinal supply of such depths''
implies $1/2\in\mathcal A$.  The local identities advance one witness; the
compactness statements consume a cofinal supply.  No result below proves that
supply, so one verified window or seam is not an unbounded theorem.

\subsubsection{Finite residuals for the half-value problem}\label{res:halfcylindershadow}

For admissible depth-$N$ word $a$, let $A=\{d:a_d=1\}$, with terminal carry
$K_N$ and future tail $T_N$.  Exactly $\frac12-X_A=(K_N-T_N)/2^N$
(\lword{HalfCylinderFiniteShadow.lean}{55}{halfStripAdmissible_residual_eq}{residual identity}).
A first even seam has sole missing carry $H=2\delta-c$; hence $[1,B]$ is fully
reachable iff $H<1$ or $B<H$
(\lword{HalfCylinderFiniteShadow.lean}{807}{evenSeam_fullIntegerStrip_iff_escape}{seam equivalence}).
Raw safety is weaker: rank-three $A=\{2,3\}$ has residual $1/42<1/32$, but its
depth-six strip $[1,8]$ contains $H=3$.  After a skip at rank $k+1$, freeze
the selected support $A$ through $k$ and let $K$ be its centred carry.  The
sole input is that $m_0=-K$, $m_{J+1}=2m_J+f_A(k+J+2)$ reaches $m_J\geq0$
for some $J\leq k+1$; the first $J$ future rows then cover $K$.  This unproved
hypothesis implies $1/2$-membership
(\lword{HalfCylinderFiniteShadow.lean}{1256}{half_mem_mersenneAchievementSet_of_governedFrozenMarginProducer}{the first-passage implication}),
not Erd\H{o}s~\#257.

\subsubsection{Fixed-coefficient rewind}\label{res:fixedcoeffrewind}

For coefficient $c$, the inverse parent of carry $k$ is
$T_c(k)=\lfloor(k+c+1)/2\rfloor=\lceil(k+c)/2\rceil$.  For a list
$\mathbf c$ of $L$ rows, write $R_{\mathbf c}$ for their composition,
$O_{\mathbf c}$ for its accumulated offset, and $D=2^L$.  Then
$R_{\mathbf c}(k)=\lfloor(k+O_{\mathbf c})/D\rfloor$
(\lword{HalfCarryCeilParentContraction.lean}{49}{rewind_eq_closedForm}{rewind closed form}).
Put $\theta=(1+O_{\mathbf c})\bmod D$.  If $1\leq B\leq D$, every target in
$[1,B]$ has one common ancestor exactly when $\theta+B-1<D$
(\lword{HalfCarryCeilParentContraction.lean}{99}{rewind_endpoint_eq_iff_phase_fit}{phase-fit equivalence}).
Otherwise the endpoints differ by one and every target maps to one of those
adjacent ancestors
(\lword{HalfCarryCeilParentContraction.lean}{180}{rewind_singleton_or_isSeamPair}{singleton-or-seam dichotomy}).
The width bound does not select the alternative; this finite lemma supplies
neither a seam exclusion nor an Erd\H{o}s~\#257 conclusion.

\subsubsection{Selected half-carry windows}\label{res:selectedhalfwindow}

At depth $N$, a selected window of radius $R$ assigns each $k\in[1,R]$ an
admissible Boolean word with terminal carry $k$.  If such windows exist for
every large $N$, with $R\geq1$, choosing $k=1$ gives cofinal finite witnesses
(\lword{HalfCarrySelectedWindow.lean}{831}{cofinalAdmissibility_of_selectedHalfWindows}{cofinality});
compactness then gives an infinite support with sum $1/2$
(\lword{HalfCarrySelectedWindow.lean}{864}{exists_infinite_support_half_of_selectedHalfWindows}{half-value conclusion}).
The unproved input is a coherent one-row history: the representatives either
agree on every divisor of $N+1$, making the next coefficient common, or the
adjacent outcome realises the canonical one-hole seam
(\lword{HalfCarrySelectedWindow.lean}{236}{SelectedHalfWindow.HistoryAlternative}{history alternative}).
No such history is supplied at all large depths; hence no cofinal window
supply and no proof of Erd\H{o}s~\#257 follow.

\subsubsection{A realised rewind criterion}\label{res:rewindphase}

For a realised carry-one history, list the suffix bits newest first and put
$X=\sum_{j<L}b_j2^j$, so the newest row is the low bit.  With $D=2^L$,
$\theta=D-1-X$
(\lword{HalfCarryRewindPhase.lean}{39}{phase_eq_denom_sub_one_sub_suffixNumeral}{suffix-phase identity}).
Thus $[1,B]$ rewinds to one ancestor iff $B-1\leq X$
(\lword{HalfCarryRewindPhase.lean}{65}{rewind_interval_singleton_iff_suffixNumeral_covers}{suffix-coverage equivalence}).
Equivalently, for frozen-prefix carry $Z=1+X$, this is $B\leq Z$
(\lword{HalfCarryRewindPhase.lean}{352}{rewind_supportCoeffHistory_singleton_iff_prefixCarry_covers}{prefix-carry equivalence}).
No theorem proves either coverage inequality at all large depths; hence no
proof of Erd\H{o}s~\#257 follows.

\subsubsection{Localised protected seams}\label{res:protectedseam}

Here ``realised at depth $M$'' is only forward: each seam-reachable carry
$k\in[3,27]$ comes from an admissible word
(\lword{HalfCarryProtectedSeamConsumer.lean}{18}{ProtectedEvenSeamRealizedAt}{forward realisation}).
The unique hole is $H=2\delta-c$; choose $k=4$ if $H=3$, and $k=3$
otherwise.  This surviving carry gives cofinal admissibility from cofinally
many such seams by
\lword{HalfCarryProtectedSeamConsumer.lean}{31}{cofinalAdmissibility_of_cofinalProtectedEvenSeamRealization}{the protected-seam cofinality implication},
then compactness gives an infinite support with sum $1/2$,
\lword{HalfCarryProtectedSeamConsumer.lean}{53}{exists_infinite_support_half_of_cofinalProtectedEvenSeamRealization}{by compactness}.
The cofinal supply is unproved, so Erd\H{o}s~\#257 remains open.

\subsubsection{Combining the two cofinal half-carry criteria}\label{res:cofinalwindowseam}

For every $N$, the hypothesis gives an $M\geq N$ with a selected window or a
realised seam.  The window uses carry $1$; the seam uses $4$ if its hole is
$3$, otherwise $3$.  Either gives an admissible word, hence cofinal admissibility
(\lword{HalfCarryCofinalWindowOrSeamConsumer.lean}{32}{cofinalAdmissibility_of_cofinalSelectedWindowOrProtectedSeam}{fan-in})
and, by compactness, an infinite support summing to $1/2$
(\lword{HalfCarryCofinalWindowOrSeamConsumer.lean}{58}{exists_infinite_support_half_of_cofinalSelectedWindowOrProtectedSeam}{half-value}).
Branches may alternate.  The hypothesis is unproved;
Erd\H{o}s~\#257 is open.

\subsubsection{Integer-greedy first-wrap reduction}\label{res:integergreedy}

For $s\geq3$, put $C_s=2^{2s-1}-2^s$ and
$B_s=2\lfloor\sqrt{2s}\rfloor+4$.  Starting at $C_s$, subtract each truncated
seam weight in descending order whenever possible; let $r_s$ be the final
remainder.  A positive subset-sum defect at most $B_s$ exists exactly when
$1\leq r_s\leq B_s$
(\lword{HalfCylinderIntegerGreedy.lean}{654}{seam_smallPositiveDefect_iff_integerGreedyRemainder}{exact equivalence}).
Each weight clears its tail by at least $2^{s+1}>B_s$, so this one greedy test
suffices.  The downstream escape hypothesis asks for $B_s<r_s$ at every
genuine skipped rank
(\lword{HalfCylinderFullShellSeamBridge.lean}{687}{HalfGreedySkippedSeamEscape}{escape hypothesis}).
It is unproved; Erd\H{o}s~\#257 remains open.

\subsubsection{Exact finite recurrence for the greedy seam}\label{res:concreteseam}

The finite seam construction realises the adjacent cut, and its next word is
the integer-greedy choice:
\lword{HalfCylinderConcreteSeamAdapter.lean}{667}{seamCutNextWord_eq_greedyWord_succ}{the seam-word recurrence}.
This is a finite recurrence, not the missing escape estimate.

\subsection{Exact final-skip reduction and its residual cells}
\label{app:half-final-skip}

This is the only subsection that feeds the main final-skip reduction
directly.  It first identifies nonmembership with a final skip and membership
with cofinal terminal zeros.  It then separates the terminal cells already
excluded from the two unresolved middle coordinates and the unresolved
future-tail inequality.  The later conditional formulas
are conditional ways to attack those residual obligations; they do not add a
second conclusion.

\subsubsection{Fatal gaps and eventual right tails}\label{app:fatalright-detail}

Call the seam eventually right if some $S\geq5$ has
$G_{s+1}=G_s\cup\{s\}$ for every $s\geq S$: all later words append
\texttt{true}.  Call a finite prefix $u$ through rank $d$ a fatal half gap if
$V(u)+T_{d+1}<1/2<V(u)+w_{d+1}$.  Skipping rank $d+1$ then falls short even
after the full tail, while taking it overshoots, so every continuation misses
$1/2$.

For the concrete integer seam words, eventual right extension is equivalent
to such a gap,
\lword{HalfCylinderFatalGapRightTail.lean}{781}{seamGreedyEventuallyRight_iff_existsFatalHalfGap}{the fatal-gap equivalence},
to nonmembership of $1/2$ in the Mersenne achievement set,
\lword{HalfCylinderFatalGapRightTail.lean}{787}{seamGreedyEventuallyRight_iff_half_not_mem}{the nonmembership equivalence},
and to a final skipped exponent in the real greedy orbit,
\lword{HalfCylinderFatalGapRightTail.lean}{802}{seamGreedyEventuallyRight_iff_exists_isLastHalfGreedySkip}{the final-skip equivalence}.
This classifies the nonmembership branch exactly.  It does not decide which
branch holds or prove universal Erd\H{o}s~\#257.

\paragraph{First final-middle cell.}  At a putative final skip $D$, let $u$ be
the selected prefix and $B_D=u\cup\{n:n>D\}$ the \emph{lazy support}: it
skips $D$, takes every later rank, and has value $V(u)+T_D<1/2$.  The finite
transition carry is $C_D=4R_{D,D}-P_D-4$, and the exact swap makes its centred
carry $C_D+3$ at row $2D+1$.  For $C_D=-3$ this is zero; forced divisor
counts at least $1$ and $2$ in the next rows make it negative at $2D+3$,
contradicting finite-level nonnegativity below $1/2$,
\lword{HalfCylinderFinalMiddleCellEscape.lean}{587}{finalMiddleCell_neg_three_not_last}{the middle-\(-3\) exclusion}.
Only $-3$ is excluded; $-2$, $-1$, and the right-tail branch remain open.

\paragraph{Second final-middle cell.}  The next cell, with transition carry
$-2$, is not eliminated.  Here its \emph{phase} means the triple of residue
classes of the putative last skip \(D\) modulo \(21\), \(10\), and \(13\).
The possible phases are nevertheless sharply constrained.  Uniformly for
seam rows $s\ge27$, the greedy decisions through rank $26$ select exactly
the ranks
\[
  2,3,6,7,14,20,21,26,
\]
by
\lword{HalfFinalMiddlePhaseSieve.lean}{1190}{seamGreedyBits_head2367_14_20_21_26}{the exact rank-(26) prefix}.

If \(D\ge27\) lies in this middle cell and is followed by an all-right
suffix, then
\[
 D\bmod21\in\{11,14,17,20\},\qquad D\bmod10\ne8,
 \qquad D\bmod13\ne11,
\]
the class $D\bmod10=7$ cannot pair with $D\bmod21=11$, and the class
$D\bmod13=10$ cannot pair with $D\bmod21=11$.  The orbit theorem is
\lword{HalfFinalMiddlePhaseSieve.lean}{1913}{finalMiddleCell_neg_two_forces_phase_refinement_through_twentySix}{the final-middle phase sieve}.
The three moduli are pairwise coprime, so their phase triples correspond to
the \(21\cdot10\cdot13=2730\) residue classes modulo \(2730\).  Exactly
\(412\) survive; this finite count is
\lword{HalfFinalMiddlePhaseSieve.lean}{2050}{finalMiddleTwentySixPhaseSurvivors_card}{the exact survivor count}.
Thus a hypothetical final $C_D=-2$ transition lies in one of $412$ classes
modulo $2730$, a finite necessary condition rather than an exclusion; the
$-1$ cell is untouched.  Even a survivor density tending to zero would not
suffice, since nested nonempty residue sets can have a nonempty inverse limit.
This route needs an empty finite sieve or an Archimedean contradiction.

\paragraph{The two remaining inputs.}  The criterion requires
$\Theta_D<C_D$ at every actual middle transition $D\ge13$ except $C_D=-3$.
Since $\Theta_D\ge0$, exactly two open tasks remain: exclude $C_D=-2,-1$,
and prove $\Theta_D<C_D$ at every remaining transition with $C_D\ge0$.
Together they imply half-membership,
\lword{HalfCylinderLastProducerContradiction.lean}{387}{half_mem_mersenneAchievementSet_of_middleProducerTailEscapeExceptNegThree}{middle-tail implication};
a separate square-root lower bound also suffices,
\lword{HalfCylinderLastProducerContradiction.lean}{409}{half_mem_mersenneAchievementSet_of_middleProducerSqrtEscape}{square-root condition}.
Neither input is proved.

A stronger finite hypothesis puts $g_D=\#G_D$.  Since
$A_D=G_D\cup\{D\}$ has $g_D+1$ elements, $\Theta_D\le g_D+1$.
Using $C_D=4R_{D,D}-P_D-4$, the condition
$g_D+P_D+5<4R_{D,D}$ is exactly $g_D+1<C_D$.  At every late middle row it
therefore excludes the negative cells and proves tail dominance.  This is
\lword{HalfCylinderMiddleCarryLowerBound.lean}{111}{SeamMiddleProducerCardEscape}{the cardinality escape condition},
and its membership consequence is
\lword{HalfCylinderMiddleCarryLowerBound.lean}{1503}{half_mem_mersenneAchievementSet_of_middleProducerCardEscape}{the prefix-cardinality bound};
it follows in turn from $R_{D,D}\ge D$ at every late middle row,
\lword{HalfCylinderMiddleCarryLowerBound.lean}{1510}{half_mem_mersenneAchievementSet_of_middleProducerRowEscape}{row-scale sufficient condition}.
Both hypotheses are unproved.

\paragraph{Finite upper-reset band certificates.}  At the upper reset
$G_{d+1}=H_d$, let $O_d$ be the upper overshoot, $A_d$ its divisor pulse,
$Q_d=4O_d+A_d$, and $B_{d,j}=2^{d-j+1}$.  For $13\le d\le30$ and $j\le d$,
Lean proves
$B_{d,j}<Q_d$ or $Q_d+2(d+j)\le B_{d,j}$,
\lword{HalfCylinderUpperResetBandCertificates.lean}{78}{seamUpperResetDyadicBandEscape_through_thirty}{the finite upper-reset certificate}.
Thus $Q_d$ avoids the danger band $Q_d\le B_{d,j}<Q_d+2(d+j)$.
For fixed $d$, one exact $R_{d+1,d+1}$ certifies every $j$; no $d>30$ is
checked, so the global band-escape hypothesis remains open.

\subsubsection{Positive half-membership classification}\label{app:halfmembership-detail}

The complementary branch has an equally exact seam formulation.  Membership
of $1/2$ in the Mersenne achievement set is equivalent to the occurrence,
beyond every bound, of a successor terminal bit equal to zero,
\lword{HalfCylinderHalfMembershipClassification.lean}{126}{half_mem_mersenneAchievementSet_iff_unboundedTerminalFalse}{the unbounded terminal-zero equivalence};
equivalently, it is equivalent to a cofinal sequence of such terminal zeros,
\lword{HalfCylinderHalfMembershipClassification.lean}{203}{half_mem_mersenneAchievementSet_iff_cofinalTerminalFalse}{the cofinal terminal-zero equivalence}.
It is also equivalent to a cofinal integer-seam sequence with skipped ranks
tending to infinity,
\lword{HalfCylinderHalfMembershipClassification.lean}{213}{half_mem_mersenneAchievementSet_iff_exists_unboundedSkippedRanksAlong}{the unbounded skipped-rank equivalence}.
These equivalences isolate the required cofinal skipped-rank condition; they
do not establish that condition.

\subsubsection{Largest skipped ranks}\label{res:largestskip}

At row $s$, let $d$ be the largest skipped rank and put
$u=G_s\cap\{2,\ldots,d-1\}$.  Then
$G_s=u\cup\{d+1,\ldots,s-1\}$.  Write $L=\sum_{e\in G_s}q_{s,e}$ and
$U=\sum_{e\in u\cup\{d\}}q_{s,e}$: $U$ replaces the filled suffix above $d$
by $d$.  In the late range $d>2s/3$ (equivalently $2s<3d$), its exact gap is
\[
  3L+\bigl(3\cdot2^{s+1}+2\cdot4^{s-d}+4\bigr)=3U
\]
in division-free form,
\lword{HalfCylinderLargestSkipGap.lean}{259}{three_mul_largestSkipLowerWeight_add_exactLateGap_eq_upperWeight}{the exact late-gap identity}.

The branch recurrence is separate.  An upper or middle step to row $s+1$
skips the new rank $s$, making it the largest skip; a right step selects $s$
and preserves $d$:
\lword{HalfCylinderLargestSkipGap.lean}{328}{seamGreedyWord_succ_isLargestFalseRank_terminal_of_upperOrMiddle}{the terminal largest-skip branch}
and
\lword{HalfCylinderLargestSkipGap.lean}{339}{IsLargestFalseRank.seamGreedyWord_succ_of_rightBranch}{right-branch persistence}.
Global propagation still requires an upper or middle step whenever $2s<3d$
but $2(s+1)<3d$ fails.  This first-crossing statement is not proved.

\subsubsection{Boundary-pulse normalisation}\label{res:boundarypulse}

For a selected rank $e$, its row-$s$ pulse counts two if $e\mid2s+1$ and one
if $e\mid2s+2$; the word pulse sums these counts.  If $d$ is a largest skip
with $2s<3d$, every filled suffix rank $d<e<s$ divides neither new index, so
the whole word pulse comes from the selected prefix below $d$:
\lword{HalfCylinderBoundaryPulse.lean}{85}{exists_lowerPrefix_wordPulse_eq_of_largestFalse_late}{the lower-prefix pulse localisation}.
At the first crossing $2s<3d\leq2(s+1)$, either $3d=2s+1$ or $3d=2s+2$;
the twice- and once-counted alternatives give pulses two and one:
\lword{HalfCylinderBoundaryPulse.lean}{150}{rowPulse_boundary_of_late_firstCrossing}{the first-crossing pulse classification}.
The required transition-carry estimate remains unproved.

\subsubsection{Largest-skip induction}\label{res:largestskipinduction}

Assume that for every $s\geq14$ and largest skip $d$ with $2s<3d$, either
$2(s+1)<3d$ or the step to row $s+1$ is upper or middle.  This hypothesis is
unproved.  Lean checks the base $s=14$, $d=13$ directly, including $28<39$,
\lword{HalfCylinderLargestSkipInduction.lean}{73}{largestSkipLateAt_fourteen}{the row-14 base}.
Induction with the branch recurrence then gives a late largest skip at every
row $s\geq14$,
\lword{HalfCylinderLargestSkipInduction.lean}{120}{largestSkipLateAt_of_stepSocket}{the row-by-row induction}.
Those ranks tend to infinity, so $1/2$ belongs to the Mersenne achievement set,
\lword{HalfCylinderLargestSkipInduction.lean}{167}{half_mem_mersenneAchievementSet_of_largestSkipLateStepSocket}{largest-skip membership criterion}.

\subsubsection{The fixed-tail survival criterion}\label{res:fixedtailsocket}

Here $r_M(1/2)=1/2-\sum_{n\leq M}\varepsilon_n(1/2)w_n$ is the post-rank-$M$
remainder, and $T_M=\sum_{n>M}w_n$ is all future mass.  An actual skip $M$ is
final exactly when $T_M<r_M(1/2)$: all later mass then falls short,
\lword{HalfCylinderFixedTailSocket.lean}{22}{isLastHalfGreedySkip_iff_skip_and_fatal}{the final-skip equivalence}.
Negating gives $r_M(1/2)\leq T_M$ for all $M\in\mathcal S(1/2)$,
\lword{HalfCylinderFixedTailSocket.lean}{57}{no_lastHalfGreedySkip_iff_every_skip_survives}{the no-final-skip equivalence};
since $1/2\in\mathcal A$ exactly when there is no final skip, this is the
\lword{HalfCylinderFixedTailSocket.lean}{73}{half_mem_iff_every_actual_skip_survives}{membership equivalence}.
The equivalence is proved; the orbit inequalities are not.

\subsubsection{Residual identity for the middle-transition support}\label{res:producercarry}

For the middle-transition coordinates $A_D,C_D,\Theta_D$ defined above, the
generic finite-support theorem reads
$1/2-X_{A_D}=2^{-(2D+2)}(C_D-\Theta_D)$,
\lword{HalfCylinderProducerCarrySocket.lean}{27}{producerCarry_residual_identity}{the residual identity}.
Since the scale is positive, $C_D>\Theta_D\iff X_{A_D}<1/2$ and
$C_D<\Theta_D\iff X_{A_D}>1/2$,
\lword{HalfCylinderProducerCarrySocket.lean}{48}{producerCarry_gt_tail_iff_supportSeries_lt_half}{the below-half equivalence}
and
\lword{HalfCylinderProducerCarrySocket.lean}{78}{producerCarry_lt_tail_iff_half_lt_supportSeries}{the above-half equivalence}.
Tail nonnegativity makes $C_D<0$ sufficient for $X_{A_D}>1/2$,
\lword{HalfCylinderProducerCarrySocket.lean}{145}{producerCarry_neg_forces_half_lt_supportSeries}{the negative-carry corollary}.
A carry at most $-8$ strengthens this to $1/2+(w_D-T_D)<X_{A_D}$,
\lword{HalfCylinderProducerCarrySocket.lean}{190}{producerCarry_le_neg_eight_forces_gap_margin}{the full-gap corollary}.
These criteria do not prove that the required carries occur.

\subsubsection{Quarter-band endpoint cells}\label{res:quarterband}

At row $s$ put $H=2^{s-1}$, and let $R,U,R',p^-,p^+$ denote the adjacent
cut's remainder, overshoot, successor remainder, and two pulses.  Its gap is
$4H$; on either non-carry branch, failure of $R'<8H$ is equivalent to:
\[
 \begin{array}{ll}
 \text{middle:}&\exists k:\ R=H+k,\quad p^-\leq4k<p^-+4,\\
 \text{right:}&\exists k:\ R=3H+k,\quad p^-+4\leq4k,\quad
 4k+16H<4(R+U)+p^+.
 \end{array}
\]
\lword{HalfCylinderQuarterBandEndpoints.lean}{153}{seamAdjacentCut_middleNextUpperFailure_iff_pulseCell}{the middle-cell equivalence};
and
\lword{HalfCylinderQuarterBandEndpoints.lean}{175}{seamAdjacentCut_rightNextUpperFailure_iff_pulseCell}{the right-cell equivalence}.
The middle-row $k$ is unique,
\lword{HalfCylinderQuarterBandEndpoints.lean}{34}{pulseCell_unique}{the cell uniqueness}.
These equivalences classify failure but prove no concrete-orbit avoidance.

\subsubsection{Right-branch deficit escape}\label{res:resetdeficit}

Write $R_s=R_{s,s}$ for the final integer remainder.  At a largest skip $d$
with $2s<3d$, a right successor forces the explicit window
\[
 2^{s+1}+4\leq4R_s,\qquad
 12R_s+3\cdot2^{s+1}<
 4\bigl(3\cdot2^{s+1}+2\cdot4^{s-d}+4\bigr)+6(s-2),
\]
\lword{HalfCylinderResetDeficitEscape.lean}{200}{rightBranch_remainder_window}{the right-branch window}.
If $p_s^-$ is the below pulse, the same branch obeys
$R_{s+1}+2^{s+1}+p_s^-+4=4R_s$,
\lword{HalfCylinderResetDeficitEscape.lean}{218}{rightBranch_remainder_succ_eq}{the exact recurrence}.
When $R_s\leq2^s$, put $\delta_s=2^s-R_s$.  A right step makes
$\delta_{s+1}\geq4\delta_s$ while the reference scale only doubles.
Consequently, if $2^s\leq2^k\delta_s$, some $t\in[s,s+k]$ is upper or middle,
\lword{HalfCylinderResetDeficitEscape.lean}{261}{upperOrMiddle_within_of_two_pow_deficit}{the deficit escape bound}.
These formulas are proved, but they do not show that the actual orbit avoids
the window.

\subsubsection{Alignment of seam and transition carries}\label{res:seamproducer}

Let $P$ omit $s$, put $A=P\cup\{s\}$, and let $\delta$ be the half-carry of
$A$ at row $2s-2$.  With first-wrap hole
$H=2\delta-f_P(2s)$ and paired pulse
$\Pi=2f_A(2s+1)+f_A(2s+2)$, the transition carry satisfies
$C(A,s)=4(H-1)-\Pi$,
\lword{HalfCylinderSeamProducerAlignment.lean}{24}{producerCarry_insert_eq_four_mul_shiftedHole_sub_pulse}{seam--carry alignment}.
Thus two coefficient rows align $H-1$ with the transition-carry coordinate; the
identity supplies no sign, size, or occurrence estimate.

\subsection{Finite evidence, coordinate changes, and eliminated special cases}
\label{app:half-evidence}

Nothing here decides an open question.  The bases are bounded evidence for
A.2--A.3, the coordinate changes translate finite formulations, and the
counterexamples rule out shortcuts.  Each item must be read through its
earlier criterion.

\subsubsection{A verified initial selected window}\label{res:selectedwindowbase}

At depth $18$, twelve checked words cover every terminal carry in $[1,12]$
inside the half strip,
\lword{HalfCarrySelectedWindowBase.lean}{79}{depth18TableCheck_eq_true}{the depth-\(18\) table verification};
they define
\lword{HalfCarrySelectedWindowBase.lean}{129}{depth18SelectedHalfWindow}{the depth-\(18\) selected window}.
Their common depth-$13$ restriction gives the first next-row certificate,
\lword{HalfCarrySelectedWindowBase.lean}{172}{depth18SelectedHalfWindow_nextRowDivisorAgreement}{the depth-\(18\) divisor-agreement certificate}.
This finite base supplies no windows at unbounded depths.

\subsubsection{Rewind-to-seam equivalence}\label{res:rewindseambridge}

For a selected depth-$N$ window $W=(a^{(h)})_{1\leq h\leq R}$, suppose the
depth-$M$ rewind has adjacent ancestors, denominator $D$, and phase $\phi$;
put $\delta=D-\phi$.  If $M\leq N$, $(N+1)/2\leq M$, $R\leq D$, and the
ancestor coefficients at $N+1$ are $c+1,c$, then
$f_{a^{(h)}}(N+1)=c+1$ for $h\leq\delta$ and $c$ otherwise,
\lword{RewindSeamOperationalBridge.lean}{140}{nextRow_profile_of_rewindSeam}{the profile}.
With $27\leq S\leq2\lfloor\sqrt{N+1}\rfloor+4$ and $S+c+1\leq2R$,
\lword{RewindSeamOperationalBridge.lean}{217}{protectedEvenSeamRealizedAt_succ_of_profile}{a protected seam occurs at \(N+1\)}.
The theorem assumes, rather than proves, every displayed hypothesis.

\subsubsection{Selected suffix cylinders}

Put
\[
 H_N=2\lfloor\sqrt N\rfloor+4.
\]
For a Boolean word \(a=(a_0,\ldots,a_N)\), let
\[
 f_a(n)=\#\{d\leq N:d\mid n,\ a_d=1\},\qquad
 K_0(a)=1,\qquad K_{r+1}(a)=2K_r(a)-f_a(r+2).
\]

\begin{definition}[selected window and suffix cylinder]
\label{def:selected-suffix-cylinder}
A \emph{selected window} of depth \(N\) and radius \(R\) is a family
\((a^{(k)})_{1\leq k\leq R}\) of Boolean words with
\(a^{(k)}_0=a^{(k)}_1=0\) such that
\[
 1\leq K_{n-1}(a^{(k)})\leq H_n
 \quad(1\leq n\leq N),
 \qquad K_{N-1}(a^{(k)})=k.
\]
For \(M\leq N\), define the suffix numeral
\[
 \nu_{M,N}(a)=\sum_{j=M+1}^{N}a_j\,2^{N-j}.
\]
The window has a \emph{suffix cylinder at cutoff \(M\) with endpoint \(E\)}
if all \(a^{(k)}\) have the same prefix through \(M\) and
\[
 \nu_{M,N}(a^{(k)})+k=E
 \qquad(1\leq k\leq R).
\]
A \emph{profiled gap stage} permits one missing carry interval
\([g_-,g_+]\subseteq[1,H_N]\): lower and upper suffix-cylinder families
represent \(1,\ldots,g_--1\) and \(g_++1,\ldots,H_N\), with independent fixed
prefixes through \(M\).  Its upper endpoint \(E_+\) covers \(H_N\) and satisfies
\(E_+=g_++2^{N-M}\); prefix adjacency is a separate transition hypothesis.
\end{definition}

Reader map: \(R\) lists the represented terminal carries \(k\);
\(K_{n-1}\) is the carry at row \(n\), bounded by the strip width \(H_n\).
The cutoff \(M\) separates a shared past from the variable suffix, while
\(\nu_{M,N}(a)+k=E\) makes every representative meet the same endpoint.
A full cylinder covers the strip; a profiled gap records its one uncovered
interval, with separate fixed pasts below and above it.

The logical roles differ.  Depth \(18\) is checked data.  The one-row and
gap updates are conditional on their coefficient inequalities; the
quarter-band bound is only necessary for swallowing a gap.  Only cofinal
full stages---not a finite run---imply half-membership.

\begin{proposition}[verified base and one-row cylinder update]
\label{res:selectsuffix}
The depth-\(18\) selected window has \(R=12\) and a suffix cylinder at cutoff
\(M=13\) with endpoint \(E=17\).

More generally, let a depth-\(N\), radius-\(R\) selected window
\((a^{(k)})_{1\leq k\leq R}\) have a suffix cylinder at cutoff \(M\leq N\)
with endpoint \(E\), where \(1\leq N\) and \(R\leq E\).  Let
\(1\leq R'\leq H_{N+1}\).  If
\[
 f_{a^{(k)}}(N+1)=C\qquad(1\leq k\leq R)
\]
and
\[
 R'+C\leq2R,
\]
then the radius-\(R'\) successor remains a suffix cylinder, with endpoint
\[
 E'=2E-C.
\]
\end{proposition}
\noindent The base and endpoint update are the
\lword{SelectedSuffixCylinder.lean}{88}{depth18SelectedHalfWindow_hasSuffixCylinderAt_thirteen}{depth-\(18\) cylinder}
and
\lword{SelectedSuffixCylinder.lean}{136}{SelectedHalfWindow.hasSuffixCylinderAt_stepExplicit}{endpoint recurrence}.

\begin{proposition}[half-divisor-horizon dichotomy]\label{res:feedbackdichotomy}
Let a full suffix-cylinder stage have depth \(N\), cutoff \(M\), with
\(1\leq N\), \(M+1\leq N\), and
\[
 N+1=2(M+1).
\]
At depth \(N+1\), either a full stage exists at cutoff \(M+1\), or the
exceptional carries form a two-sheet profiled gap stage.  If
\(27\leq H_{N+1}\), the residual alternative also realises a protected
one-hole seam.
\end{proposition}
\noindent Formalised by the
\lword{SuffixCylinderGlobalProducer.lean}{174}{CylinderStage.feedbackStep_or_protectedSeam}{half-divisor-horizon dichotomy}
and
\lword{SuffixCylinderInStrip.lean}{665}{CylinderStage.feedbackAdvance_or_inStripTwoSheet}{two-family gap alternative}.

\begin{proposition}[profiled-gap transition]\label{res:profiledgap}
Suppose a profiled gap stage has lower and upper next coefficients
\(C_-\) and \(C_+\), respectively.  If
\[
\begin{gathered}
 1\leq 2g_- -C_- -1,\qquad C_+\leq2g_+,\\
 2g_- -C_- -1\leq 2g_+ -C_+\leq H_{N+1},\\
 H_{N+1}+C_+\leq2H_N,
\end{gathered}
\]
then the stage advances one row with exact child gap
\[
 [g'_-,g'_+]
   =[\,2g_- -C_- -1,\;2g_+ -C_+\,].
\]
At a cutoff boundary \(N=2M+1\), \(M\geq4\), the lower and upper prefixes
acquire the forced bits \(0\) and \(1\); their consecutiveness is preserved.
Under the displayed child-gap conditions at the next row, this promoted
stage therefore advances with coefficients determined by its retained
prefixes.
\end{proposition}
\noindent Formalised by the
\lword{SuffixCylinderProfiledGap.lean}{124}{ProfiledGapStage.advanceOfNextCoeffProfile}{two-coefficient gap recurrence}
and
\lword{SuffixCylinderProfiledAutoStep.lean}{110}{ProfiledGapStage.promoteThenAdvance_hasAdjacentPrefixes}{promote-and-advance theorem}.

\begin{proposition}[two-row full-gap width bound]\label{res:cylinderswallow}
Suppose that after a two-row update the child gap has lower endpoint at most
\(1\) and upper endpoint at least \(B\), so neither tracked sheet supplies a
canonical parent for any protected carry \(1,\ldots,B\).  If the lower
coefficient pulse is at most the upper pulse plus \(2\), then
\[
 B\leq4(g_+-g_-+1)+2. \tag{A.1}
\]
This is a necessary gap-width condition, not a contradiction.
\end{proposition}
\noindent Formalised as the
\lword{SuffixCylinderProfiledSwallow.lean}{254}{nextBound_le_four_gapWidth_add_two_of_macro_swallow}{two-row gap-width bound}.

\begin{corollary}[cofinal suffix-cylinder criterion]\label{res:cofinalcylinder}
If full suffix-cylinder stages exist at arbitrarily large depths, then
there is an infinite set \(A\subseteq\Npos\) such that
\[
 \sum_{n\in A}\frac1{2^n-1}=\frac12.
\]
\end{corollary}
\noindent This is the
\lword{SuffixCylinderTerminalOnlyBridge.lean}{148}{exists_infinite_positive_support_half_of_cofinalCylinderStages}{positive-support cofinal-cylinder criterion}.

\paragraph{Status and relation to prior work.}
The two transition propositions do not supply their coefficient inequalities,
and the finite base does not supply the cofinal family required by
Corollary~\ref{res:cofinalcylinder}.  The
strict-tail achievement-set literature supplies the surrounding Cantor-set
geometry~\cite{kovactao}, not this finite carry recurrence; the comparison
record makes no novelty claim for the recurrence itself.

\subsubsection{Verified finite suffix-cylinder stages}\label{res:prefeedback}

The finite record has three milestones.
\begin{enumerate}[label=\textup{(\arabic*)},leftmargin=2.2em]
\item \emph{Depth \(27\).}  Before the first half-divisor horizon, a full
selected window exists at every depth from \(18\) through \(27\), and the
depth-\(27\) endpoint covers the whole protected strip:
\lword{SelectedSuffixCylinderPreFeedback.lean}{118}{nonempty_stage_of_eighteen_le_of_le_twentySeven}{stages through depth \(27\)}
and
\lword{SelectedSuffixCylinderPreFeedback.lean}{147}{depth27SelectedHalfWindow_endpoint_covers_strip}{depth-\(27\) strip coverage}.

\item \emph{Depth \(29\).}  The threshold-stage continuation still covers the
strip and has a common suffix cylinder at every cutoff from \(14\) through
\(25\), while cutoff \(26\) is explicitly not common:
\lword{SuffixCylinderThreshold.lean}{700}{depth29SelectedHalfWindow_endpoint_covers_strip}{depth-\(29\) strip coverage}
and
\lword{SuffixCylinderThreshold.lean}{806}{depth29SelectedHalfWindow_not_commonRestrictionAt_twentySix}{cutoff-\(26\) obstruction}.

\item \emph{Depth \(52\).}  An independent exact run reaches depth \(51\)
with endpoint \(51{,}327{,}745\), beyond the next head-bit threshold,
and the promoted successor at depth \(52\) is checked:
\lword{SuffixCylinderStage51.lean}{536}{fullCylinderStage51_feedback_headThreshold_margin}{depth-\(51\) threshold margin}
and
\lword{SuffixCylinderTerminalOnlyBridge.lean}{118}{fullCylinderStage52_halfTerminalOnlyStripWitness}{depth-\(52\) terminal-strip witness}.
\end{enumerate}
These are finite milestones, not an unbounded induction; they do not provide
the arbitrarily large stages required by
Corollary~\ref{res:cofinalcylinder}.

\subsubsection{Half-divisor unit drop}\label{res:halfdivisordrop}

Reader map: the terminal half-divisor bit selects whether the row coefficient
is \(C\) or \(C+1\).  Since the carry update is \(K'=2K-C\), switching that
bit from \(0\) to \(1\) drops the successor carry by exactly one:
\lword{HalfDivisorUnitDrop.lean}{20}{supportCoeff_extend_true_eq_false_add_one_at_double}{the terminal-bit unit-drop identity}.
An explicitly identified adjacent boundary pair packages the same one-unit
split at a shared cutoff,
\lword{HalfDivisorUnitDrop.lean}{35}{supportCoeff_boundaryPair_unitDrop_at_double}{the boundary-pair unit-drop identity}.
The inverse-parent recurrence transfers such a pair to the next protected
seam while retaining the exact unit-drop witness,
\lword{RewindHalfDivisorAdapter.lean}{86}{rewindBaseUnitDropAt_of_halfDivisorBoundaryPair}{the rewind unit-drop transfer}
and
\lword{RewindHalfDivisorAdapter.lean}{147}{protectedEvenSeamRealizedAt_succ_of_fullStrip_rewindHalfDivisorBoundaryPair}{the protected-seam transfer}.
This is the local coefficient input for the rewind-seam profile, not a global
construction of such boundary pairs.

\subsubsection{A finite rewind-boundary obstruction}\label{res:rewindboundarycounterexample}

The next example separates numerical seam data from the structural boundary
predicate.  A concrete depth-$26$ two-word window has, at depth \(27\),
rewind history \([1]\), a width-two seam, and the doubled-row identity, yet no
depth-\(13\) half-divisor boundary-pair witness:
\lword{RewindBoundaryPairCounterexample.lean}{169}{depth27_actualRewindSeam_not_boundaryPair}{the depth-\(27\) boundary-pair counterexample}.
Thus the protected-window buffer in the compactness criterion is a genuine
hypothesis, not a consequence of the scalar seam facts alone.

\subsubsection{Explicit carry coordinates at the final non-right transition}
\label{res:producercarrycoordinates}

At an adjacent cut, the \emph{below} and \emph{above} row words are the
neighbouring subset sums immediately below and above the target.  Their
unsigned distances from it are the remainder and overshoot; their pulses
record the paired divisor-incidence contribution of the next two rows.
After adjoining the terminal rank \(s\ge5\), the below support has transition
carry
\[
  4\,\mathrm{remainder}-\mathrm{belowPulse}-4,
\]
whereas the above support has transition carry
\[
  -\bigl(4\,\mathrm{overshoot}+\mathrm{abovePulse}+4\bigr).
\]
The generic row-word identity is
\lword{HalfCylinderProducerLowerBound.lean}{105}{producerCarry_insert_seamWordSupport_eq}{row-word transition identity},
with the below and above specialisations at
\lword{HalfCylinderProducerLowerBound.lean}{150}{producerCarry_insert_seamBelowSupport_eq_middleCoordinate}{middle-coordinate identity}
and
\lword{HalfCylinderProducerLowerBound.lean}{183}{producerCarry_insert_seamAboveSupport_eq_neg_upperCoordinate}{upper-coordinate identity}.
Thus the upper carry is negative, while the middle sign depends on the
competition between remainder and pulse.  The identities supply no uniform
exponential lower bound, sign persistence at unbounded ranks, or universal
escape mechanism.

\subsubsection{Incompatibility with Campbell's prime-progression bound}
\label{res:campbellboundary}

Crandall asked whether the binary block \(11\) occurs infinitely often in
the Erd\H{o}s--Borwein constant; Campbell answered this affirmatively using
a Chinese-remainder construction and a prime-progression estimate extracted
from Alford--Granville--Pomerance~\cite{crandall,alfordgranvillepomerance,campbell}.
Campbell chooses
exponent \(\delta=1/4\): at prime-search height \(X\), the progression
modulus \(d\) must satisfy \(d\le X^{1/4}\), equivalently \(d^4\le X\).

For the proposed last-skip synchronization, \(P\) is the fatal window's
prime cap, \(t\) its unavoidable phase period, and \(d\) the combined
progression modulus.  The window gives \(P\le4t\), while absorbing the
period requires \(t\le d\).  Taking \(X=P\), the four requirements
\(2\le d\), \(t\le d\), \(P\le4t\), and \(d^4\le P\) are inconsistent.
Formalised:
\lword{CampbellShiftSynchronization.lean}{299}{campbellQuarterExponent_periodFreeze_noSync}{the period-freeze incompatibility}
and
\lword{CampbellShiftSynchronization.lean}{320}{campbellPhaseFreeze_parameters_impossible}{the parameter incompatibility}.

The associated shifted-zero condition is guarded by a putative last skip: it
requires every such endpoint to violate its own dyadic cylinder.  It is
therefore exactly equivalent both to infinitely many actual greedy skips and
to $1/2\in\mathcal A$:
\lword{CampbellShiftSynchronization.lean}{471}{shiftWindowZero_iff_greedyMersenneSkippedSupport_infinite}{the shifted-zero/skip equivalence}
and
\lword{CampbellShiftSynchronization.lean}{488}{shiftWindowZero_iff_half_mem_mersenneAchievementSet}{the shifted-zero/half-membership equivalence}.
It is an endpoint restatement, not an independent Campbell input, and does
not decide half-membership.

}

\subsection{What remains open}

The two directions remain separate.
\begin{enumerate}[label=\textup{(\alph*)},leftmargin=2.2em]
\item The universal problem is to prove \(X_A\notin\Q\) for every infinite
\(A\subseteq\Npos\).  Theorem~\ref{res:support} proves several
families, while Theorem~\ref{res:rigidity-main} and its four corollaries
constrain a hypothetical rational support.  Neither route supplies the
universal quantifier.
\item The distinguished half-value route asks whether \(1/2\in\mathcal A\).
Theorem~\ref{res:halfmembership} identifies this with infinitely many greedy
skips, while Theorem~\ref{res:fatalright} identifies nonmembership with a
final skip.  Proposition~\ref{res:finalskipexclusions} proves that this final
transition cannot be upper and that a final middle transition cannot have
\(C_D=-3\).  Corollary~\ref{res:taildominance} would prove membership if no
actual middle transition had \(C_D=-2\) or \(-1\), and if every remaining
middle transition with \(C_D\geq0\) satisfied the future-tail
inequality~\textup{(4.2)}.  Neither statement is proved.
\end{enumerate}
Appendix~\ref{app:half-technical} records the exact local reductions,
conditional compactness criteria, finite certificates, counterexamples to
stronger local assertions, and arguments that rule out several proposed
strategies.  These results narrow the remaining cases; they decide neither
open \#257 question.

\section{The totient constant $S$ (Erd\H{o}s \#249)}\label{sec:249}

Whether
\[
S=\sum_{n\ge1}\frac{\ph(n)}{2^n}
\]
is irrational remains open.  The argument has four logically distinct
layers.  First, a finite Farey computation excludes every rational
denominator up to an explicit bound.  Second, irrationality is characterised
exactly by the existence of finite certificates against every possible
eventual binary period.  Third, many such certificates are verified at
bounded parameters.  Fourth, and still missing, is a theorem supplying them
at unbounded parameters.  Only the fourth layer would settle the problem.

The main section states this reduction.  Appendix~\ref{app:totient-normal}
gives equivalent forms of the missing assertion that certificates occur at
arbitrarily large scales.  Appendix~\ref{app:totient-finite} collects finite
estimates and checked certificates; any conclusion of irrationality there
assumes, rather than proves, an unbounded supply.  Appendix~
\ref{app:totient-boundaries} records candidate sufficient conditions and
shows why several natural shortcuts fail.  Appendix~\ref{app:totient-rank}
proves exact dyadic rank growth but not the missing implication from that
growth to irrationality.  Only the first appendix is an exact reformulation
of the open fourth layer.

\subsection{Unconditional: no small denominator}\label{sec:uncond}

\begin{theorem}[denominator exclusion]\label{res:farey}
Set
\[
\Qzero:=79\,639\,646\,646\,701\,375\,323\,355\,774\,875\,831\,053
       \approx7.96\times10^{34}.
\]
For every integer $p$ and every $q$ with $1\le q\le\Qzero$,
\[
S \;\ne\; \frac{p}{q}.
\]
Equivalently, if $S$ is rational then its reduced denominator exceeds that bound.
\end{theorem}
\noindent Formalised as the
\lword{CertificateKernel.lean}{18056}{tsum_totient_div_pow_two_ne_ratCast_of_den_le_79639646646701375323355774875831053}{denominator-exclusion theorem}.
For a fixed window length $K$, the checked prefix and a rigorous tail bound
define a finite gap certificate.  If a denominator $q$ failed that certificate,
elementary rearrangement would put some rational $r/q$ in a specific short
interval.  The larger bounds bracket that interval by fractions
$a/b<c/d$ satisfying $bc-ad=1$.  The mediant lemma says that every rational
strictly between such neighbours has denominator at least $b+d$.  Thus no
$q<b+d$ can fail the certificate; for each such $q$, the corresponding scaled
tail lies strictly between consecutive integers, whereas $S=m/q$ would force it
to be an integer.

The finite exclusion ladder records the successive verified thresholds
\[
4838 \;\to\; 4\,194\,304 = 2^{22} \;\to\; 2.49\times10^{17}\ (K{=}120)
\;\to\; 7.96\times10^{34}\ (K{=}240).
\]
The mediant framework is historically associated with Farey's 1816 note on
vulgar fractions~\cite{farey1816}; every exclusion above is Lean-checked in the
formal-source checkpoint.  The first threshold is a direct low-carry check,
$2^{22}$ is retained as a named intermediate corollary, and the two largest
thresholds use the Farey separation. See
\lword{CertificateKernel.lean}{15347}{tsum_totient_div_pow_two_ne_ratCast_of_den_le_4838}{denominator exclusion through 4838}
and \lword{CertificateKernel.lean}{18029}{tsum_totient_div_pow_two_ne_ratCast_of_den_le_4194304}{the intermediate 4194304 bound},
then \lword{GapFareyBound.lean}{88}{gap_check_window_1_120_le_248672326362367909}{the 120-term Farey bound} and
\lword{GapFareyBound.lean}{176}{gap_check_window_1_240_le_79639646646701375323355774875831053}{240-term Farey bound}
for the two Farey windows.

This is an explicit lower bound on the denominator of any rational
representation, not a proof of irrationality. A rational with a larger
denominator is not excluded by it.

\begin{remark}[one bound, four representations]
Because the Lambert identities of Section~\ref{sec:ladder} are equalities, the same
finite record transfers verbatim to three other representations of the constant:
to the Lambert representation $\sum_d \alpha(d)/(2^d-1)$
(\lword{CertificateKernel.lean}{18145}{tsum_primWeight_div_two_pow_sub_one_ne_int_div_of_den_le_79639646646701375323355774875831053}{Lambert-representation denominator transfer}),
to the signed M\"obius-square constant $T=\sum_d\mu(d)/(2^d-1)^2$ at half the bound
(\lword{CertificateKernel.lean}{18159}{tsum_moebius_div_two_pow_sub_one_sq_ne_int_div_of_den_le_39819823323350687661677887437915526}{signed-square denominator transfer}),
and to the coprimality probability of Section~\ref{sec:geom} at half the bound.
Half appears because $S=\tfrac12+T$ turns a denominator $d$ for $T$ into $2d$ for $S$.
\end{remark}

\subsection{The geometric picture: $S$ as coprime-pair mass}\label{sec:geom}

Let $X,Y$ be independent fair-coin waiting times, $\Pr(X=n)=2^{-n}$ for $n\ge1$.
Then $\Pr(d\mid X)=1/(2^d-1)$, so the Lambert transform is a divisor calculus of
$X$: $L(f)=\mathbb E[(f*\mathbf 1)(X)]$.  The identities give $L(\mu)=\Pr(X=1)=1/2$,
$L(\ph)=\mathbb E[X]=2$, $L(\mathbf 1)=\mathbb E[\tau(X)]=E$, and
$L(\alpha)=\mathbb E[\ph(X)]=S$.

Counting visible lattice points $(a,b)$ with $a+b=n$, $a\ge1$, $b\ge0$, and $\gcd(a,b)=1$ gives
exactly $\ph(n)$, uniformly in $n$, so $S$ is itself a coprime-pair mass.

\begin{proposition}[coprimality probability]\label{res:coprime}
\[
S \;=\; \tfrac12 \;+\; \Pr\!\big(\gcd(X,Y)=1\big)
\;=\; \tfrac12 \;+\!\!\sum_{\substack{a,b\ge1\\ \gcd(a,b)=1}}\!\! 2^{-(a+b)}.
\]
\end{proposition}
\noindent Formalised: \lword{CertificateKernel.lean}{18229}{totient_series_eq_half_add_visible_coprime_pairs}{coprime-pair representation},
with the underlying lattice identity at
\lword{CertificateKernel.lean}{18216}{tsum_visible_coprime_pairs_eq_totient_series}{visible-lattice-point identity}.
Base $2$ is the one point where $\Pr(X=a)\Pr(Y=b)=2^{-(a+b)}$ needs no normalising
constant.  Restricting the half-open lattice mass to positive waiting times
removes $(1,0)$, producing the boundary term $1/2$.  Thus \#249 asks whether two independent
fair-coin waiting times are coprime with irrational probability.
After reducing $(X,Y)$ by their gcd, the same distribution has an exact
Stern--Brocot cylinder decomposition.
Appendix~\ref{app:lambert-probability} states the resulting probability law
and the sharp Fibonacci stability of its alternating runs.  These are
unconditional structural theorems inside one representation of \(S\); they
do not supply the denominator survival needed for irrationality.

\subsection{Exact certificate characterisation}\label{sec:reduction}

The reduction uses only the elementary bound $\ph(n)\le n$, geometric tail
estimates, and finite integer arithmetic.
For $t\ge1$, write $M_t=\lcm(1,\ldots,t)$.

\paragraph{Why tail differences detect rationality.}
Since \(2^NS=\sum_{n\le N}\ph(n)\,2^{N-n}+R_N\), each scaled tail
\(R_N\) is \(2^NS\) minus an integer.  Suppose \(S=p/q\) with
\(q=2^cv\) and \(v\) odd.  For every multiple \(h\) of the
multiplicative order of \(2\) modulo \(v\) and every \(N\geq c\),
\[
 R_{N+h}-R_N
 =\frac{2^N\bigl(2^h-1\bigr)p}{q}-k
 \in\Z
\]
for some integer \(k\), because \(v\mid2^h-1\) and \(2^c\mid2^N\).
Rationality therefore supplies one period \(h\) all of whose tail
differences at large \(N\) are integral; conversely, a single integral
positive-shift difference already forces \(S\) to be rational.  A finite
certificate refutes one candidate integrality, and refuting every period
at arbitrarily large \(N\) refutes rationality outright.  This is the
mechanism behind the equivalences of this subsection.

\begin{definition}[tail differences and finite certificates]
\label{def:totient-certificate}
For $N\ge0$, define the scaled tail
\[
R_N \;=\; \sum_{m\ge1}\frac{\ph(N+m)}{2^m}.
\]
For $h\in\Npos$ and $N,L\in\N$, define
\[
\Delta_{h,N,L}=\sum_{j=0}^{L-1}\big(\ph(N+h+1+j)-\ph(N+1+j)\big)\,2^{\,L-1-j}\in\Z,
\]
with the sum empty when \(L=0\),
and say that $(h,N,L)$ is a \emph{finite certificate} when
\[
\mathsf{Sep}(h,N,L):\quad
N+h+L+2 \;<\; \big(\Delta_{h,N,L}\bmod 2^L\big) \;<\; 2^L-(N+h+L+2).
\]
\end{definition}

Here \(x\bmod 2^L\) denotes the least residue in
\(\{0,\ldots,2^L-1\}\).  The integer $\Delta_{h,N,L}$ contains the first
$L$ binary places of $R_{N+h}-R_N$; the two margins bound the omitted tail
using $\ph(n)\le n$.  Thus each instance of the predicate is decidable by
finite integer arithmetic, while its conclusion concerns the infinite tail
difference.  The strict inequalities are impossible when \(L=0\), so every
successful certificate has positive depth.
Heuristically, the two margins together occupy a proportion of about
\(2(N{+}h{+}L{+}2)/2^L\) of the modulus, which is small as soon as
\(2^L\) is large compared with \(N+h+L\); a certificate can fail only
when the residue lies exceptionally close to \(0\) or \(2^L\), that is,
when the tail difference is exceptionally close to an integer.  Nothing
in this paper converts that observation into the required supply at
arbitrarily large \(N\).

\begin{proposition}[certificate completeness]\label{res:complete}
For every $h\ge1$ and $N\ge0$,
\[
\bigl(\exists L,\ \mathsf{Sep}(h,N,L)\bigr)
\quad\Longleftrightarrow\quad
R_{N+h}-R_N\notin\Z.
\]
\end{proposition}
\noindent This is the
\lword{LcmConeFlatness.lean}{316}{exists_certifiedKill_iff_tail_diff_notMem_int}{certificate-completeness equivalence}.
Thus the finite predicate loses no information at a fixed pair \((h,N)\).
Searching \(L=1,2,\ldots\) is therefore a complete positive
semi-decision procedure for non-integrality at fixed \((h,N)\): it halts
exactly when the difference is non-integral.  When the difference is integral,
completeness says that no depth succeeds, so this search does not halt.  It is
not a two-sided decision procedure for an infinite tail.

\begin{theorem}[pointwise certificate characterisation]
\label{res:pointwisecomplete}
The following are equivalent:
\[
\begin{aligned}
 &S\notin\Q,\\
 &\forall\,h\ge1\ \forall\,N\ge0,\qquad R_{N+h}-R_N\notin\Z,\\
 &\forall\,h\ge1\ \forall\,N\ge0\ \exists\,L,\qquad
      \mathsf{Sep}(h,N,L).
\end{aligned}
\]
\end{theorem}
\noindent Formalised:
\lword{LcmConeFlatness.lean}{386}{irrational_totient_series_iff_all_tail_diffs_nonintegral}{all-differences equivalence},
and
\lword{LcmConeFlatness.lean}{399}{irrational_totient_series_iff_pointwise_certificates}{pointwise certificate equivalence}.
The converse direction uses the fact that integrality of even one
positive-shift tail difference forces $S$ to be rational
(\lword{LcmConeFlatness.lean}{357}{not_irrational_of_tail_diff_mem_int}{integral-difference rationality criterion}).

Here $h$ tests a candidate rational period, $N_0$ means ``after every cutoff'',
$N\geq N_0$ chooses the location, and $L$ is the finite binary depth checked
there.  A bounded list cannot cover every $h$ and every $N_0$.

\begin{theorem}[exact tail-period characterisation]\label{res:red21}
$S$ is irrational if and only if, for every period $h\ge1$ and every threshold
$N_0$, there exist $N\ge N_0$ and $L$ with $\mathsf{Sep}(h,N,L)$.
\end{theorem}
\noindent Formalised as the
\lword{LcmConeFlatness.lean}{412}{irrational_totient_series_iff_certificate_supply}{certificate-supply equivalence}.

Every multiple of an eventual period is again a period, and every fixed
\(h\) divides \(M_t\) for all \(t\geq h\).  The least common multiple
therefore provides an exact one-parameter form: the single diagonal
family \((M_t,M_t)\) tests every candidate period at once.

\begin{theorem}[exact diagonal characterisation]\label{res:diag}
The following are equivalent:
\[
\begin{aligned}
 &S\notin\Q;\\
 &\forall\,t_0\in\N\ \exists\,t\geq t_0\ \exists\,L,\qquad
   \mathsf{Sep}(M_t,M_t,L).
\end{aligned}
\]
\end{theorem}
\noindent Formalised as the
\lword{LcmConeFlatness.lean}{426}{irrational_totient_series_iff_lcm_diagonal_certificate_supply}{exact lcm-diagonal characterisation}.
This replaces the two free parameters \(h,N\) by \((M_t,M_t)\);
producing such certificates at arbitrarily large \(t\) is still open.

\begin{example}[verified finite range]\label{res:deposits}
Lean checks:
\begin{itemize}[leftmargin=1.4em]
\item $\mathsf{Sep}(h,12,16)$ for every $1\le h\le8$;
\item a fixed-window certificate for every $1\le h\le16$;
\item diagonal certificates $\mathsf{Sep}(M_t,M_t,L)$ at $t=1$ and every
lcm-height change endpoint $2\le t\le64$ (that is, $M_{t-1}<M_t$), namely
\[
t\in\{1,2,3,4,5,7,8,9,11,13,16,17,19,23,25,27,29,31,32,37,41,43,
47,49,53,59,61,64\};
\]
\item three finite menus of tails that cannot all share one fractional part.
\end{itemize}
\end{example}
\noindent These four classes are formalised by the
\lword{TotientTailPeriodKiller.lean}{404}{certifiedKill_all_small}{small-window certificate family},
\lword{CarrySurvivorExtinction.lean}{574}{certifiedKill_all_upto_sixteen}{periods through sixteen},
\lword{DiagonalPincerCertificatesT64.lean}{1967}{certifiedKill_diagonal_all_imported_through_t64}{28 imported diagonal certificates through scale 64},
and
\lword{LcmConeNonflat.lean}{506}{coneNonflatCert_cells}{cone-cell certificate table}.

For example, at $(h,N,L)=(1,12,16)$,
\[
\Delta_{1,12,16}=-143140,\qquad
\Delta_{1,12,16}\bmod 2^{16}=53468,
\]
while the omitted-tail allowance is $12+1+16+2=31$.  Hence
\[
31<53468<2^{16}-31,
\]
which is exactly $\mathsf{Sep}(1,12,16)$ and therefore proves
$R_{13}-R_{12}\notin\Z$.  Pairwise rows name a non-integral difference; joint
rows prove that one exists within the menu.  They do not form an unbounded family.

\subsection{What remains open}\label{sec:249-open}

By Theorem~\ref{res:red21}, the unresolved statement is precisely
\[
\forall h\ge1\ \forall N_0\ge0\ \exists N\ge N_0\ \exists L\ge1,\qquad
\mathsf{Sep}(h,N,L).
\tag{5.1}\label{eq:unbounded-certificate-supply}
\]
The diagonal condition in Theorem~\ref{res:diag} is an equivalent
one-parameter form of this statement.  The verified scales in
Example~\ref{res:deposits} form a finite set; they do not give certificates
for every period after every cutoff.  No theorem in the formal source proves
\eqref{eq:unbounded-certificate-supply}; consequently no irrationality claim
for $S$ is made here.

\begin{remark}[finite inspection cannot establish the supply]
The finite status of the verified scales is forced, not cautious.  Fix a
bound \(B\ge1\) and a period \(P>B\), and define \(\gamma(n)=\ph(n)\) for
\(n\le B\), while for \(n>B\) let \(\gamma(n)=n-1\) if \(P\mid n\) and
\(\gamma(n)=n\) otherwise.  Then \(0\le\gamma(n)\le n\), and
\[
 \sum_{n\ge1}\frac{\gamma(n)}{2^n}
 =\sum_{n\le B}\frac{\ph(n)}{2^n}+\frac{B+2}{2^B}-\frac1{2^P-1}
\]
is rational, and its reduced denominator has odd part exactly \(2^P-1\).  Every
finite computation that inspects only coefficients with index at most \(B\)
--- in particular every certificate \(\mathsf{Sep}(h,N,L)\) with
\(N+h+L\le B\) --- returns the same result for \(\gamma\) as for \(\ph\),
and choosing \(P\) with \(2^P-1>\Qzero\) makes the splice consistent with
the denominator exclusion as well.  This elementary splice is not formalised
in the source and proves nothing about \(\ph\) itself; it shows that a proof
of~\eqref{eq:unbounded-certificate-supply} must use totient values at
unbounded indices, so no extension of the finite record in
Example~\ref{res:deposits} can establish the supply.
\end{remark}

Appendix~\ref{app:totient-normal} records the exact reformulations of this
open obligation.  Appendix~\ref{app:totient-finite} records fixed-scale
enclosures, projections, and certificates, followed by conditional
implications that assume them at arbitrarily large scales.  Appendices~
\ref{app:totient-boundaries}--\ref{app:totient-rank} delimit candidate
unbounded routes.  None proves \eqref{eq:unbounded-certificate-supply}.

\long\def\TotientTechnicalAppendix{%

\subsection{Exact cofinal normal forms and cone consequences}
\label{app:totient-normal}

Put $M_t=\lcm(1,\ldots,t)$.  This subsection has two routes with different
logical status.  The cone route starts from a necessary consequence of
rationality: for large $t$, the tails at the locations
$M_t,2M_t,3M_t,\ldots$
have one fractional part.  A finite joint certificate can contradict this
prediction at finitely many locations, but no converse completeness theorem
is claimed.  The diagonal route instead gives an exact decomposition of
$R_{2M_t}-R_{M_t}$: integrality is equivalent to one residue-class condition
at each $t$, and missing that class cofinally is equivalent to the
irrationality of $S$.

\paragraph{Cone flatness and completeness.} For fixed $t$, call the tail
locations \(\{kM_t:k\ge1\}\) the \emph{lcm cone}.  Rationality eventually
forces one fractional part across this sequence.

\begin{proposition}[cone flatness]\label{res:flat}
If $S$ is rational, then for every sufficiently large $t$ there is
$\theta_t\in[0,1)$ such that the fractional part of the tail $R_{kM_t}$ equals
$\theta_t$ for every $k\ge1$.
\end{proposition}
\noindent Formalised: \lword{LcmConeFlatness.lean}{90}{rational_totient_series_forces_lcm_cone_flatness}{cone flatness under rationality}.

Proposition~\ref{res:complete} states fixed-parameter certificate
completeness, and Theorem~\ref{res:pointwisecomplete} gives its global
pointwise form.  Their additional role here is cone-level: a joint finite
refuter tests several multipliers $k$ at once rather than one pair.  It refutes the
single-fractional-part model that rationality predicts on a finite sample
(\lword{LcmConeNonflat.lean}{126}{exists_nonintegral_pair_of_coneNonflatCert}{finite cone-nonflatness refuter}).
This finite cone predicate is sound in the direction just stated: a
certificate produces a non-integral pair and hence contradicts the sampled
flatness model.  No converse is proved saying that every non-flat finite
sample, or every irrational value of \(S\), must be detected by this joint
predicate.  It is therefore a refuter, not a complete decision procedure.

\paragraph{An exact target interval.} Put $H_t=M_t$ and
\[
 D_t=R_{2H_t}-R_{H_t}.
\]
Write the explicitly summed M\"obius contribution as the reduced rational
$Q_t=a_t/d_t$ and call it the \emph{M\"obius shadow}.  Define the remaining
term $F_t=D_t-Q_t$ to be the \emph{foreign defect}; it contains exactly the
channels absent from that explicit sum, so no part of $D_t$ is discarded.
The \emph{full target} is the residue class of $-a_t$ modulo $d_t$:
\[
 \mathsf{Hit}(t)
 \quad\Longleftrightarrow\quad
 \exists k\in\Z,\qquad d_tF_t=-a_t+d_tk.
\]
Since $D_t=a_t/d_t+F_t$, this is exactly the condition $D_t\in\Z$; no
complementary term is discarded or assigned a conjectural sign.

\begin{proposition}[diagonal integrality--target equivalence]\label{res:pincer}
For every $t$, diagonal integrality is equivalent to the foreign diagonal
defect meeting the full target.
\end{proposition}
\noindent This is the
\lword{DiagonalPincerDecomposition.lean}{215}{diagonal_int_iff_foreignDiagonalDefect_hits_fullTarget}{full-target equivalence}.

\begin{corollary}[exact full-target avoidance characterisation]
\label{res:fulltargetavoidance}
The following are equivalent:
\[
\begin{aligned}
 &S\notin\Q;\\
 &\forall\,t_0\in\N\ \exists\,t\geq t_0,\qquad
   \neg\mathsf{Hit}(t).
\end{aligned}
\]
\end{corollary}
\noindent The reverse implication is the
\lword{DiagonalPincerDecomposition.lean}{290}{irrational_totientSeries_of_full_target_avoidance_supply}{unbounded target-avoidance criterion}.
For the forward implication, Theorem~\ref{res:diag} supplies cofinally many
non-integral diagonal differences, and
Proposition~\ref{res:pincer} converts each one into a missed full target.
Thus the proposition is the fixed-scale equivalence and the corollary is its
exact global form.  The cofinal avoidance statement remains open.

\subsection{Fixed-scale mechanisms and conditional cofinal implications}
\label{app:totient-finite}

This subsection contains four logical statuses.  The squared-Mersenne
enclosure and the projected-separation tests are sufficient only: separation
forces a missed target, with no converse asserted.  The prime-power unit-gap
characterisation, channel decompositions, shadow-denominator formula, and
residue--projection equivalence are exact at fixed parameters; except where
stated, they are not characterisations of target avoidance.  Any corollary
concluding irrationality adds an unproved cofinal repetition hypothesis.  The
tabulated rows, primality facts, and diagonal certificates through $t=64$
are verified finite instances.  No result here supplies the missing cofinal
quantifier in Appendix~\ref{app:totient-normal}.

\paragraph{A one-way analytic enclosure.}
For $H\ge1$, put $\mathcal D_H=R_{2H}-R_H$,
$c_H=2^H(2^H-1)$, and
$\Phi_N=\sum_{0\le n\le N}\ph(n)2^{N-n}$.  The binary shift identity gives
$\mathcal D_H=c_HS+\Phi_H-\Phi_{2H}$.  Truncate the M\"obius-square identity
of Proposition~\ref{res:lens} after $D$ terms.

\begin{proposition}[squared-Mersenne enclosure]\label{res:squareenclosure}
For $D\ge0$, the rational centre and its exact error are
\[
 \begin{aligned}
 C_{H,D}&=\Phi_H-\Phi_{2H}
   +c_H\left(\frac12+\sum_{1\le d\le D}
      \frac{\mu(d)}{(2^d-1)^2}\right),\\
 \mathcal D_H-C_{H,D}
   &=c_H\sum_{d>D}\frac{\mu(d)}{(2^d-1)^2}.
 \end{aligned}
\]
The second identity is exact, and
$|\mathcal D_H-C_{H,D}|\le4c_H/[3(2^{D+1}-1)^2]$.  Hence separation of
$C_{H,D}$ from every integer by more than this bound forces
$\mathcal D_H\notin\Z$ and full-target avoidance.  No converse is asserted.
\end{proposition}
\noindent Formalised:
\lword{SquaredMersenneDiagonalEnclosure.lean}{116}{scaleDiagonalTailDifference_sub_lambertProjectedDiagonal}{squared-Mersenne tail identity}
and
\lword{SquaredMersenneDiagonalEnclosure.lean}{237}{scaleFullTarget_miss_of_lambert_projected_separation}{Lambert-projected separation criterion}.

\paragraph{A reduced-denominator finite predicate.}  Separately, there is a
stricter finite condition for an odd reduced denominator
$u$.  Its predicate $\mathsf{UnitGap}(u,N,K)$ asks that
every candidate integer in the associated dyadic tail interval be non-coprime
to $u$.  At a prime power $u=p^e$, the condition is equivalent to the exact
alternative that the interval has no candidate, or that it has one candidate
and $p$ divides that candidate.  Thus such an interval has at most one
candidate.

\begin{proposition}[prime-power unit-gap ceiling]\label{res:unitgap}
For every prime power $p^e$ with $e>0$, a
$\mathsf{UnitGap}(p^e,N,K)$ has candidate count at most
one.  The row $(u,N,K)=(3,3,5)$ satisfies this predicate, although its
corresponding ordinary empty-gap inequality fails.
\end{proposition}
\noindent Formalised:
\lword{PrimitiveWeightCertificate.lean}{54}{reducedDenominatorUnitGapCert_prime_pow_iff}{prime-power unit-gap characterisation},
\lword{PrimitiveWeightCertificate.lean}{111}{reducedDenominatorCandidateCount_prime_pow_le_one}{prime-power one-candidate bound},
\lword{PrimitiveWeightCertificate.lean}{151}{reducedDenominatorUnitGapCert_three_window_3_5}{three-window unit-gap certificate}, and
\lword{PrimitiveWeightCertificate.lean}{157}{not_ordinary_gap_certificate_three_window_3_5}{three-window ordinary-gap failure}.
This is a finite certificate-level strengthening.  We do not establish that
rationality of $S$ supplies these predicates at unbounded parameters.

\paragraph{Verified finite certificate families.} Four bounded certificate
families are machine-checked.  Each successful $\mathsf{Sep}$ row is a
kernel-checked proof of one stated non-integral tail difference, not a numerical
experiment; the joint cone table uses its own sound finite refuter.

\begin{example}[verified finite instances, expanded record]\label{app:deposits-detail}
The following are checked by the Lean kernel:
\begin{itemize}[leftmargin=1.4em]
\item $\mathsf{Sep}(h,12,16)$ for every period $h\in[1,8]$;
\item finite certificates for every period $h\in[1,16]$ at a fixed window;
\item diagonal certificates $\mathsf{Sep}(M_t,M_t,L)$ at $t=1$ and the 27
strict lcm-jump scales through $64$, namely
\[t\in\{1,2,3,4,5,7,8,9,11,13,16,17,19,23,25,27,29,31,32,37,41,43,
47,49,53,59,61,64\};\]
\item a tabulated joint cone-vertex certificate.
\end{itemize}
\end{example}
\noindent The four validation classes are formalised, in order, by the
\lword{TotientTailPeriodKiller.lean}{404}{certifiedKill_all_small}{small-window certificate family},
\lword{CarrySurvivorExtinction.lean}{574}{certifiedKill_all_upto_sixteen}{periods through sixteen},
\lword{DiagonalPincerCertificatesT64.lean}{1967}{certifiedKill_diagonal_all_imported_through_t64}{28 imported diagonal certificates through scale 64},
and \lword{LcmConeNonflat.lean}{506}{coneNonflatCert_cells}{cone-cell certificate table}.
For the first three classes, fixed-parameter completeness converts each successful
$\mathsf{Sep}$ row into the stated non-integral tail difference.  The fourth uses
the separate cone-refuter soundness theorem.  None is evidence that certificates
occur at unbounded scales.

The first row of this table is worked in full following
Example~\ref{res:deposits}; its underlying implication is
\lword{TotientTailPeriodKiller.lean}{262}{tail_diff_notMem_int_of_certifiedKill}{certificate soundness}.
No row asserts that such differences occur at unbounded scales.

\paragraph{Exact finite channel split and sufficient projection test.}  There is a second finite
presentation of the same obstruction.  For a cutoff $D$, split the finite
M\"obius residue diagonal into the channels $d\leq D$ which divide $H$ and
the remaining channels.  The divisor part is exactly the truncated M\"obius
shadow, while every retained nondivisor channel is explicit.  Above the
stable cutoff $2H$, a finite window of omitted channels is bounded by
\[
 c_H\left(\frac{2}{2^D}+\frac{4}{3\,4^D}\right),
\]
where $c_H$ is the diagonal coefficient.  Thus, if the actual foreign defect
agrees with this projection within the displayed bound, and the projected
quantity is farther than that bound from every integer, then the full target
is missed.

\begin{proposition}[finite foreign-residue projection]\label{res:foreignresidue}
Let \(F_H^{(D)}\) be the retained nondivisor projection at cutoff \(D\), let
\(F_H\) be the actual foreign defect, and let \(Q_H\) be the scaled explicit
divisor shadow.  The finite residue diagonal is exactly the sum of
\(F_H^{(D)}\) and its divisor-channel complement.  For \(2H\leq D\), put
\[
 B_{H,D}=c_H\left(\frac{2}{2^D}+\frac{4}{3\,4^D}\right).
\]
If
\[
 |F_H-F_H^{(D)}|\leq B_{H,D}
 \quad\text{and}\quad
 |Q_H+F_H^{(D)}-z|>B_{H,D}\quad\text{for every }z\in\Z,
\]
then the full target at height \(H\) is not attained.
\end{proposition}
\noindent Formalised:
\lword{ActualForeignResidueProjection.lean}{307}{finiteResidueDiagonal_eq_projectedForeign_add_divisor}{foreign--divisor channel split},
\lword{ActualForeignResidueProjection.lean}{275}{abs_foreignTailWindow_le_foreignComplementBound}{omitted-channel bound},
and
\lword{ActualForeignResidueProjection.lean}{348}{scaleFullTarget_miss_of_projected_separation}{projected-separation criterion}.
The passage from finite windows to the analytic foreign complement, including
the corresponding kernel identity, is not supplied here.  Nor is an
unbounded family of separated projections; these are parts of the open
certificate-supply problem.

Second, the rational M\"obius shadow has an exact reduced denominator.  If
$r_t$ is the square-free kernel of $H_t$, $s_t=H_t/r_t$, and
$J(r_t)=\prod_{q}(q^2-1)$ is the odd Jordan scalar, the product running over
the odd prime divisors $q$ of $r_t$, then
\[
 d_t=
 \frac{2^{r_t}-1}
 {\gcd\!\bigl(2^{r_t}-1,\,s_t\lvert J(r_t)\rvert\bigr)}.
\]
For $t\geq5$, the product of the Mersenne factors $2^p-1$ over primes
$t/2<p\leq t$ divides $d_t$; in particular
$d_t\geq2^{\lfloor t/2\rfloor}$.  This is a
denominator theorem for $Q_t$ only.  Cancellation by $F_t$ is exactly what the
full-target condition retains.

\begin{proposition}[Mersenne-shadow denominator]\label{res:shadowden}
Let \(r_t\) be the square-free kernel of \(H_t\), let \(s_t=H_t/r_t\), and
let \(J(r_t)\) be the odd Jordan scalar.  The reduced denominator \(d_t\) of
the scaled M\"obius shadow is
\[
 d_t=
 \frac{2^{r_t}-1}
 {\gcd\!\bigl(2^{r_t}-1,\,s_t\lvert J(r_t)\rvert\bigr)}.
\]
If \(t\geq5\), then
\[
 \prod_{\substack{p\ \mathrm{prime}\\ t/2<p\leq t}}(2^p-1)
 \ \mid\ d_t,
 \qquad d_t\geq2^{\lfloor t/2\rfloor}.
\]
\end{proposition}
\noindent Formalised:
\lword{MersenneShadowDenominatorGrowth.lean}{99}{exists_upperHalf_mersenne_channel_dvd_den_with_bounds}{upper-half Mersenne divisor bound}
and
\lword{MersenneShadowDenominatorGrowth.lean}{147}{lcmHeight_scaledMobiusShadow_den_exact}{exact shadow denominator}.

Finally, the remaining diagonal error can be represented by a finite integer
projection and then by binary suffix data.  At depth $L$, let
\[
 P_L(M)=\sum_{j=0}^{L-1}\ph(M+1+j)\,2^{L-1-j}
\]
be the length-$L$ window numerator underlying
Definition~\ref{def:totient-certificate}, and define
\[
 W_{t,L}=P_L(2H_t)-P_L(H_t),\qquad
 \eta_{t,L}=W_{t,L}\bmod 2^L,
\]
with the residue chosen in $[0,2^L)$.  The finite
residue certificate is the asymmetric safe interval
\[
 H_t+L+2<\eta_{t,L}<2^L-(2H_t+L+2).
\]
The two unequal margins are the tail bounds at the two cone vertices $H_t$ and
$2H_t$; replacing them by a symmetric slogan would lose part of the checked
statement.

\begin{corollary}[finite residue projection]\label{res:freshloss}
For every \(t,L\in\N\), the diagonal projection certificate is equivalent to
\[
 H_t+L+2<\eta_{t,L}<2^L-(2H_t+L+2),
\]
and either condition forces full-target avoidance at height \(t\).
Consequently,
\[
 \forall t_0\in\N\;\exists t\geq t_0\;\exists L\in\N,\qquad
 H_t+L+2<\eta_{t,L}<2^L-(2H_t+L+2)
\]
implies that \(S\) is irrational.
\end{corollary}
\noindent Formalised:
\lword{DiagonalFreshLossBridge.lean}{61}{diagonalFreshLossResidueCert_iff_projectionCert}{residue--projection equivalence}
and
\lword{DiagonalFreshLossBridge.lean}{191}{irrational_totientSeries_of_diagonalFreshLossProjectionSupply}{unbounded projection criterion}.
The finite equivalence is only between the residue inequality and the
projection certificate.  Neither condition is asserted to be necessary for
full-target avoidance or for irrationality.
The stronger adjacent-suffix supplies formalised in the same module imply the
displayed cofinal supply; they are not additional conclusions of one finite
certificate.
Finite instances of this implication are proved, but no unbounded projection
or suffix supply is established.

\paragraph{Primality certificates at changing-lcm scales through $t=64$.}\label{res:t64closure}
A finite family of Lucas certificates proves the primality facts required by
the diagonal certificates at every changing-lcm scale through $t=64$;
for example,
\lword{DiagonalPincerPrimeCertificates/ClosureT64.lean}{51759}{prime_lucas_2364533768205644535022723273}{representative Lucas primality certificate}.
This is a finite theorem family, not an unbounded certificate supply or an
irrationality proof.

\paragraph{The diagonal certificate at $t=64$.}\label{res:t64endpoint}
Using those primality facts, Lean checks the individual diagonal certificate
at $t=64$,
\lword{DiagonalPincerCertificateT64Endpoint.lean}{1928}{certifiedKill_diagonal_t64}{scale-64 diagonal certificate}.
This is one finite certificate, not an unbounded supply.

\subsection{Candidate cofinal inputs and no-go boundaries}
\label{app:totient-boundaries}

The results below have three distinct roles, and are not consecutive proof
steps.  The unproved three-rank band is a candidate cofinal sufficient input.
The affine equivalence, fresh-prime decomposition, and one-bit lift are exact
identities that expose quantities an unbounded argument would have to control;
the zero affine orbit shows only that uniform correction bounds do not provide
that control.  The parity, prime-adjunction, scalar-localisation, and square-CRT
results independently rule out four further shortcuts.  None of these no-go
results rules out the whole strategy, supplies the missing cofinal input, or
uses another no-go result as a premise.

\paragraph{Candidate cofinal input: a three-rank odd window.}\label{res:flexibleoddwindow}
At an analytically admissible odd power-of-two rank, centrality at any one of
the canonical odd rank or its next two base-four successors yields the
required adjacent-suffix gap condition,
\lword{DiagonalFlexibleOddWindowSupply.lean}{93}{diagonalAdjacentSuffixGapSupply_of_powerTwoOddGuardThreeRankBand}{three-rank suffix-gap implication}.
Thus a cofinal three-rank band supply implies irrationality,
\lword{DiagonalFlexibleOddWindowSupply.lean}{136}{irrational_totientSeries_of_powerTwoOddGuardThreeRankBandSupply}{cofinal three-rank criterion};
that supply is not proved.

\paragraph{Exact affine form and bounded zero-orbit obstruction.}\label{res:flexibleoddaffine}
With the stated short-correction bounds, the three central-band predicates are
equivalent to one three-scale affine escape condition,
\lword{DiagonalFlexibleOddWindowAffine.lean}{236}{threeRankBandAt_iff_threeScaleAffineEscapeAt_of_shortCorrections}{three-scale affine equivalence}.
A zero-start affine orbit shows that a uniform correction envelope alone does
not force escape at any finite number of ranks,
\lword{DiagonalFlexibleOddWindowAffine.lean}{225}{zeroAffineOrbit_stays_in_scaled_edge}{zero-orbit obstruction}.

\paragraph{Exact fresh-prime deficit decomposition.}  The literal foreign channel
also has an endpoint form suited to curvature tests.  Retain only the prime
support already present at $H$ in the two endpoints $H+s$ and $2H+s$, and
write the resulting discrepancy from the actual totient as an endpoint
deficit.  Each such deficit is nonnegative.  The actual diagonal increment
then equals the old-prime increment plus the lower deficit minus the upper
deficit.  The second and five-point curvature operators preserve this exact
split; the latter has coefficient row $[-8,4,2,1,1]$.  Consequently, the old
curvature margin loses at most the stated adverse weighted deficit.

\begin{proposition}[fresh-prime deficit decomposition]\label{res:freshprimedeficit}
The endpoint fresh deficits are nonnegative, give an exact decomposition of
the foreign channel, and bound the loss from the old-prime curvature margins.
\end{proposition}
\noindent Formalised:
\lword{FreshPrimeDeficitDecomposition.lean}{91}{endpointFreshDeficit_nonneg}{endpoint-deficit nonnegativity},
\lword{FreshPrimeDeficitDecomposition.lean}{184}{finiteForeignChannelIncrement_eq_endpointFreshDeficit_sub}{foreign-increment decomposition},
and
\lword{FreshPrimeDeficitDecomposition.lean}{269}{oldFivePoint_sub_secondBranchAdverseDeficit_le_actual}{adverse-deficit curvature bound}.
This supplies the exact correction term for the finite evaluations.  It supplies
neither a cofinal deficit bound nor an irrationality criterion.

\paragraph{Exact one-bit lifting.}  The power-of-two signed-margin condition admits a
separate exact arithmetic reduction.  If a signed difference is already
divisible by $2^k$, lifting its congruence to $2^{k+1}$ is equivalent to
evenness of the quotient after division by $2^k$.  Failure of the lift gives
the unique translated residue class.  In particular, a displayed difference
$P-N=16c$ agrees modulo $32$ exactly when $c$ is even; otherwise the residue
is the offset class by $16$.

\begin{proposition}[one-bit signed-margin lift]\label{res:powertwobitlift}
The fifth-bit question after fourth-bit agreement is exactly one cofactor
parity test, with the complementary offset class characterising failure.
\end{proposition}
\noindent Formalised:
\lword{PowerTwoBitLift.lean}{34}{two_pow_succ_dvd_iff_quotient_even}{one-bit divisibility lift},
\lword{PowerTwoBitLift.lean}{111}{modEq_thirtyTwo_iff_even_difference_cofactor}{mod-32 cofactor criterion},
and
\lword{PowerTwoBitLift.lean}{129}{not_modEq_thirtyTwo_iff_offset_sixteen}{offset-16 complement}.
The centred-state formulation identifies the signed-margin interval
condition with failure of this next lift, subject to its stated correction and margin
budget:
\lword{PowerTwoCenteredBitLift.lean}{121}{actualPenultimateSignedMargin_centeredLift_iff_not_twice_dvd}{centred-lift equivalence}
and
\lword{PowerTwoCenteredBitLift.lean}{159}{actualPenultimateSignedMargin_centeredThirtyTwo_iff_not_thirtyTwo_dvd}{centred mod-32 equivalence}.
This is a finite arithmetic equivalence.  It does not supply the required
cofactors at unbounded scales and gives no irrationality conclusion.

\paragraph{Independent no-go: parity is not a margin certificate.}  The preceding lift does not
settle the reduced signed-margin interval condition.  For a positive even
modulus \(m\) and integral radius parameter \(a\) with
\(16<a<8m-16\), each parity class contains both a residue inside the interval
and one outside it.  The same statement applies to a positive
power-of-two modulus.  Thus even a fifth-bit computation, by itself, cannot
decide the required margin condition.

\begin{proposition}[parity does not decide the reduced margin condition]
\label{res:paritysocket}
Let \(m,a\in\Z\), with \(m>0\) even and
\[
 16<a<8m-16,
\]
and put
\[
 \mathcal M_{m,a}
 =\{y\in\Z:\exists k\in\Z,\ 16|y-km|<a\}.
\]
Each parity class contains both a member and a nonmember of
\(\mathcal M_{m,a}\).  In particular, the conclusion holds for
\(m=2^j\) whenever \(j\geq1\).
\end{proposition}
\noindent Formalised:
\lword{Totient827ParitySocketCeiling.lean}{73}{reducedMarginSocket_not_decided_by_parity}{parity obstruction}
and
\lword{Totient827ParitySocketCeiling.lean}{110}{powerTwo_reducedMarginSocket_not_decided_by_parity}{power-of-two parity obstruction}.
The result is structural and proves no signed-margin certificate at any
particular scale; it excludes parity as a sufficient standalone criterion.

\paragraph{Independent no-go: prime adjunction.} One tempting way to seek an
unbounded obstruction is to compare the full target at $H$, $pH$, $qH$, and
$pqH$.  The following affine transport law rules out this route before any
projection is introduced.
\begin{proposition}[prime-adjunction transport collapse]\label{res:primeadjunction}
For every positive $H$ and $k$, the actual diagonal satisfies
\[
 D_{kH}=Q_k(H)D_H+Z_k(H),
 \qquad Q_k(H)\in\N,\quad Z_k(H)\in\Z.
\]
Hence integrality at $H$ propagates to every multiple, and the four target
hits at the prime-adjunction diamond are equivalent to the single hit at its
root.  The corresponding transport of the complete foreign defect has zero
diamond curvature: the two transport compositions around the diamond agree
exactly.
\end{proposition}
\noindent Formalised:
\lword{FullTargetPrimeAdjunctionNoGo.lean}{196}{fullTarget_primeAdjunction_diamond_iff_root}{the prime-adjunction diamond collapse}
and
\lword{FullTargetPrimeAdjunctionNoGo.lean}{247}{foreignTransportCorrection_primeAdjunction_flat}{the foreign-correction flatness}.
Thus explicit-shadow fibres alone cannot create a transport obstruction.  A
future positive route would need a separately defined projection together with
a theorem controlling what that projection discards.  This is a proved
obstruction to one strategy, not progress on the unbounded-supply proposition.

\begin{proposition}[scalar-localisation obstruction]\label{res:adelic}
Let $x\in\Q$ and $c\in\Z$.  If $H\mid\operatorname{den}(x)$ and the reduced denominator
of $cx$ divides $H$, then
\[
 \frac{\operatorname{den}(x)}{H}\mid |c|.
\]
\end{proposition}
\noindent Formalised:
\lword{AdelicHeightObstruction.lean}{23}{scalarLocalization_complement_dvd}{complementary-denominator divisibility}
and \lword{AdelicHeightObstruction.lean}{56}{scalarLocalization_integer_eq_mul_num}{scalar-localisation identity}.
Thus a scalar localisation can move a complementary denominator factor into
its coefficient, but cannot erase it.  This rules out a scalar-denominator
shortcut of that form.  It neither proves full-target avoidance nor supplies
any certificates, so it gives no irrationality criterion for $S$.

\begin{proposition}[square-CRT correction suppression]\label{res:squarecrt}
Let $E$ be a finite family indexed by distinct primes $p_i$, with prescribed
anchors $A_i$ and residues $a_i$.  There is a bounded common base $n$ such
that, for each $i\in E$ and each shift $h$ with $p_i\nmid a_i+h$, one has
\[
 n=A_i+p_i(a_i+p_i t)\quad\text{and}\quad
 \ph(n+p_i h-A_i)=(p_i-1)\ph(a_i+p_i t+h)
\]
for a suitable $t\in\N$.  In particular, if every $p_i$ exceeds a fixed
horizon $J$, the choice $a_i=0$ works simultaneously for all $1\le h\le J$.
\end{proposition}
\noindent Formalised:
\lword{SquareCRTCube.lean}{198}{exists_squareCRT_clean_totient_family}{finite clean totient family}
and \lword{SquareCRTCube.lean}{228}{exists_squareCRT_clean_horizon}{clean horizon family}.
The finite mechanism does not force a cube coefficient to be nonzero: there
is a checked clean two-step cube with both displayed coefficients zero
(\lword{SquareCRTCube.lean}{285}{squareCRT_clean_block_can_vanish}{vanishing clean block}) and a
separate checked clean cube with second coefficient $-4$
(\lword{SquareCRTCube.lean}{303}{squareCRT_clean_block_can_be_nonzero}{nonzero clean block}).
Thus the result gives neither an unbounded certificate supply nor an
irrationality criterion for $S$.

\subsection{Unconditional dyadic rank and the rationality-side carry bound}
\label{app:totient-rank}

The results below again have different logical roles.  The dyadic
totient-kernel rank theorem is unconditional: its CRT--Dirichlet construction
gives an explicit separated minor at every level.  Under rationality, this
forces a tempered carry orbit of unbounded section rank; no theorem supplies
the arithmetic upper bound needed for a contradiction, and rank alone cannot
do so.  The later determinant identities and even-residue reduction are exact
finite inputs, while the compressed-adjoint and residual-gauge results delimit
particular certificate strategies.  None supplies the missing arithmetic
coupling or proves that (S) is irrational.

\begin{theorem}[unconditional dyadic totient-kernel rank]
\label{res:dyadictotient}
For every $e\geq0$, the canonical level-$e$ dyadic totient-kernel family is
linearly independent over $\Q$ and its span has dimension $2^e+1$.
Consequently, the span of all sections
\[
 n\longmapsto \ph(2^j n+r),\qquad j\geq0,\quad 0\leq r<2^j,
\]
is not finite-dimensional over $\Q$.
\end{theorem}
\noindent Formalised by the all-level minor construction
\lword{TotientMahlerDefect.lean}{847}{exists_separatedMinorCertificate_totientAffineOddFamily}{CRT--Dirichlet separated minors},
the independence theorem
\lword{TotientMahlerDefect.lean}{900}{linearIndependent_canonicalTotientKernelFamily}{canonical kernel independence},
the exact rank formula
\lword{TotientMahlerDefect.lean}{954}{finrank_canonicalTotientKernel_eq}{exact canonical rank},
and
\lword{TotientMahlerDefect.lean}{1014}{not_finiteDimensional_span_fullTotientKernel}{infinite full-kernel dimension}.
The theorem concerns the coefficient kernel.  It does not imply that $S$ is
irrational.

\paragraph{Relation to prior work.}
The qualitative conclusion is not new.  Finite-dimensionality of the span of
all sections is \(2\)-regularity of the sequence in the sense of Allouche
and Shallit~\cite{alloucheshallit}, and Coons proved in 2010 that \(\ph\) is
not \(k\)-regular for any \(k\ge2\), by counting poles of
\(\zeta(s-1)/\zeta(s)\) against the meromorphic continuation available to
the Dirichlet series of a regular sequence~\cite{coons2010}.  In particular
the infinite-dimensionality of the full kernel span follows from Coons's
theorem.  What Theorem~\ref{res:dyadictotient} adds is quantitative and
constructive: the exact canonical rank \(2^e+1\) at every level, with an
explicit separated minor, proved by elementary CRT--Dirichlet arithmetic
that the formal source checks without analytic continuation.
Proposition~\ref{res:carryrank} uses exactly this level-by-level form.

\begin{proposition}[carry-section rank forced by rationality]\label{res:carryrank}
If the totient coefficient series were rational, there would be a
positive-multiplier tempered integral carry orbit whose canonical
carry-section span has dimension at least $2^e-1$ for every $e$.
\end{proposition}
\noindent Formalised:
\lword{TotientCarryKernelRigidity.lean}{211}{finrank_canonicalCarryKernel_ge_of_linearIndependent}{the carry-kernel rank lower bound}
and
\lword{TotientCarryKernelRigidity.lean}{284}{not_irrational_totientSeries_implies_unbounded_carryRank_unconditional}{the rationality-to-unbounded-rank implication}.
No theorem bounds the dyadic section rank of every rationality-supplied
tempered carry.  The proposition therefore identifies an infinite
rationality-side barrier, not an irrationality criterion for $S$.

\begin{remark}[rank alone cannot decide rationality]
No converse dimension bound is available at this generality: the same class
of coefficient streams satisfying $0\leq\gamma(n)\leq n$ contains rational
series whose tempered orbits have
infinite-dimensional section span.  Choose Boolean values
\(a_N\in\{0,1\}\) with \(a_0=a_1=1\) whose dyadic sections satisfy no
nontrivial rational linear relation (a routine diagonal choice on sparse
index blocks); set \(u_N=N+a_N\) and \(\gamma(N+1)=2u_N-u_{N+1}\).  Then
\(0\le\gamma(n)\le n\), the orbit \(u\) is tempered, and
Theorem~\ref{res:rigidity} evaluates
\(\sum_{n\ge1}\gamma(n)2^{-n}=u_0=1\in\Q\); yet the sections of \(u\) span
an infinite-dimensional space, because the sections of \(N\mapsto N\) span
only two dimensions.  This elementary observation is not formalised in the
source.  It fixes the shape of the missing ingredient: a route through
Proposition~\ref{res:carryrank} must couple carry-section rank to the
arithmetic size of a totient carry, since dimension alone is compatible
with rationality.
\end{remark}

\begin{proposition}[exact rectangular determinant expansion]\label{res:signedmoment}
For matrices $M:\iota\times\kappa\to R$ and
$N:\kappa\times\iota\to R$ over a commutative ring, the determinant of
$MN$ is the sum over all maps $p:\iota\to\kappa$ of the determinant of the
corresponding selected-column matrix times the diagonal product
$\prod_i N_{p(i),i}$.
\end{proposition}
\noindent This is the
\lword{SignedQMomentObstruction.lean}{28}{det_mul_rectangular}{rectangular determinant expansion}.

\begin{proposition}[exact dominant-dyadic-term test]\label{res:dominantdyadic}
If one selected integer numerator is odd and its dyadic exponent is strictly
larger than every other selected exponent, then the common numerator after
dyadic denominator clearing is odd and hence nonzero.
\end{proposition}
\noindent Formalised by
\lword{SignedQMomentObstruction.lean}{77}{scaled_dyadic_sum_odd}{dominant dyadic parity}, and
\lword{SignedQMomentObstruction.lean}{95}{scaled_dyadic_sum_ne_zero}{dominant dyadic nonvanishing}.
These are finite algebraic inputs to a signed-moment route.  They do not show
that any actual totient Hankel determinant is nonzero, and they give no
irrationality criterion for $S$.

The formal proof also exposes the generic theorem that any supplied separated
minor gives the corresponding rank,
\lword{TotientMahlerDefect.lean}{101}{finrank_canonicalTotientKernel_eq_of_certificate}{separated-minor rank criterion}.
The two generic algebraic propositions above remain useful as possible inputs
to other determinant routes, but the totient-kernel minor itself is supplied
at every level by Theorem~\ref{res:dyadictotient}.

\begin{proposition}[exact even-residue reduction]\label{res:evenresidue}
Every even residue channel in the canonical dyadic totient-kernel family is
exactly a scalar multiple of its odd-core channel.
\end{proposition}
\noindent Formalised as the
\lword{TotientMahlerDefect.lean}{159}{totientKernel_even_residue_reduce}{even-residue reduction}.
This removes duplicate channels before the separated minor is constructed.

\begin{proposition}[compressed-adjoint size obstruction]\label{res:compressedadjoint}
If positive integers $Q,v$ and integers $A,b$ satisfy $QvA=b$ and
$|b|<Qv$, then $A=0$.
\end{proposition}
\noindent Formalised as the
\lword{TotientMahlerDefect.lean}{1051}{false_of_compressedAdjointCertificate}{compressed-adjoint contradiction}.
This finite incompatibility does not construct the data required by a
compressed-adjoint route.

\begin{proposition}[residual-gauge obstruction for monomial minors]\label{res:residualgauge}
Let $z_j,W_j\ne0$ and $M_{ij}=W_jz_j^{e_i}$.  Then
\(det M=\det(z_j^{e_i})\prod_jW_j\), so the column weights preserve
determinant nonvanishing.  If $e_i=1$, choosing $W_j=z_j^{-1}$ makes row $i$
identically $1$.  Allowing row-dependent weights
$W_{ij}=z_j^i/z_j^{e_i}$ reconstructs a consecutive-power matrix whose
zeroth row is identically $1$.
\end{proposition}
\noindent Formalised:
\lword{ResidualGaugeObstruction.lean}{55}{det_residualMonomialMatrix_ne_zero_iff}{residual-gauge determinant equivalence},
\lword{ResidualGaugeObstruction.lean}{94}{locked_reconstruction_preserves_nonzero_minor}{locked-reconstruction nonvanishing}, and
\lword{ResidualGaugeObstruction.lean}{145}{rowReconstruction_zero_row_one}{reconstructed zero row}.
Thus rank, determinant, or conditioning information alone cannot exclude the
target without an arithmetic constraint on the residual weights.  Such a
coupled criterion remains possible; the proposition gives no irrationality
criterion for $S$.

\paragraph{Appendix consequence.}
Nothing in this appendix strengthens the fixed-parameter completeness of
Proposition~\ref{res:complete} into an unbounded supply.  Its remaining condition is
therefore the still-open quantifier in Theorem~\ref{res:red21}:
\[
 \forall h\geq1\;\forall N_0\;\exists N\geq N_0\;\exists L\geq1,\qquad
 \mathsf{Sep}(h,N,L).
\]
The one-scale mechanisms and conditional implications in
Appendix~\ref{app:totient-finite}, the candidate inputs and no-go boundaries
in Appendix~\ref{app:totient-boundaries}, and the conditional rank routes in
Appendix~\ref{app:totient-rank} delimit possible approaches to this
statement.  None provides the required certificates after every cutoff.

}

\section{One transform and the limit of finite evidence}
\label{sec:architecture}

\noindent This section maps the two open reductions and the limit of finite
evidence.  It adds organisation, not theorems; its closing elementary
observation is explicitly unformalised.

\subsection{One transform, one mechanism, three questions}
\label{sec:one-transform}

The Lambert rearrangement of Section~\ref{sec:ladder} factors the common
transform through two maps, the divisor transform followed by binary
evaluation:
\[
 f\;\longmapsto\;f*\mathbf 1\;\longmapsto\;
 \sum_{m\ge1}\frac{(f*\mathbf 1)(m)}{2^m}\;=\;L(f).
\]
Here \((f*\mathbf 1)(m)=\sum_{d\mid m}f(d)\) is the \emph{divisor stream}:
the first map sums over divisors, and the second reads those sums as
binary-series coefficients.  The primitive-conductor input \(f=\alpha\) gives
\(L(\alpha)=S\) with divisor stream \(\alpha*\mathbf 1=\ph\)
(Proposition~\ref{res:lift}); a support indicator \(f=\mathbf 1_A\) gives
\(L(\mathbf 1_A)=X_A\) with divisor stream \(f_A\)
(Appendix~\ref{app:carry-boolean}); and the comparison inputs
\(\mu\), \(\mathbf 1\), and \(\ph\) give the comparison values listed in
Section~\ref{sec:ladder}.

For any \(\N\)-valued stream \(\gamma\) with \(\gamma(n)\le n\), including
\(\ph\) and every \(f_A\), Theorem~\ref{res:rigidity} says that
\(\sum_n\gamma(n)2^{-n}\) is rational exactly when an integer sequence
\(u(N)\) and positive integer $v$ satisfy
\(v\gamma(N+1)=2u(N)-u(N+1)\) and \(u(N)=o(2^N)\).  This exact recurrence
plus subexponential bound is what \emph{tempered integer orbit} means here.
Tail differences (Section~\ref{sec:reduction}) and carry states
(Theorem~\ref{res:rigidity-main}) are two coordinates for that condition.
M\"obius inversion recovers the underlying input: \(\alpha=\ph*\mu\) for
\#249 and \(\mu*f_A=\mathbf 1_A\) for \#257
(Proposition~\ref{res:boolmob}).

The three open questions have different quantifiers, even though they use
this one map.  Write \(\mathcal I_\infty\) for the indicators
\(\mathbf 1_A\) of infinite supports \(A\subseteq\{2,3,\ldots\}\).  Then:
\begin{enumerate}[label=\textup{(\roman*)},leftmargin=2.2em]
\item Erd\H{o}s~\#249 fixes one input and asks for the status of its one
value: \(L(\alpha)=S\notin\Q\)?
\item Universal \#257 asks whether every permitted value is irrational:
\(L(\mathcal I_\infty)\cap\Q=\varnothing\).  Its negation is one infinite
support $A$ with $X_A\in\Q$.
\item The half-value route fixes one rational target and asks an existence
question:
\(L^{-1}(1/2)\cap\mathcal I_\infty\ne\varnothing\).  Here
\(L^{-1}(1/2)\) is the \emph{fibre}, meaning all inputs mapped to $1/2$.
Theorem~\ref{res:greedy} allows at most one positive-exponent support in that
fibre, namely the canonical greedy orbit.
\end{enumerate}
Question (iii) would refute (ii), but failure at the particular target
$1/2$ would not prove (ii): another rational target could still be reached.
Figure~\ref{fig:transform} marks these three ways of interrogating the map;
it supplies no theorem transferring status between them.

\begin{figure}[H]
\centering
\small
\begin{tikzpicture}[>={Stealth[length=1.8mm]}, yscale=0.92]
% three aligned columns: inputs right-aligned, streams in a fixed-width
% centre column, outcomes left-aligned, so all ten arrows share two rails
  \node[font=\footnotesize, anchor=east] at (1.5,1.15) {input $f$};
  \node[font=\footnotesize] at (5.3,1.15) {divisor stream $f*\mathbf 1$};
  \node[font=\footnotesize, anchor=west] at (9.0,1.15) {value $L(f)$ and status};
  \draw[->] (1.62,1.15) -- node[above,font=\scriptsize]{divisor transform} (3.62,1.15);
  \draw[->] (6.98,1.15) -- node[above,font=\scriptsize]{binary evaluation} (8.88,1.15);
  \node[anchor=east] (r1a) at (1.5,0)    {$\alpha=\ph*\mu$};
  \node[anchor=east] (r2a) at (1.5,-0.9) {$\ph$};
  \node[anchor=east] (r3a) at (1.5,-1.8) {$\mathbf 1$};
  \node[anchor=east] (r4a) at (1.5,-2.7) {$\mu$};
  \node[anchor=east] (r5a) at (1.5,-3.6) {$\mathbf 1_A$, $A$ infinite};
  \node[minimum width=2.7cm] (r1b) at (5.3,0)    {$\ph$};
  \node[minimum width=2.7cm] (r2b) at (5.3,-0.9) {$\mathrm{Id}$};
  \node[minimum width=2.7cm] (r3b) at (5.3,-1.8) {$\mathbf 1*\mathbf 1$};
  \node[minimum width=2.7cm] (r4b) at (5.3,-2.7) {$1,0,0,0,\ldots$};
  \node[minimum width=2.7cm] (r5b) at (5.3,-3.6) {$f_A$};
  \node[anchor=west] (r1c) at (9.0,0)    {$S$: open \textup{(i)}};
  \node[anchor=west] (r2c) at (9.0,-0.9) {$2$: rational};
  \node[anchor=west] (r3c) at (9.0,-1.8) {$E$: irrational};
  \node[anchor=west] (r4c) at (9.0,-2.7) {$1/2$: rational};
  \node[anchor=west] (r5c) at (9.0,-3.6) {$X_A$: ever rational? \textup{(ii)}};
  \foreach \r in {1,2,3,4,5}{
    \draw[->] (r\r a) -- (r\r b);
    \draw[->] (r\r b) -- (r\r c);
  }
  \draw[->,densely dashed,shorten <=1mm,shorten >=1mm]
    ([xshift=2.5mm]r5c.east) to[bend right=52]
    node[right,font=\scriptsize,xshift=1mm]{(iii)?}
    ([xshift=2.5mm]r4c.east);
\end{tikzpicture}
\caption{The one transform behind the paper, with the statuses of
Section~\ref{sec:ladder}.  Labels (i)--(iii) are the three questions of
Section~\ref{sec:one-transform}: \#249 fixes the top input and asks about the single image
point \(S\); universal \#257 restricts the domain to indicators of
infinite supports and asks whether its image ever meets \(\Q\); the
half-value route fixes the value \(1/2\), reached by the signed
M\"obius input through the stream \(1,0,0,\ldots\), and asks whether an
infinite Boolean input reaches it as well.  The figure organises
statements proved in Sections~\ref{sec:ladder}--\ref{sec:eb} and
Appendix~\ref{sec:carry}; it asserts nothing further.}
\label{fig:transform}
\end{figure}

\subsection{Why no status transfers}

The transform alone does not determine status: Section~\ref{sec:ladder} gives
rational \(L(\ph)=2\) and \(L(\mu)=1/2\), irrational
\(L(\mathbf 1)=E\), transcendental \(L(\mathrm{Id})\), and open
\(L(\alpha)=S\).

Nor do simple input properties distinguish them: \(\ph\) and \(\mathbf 1\)
are positive, multiplicative, and at most linear, yet their values are
rational and irrational.  Every finite Boolean support has rational value,
whereas Theorem~\ref{res:support} gives infinite supports with irrational
values.  Appendix~\ref{app:carry-boolean} works out the finite support
\(\{2,3\}\), value \(10/21\), and its period-six carry orbit.

Progress must therefore use input-specific arithmetic: exact totient values
for \#249, Boolean M\"obius inversion plus infinitude for universal \#257,
or greedy divisor feedback for the half-value fibre.
Appendix~\ref{app:totient-boundaries} rules out several generic shortcuts.

\subsection{The shared bottleneck}

Both reductions require finite events beyond every proposed cutoff, but the
conclusions point in opposite directions.  In literal quantifiers,
\[
\begin{aligned}
 S\notin\Q
 &\quad\Longleftrightarrow\quad
 \forall h\ge 1\;\forall N_0\;\exists N\ge N_0\;\exists L\;
 \mathsf{Sep}(h,N,L),\\
 \frac12\in\mathcal A
 &\quad\Longleftrightarrow\quad
 \forall N_0\;\exists s\ge\max\{N_0,6\}\;(\tau_s=0).
\end{aligned}
\]
The first line requires separating certificates after every cutoff for each
shift $h$; the second requires terminal zero bits after every row cutoff.
The events belong to different constructions and imply opposite kinds of
conclusion: irrationality in the first line, membership in the second.
Proposition~\ref{res:complete}, Theorem~\ref{res:pointwisecomplete}, and
Theorem~\ref{res:red21} give the first equivalence; Theorem~\ref{res:halfmembership}
gives the second.  Neither required cofinal supply is known.  By contrast,
the classical full-support argument uses Bertrand--CRT and averaging to
construct witnesses at every precision.  Other coordinate identities and
rational-support filters do not change either missing quantifier.

Thus useful progress must control \(\mathsf{Sep}\)-certificates at unbounded
parameters for \#249, or produce terminal zeros beyond every row cutoff (or
an eventually right greedy seam) for the half-value route.  More fixed-scale
computation alone changes neither open statement.

\paragraph{What finite witnesses can decide.}
Here a \emph{finite witness} is exactly checkable data whose conclusion does
not assume how the uninspected tail behaves; its force depends on the claim.
For the half-value route, a finite calculation can settle nonmembership.
By Theorem~\ref{res:fatalright}, \(1/2\notin\mathcal A\) is equivalent to the
existence of a fatal half-gap.  Exact rational evaluation of the greedy
prefix, together with the rational tail enclosures of
Proposition~\ref{res:death}, can verify such a gap after finitely many steps.
The gap excludes every support agreeing with that prefix.  The reverse
direction has no finite certificate here: survival through any finite row is
inconclusive, while membership requires \(\tau_s=0\) beyond every row cutoff.

For \(S\), each \(\mathsf{Sep}\)-certificate has narrower scope: it proves
that one specified tail difference is non-integral.  Theorem~\ref{res:red21}
requires certificates after every cutoff for every positive shift, so no
single certificate proves irrationality.  Rationality forces eventual
integrality, but failure to find a certificate at one truncation proves
neither integrality nor rationality: a greater binary depth \(L\) may still
certify the same tail difference.  Thus finite discovery can certify
half-value nonmembership, but not half-value membership or either status of
\(S\) within these certificate systems.

\begin{remark}[why finite agreement does not decide the open cases]
The following two elementary countermodels are not formalised in the source.
Both alter the object beyond a finite inspected range; neither makes a claim
about the actual totient stream or the actual Mersenne weights.

For \#249, any fixed finite family of certificate computations inspects only
finitely many coefficients.  The splice in Section~\ref{sec:249-open}
constructs a stream \(\gamma\) agreeing with \(\ph\) on all of them, still
satisfying \(0\leq\gamma(n)\leq n\), whose binary series is rational.  The
same finite output is therefore compatible with rationality.  A proof of the
missing supply must use totient values beyond every fixed bound.

For the half-value route, fix any finite prefix of greedy decisions and
replace only the uninspected weights by a ratio-one-third geometric tail.
If its total mass is below the pending residual, the complete modified tail
falls short and the last inspected skip is final.  If its total mass equals
the residual, that tail realises the target.  Both choices preserve strict
tails and every inspected decision.  Thus a finite prefix and strict-tail
geometry do not decide membership; a proof must use the specific Mersenne
arithmetic of the tail.
\end{remark}

\section{Formal verification and reproducibility}\label{sec:verify}

\noindent This section records proof authority and reproducibility; it does
not add a mathematical conclusion to either reduction.

\paragraph{Proof authority and source identity.}
The Lean~4 proof assistant~\cite{lean} and Mathlib library~\cite{mathlib} check
the named results.  The source, exact informal-to-formal links, and build
instructions are available in the
\href{https://github.com/wcook04/plectis-lean-erdos249-257}{public repository}.
This paper pins formal-source revision \texttt{\commitshort}, published under
the tag
\href{https://github.com/wcook04/plectis-lean-erdos249-257/tree/formal-source-2026-07-18}{\texttt{formal-source-2026-07-18}}.
The blue links resolve to immutable source files at that checkpoint.
The release also pins the Lean and Mathlib
versions; the toolchain is Lean \texttt{v4.29.1}.

\paragraph{Reproduction commands and artefact identities.}
A cold clone at the published revision is checked with \texttt{lake build}.
The release-validation command also checks that every registered claim resolves
to its stated Lean declaration and manuscript location, that the PDF matches
the committed source, and that the public source pin resolves to the recorded
revision.  The resulting report records release validation; it is not
mathematical proof authority.  The repository records SHA-256 identities for
the claims file, Lean-declaration index, manuscript source, rendered PDF, and
generated navigation files, together with their regeneration commands.  No
hardware-independent build time is claimed;
elapsed time depends on the Lean cache and machine, and is not part of the
mathematical claim.

\paragraph{What the computation establishes.}
Lean elaborates each declaration into a proof term, which the kernel checks;
tactics construct these terms but are not separate proof authority.
\texttt{norm\_num} constructs concrete arithmetic proofs, and \texttt{decide}
reduces a decidable finite proposition to obtain a proof.  The
\lword{HalfFinalMiddlePhaseSieve.lean}{2050}{finalMiddleTwentySixPhaseSurvivors_card}{finite survivor-count theorem}
uses \texttt{decide} to count 412 phases.  Its local \texttt{maxRecDepth}
setting only raises the permitted unfolding depth; the proposition and logical
rules are unchanged.  It does not use compiler-backed native evaluation, and
its adjoining \texttt{\#print axioms} audit
reports neither \texttt{Lean.trustCompiler} nor \texttt{sorryAx}, Lean's
unfinished-proof axiom.  Generated Lean certificate declarations receive the
same check.  An external dataset, Python release script, generated index, or
this manuscript does not.  Appendix~\ref{app:index} connects each displayed
result to its exact formal statement.

\paragraph{Section consequence.}
At the published formal-source checkpoint, the linked declarations and their
checked proof terms certify the stated proved, conditional, finite, and
equivalence results.  This verification does not change the mathematical status of the
unresolved statements: the open problems collected in
Section~\ref{sec:further} remain open.

\section{Exact remaining problems and further questions}\label{sec:further}

Section~\ref{sec:spines} gives the exact reductions.  Here we record two
concrete sufficient criteria and then state the unresolved questions.
Erd\H{o}s \#249 has one open obligation in two equivalent forms.  Erd\H{o}s
\#257 has a universal statement and a counterexample route pointing in
opposite directions: half-value membership would refute the universal
statement, while nonmembership would decide only the half-value question.

\long\def\FurtherLambertAppendix{%

\subsection{Weighted Lambert comparison families}
\label{sec:additional-lambert}
\label{app:lambert-weighted}

This subsection changes the coefficient input.  It settles or classifies
eventually-periodic comparison series and gives a nonzero period-four input
whose rearranged coefficients vanish on one residue class.  None has the
totient weight defining \(S\), so no status transfers to Erd\H{o}s~\#249.

\paragraph{Periodic weights.}\label{sec:periodicweights}
For \(L_b(\gamma)=\sum_{a\ge1}\gamma(a)/(b^a-1)\), the checked cases are:
\begin{theorem}[eventually-periodic nonnegative weights]
\label{res:periodicnonnegative}
For every $b\ge2$, an eventually-periodic nonnegative rational weight that is
not identically zero on its eventual period gives an irrational series.
\end{theorem}
\noindent Formalised:
\lword{CertificateKernel.lean}{12483}{irrational_ratWeightSeries_eventuallyPeriodic}{eventually-periodic nonnegative criterion}.
This is contained in a broader theorem of Luca--Tachiya~\cite{lucatachiya}.
\begin{theorem}[periodic signed-weight dichotomy]\label{res:periodicsigned}
For a periodic integer weight \(\gamma\), either \(L_b(\gamma)\) is irrational,
or \(b^kL_b(\gamma)\in\Z\) for some \(k\ge0\); equivalently, its base-\(b\)
expansion terminates.
\end{theorem}
\noindent Formalised:
\lword{CertificateKernel.lean}{13847}{irrational_or_bpow_mul_eq_intCast_intWeightedErdosSeries_periodic}{periodic signed-weight dichotomy}.
\begin{proposition}[period-four example]\label{res:periodfourobstruction}
For the period-four weight \(\gamma=1,0,-1,0,\ldots\), its coefficient after
Lambert rearrangement, \(c(n)=\sum_{d\mid n}\gamma(d)\), vanishes whenever
\(n\equiv3\pmod4\).
\end{proposition}
\noindent Formalised:
\lword{CertificateKernel.lean}{13898}{intWeightedCoeff_periodFourSignWeight_eq_zero_of_mod_four_eq_three}{period-four coefficient obstruction}.
Thus one vanishing coefficient class does not force a zero input; the
nonperiodic totient-derived weight \(\alpha=\ph*\mu\) is not covered.

\subsection{Finite algebraic boundaries and squared-Lambert coordinates}
\label{app:lambert-algebraic}

This subsection keeps the relevant coordinates but separates three roles:
finite denominator-clearing obstructions, the exact M\"obius-square identity
for \(S\), and comparison rows in the squared-Lambert calculus.  The identity
changes coordinates; the finite propositions delimit algebraic methods; and
the status of one squared-Lambert row does not pass to another.

\paragraph{Finite denominator-clearing boundaries.}\label{sec:primitiveindex}
Put \(a_n=(\ph*\mu)(n)/n\).  Let $T_N$ be the product of $4$, every odd
prime $p\le N$, and one further factor $p$ for each $p^2\le N$.
\begin{proposition}[finite denominator obstruction]
\label{res:primitiveindex}
For $N\ge4$, if a positive integer $D$ makes $Da_n$ integral for every
$1\le n\le N$, then $T_N\mid D$.  Hence no fixed positive $D$ clears every
$a_n$.
\end{proposition}
\noindent Formalised:
\lword{PrimitiveDeterminantLift.lean}{148}{twoTierPrimorial_dvd_of_integral_jet}{two-tier primorial obstruction}.

For \(d\ge1\), define the pointed periodic atom
\[
 \omega_d(N)=\frac{2^{\,N\bmod d}}{2^d-1}.
\]
It satisfies the exact carry law
\[
 2\omega_d(N)-\omega_d(N+1)=\mathbf1_{d\mid N+1}.
\]

\begin{proposition}[finite Mersenne-atom determinant identity]
\label{res:mersennetailatoms}
Fix \(m\ge1\), conductors \(d_j\ge1\), indices \(n_{ij}\ge0\), and integer
column weights \(w_j\), and set
\[
 W_{ij}=w_j\omega_{d_j}(n_{ij}),
 \qquad U_{ij}=2^{\,n_{ij}\bmod d_j}.
\]
Then
\[
 \left(\prod_{j=1}^m(2^{d_j}-1)\right)\det W
 =\left(\prod_{j=1}^m w_j\right)\det U\in\Z.
\]
If the integer on the right is nonzero, then the absolute value of the
cleared determinant on the left is at least \(1\).
\end{proposition}
\noindent Formalised:
\lword{MersenneTailAtoms.lean}{28}{mersenneTailAtom_step}{Mersenne-atom carry law},
\lword{MersenneTailAtoms.lean}{158}{weightedMersenneAtom_denominatorProduct_mul_det}{determinant-clearing identity},
and
\lword{MersenneTailAtoms.lean}{181}{one_le_abs_weightedMersenneAtom_cleared_det}{cleared-determinant lower bound}.
Thus the displayed product makes the determinant integral and gives only the
lower bound $1$ when nonzero.  It is not an irrationality criterion for \(S\).

\begin{proposition}[M\"obius-square identity for \(S\)]\label{res:lens}
\[
S \;=\; \tfrac12 \;+\; \sum_{d\ge1}\frac{\mu(d)}{(2^d-1)^2}.
\]
\end{proposition}
\noindent Formalised: \lword{CertificateKernel.lean}{18126}{totient_series_eq_half_add_moebius_mersenne_square}{M\"obius-square identity}.
Writing \(T=\sum_{d\ge1}\mu(d)/(2^d-1)^2\), Proposition~\ref{res:coprime}
gives \(T=\Pr(\gcd(X,Y)=1)=S-\tfrac12\), so \(S\) is irrational exactly
when \(T\) is.  By contrast,
\(R_N=2^NS-\sum_{n\le N}\ph(n)2^{N-n}\) is an integer translate of a
scaled copy of \(S\).  Tail differences encode rationality as explained in
Section~\ref{sec:reduction}; they are not further constants equal to \(T\).

\paragraph{Squared Lambert calculus.}\label{sec:squared}
For \(L_2(f)=\sum_{d\ge1}f(d)/(2^d-1)^2\), the denominator is the probability
that \(d\) divides two independent geometric waiting times.  The formal
transfer theorem,
\lword{GcdMomentCalculus.lean}{105}{tsum_lambert_linear_weight_sq_pure}{squared Lambert transfer},
therefore gives a divisor calculus for their gcd.
\begin{proposition}[squared Lambert identity]\label{res:gcdmoment}
At \(r=1/2\), the three rows \(f=\mu,\mathbf 1,\ph\) yield respectively
\[
 L_2(\mu)=S-\tfrac12,\qquad
 L_2(\mathbf 1)=\sum_{n\ge1}\frac{\sigma(n)-\tau(n)}{2^n},\qquad
 L_2(\ph)=\sum_{n\ge1}\frac{P(n)-n}{2^n},
\]
where \(P\) is Pillai's gcd-sum function.
\end{proposition}
\noindent Formalised: \lword{GcdMomentCalculus.lean}{216}{tsum_one_div_mersenne_sq_eq_sigma_sub_tau_series}{constant-weight squared-Lambert identity} and
\lword{GcdMomentCalculus.lean}{235}{tsum_totient_div_mersenne_sq_eq_gcd_moment_series}{totient gcd-moment identity}.
Write
\(\zeta_q(1)=\sum_{n\ge1}q^n/(1-q^n)\) and
\(\zeta_q(2)=\sum_{n\ge1}nq^n/(1-q^n)\).
Postelmans--Van Assche prove that \(1,\zeta_q(1),\zeta_q(2)\) are linearly
independent over \(\Q\) whenever \(1/q\) is an integer at least
\(2\)~\cite{postelmansvanassche}.  At \(q=1/2\), their difference is the middle
row, so that row is irrational.  This does not transfer to the M\"obius row,
which is exactly the open constant \(S-\tfrac12\).

\subsection{Probability-coordinate geometry: reduced slopes and cylinders}
\label{app:lambert-probability}

These results live entirely inside the coprime-pair probability coordinate.
They normalise reduced directions and quantify Stern--Brocot cylinder and run
geometry.  They are unconditional geometric consequences and do not imply
the open certificate supply.

\paragraph{Reduced directions and cylinder masses.}
Divide the fair-coin pair by its gcd.  For positive coprime \(a,b\), put
\[
  \nu(a,b)=\frac1{2^{a+b}-1},
  \qquad
  M(a,b)=\frac1{(2^a-1)(2^b-1)}.
\]
The first quantity is the probability that the reduced direction is
\((a,b)\); the second is the mass of the Stern--Brocot cylinder rooted at
\((a,b)\).  Let \(M_d(a,b)\) be the sum, over the first \(d\) generations
below that root, of the node contribution \((2^{u+v}-1)^{-1}\).  The children
of \((u,v)\) are \((u+v,v)\) and \((u,u+v)\).

\begin{proposition}[reduced-direction normalisation]\label{res:directionnormalization}
The reduced directions form a probability distribution,
\[
  \sum_{\substack{a,b\ge1\\(a,b)=1}}\nu(a,b)=1.
\]
\end{proposition}
\noindent This is the
\lword{GcdMomentCalculus.lean}{349}{tsum_pos_coprime_inv_mersenne_eq_one}{reduced-direction mass one}.

\begin{theorem}[Stern--Brocot cylinder recursion and convergence]
\label{res:sternbrocotcylinders}
At every positive node,
\[
  M(a,b)=\frac1{2^{a+b}-1}+M(a+b,b)+M(a,a+b),
\]
and, uniformly in \(a,b\) and \(d\),
\[
  0\le M(a,b)-M_d(a,b)
  \le \left(\frac23\right)^d M(a,b).
\]
Consequently \(M_d(a,b)\to M(a,b)\); at the root \(M(1,1)=1\).
\end{theorem}
\noindent Formalised:
\lword{GcdMomentCalculus.lean}{474}{cylinderMass_split}{cylinder mass recursion},
\lword{GcdMomentCalculus.lean}{525}{sternBrocotDepthMass_error}{depth-$d$ remainder bound}, and
\lword{GcdMomentCalculus.lean}{560}{tendsto_sternBrocotDepthMass}{cylinder convergence}.
The recursion is the elementary rational identity obtained after setting
\(A=2^a\), \(B=2^b\); the quantitative point is that the two children retain
at most two thirds of their parent's mass.

\paragraph{Run lengths and arithmetic height.}
Write an alternating Stern--Brocot word as \(r\) nonempty runs of lengths
\(n_i=1+e_i\), with \(e_i\ge0\).  Starting with \(P_\varnothing=(1,1)\), define
\[
  P_{(n_1,\ldots,n_r)}=(n_1A+B,A)
  \quad\text{when}\quad P_{(n_2,\ldots,n_r)}=(A,B),
  \qquad H(n_1,\ldots,n_r)=P_1+P_2.
\]
Thus \(H\) is the arithmetic height of the primitive pair reached by the
word.  Let \(F_0=0,F_1=1\).

\begin{theorem}[Fibonacci height bound and exact run-length defects]
\label{res:sternbrocotruns}
Every \(r\)-run word satisfies
\[
 H(1+e_1,\ldots,1+e_r)
 \ge F_{r+3}+F_{r+1}\sum_{i=1}^r e_i,
\]
and \(H(1,\ldots,1)=F_{r+3}\).  A defect confined to position
\(i\in\{0,\ldots,r-1\}\) has the exact size
\[
 H(\underbrace{1,\ldots,1}_{i},1+e,
   \underbrace{1,\ldots,1}_{r-i-1})
 =F_{r+3}+F_{i+2}F_{r-i+1}e.
\]
More generally, the excess over $F_{r+3}$ is a polynomial in the $e_i$, with
nonnegative coefficients and degree at most one in each variable; the checked
recurrence gives every mixed term.
\end{theorem}
\noindent Formalised:
\lword{SternBrocotRunGeometry.lean}{185}{runHeight_fib_lower}{Fibonacci height lower bound},
\lword{SternBrocotRunGeometry.lean}{263}{runHeight_defectRunLengths}{multiaffine run-height formula},
\lword{SternBrocotRunGeometry.lean}{343}{runHeight_defect_fib_sum_lower}{aggregate defect lower bound},
\lword{SternBrocotRunGeometry.lean}{413}{runHeight_oneSiteDefect}{one-site defect identity}.

\begin{proposition}[one-run probability mass]\label{res:onerunmass}
The two orientations with one nonempty Stern--Brocot run have total mass
\[
 \frac12.
\]
\end{proposition}
\noindent This is the
\lword{SternBrocotRunGeometry.lean}{467}{oneRunMass_eq_half}{one-run mass}.

\begin{proposition}[unit-run bookkeeping exponent]\label{res:rundenominator}
Along the all-unit \(r\)-run, let
\(E_r=\sum_{j=0}^{r-1}F_{j+2}\) be the sum of the invariant-coordinate
exponents.  Then \(E_r=F_{r+3}-2\).
\end{proposition}
\noindent This is the
\lword{SternBrocotRunGeometry.lean}{491}{naturalRunDenominatorExponent_add_two}{unit-run exponent sum}.
This identity does not make $E_r$ a reduced-denominator exponent for the
probability mass or for \(S\).

The Lambert and M\"obius identities used above are classical
~\cite{apostol,merca2017,mercaschmidt}, and the mediant language goes back to
Farey~\cite{farey1816}.  Postelmans--Van Assche supply the cited irrationality
of the constant-weight squared-Lambert row, not these cylinder or run
statements.  We claim the displayed Lean-checked identities and estimates,
not priority for their assembled package.  The bookkeeping exponent
\(E_r=F_{r+3}-2\) is two below the minimum height $F_{r+3}$.  No theorem here
proves that a run tail survives denominator clearing.
}

\subsection{Two concrete scalar criteria}

Two scalar conditions give concrete sufficient routes toward the two open
directions.  Both are exact arithmetic statements; neither is proved at the
unbounded range needed for its final conclusion.

For \#249, retain $M_t=\lcm(1,\ldots,t)$ and write
\[
 P_L(N)=\sum_{j=0}^{L-1}\ph(N+1+j)\,2^{L-1-j}.
\]
Thus \(P_L(N)\) is the base-two weighted block of the next \(L\) totient
values after \(N\).  At scale \(t\), choose one logarithmic depth and compare
the blocks after \(M_t\) and \(2M_t\):
\[
 m_t=\lfloor\log_2 M_t\rfloor+10,\qquad
 d_t=\bigl(P_{m_t}(2M_t)-P_{m_t}(M_t)\bigr)\bmod 2^{m_t},
\]
where $0\le d_t<2^{m_t}$.  The integer \(d_t\) is their block-difference
residue.  Define
\[
  \sigma_t=\min\bigl(d_t-2^{m_t-5},
    (2^{m_t}-2^{m_t-5})-d_t\bigr)\in\Z.
\]
This is the signed distance from the nearer edge of the central band below:
it is nonnegative exactly on that band and negative outside it.

\begin{proposition}[fixed-depth central-band equivalence]\label{res:slackfrontier}
For every $t\ge1$, the residue \(d_t\) lies in the central band
$[2^{m_t-5},2^{m_t}-2^{m_t-5}]$ if and only if $\sigma_t\ge0$.
\end{proposition}
\noindent This fixed-scale equivalence and its beyond-every-cutoff strict-jump
form are the
\lword{DiagonalFreshLossBridge.lean}{2469}{canonicalAdjacentSuffixCentral_iff_slack_nonneg}{canonical central-band equivalence}
and
\lword{DiagonalFreshLossBridge.lean}{2559}{canonicalAdjacentSuffixJumpCentralSupply_iff_slackSupply}{strict-jump slack equivalence}.

\begin{corollary}[strict-jump slack criterion]
\label{res:slackcriterion}
If, for every $t_0$, there is a $t\ge\max\{3,t_0\}$ such that
\[
 M_t<M_{t+1}\qquad\text{and}\qquad \sigma_t\ge0,
\]
then $S$ is irrational.
\end{corollary}
\noindent This implication is the
\lword{DiagonalFreshLossBridge.lean}{2731}{irrational_totientSeries_of_canonicalAdjacentSuffixJumpSlackSupply}{strict-jump slack irrationality criterion}.
The proposition is a fixed-\(t\) equivalence; the corollary gives only
sufficiency.  No converse is known: completeness may choose a depth from the
tail difference, whereas this criterion fixes \(m_t\).  Persistent negative
slack at that depth refutes only this route and does not decide whether \(S\)
is rational.

For \#257, write \(r_n=r_n(1/2)\) for the unrepresented remainder after rank
\(n\).  The exact tail expansion and second-scale normalisation are
\[
 T_n=2^{-n}+\frac13\,4^{-n}+E_n,
 \qquad u_n=4^n(2r_n-2^{-n}),
\]
where \(0\le4^nE_n\le(2/7)2^{-n}\).  The identity
\(u_n-1/3=4^n(2r_n-T_n)+4^nE_n\) explains the centre \(1/3\): distance from it
is the scaled residual--tail imbalance, up to the bounded error.  Under
\(r_n\le T_n\), the next step survives exactly when
\(\lvert2r_n-T_n\rvert\ge w_{n+1}-T_{n+1}\).  A sufficient condition is
\[
  \frac16+\frac{37}{56}2^{-n}\le \left|u_n-\frac13\right|.
\]
We call this the \emph{second-channel separation condition}.  Here
\(37/56=2/7+3/8\) budgets the higher-tail and gap errors, respectively.
The absolute value is needed because \(u_n<0\) at \(n=3,4\).

\begin{proposition}[finite second-channel separation]\label{res:secondfrontier}
The second-channel separation condition is decidable over $\Q$ and holds for
$1\le n\le6$.
\end{proposition}
\noindent This is the
\lword{GreedyAchievementSet.lean}{2723}{halfSecondChannelSeparatedRat_of_pos_le_six}{finite second-channel separation}.

\begin{corollary}[second-channel separation criterion]
\label{res:secondcriterion}
If the second-channel separation condition holds for every \(n\ge7\), then
\(1/2\) belongs to the Mersenne achievement set \(\mathcal A\).
\end{corollary}
\noindent This implication is the
\lword{GreedyAchievementSet.lean}{2901}{half_mem_mersenneAchievementSet_of_secondChannelSeparationRat_from_seven}{conditional from-seven membership criterion}.
The finite proposition supplies \(1\le n\le6\), but no case with \(n\ge7\);
no converse is proved.  Membership itself has the exact characterisation
\[
  1/2\in\mathcal A
  \quad\Longleftrightarrow\quad
  \text{the canonical greedy orbit omits infinitely many exponents}.
\]
This is formalised at
\lword{GreedyAchievementSet.lean}{2514}{half_mem_mersenneAchievementSet_iff_greedySkippedSupport_infinite}{half-membership--infinite-skip equivalence}.
The forward implication uses the
\hyperref[res:full]{full-support irrationality theorem}; for the reverse,
if a residual ever exceeds the remaining tail, every later exponent is
selected; hence only finitely many omissions remain.  The \(n\ge7\) hypothesis
is open.  If proved, it would yield an infinite support of rational value
\(1/2\), refuting rather than proving the universal \#257 statement.

\paragraph{Dependency direction.}
The two criteria have the proved logical form
\[
\begin{array}{c}
 \text{after every cutoff, nonnegative slack}\\[-1pt]
 \text{occurs at a strict lcm jump}\\
 \Downarrow\\
 \text{unbounded \#249 certificate supply}\\
 \Downarrow\\
 S\notin\Q
\end{array}
\qquad\qquad
\begin{array}{c}
 \text{second-channel separation}\\[-1pt]
 \text{for every }n\ge7\\
 \Downarrow\\
 \dfrac12\in\mathcal A\\
 \Updownarrow\\
 \mathcal S(1/2)\text{ is infinite}.
\end{array}
\]
Neither premise is proved.  The range \(1\le n\le6\) is only a finite prefix
of the second.  The left conclusion settles \#249; the right would refute,
not prove, the universal \#257 statement.

\subsection{The unresolved statements}

\paragraph{Erd\H{o}s \#249: one obligation in two equivalent forms.}
\begin{problem}[Erd\H{o}s \#249]
Prove that
\[
S=\sum_{n\ge1}\frac{\ph(n)}{2^n}
\]
is irrational.
\end{problem}

By Theorem~\ref{res:red21}, the preceding problem is equivalent to the
following exact quantifier statement.

\begin{problem}[unbounded certificate supply]
Prove~\eqref{eq:unbounded-certificate-supply}.
Equivalently, prove the lcm-diagonal supply in
Theorem~\ref{res:diag}.  The finite scales in
Example~\ref{res:deposits} do not establish either unbounded form.
\end{problem}

These are not two independent open problems: Theorem~\ref{res:red21} and
Theorem~\ref{res:diag} identify the second as an exact certificate form of
the first.

\paragraph{Erd\H{o}s \#257: universal statement and counterexample route.}
\begin{problem}[universal Erd\H{o}s \#257]
Prove that
\[
  \sum_{n\in A}\frac{1}{2^n-1}
\]
is irrational for every infinite set $A\subseteq\Npos$, rather than only for the
support families in Theorem~\ref{res:support}.
\end{problem}

\begin{problem}[the half-value counterexample route]
Decide whether $1/2\in\mathcal A$.  By
Theorem~\ref{res:halfmembership}, this is equivalent to the canonical greedy
orbit omitting infinitely many exponents.  A positive answer would give an
infinite support of rational value $1/2$ and refute the universal \#257
statement; a negative answer would decide only this distinguished
point-membership question.
\end{problem}

\paragraph{Conclusion: what the reductions change.}
The paper supplies a checked map and reusable finite tools, not the unbounded
mechanism that would decide either direction.  The reductions change form,
not status.  For \#249 the one missing theorem is
unbounded certificate supply, equivalently in pointwise or lcm-diagonal form.
For the half-value route to \#257, arbitrary supports give way to one greedy
orbit; in final-skip coordinates it remains to exclude cells \(-2,-1\) and
prove the future-tail inequality at the remaining nonnegative-carry middle
transitions.  Supply beyond every cutoff proves \#249; terminal zero beyond
every row cutoff refutes universal \#257; eventual right behaviour proves
only \(1/2\notin\mathcal A\).

\paragraph{Two refinements not pursued.}
Certificate completeness gives a least successful depth at every
non-integral pair, but no growth bound in \((h,N)\); such a bound would decide
each fixed pair.  Separately, the zero-window theorem is inert when a support
contains a small element (Appendix~\ref{app:carry-sublog}); coverage outside
every finite core would remove this.  We have not pursued either question.

\pdfbookmark[1]{Statements and declarations}{statements-declarations}\subsection*{Statements and declarations}

\paragraph{Artefact and data availability.} The
\href{\sourceurl}{archived formal-source revision} contains the Lean sources,
pinned toolchain, library manifest, and generated certificate data used in the
verification. Repository metadata and this manuscript provide navigation
rather than proof authority; Section~\ref{sec:verify} gives the verification
route.

\paragraph{Use of AI assistance.} Large-language-model assistants were used
during development for proof drafting, refactoring, and prose editing. Every
registered formal claim is linked to a declaration in the archived source
revision. Lean checks each proof term against the pinned library; the project
contains no proof placeholders or project-defined axioms.

\paragraph{Funding and competing interests.} This work received no external
funding. The author declares no competing interests.

\paragraph{Acknowledgements.} The problem numbering and status follow the
Erd\H{o}s Problems catalogue maintained by Thomas Bloom~\cite{erdosproblems}.

\appendix
\section{Technical reductions for the half-value branch}
\label{app:half-technical}

This appendix records the local arithmetic used to reach
Theorems~\ref{res:halfmembership} and~\ref{res:fatalright},
Proposition~\ref{res:finalskipexclusions}, and
Corollary~\ref{res:taildominance}.  Its statements have four statuses: exact
identities or equivalences; implications with an unproved hypothesis; exact
finite calculations; and counterexamples showing that a proposed local
argument is insufficient.  None changes the two open conclusions at the end
of Section~\ref{sec:eb}.  Readers concerned only with the main reductions may
skip the appendix.

The parts are ordered by their role in the greedy argument.
Appendix~\ref{app:half-localizer} computes the residual interval for one
take--skip--take pattern, without showing that the half-value orbit enters or
avoids it.  Appendix~\ref{app:half-conditional-producers} gives one
conditional irrationality criterion for a structured support, several
sufficient hypotheses for \(1/2\in\mathcal A\), and the local identities they
use.  Appendix~\ref{app:half-finite-cofinal} separates a finite witness from
the unbounded supply of such witnesses needed for membership.
Appendix~\ref{app:half-final-skip} identifies nonmembership with a last greedy
skip, excludes some possible last transitions, and states the two cell
exclusions and tail inequality still missing.  Appendix~\ref{app:half-evidence}
contains finite checks, changes of coordinates, and counterexamples to
shortcuts.

The two routes used below have the logical form
{\setlength{\arraycolsep}{3pt}
\[
\begin{array}{rclcl}
\text{A.2--A.3 unbounded-depth witnesses}
&\Longrightarrow& \frac12\in\mathcal A
&\Longrightarrow& \text{infinite support summing to }\frac12,\\
\text{A.4 exclusions and tail bound}
&\Longrightarrow& \text{no final skip}
 &\Longleftrightarrow& \frac12\in\mathcal A.
\end{array}
\]
}
The displayed implications are proved, and the final equivalence is exact.
Both leftmost hypotheses are open.  Thus the appendix explains what would
suffice and why several finite arguments fail; it does not decide whether
\(1/2\in\mathcal A\).

Throughout this appendix:
\begin{itemize}[leftmargin=1.6em]
\item a \emph{take} (respectively a \emph{skip}) at rank \(n\) is the
greedy decision \(\varepsilon_n=1\) (respectively \(\varepsilon_n=0\));
\item the \emph{finite greedy seam at row \(s\)} is the Boolean word of
Definition~\ref{def:terminal-bit}; an unqualified \emph{seam} in this
appendix is an adjacent pair of such words at a branch boundary, and the
compounds seam pair, one-hole seam, first-wrap seam, and rewind seam carry
this pair sense;
\item the \emph{square-root strip} at depth \(N\) is the carry interval
\([1,H_N]\) with \(H_N=2\lfloor\sqrt N\rfloor+4\); an interval of carries
kept available by a construction is \emph{protected};
\item a \emph{hole} of a seam is the missing carry value in an otherwise
covered protected interval;
\item a \emph{rewind} is the inverse finite recurrence of a seam;
\item the \emph{transition carry} is the integer coordinate appearing in
the exact residual identity, and the \emph{remainder} of a row is its
integer greedy residual;
\item the branches are \emph{upper} ($G_{s+1}=H_s$), \emph{middle}
($G_{s+1}=G_s$), and \emph{right} ($G_{s+1}=G_s\cup\{s\}$);
\item the \emph{pulse} of a rank at row \(s\) counts its divisor
incidences at \(2s+1\) and \(2s+2\), the first counting twice; the pulse
of a word is the sum over its ranks.
\end{itemize}
These terms denote mathematical objects, not stages of a proof search.
Module-local coordinates not listed here (for example the centred, lazy,
or first-wrap variants) are defined in the linked formal statements.

\HalfValueTechnicalAppendix

\paragraph{Appendix consequence.}
This appendix supplies the local implications behind
Theorems~\ref{res:halfmembership} and~\ref{res:fatalright},
Proposition~\ref{res:finalskipexclusions}, and
Corollary~\ref{res:taildominance}.  It does not establish the corollary's
global hypothesis.  A sufficient hypothesis is either a cofinal window, seam, or
terminal approximation producing \(1/2\in\mathcal A\), or both an exclusion
of every actual middle transition with \(C_D=-2,-1\) and a proof of
\(\Theta_D<C_D\) at every remaining non-\((-3)\) middle transition.  The
bounded checks and local no-go theorems provide neither route.

\section{Auxiliary rationality criteria from binary carry systems}\label{sec:carry}

This appendix supplies the general results assembled in
Theorem~\ref{res:rigidity-main} and
Corollaries~\ref{res:rigidity-unbounded-main}--\ref{res:rigidity-dyadic-mass-main},
and the detailed greedy geometry cited after Theorem~\ref{res:greedy}.
It has two independent roles.  Appendices~\ref{app:carry-normal-form}
and~\ref{app:carry-boolean} give exact coordinate changes: rationality is
equivalent first to a \emph{tempered tail orbit}, an integer recurrence with
a vanishing boundary condition, and then to a Boolean--M\"obius certificate,
an integer carry orbit whose quotient stream reconstructs a Boolean support.
Appendices~\ref{app:carry-sublog} and~\ref{app:carry-mass} use those coordinates
only forward, to derive necessary divisor-coverage, denominator-wrap,
reciprocal-mass, and carry-growth conditions.  A \emph{wrap} occurs when
doubling a least residue crosses its odd modulus; the reciprocal mass is
\(\rho(A)=\sum_{a\in A}1/a\).

Appendix~\ref{app:carry-geometry} is separate.  It studies the achievement
set itself: a finite nonmembership certificate proves exclusion, whereas survival
through a finite depth proves nothing.  The following display records these
statuses before introducing the detailed coordinates.

The dependency chain is
\[
\begin{array}{rcl}
 \text{a rational support value}
 &\Longleftrightarrow& \text{a tempered integral tail orbit}
 \quad\text{(B.1)},\\
 \text{a specified support fraction \(p/q\)}
 &\Longleftrightarrow& \text{a Boolean--M\"obius carry certificate}
 \quad\text{(B.2)},\\
 \text{a nonempty rational support}
 &\Longrightarrow& \text{sublogarithmic zero gaps}\quad\text{(B.3)},\\
 \text{a nonempty support, odd part }v>1
 &\Longrightarrow& \text{wrap and reciprocal-mass bounds}
 \quad\text{(B.4)},\\
 \text{an infinite rational support}
 &\Longrightarrow& \text{unbounded carry states}
 \quad\text{(B.4)},\\
 \text{an infinite dyadic-denominator support}
 &\Longrightarrow& \text{the reciprocal-mass alternative}
 \quad\text{(B.4).}
\end{array}
\]
The last four lines list independent necessary conditions, at the weakest
hypotheses under which they are proved, not converses and not
contradictions.  Appendix~\ref{app:carry-geometry} is orthogonal to this
rationality chain:
finite greedy failure proves nonmembership in the achievement set, whereas
survival through any finite depth proves nothing.

The prior-art boundary is theorem-family-specific. The integral carry
recurrence has a direct antecedent in Wang~\cite{wang2026};
the strict-tail geometry is recorded by Kova\v{c}--Tao~\cite{kovactao};
M\"obius inversion, repetend algebra, and divisor averaging are classical
(see, for example, Apostol~\cite{apostol}). We make no claim of mathematical
priority for the converse/rigidity, certificate-normal-form, or coupled
reciprocal-mass/collision statements collected here.

\subsection{Exact normal form: tail-orbit rigidity}
\label{app:carry-normal-form}

This subsection supplies the generic equivalence used in
Appendix~\ref{app:carry-boolean}.
It converts rationality into an integer recurrence with a boundary condition;
it does not itself exclude such an orbit or impose the Boolean support
constraint.

For a coefficient sequence $\gamma:\N\to\N$ with $\gamma(n)\le n$, write
\[
X_\gamma=\sum_{n\ge1}\frac{\gamma(n)}{2^n},
\qquad
T_\gamma(N)=\sum_{j\ge1}\frac{\gamma(N+j)}{2^j},
\]
the series and its scaled tail after $N$ binary digits. Call an integer
sequence $u$ a \emph{tempered orbit} for $\gamma$ with multiplier $v\ge1$ if
\[
u(N+1)=2\,u(N)-v\,\gamma(N+1)\ \text{ for all } N,
\qquad\text{and}\qquad
u(N)/2^N\to0 .
\]
The recurrence is exact binary long division against the coefficient stream;
the boundary condition forbids the homogeneous solution. Wang uses the
integral carry recurrence forced by rationality for the weighted binary
special case $\gamma(n)=n d_n$~\cite{wang2026}. The theorem below extends that
forward mechanism to arbitrary nonnegative $\gamma(n)\le n$; its explicit converse
and subexponential-orbit rigidity are the local additions, with no priority
claim attached. Wang's positive-density theorem and its Erd\H{o}s
\#260 corollary are not used or formalised here.

\begin{theorem}[tail-orbit rigidity]\label{res:rigidity}
Let $\gamma(n)\le n$ for all $n$. Every tempered orbit is the scaled tail,
$u(N)=v\,T_\gamma(N)$ for all $N$; and $X_\gamma$ is rational if and only if a tempered
orbit exists for some multiplier $v\ge1$.
\end{theorem}
\noindent Formalised:
\lword{GenericTailOrbitRigidity.lean}{338}{temperedBinaryOrbit_eq_scaledTail}{the scaled-tail orbit identity}
(rigidity) and
\lword{GenericTailOrbitRigidity.lean}{425}{binaryCoeffSeries_rational_iff_exists_temperedBinaryOrbit}{the rationality--orbit equivalence}
(the criterion), restated in Mathlib's irrationality vocabulary at
\lword{GenericTailOrbitRigidity.lean}{434}{not_irrational_binaryCoeffSeries_iff_exists_temperedBinaryOrbit}{the non-irrationality--orbit equivalence}.
The engine is one observation: a real orbit with $d(N+1)=2\,d(N)$ and
$d(N)=o(2^N)$ vanishes identically
(\lword{GenericTailOrbitRigidity.lean}{314}{doublingOrbit_eq_zero_of_tempered}{zero-tail rigidity}).

The tempered boundary is essential: the
homogeneous parasite $N\mapsto2^N$ can be added to any orbit without
disturbing the recurrence, so positivity of an orbit certifies nothing by
itself, and positivity is nowhere advertised as an equivalent criterion. The
linear-growth hypothesis covers both constants introduced in
Section~\ref{sec:intro}:
$\ph(n)\le n$ for the \#249 coefficients, and $f_A(n)\le\tau(n)\le n$ for the
\#257 support coefficients $f_A(n)=\#\{a\in A:a\mid n\}$.  The results below
instantiate the criterion on supports.  Its separate totient
carry-rank consequence belongs to the \#249 rank route in
Appendix~\ref{app:totient-rank}.

\subsection{Exact \#257 coordinates: Boolean--M\"obius carry certificates}
\label{app:carry-boolean}

This subsection has one preliminary boundary result and then one exact
equivalence chain.  The boundary result rules out extracting the desired
Boolean support directly from the signs in \(L(\mu)=1/2\); it is not a premise
of the carry equivalence that follows.

\begin{proposition}[M\"obius-sign support no-go]\label{res:mobiussignnogo}
Let
\[
 A_-=\{d\ge2:\mu(d)=-1\},
 \qquad A_+=\{d\ge2:\mu(d)=1\}.
\]
Then
\[
 \sum_{d\in A_-}\frac1{2^d-1}
 =\frac12+\sum_{d\in A_+}\frac1{2^d-1}
 \ge\frac12+\frac1{63}>\frac12.
\]
In particular, the signed identity \(L(\mu)=1/2\) cannot be converted into a
Boolean support of value \(1/2\) merely by selecting the negative M\"obius
indices.
\end{proposition}
\noindent Formalised:
\lword{MobiusSignSupportNoGo.lean}{111}{tsum_negativeMobius_eq_half_add_positiveMobiusTail}{the negative-M\"obius decomposition}
and
\lword{MobiusSignSupportNoGo.lean}{150}{half_add_one_div_sixty_three_le_tsum_negativeMobius}{the strict lower bound}.
The term \(d=6\) supplies the displayed \(1/63\).  This rules out one
sign-truncation route only; it does not exclude another infinite Boolean
support with value \(1/2\).

\paragraph{The exact coordinate chain.}
At base $2$ the support series of Section~\ref{sec:eb} is itself a binary
coefficient series, $\sum_{n\in A}1/(2^n-1)=X_{f_A}$
(\lword{BooleanMobiusCarry.lean}{376}{erdosSupportSeries_two_eq_binaryCoeffSeries}{the support-series/binary-series identity}).
Boolean M\"obius inversion first identifies the support from its divisor
coefficients; tail-orbit rigidity then identifies rationality with a tempered
integer orbit.  Theorem~\ref{res:carrycert} is exactly the composition of
those two equivalences.  The divisor transform and M\"obius inversion are
classical Lambert-series tools~\cite{apostol,merca2017,mercaschmidt}; the
contribution here is their Lean-checked composition with the tempered-orbit
criterion.

\begin{proposition}[Boolean M\"obius inversion, both directions]\label{res:boolmob}
On positive integers, $f_A=\mathbf 1_A*\mathbf 1$ and $\mu*f_A=\mathbf 1_A$, where
$\mathbf 1_A$ is the indicator of $A$. Conversely, every integer-valued
arithmetic function $f$ whose M\"obius transform is Boolean,
$(\mu*f)(n)\in\{0,1\}$ for all $n\ge1$, satisfies $f=f_B$ for the support
$B=\{n\ge1:(\mu*f)(n)=1\}$ selected by that transform.
\end{proposition}
\noindent Formalised:
\lword{BooleanMobiusCarry.lean}{76}{positiveSupportBitAF_mul_zeta}{the positive-support divisor identity},
\lword{BooleanMobiusCarry.lean}{94}{moebius_mul_supportCoeffAF}{the M\"obius inversion identity}, and, with no
search or finiteness hypothesis, the converse
\lword{BooleanMobiusCarry.lean}{180}{eq_supportCoeffAF_booleanMobiusSupport_of_boolean}{the Boolean reconstruction theorem}.

On the carry side, Theorem~\ref{res:rigidity} applies at $\gamma=f_A$: the support
series is rational exactly when a tempered carry orbit for $f_A$ exists
(\lword{BooleanMobiusCarry.lean}{383}{erdosSupportSeries_rational_iff_exists_temperedCarry}{the rationality--carry equivalence});
a displayed fraction $p/q$ corresponds to an orbit $U$ with multiplier $q$ and
initial value $U(0)=p$
(\lword{BooleanMobiusCarry.lean}{466}{support_fraction_iff_exists_temperedCarry}{the support-fraction carry equivalence});
and exact division recovers the coefficient from the orbit,
$\bigl(2U(N)-U(N+1)\bigr)/q=f_A(N+1)$
(\lword{BooleanMobiusCarry.lean}{680}{temperedCarry_quotient_eq_supportCoeff}{carry-quotient recovery}).
The elementary divisor-pair bound $\tau(n)\le2\lfloor\sqrt n\rfloor$
confines the tails to a square-root strip,
$T_{f_A}(N)\le2\sqrt N+4$.  For a nonempty support the orbit states are
positive and satisfy $U(N)\le q\,(2\sqrt N+4)$
(\lword{BooleanMobiusCarry.lean}{705}{temperedCarry_le_denominator_mul_two_sqrt_add_four}{the carry upper bound});
and the strip is strong enough to re-derive the tempered boundary, so the
analytic side condition can be traded for one inequality. The two coordinate
changes compose.

\begin{theorem}[certificate-level equivalence]\label{res:carrycert}
Fix an integer $p$ and an integer $q\ge1$. There is a nonempty support $A$ of
positive integers with $\sum_{n\in A}1/(2^n-1)=p/q$ if and only if there is an
integer sequence $U$ with $U(0)=p$ such that: every $U(N)$ is positive;
$U(N)\le q\,(2\sqrt N+4)$; $q$ divides $2U(N)-U(N+1)$ for every $N$; and the
M\"obius transform of the quotient sequence $\bigl(2U(N)-U(N+1)\bigr)/q$ is
Boolean. The support is not guessed: it is reconstructed from the certificate
by Proposition~\ref{res:boolmob}.
\end{theorem}
\noindent Formalised as
\lword{BooleanMobiusCarry.lean}{948}{exists_normalized_support_fraction_iff_exists_booleanMobiusCarry}{the normalised support-fraction certificate equivalence}.

\begin{example}[the support $\{2,3\}$]
The support $\{2,3\}$ has value
$\tfrac13+\tfrac17=\tfrac{10}{21}$
(\lword{BooleanMobiusCarry.lean}{1054}{erdosSupportSeries_support23_eq_ten_div_twenty_one}{the two-point series value}),
its orbit is the pure period-six cycle $10,20,19,17,13,26$ with multiplier
$21$ (\lword{BooleanMobiusCarry.lean}{1103}{carryOrbit23_isTempered}{the tempered two-point orbit}), and the
M\"obius transform of the carry quotient recovers exactly $\{2,3\}$
(\lword{BooleanMobiusCarry.lean}{1140}{mobius_carryOrbit23_recovers_support}{support recovery}).
\end{example}

These are coordinates, not by themselves a strategy: nothing here excludes
infinite Boolean carry paths, and no finite search is promoted to a
completeness argument.

\subsection{Necessary condition I: sublogarithmic divisor coverage}
\label{app:carry-sublog}

This theorem is a one-way consequence of a rational support value and is used
in Corollary~\ref{res:rigidity-sublog-main}.  It limits zero gaps in the
divisor-coefficient stream; it neither reconstructs a support nor contradicts
the existence of an infinite rational support.

The coefficient $f_A(n)$ counts the elements of $A$ dividing $n$. Say that a
zero window of length $h$ starts after $c+N$ when
$f_A(c+N+j+1)=0$ for every $0\le j<h$. Thus no integer in that interval is
divisible by an element of the support.

\begin{theorem}[sublogarithmic zero windows]\label{res:sublog}
Let $A$ contain a positive integer and suppose
\[
  \sum_{a\in A}\frac1{2^a-1}=\frac{p}{2^c v},
  \qquad p\in\Z,\quad c\in\N,\quad v\ge1.
\]
For every $\varepsilon>0$ there is a constant $B\ge0$, depending only on
$\varepsilon,c,$ and $v$, such that every zero window after $c+N$ satisfies
\[
  h\le \varepsilon\log_2(N+1)+B,
  \qquad \log_2 x:=\frac{\log x}{\log 2}.
\]
The bound is uniform in $A,p,N,$ and $h$.
\end{theorem}
\noindent Formalised:
\lword{SublogDivisorCoverage.lean}{392}{supportCoeffZeroWindow_length_le_eps_logb_add}{sublogarithmic zero-window bound}.

The proof first establishes, for each integer $k\ge2$, the explicit divisor
estimate
\[
  \tau(n)^k\le \bigl(k^{2^k}\bigr)^k n.
\]
It transports the resulting $n^{1/k}$ bound through the binary tail and then
uses the exact carry recurrence across a zero window. Raising the resulting
inequality to the $k$th power absorbs the occurrence of $h$ on the right and
leaves a coefficient $1/(k-1)$ in front of the binary logarithm. Choosing $k$
after $\varepsilon$ gives the theorem. This controls gaps in divisor coverage;
it does not assert a subpower bound for the carry state itself.

For one fixed support the conclusion is elementary, since every zero window
is shorter than the least support element.  The content is uniformity:
\(B\) does not depend on \(A\), and taking \(N=0\) shows that a rational
value with denominator data \((c,v)\) bounds the least element of its
support in terms of \((\varepsilon,c,v)\) alone.  Against the half-value
fibre the constraint is inert: the greedy candidate support there contains
\(2\), so its divisor stream has no zero window of length above one and
satisfies the bound with room to spare.

\subsection{Necessary conditions II: wraps, mass, and unbounded states}
\label{app:carry-mass}

Three routes use the carry recurrence:
\[
\begin{array}{rcl}
\text{odd denominator}&\longrightarrow&
  \text{doubling-residue wraps}\longrightarrow\text{mass lower bound},\\
\text{finite common multiple}&\longrightarrow&
  \text{recurring excess spikes}\longrightarrow\text{collision correction},\\
\text{infinite support}&\longrightarrow&\text{unbounded carry states}.
\end{array}
\]
The first two routes use Ces\`aro averages when the reciprocal mass is finite;
the third needs no summability hypothesis.  These are parallel necessary
consequences of rationality, not implications from one conclusion to the
next.  The one-wrap classification identifies the extremal residue cycles
and is not a premise of any of the three routes.

Suppose now the support series equals $p/(2^c v)$ with $v$ odd and $A$
containing a positive element.  Write $q=2^cv$, and let $U$ be the tempered
orbit of Appendix~\ref{app:carry-boolean} with multiplier $q$.  Removing the
first $c$ binary places and the dyadic factor gives the positive integer state
\[
 u(N)=\frac{U(c+N)}{2^c}=v\,T_{f_A}(c+N);
\]
this is $u_N$ in Theorem~\ref{res:rigidity-main}, and it satisfies
$u(N+1)+v\,f_A(c+N+1)=2u(N)$.  Its least residue modulo $v$ is
$\bar p_N=p2^N\bmod v$, with $0\leq\bar p_N<v$.  Put
\[
 e_N=\left\lfloor\frac{u(N)}v\right\rfloor,
 \qquad u(N)=ve_N+\bar p_N;
\]
thus $e_N$ is the nonnegative integral excess above the least residue
(\lword{RationalSupportCarrySkeleton.lean}{668}{exists_shifted_odd_tail_nat_states_of_support_fraction}{shifted natural-state existence},
\lword{RationalSupportCarrySkeleton.lean}{506}{oddSupportTail_eq_residue_div_add_excess}{residue--excess decomposition}).
Doubling a least residue either stays under $v$ or wraps past it exactly once,
so each step emits a wrap digit $k_N=\lfloor2\bar p_N/v\rfloor\in\{0,1\}$, and over
any cycle length $h$ with $2^h\equiv1\pmod v$ the repetend identity holds:
\[
\sum_{N<h}\bar p_N \;=\; v\cdot w,
\qquad
w=\sum_{N<h}k_N,
\]
with $w\ge1$ when $p$ is coprime to $v>1$. The formal development records
the displayed equality as the
\lword{RationalSupportCarrySkeleton.lean}{171}{sum_doublingResidue_eq_mul_wrapCount}{wrap-count identity}
and its non-vanishing as the
\lword{RationalSupportCarrySkeleton.lean}{181}{doublingWrapCount_pos}{positive-wrap theorem}.

\begin{proposition}[one-wrap classification]\label{res:onewrap}
Let $v>1$ be odd, $p$ coprime to $v$, and $h\ge1$ with $2^h\equiv1\pmod v$.
If the cycle of $p$ wraps exactly once, then $v=2^h-1$ and the starting
residue is a power of two: $\bar p_0=2^a$ for some $a<h$.
\end{proposition}
\noindent The one-wrap cycles are thus exactly the Mersenne repetends
$2^a/(2^h-1)$. Formalised:
\lword{RationalSupportCarrySkeleton.lean}{415}{one_doublingWrap_classification}{one-wrap classification},
from the generic closed-cycle form
\lword{RationalSupportCarrySkeleton.lean}{353}{one_wrap_cycle_classification}{one-wrap cycle classification},
and instantiated at the true order $h=\operatorname{ord}_v(2)$
(\lword{RationalSupportCarrySkeleton.lean}{478}{one_oddOrder_doublingWrap_classification}{odd-order one-wrap classification}). The
classification is algebraic; a finite check of $446$ coprime starting residues
across twelve modulus/order rows is retained as kernel-checked validation, not
as the proof
(\lword{RationalSupportCarrySkeleton.lean}{2454}{orderWrapValidation446_passes}{446-start finite validation}).

The analytic bridge is Ces\`aro averaging. Write $\rho(A)=\sum_{a\in A}1/a$
for the reciprocal mass. When the
reciprocal family is summable, the mean of $f_A(1),\dots,f_A(N)$ converges to
$\rho(A)$ --- the density of the multiples of $a$, summed over the support ---
and the mean of the scaled tails has the same limit
(\lword{RationalSupportCarrySkeleton.lean}{945}{tendsto_supportCoeff_mean_reciprocalMass}{divisor-mean limit},
\lword{RationalSupportCarrySkeleton.lean}{1036}{tendsto_binaryCoeffTail_supportCoeff_mean_reciprocalMass}{coefficient-tail mean limit}).
The divergent case is carried as an explicit disjunction, not through a
default value assigned to a divergent formal sum.

\begin{proposition}[order-wrap lower bound, with exact excess mean]\label{res:orderwrap}
Let \(A\subseteq\Npos\) be nonempty and suppose
\[
 X_A=\frac{p}{2^cv},
 \qquad p\in\Z,\quad c\in\N,\quad v>1\ \text{odd}.
\]
Let \(h=\operatorname{ord}_v(2)\), let \(w\) be the number of wraps in the
least-residue cycle \(p2^N\bmod v\) for \(0\leq N<h\), and set
\[
 u(N)=vT_{f_A}(c+N),\qquad e_N=\left\lfloor\frac{u(N)}v\right\rfloor.
\]
Then $\sum_{a\in A}1/a$ diverges or $\rho(A)\ge w/h$; if moreover $p$ is
coprime to $v$, then $w\ge1$, so
$\rho(A)\ge1/\operatorname{ord}_v(2)$. In the summable case the bound is the
periodic part of an exact identity: the Ces\`aro mean of \(e_N\) converges,
and
\[
\rho(A)\;=\;\frac{w}{h}\;+\;\lim_{N\to\infty}\frac1N\sum_{n<N}e_n .
\]
\end{proposition}
\noindent The divergent-or-bounded conclusion is formalised by the
\lword{RationalSupportCarrySkeleton.lean}{1501}{support_fraction_oddOrder_wrapRatio_reciprocalMassDivergesOrAtLeast}{wrap-ratio mass alternative},
its coprime specialisation by the
\lword{RationalSupportCarrySkeleton.lean}{1517}{support_fraction_one_div_oddOrder_reciprocalMassDivergesOrAtLeast}{odd-order mass alternative},
and the automatic excess limit in the fraction-facing coordinates above by the
\lword{RationalSupportCarrySkeleton.lean}{1577}{exists_shifted_oddTailExcess_mean_of_support_fraction}{shifted excess-mean identity}.
A rational value thus ties the odd part of its denominator to the sparsity of
its support: a summable support with small $\rho(A)$ can only display
denominators whose odd part has small multiplicative order of $2$.

\begin{corollary}[collision strengthening at common multiples]\label{res:collision}
In the summable case, let $F\subseteq A$ be finite and let $L\ge1$ be a common
multiple of $F$. Then
\[
\rho(A)\;\ge\;\frac{w}{h}\;+\;\frac{\lceil(|F|-1)/2\rceil}{L}.
\]
In particular, an infinite support whose value is a dyadic rational $p/2^c$
has $\sum_{a\in A}1/a$ divergent or $\rho(A)>1$.
\end{corollary}
\noindent The mechanism is a collision: at every positive multiple of $L$ at
least $|F|$ support divisors land on the same index, so $f_A\ge|F|$ there,
while the exact carry equation $f_A(n)+e_n=k_{n-1}+2e_{n-1}$ with
$k_{n-1}\le1$ lets one step absorb at most one unit --- so the preceding
excess must spike to at least $\lceil(|F|-1)/2\rceil$, and the spikes recur
with density $1/L$. The three steps are formalised by the
\lword{RationalSupportCarrySkeleton.lean}{1963}{shiftedBooleanCollision_wrap_bound_of_common_multiple}{shifted common-multiple bound},
\lword{RationalSupportCarrySkeleton.lean}{1868}{shiftedSupportCoeff_add_oddTailExcess_succ_eq_wrap_add_two_excess}{shifted wrap--excess recurrence},
and
\lword{RationalSupportCarrySkeleton.lean}{2200}{dyadic_support_fraction_reciprocalMass_diverges_or_gt_one}{dyadic reciprocal-mass alternative}.

\begin{theorem}[carry states are unbounded over infinite supports]\label{res:unbounded}
Let $A$ be infinite with value $p/(2^c v)$, $v\ge1$. Then the positive integer
carry state $u(N)=v\,T_{f_A}(c+N)$ is unbounded: for every $B$ there is an $N$
with $u(N)>B$.
\end{theorem}
\noindent The proof picks $2B+1$ positive support elements; at a common
multiple $L$ of all of them, the local inequality $1+v\,|F|\le2\,u(L-c-1)$
pushes the state past $B$. The conclusion is the
\lword{RationalSupportCarrySkeleton.lean}{2377}{exists_unbounded_shifted_odd_tail_nat_state_of_support_fraction}{fraction-facing unbounded-state theorem};
its local engine is the
\lword{RationalSupportCarrySkeleton.lean}{2228}{one_add_mul_card_le_two_mul_shifted_state}{support-cardinality state bound}.
So a rational value over an infinite support cannot run on a bounded carry
alphabet. That rules out finite-state readings of the carry system; it does not
decide the problem. Nothing in the constraints above
prevents an infinite support from satisfying every one of these constraints.
The repetend identities and divisor averages used here are classical. We make
no claim of mathematical novelty for the coupled reciprocal-mass bounds,
collision strengthening, or global unboundedness statement.

\subsection{Independent geometry and one-sided nonmembership certificates}
\label{app:carry-geometry}

Finally, consider the set of all values a base-$2$ support series can
take. For $n\ge1$ put $w_n=1/(2^n-1)$, let $T(n)$ be the mass strictly after
exponent $n$, and let
\[
\mathcal A\;=\;\Bigl\{\,\sum_{n\in A}w_n \;:\; A\subseteq\N,\ 0\notin A\Bigr\},
\]
the \emph{Mersenne achievement set}, the exponent $0$ being normalised away as
analytically invisible. Coding by supports of positive exponents is literally
the formal support series
(\lword{GreedyAchievementSet.lean}{554}{positiveMersenneSupportValue_eq_erdosSupportSeries}{support-value identity}). In
this language the base-$2$ \#257 statement reads: the rational members of
$\mathcal A$ are exactly the finite subset sums. Nothing below decides that.
Kova\v{c} and Tao verify, for every fixed integer $t\ge2$, that
\[
  \sum_{\ell>n}\frac1{t^\ell-1}<\frac1{t^n-1},
\]
and deduce distinct Lambert subsums and a Cantor set~\cite{kovactao}. At
$t=2$ this is the relevant Lambert-series instance of Kakeya's classical
strict-tail criterion for subsum sets~\cite{kakeya1914}. The local
contribution is the checked greedy criterion, quantitative enclosure, and
finite nonmembership certificates over that known geometry; it does not
attribute those refinements to the cited sources or settle \#257.

\begin{theorem}[superincreasing geometry and greedy membership]\label{res:greedy-details}
For every $n\ge1$ the weight dominates the whole tail after it, $T(n)<w_n$,
with the two-scale gap $w_n-T(n)=\tfrac23\,4^{-n}+O(8^{-n})$ and explicit
valid $O$-constant $3$. Consequently a real $x$ belongs to $\mathcal A$
exactly when $x\ge0$ and every greedy residual of $x$ is at most the remaining
tail; and normalised support coding is injective, so each member of
$\mathcal A$ has a unique support, which the greedy recursion recovers bit by
bit.
\end{theorem}
\noindent Formalised:
\lword{GreedyAchievementSet.lean}{180}{mersenneTail_lt_weight}{strict-tail inequality},
\lword{GreedyAchievementSet.lean}{473}{mersenneGap_isBigOWith}{gap big-O estimate} (explicit bound
\lword{GreedyAchievementSet.lean}{448}{mersenneGap_asymptotic_bound}{explicit gap bound}),
\lword{GreedyAchievementSet.lean}{1445}{mem_mersenneAchievementSet_iff_greedy_survival}{greedy survival criterion},
and
\lword{GreedyAchievementSet.lean}{1527}{positiveMersenneSupportValue_injective_normalized}{normalised support-value injectivity},
transferred to the kernel series at
\lword{GreedyAchievementSet.lean}{1536}{erdosSupportSeries_two_injective_normalized}{normalised support-series injectivity}.
The nonnegativity guard in the membership criterion is necessary: without it a
negative target would survive every level while lying outside $\mathcal A$.

The binary coding also permits the global geometry to be checked directly.

\begin{proposition}[fat-Cantor geometry]\label{res:greedytopology}
The achievement set $\mathcal A$ is compact, perfect, totally disconnected,
and nowhere dense.  Its Lebesgue measure is exactly one.
\end{proposition}
\noindent Formalised:
\lword{GreedyAchievementSet.lean}{996}{volume_mersenneAchievementSet}{measure-one theorem},
\lword{GreedyAchievementSet.lean}{1620}{perfect_mersenneAchievementSet}{perfectness},
\lword{GreedyAchievementSet.lean}{1636}{isTotallyDisconnected_mersenneAchievementSet}{total disconnectedness},
and
\lword{GreedyAchievementSet.lean}{1645}{isNowhereDense_mersenneAchievementSet}{nowhere density}.
Kova\v{c}--Tao supply the strict-tail Cantor-set context cited above.  The
formalisation records these exact properties without making a novelty or
priority claim for them.

\begin{proposition}[exact rational runs and finite nonmembership certificates]\label{res:death}
The greedy recursion executed in exact rational arithmetic agrees step by step
with the real recursion under casting. For a rational target $x$, write
$r_n(x)$ for its exact rational residual after level $n$ and, for a lookahead
$\ell\ge0$, put
\[
\widehat T_{n,\ell}
  =\sum_{k=0}^{\ell-1}w_{n+k+1}+2w_{n+\ell+1}\in\Q .
\]
The sum is empty when $\ell=0$, and always
$T(n)<\widehat T_{n,\ell}$. Hence the finite inequality
$\widehat T_{n,\ell}\le r_n(x)$ soundly proves $x\notin\mathcal A$; its
level-one instance shows $3/4\notin\mathcal A$. A rational number already
known to lie in $\mathcal A$ has computable support bits.
\end{proposition}
\noindent Formalised:
\lword{GreedyAchievementSet.lean}{1130}{cast_greedyMersenneRemainderRat}{rational--real greedy agreement},
\lword{GreedyAchievementSet.lean}{1692}{mersenneTail_lt_cast_tailUpperRat}{rational tail-enclosure bound},
\lword{GreedyAchievementSet.lean}{1726}{certifiedGreedyMersenneDeath_not_mem}{finite-certificate soundness},
\lword{GreedyAchievementSet.lean}{1748}{three_fourths_not_mem_mersenneAchievementSet}{three-quarters nonmembership},
and \lword{GreedyAchievementSet.lean}{1832}{rational_member_support_bit_iff}{rational-member support-bit equivalence}.
The certificates are one-sided: survival through any finite depth proves
nothing about membership, and no decidability of membership in $\mathcal A$ is
asserted.  Neither the global geometry nor the finite nonmembership certificates
exclude rational values for either open series.

\paragraph{Appendix consequence.}
The transfer to the main text is Theorem~\ref{res:rigidity-main} followed by
the four independent
Corollaries~\ref{res:rigidity-unbounded-main}--\ref{res:rigidity-dyadic-mass-main}.
Rationality yields the exact carry recurrence; any nonempty rational support
must satisfy the sublogarithmic divisor-coverage bound, and the odd-part
wrap and mass bounds when its denominator has odd part \(v>1\); an infinite
one must in addition satisfy the unbounded-state theorem and, at a dyadic
denominator, the reciprocal-mass alternative.
Appendix~\ref{app:carry-geometry} also supplies the finite nonmembership certificates
used in Theorem~\ref{res:greedy}.  What is not proved is that the support attached to
either target violates one of these necessary conditions.  That
contradiction, not another rational normal form, is the missing premise.

\section{Technical reductions for the totient-certificate branch}
\label{app:totient-technical}

Recall the scaled tail
\(R_N=\sum_{m\geq1}\ph(N+m)/2^m\).
The predicate \(\mathsf{Sep}(h,N,L)\) retains the first $L$ binary places of
\(R_{N+h}-R_N\) and checks, by finite integer arithmetic, that their least
residue modulo $2^L$ clears the rigorous omitted-tail margins at both $0$
and $2^L$.  Thus $N$ is the tail location and $L$ the finite certificate
depth.  At a fixed pair \((h,N)\), some $L$ succeeds exactly when the
difference is non-integral.  Irrationality requires the stronger cofinal statement: for
every candidate period \(h\), a certificate must occur after every cutoff.
This appendix develops coordinates and finite mechanisms for that statement;
it does not prove the required cofinal supply
\eqref{eq:unbounded-certificate-supply}.

The exact chain is
\[
\begin{array}{rcl}
 R_{N+h}-R_N\notin\Z
 &\Longleftrightarrow&
 \exists L\geq1,\ \mathsf{Sep}(h,N,L),\\
 S\notin\Q
 &\Longleftrightarrow&
 \forall h\geq1\ \forall N_0\ \exists N\geq N_0\ \exists L\geq1,\
 \mathsf{Sep}(h,N,L).
\end{array}
\]
Appendix~\ref{app:totient-normal} gives exact coordinates for the two lines.
Appendix~\ref{app:totient-finite} separates fixed-scale mechanisms from
conditional implications that assume their cofinal repetition.
Appendix~\ref{app:totient-boundaries} records candidate sufficient inputs and
rules out several shortcuts.  Appendix~\ref{app:totient-rank} proves
unconditional dyadic rank growth but not the missing implication from rank to
the second line.  Thus only the normal forms are equivalences; the remaining
subsections provide finite results, conditional routes, or boundaries.

\TotientTechnicalAppendix

\section{Lambert, probability, and finite-algebraic complements}
\label{app:lambert-complements}

This appendix supplies the extensions cited from the Lambert-series identities
in Section~\ref{sec:ladder} and from the coprime-pair representation after
Proposition~\ref{res:coprime}.  Its new results are the periodic-weight
classifications, finite denominator obstructions, the M\"obius-square and squared
Lambert identities, and the Stern--Brocot cylinder and continuant laws.
These enlarge the two coordinate pictures; they are not additional steps in
either open proof.

The three families are parallel, not cumulative.
Appendix~\ref{app:lambert-weighted} proves or classifies weighted Lambert
series with coefficient inputs different from the open totient input.
Appendix~\ref{app:lambert-algebraic} gives exact identities and finite
algebraic boundaries for the squared and M\"obius coordinates.
Appendix~\ref{app:lambert-probability} develops unconditional geometry
internal to the coprime-pair probability coordinate.
No conclusion about a comparison input transfers irrationality to \(S\), and
none of the three families supplies the cofinal
\(\mathsf{Sep}\) supply theorem.

\FurtherLambertAppendix

\paragraph{Appendix consequence.}
The transfer is coordinate extension only.
Appendix~\ref{app:lambert-weighted} enlarges the Mersenne--Lambert comparison
family in Section~\ref{sec:ladder};
Appendix~\ref{app:lambert-algebraic} supplies the M\"obius-square and
gcd-moment coordinates; and Appendix~\ref{app:lambert-probability} refines
the coprime-pair representation in
Proposition~\ref{res:coprime}.  These are three independent coordinate
extensions.  None
implies the cofinal \(\mathsf{Sep}\) supply for \(S\) or a proof
of \(1/2\in\mathcal A\).  Those, rather than any result internal to this
appendix, remain the missing statements identified in the main text.

\section{Guide to the formal sources}\label{app:index}

Each blue mathematical phrase opens its exact Lean declaration at the pinned
public revision.  The repository's JSON claim registry records the same
declaration names for search.  Reader routes are: bounded rational exclusion,
Theorem~\ref{res:farey}; the exact \#249 certificate reduction,
Proposition~\ref{res:complete} and
Theorems~\ref{res:pointwisecomplete}--\ref{res:diag}; and the \#257 half-value
classification, Theorems~\ref{res:halfmembership} and~\ref{res:fatalright}.
The classical benchmark is Theorem~\ref{res:full}, while
Section~\ref{sec:architecture} compares the reductions.  Independent
structural results, none of which closes either problem, are
Theorem~\ref{res:dyadictotient}, the sublogarithmic zero-window corollary, and
Theorems~\ref{res:sternbrocotcylinders} and~\ref{res:sternbrocotruns}.

% Retain the exhaustive source-link manifest in the authored TeX for automated
% validation and query tooling, but keep it outside the printed paper.
\iffalse

Each row links to the declaration at the archived source revision.

\begin{center}
\footnotesize
\begin{tabular}{@{}p{7.2cm}p{7.8cm}@{}}
\toprule
informal statement & linked Lean source\\
\midrule
$\sum 1/(b^n-1)$ irrational, all $b\ge2$ & \lrefx{CertificateKernel.lean}{8000}{irrational_erdosSum_full_support}\ (\texttt{CertificateKernel})\\
Erd\H{o}s--Borwein $E$ irrational & \lrefx{CertificateKernel.lean}{8007}{irrational_erdosBorwein_series}\ (\texttt{CertificateKernel})\\
factorial support & \lrefx{CertificateKernel.lean}{5707}{irrational_erdosSum_factorial_support}\ (\texttt{CertificateKernel})\\
power-of-two support & \lrefx{CertificateKernel.lean}{5731}{irrational_erdosSum_two_pow_support}\ (\texttt{CertificateKernel})\\
pairwise-coprime supports & \lrefx{CertificateKernel.lean}{10448}{irrational_erdosSupportSeries_pairwise_coprime}\ (\texttt{CertificateKernel})\\
eventually-periodic nonnegative rational weights & \lrefx{CertificateKernel.lean}{12483}{irrational_ratWeightSeries_eventuallyPeriodic}\ (\texttt{CertificateKernel})\\
periodic signed-weight irrational-or-terminating dichotomy & \lrefx{CertificateKernel.lean}{13847}{irrational_or_bpow_mul_eq_intCast_intWeightedErdosSeries_periodic}\ (\texttt{CertificateKernel})\\
period-four residue obstruction & \lrefx{CertificateKernel.lean}{13898}{intWeightedCoeff_periodFourSignWeight_eq_zero_of_mod_four_eq_three}\ (\texttt{CertificateKernel})\\
$L(\mu)=1/2$ & \lrefx{MersenneLambertLadder.lean}{587}{tsum_moebius_div_two_pow_sub_one_eq_half}\ (\texttt{MersenneLambertLadder})\\
$L(\ph)=2$ & \lrefx{MersenneLambertLadder.lean}{575}{tsum_totient_div_two_pow_sub_one_eq_two}\ (\texttt{MersenneLambertLadder})\\
Lambert representation $L(\alpha)=S$ & \lrefx{CertificateKernel.lean}{18117}{tsum_primWeight_div_two_pow_sub_one_eq_totient_series}\ (\texttt{CertificateKernel})\\
M\"obius-square identity for $S$ & \lrefx{CertificateKernel.lean}{18126}{totient_series_eq_half_add_moebius_mersenne_square}\ (\texttt{CertificateKernel})\\
$L_2(\mathbf1)$ and $L_2(\ph)$ gcd moments & \lrefx{GcdMomentCalculus.lean}{216}{tsum_one_div_mersenne_sq_eq_sigma_sub_tau_series}, \lrefx{GcdMomentCalculus.lean}{235}{tsum_totient_div_mersenne_sq_eq_gcd_moment_series}\ (\texttt{GcdMomentCalculus})\\
Stern--Brocot cylinder recursion and rate & \lrefx{GcdMomentCalculus.lean}{474}{cylinderMass_split}, \lrefx{GcdMomentCalculus.lean}{525}{sternBrocotDepthMass_error}\ (\texttt{GcdMomentCalculus})\\
$S=\tfrac12+\Pr(\gcd=1)$ & \lrefx{CertificateKernel.lean}{18229}{totient_series_eq_half_add_visible_coprime_pairs}\ (\texttt{CertificateKernel})\\
$S\ne p/q$ for $q\le\Qzero$ & \lrefx{CertificateKernel.lean}{18056}{tsum_totient_div_pow_two_ne_ratCast_of_den_le_79639646646701375323355774875831053}\ (\texttt{CertificateKernel})\\
tail-period exact equivalence & \lrefx{LcmConeFlatness.lean}{412}{irrational_totient_series_iff_certificate_supply}\ (\texttt{LcmConeFlatness})\\
exact lcm-diagonal equivalence & \lrefx{LcmConeFlatness.lean}{426}{irrational_totient_series_iff_lcm_diagonal_certificate_supply}\ (\texttt{LcmConeFlatness})\\
cone flatness from rationality & \lrefx{LcmConeFlatness.lean}{90}{rational_totient_series_forces_lcm_cone_flatness}\ (\texttt{LcmConeFlatness})\\
certificate completeness & \lrefx{LcmConeFlatness.lean}{316}{exists_certifiedKill_iff_tail_diff_notMem_int}\ (\texttt{LcmConeFlatness})\\
joint cone refuter & \lrefx{LcmConeNonflat.lean}{126}{exists_nonintegral_pair_of_coneNonflatCert}\ (\texttt{LcmConeNonflat})\\
verified finite instances & \lrefx{TotientTailPeriodKiller.lean}{404}{certifiedKill_all_small}\ (\texttt{TotientTailPeriodKiller})\\
full-target diagonal reduction & \lrefx{DiagonalPincerDecomposition.lean}{215}{diagonal_int_iff_foreignDiagonalDefect_hits_fullTarget}\ (\texttt{DiagonalPincerDecomposition})\\
squared-Mersenne enclosure & \lrefx{SquaredMersenneDiagonalEnclosure.lean}{116}{scaleDiagonalTailDifference_sub_lambertProjectedDiagonal}\ (\texttt{SquaredMersenneDiagonalEnclosure})\\
exact Mersenne-shadow denominator & \lrefx{MersenneShadowDenominatorGrowth.lean}{147}{lcmHeight_scaledMobiusShadow_den_exact}\ (\texttt{MersenneShadowDenominatorGrowth})\\
finite residue-projection criterion & \lrefx{DiagonalFreshLossBridge.lean}{191}{irrational_totientSeries_of_diagonalFreshLossProjectionSupply}\ (\texttt{DiagonalFreshLossBridge})\\
prime-adjunction diamond no-go & \lrefx{FullTargetPrimeAdjunctionNoGo.lean}{196}{fullTarget_primeAdjunction_diamond_iff_root}\ (\texttt{FullTargetPrimeAdjunctionNoGo})\\
scalar-localisation height obstruction & \lrefx{AdelicHeightObstruction.lean}{23}{scalarLocalization_complement_dvd}\ (\texttt{AdelicHeightObstruction})\\
square-CRT correction suppression & \lrefx{SquareCRTCube.lean}{198}{exists_squareCRT_clean_totient_family}\ (\texttt{SquareCRTCube})\\
signed dyadic-moment algebra & \lrefx{SignedQMomentObstruction.lean}{28}{det_mul_rectangular}\ (\texttt{SignedQMomentObstruction})\\
infinite-dimensional dyadic totient kernel & \lrefx{TotientMahlerDefect.lean}{1014}{not_finiteDimensional_span_fullTotientKernel}\ (\texttt{TotientMahlerDefect})\\
residual-gauge obstruction for monomial minors & \lrefx{ResidualGaugeObstruction.lean}{55}{det_residualMonomialMatrix_ne_zero_iff}\ (\texttt{ResidualGaugeObstruction})\\
tail-orbit rationality criterion & \lrefx{GenericTailOrbitRigidity.lean}{425}{binaryCoeffSeries_rational_iff_exists_temperedBinaryOrbit}\ (\texttt{GenericTailOrbitRigidity})\\
Boolean carry certificate $\Leftrightarrow$ support fraction & \lrefx{BooleanMobiusCarry.lean}{948}{exists_normalized_support_fraction_iff_exists_booleanMobiusCarry}\ (\texttt{BooleanMobiusCarry})\\
one-wrap classification & \lrefx{RationalSupportCarrySkeleton.lean}{415}{one_doublingWrap_classification}\ (\texttt{RationalSupportCarrySkeleton})\\
order-wrap mass alternative and exact excess mean & \lrefx{RationalSupportCarrySkeleton.lean}{1517}{support_fraction_one_div_oddOrder_reciprocalMassDivergesOrAtLeast}, \lrefx{RationalSupportCarrySkeleton.lean}{1577}{exists_shifted_oddTailExcess_mean_of_support_fraction}\ (\texttt{RationalSupportCarrySkeleton})\\
unbounded carry states & \lrefx{RationalSupportCarrySkeleton.lean}{2377}{exists_unbounded_shifted_odd_tail_nat_state_of_support_fraction}\ (\texttt{RationalSupportCarrySkeleton})\\
sublogarithmic zero windows & \lrefx{SublogDivisorCoverage.lean}{392}{supportCoeffZeroWindow_length_le_eps_logb_add}\ (\texttt{SublogDivisorCoverage})\\
greedy membership criterion & \lrefx{GreedyAchievementSet.lean}{1445}{mem_mersenneAchievementSet_iff_greedy_survival}\ (\texttt{GreedyAchievementSet})\\
Mersenne achievement set: measure one and nowhere dense & \lrefx{GreedyAchievementSet.lean}{996}{volume_mersenneAchievementSet}, \lrefx{GreedyAchievementSet.lean}{1645}{isNowhereDense_mersenneAchievementSet}\ (\texttt{GreedyAchievementSet})\\
Campbell prime-progression incompatibility & \lrefx{CampbellShiftSynchronization.lean}{299}{campbellQuarterExponent_periodFreeze_noSync}, \lrefx{CampbellShiftSynchronization.lean}{488}{shiftWindowZero_iff_half_mem_mersenneAchievementSet}\ (\texttt{CampbellShiftSynchronization})\\
\bottomrule
\end{tabular}
\end{center}
\fi

\clearpage
\pdfbookmark[0]{References}{references}
\begin{thebibliography}{99}
\small
\raggedright
\setlength{\itemsep}{0pt}
\bibitem{erdos1948} P.~Erd\H{o}s, \emph{On arithmetical properties of Lambert series},
  J. Indian Math. Soc. (N.S.) \textbf{12} (1948), 63--66,
  \href{https://users.renyi.hu/~p_erdos/1948-04.pdf}{archive scan}.
\bibitem{erdos1968} P.~Erd\H{o}s, \emph{On the irrationality of certain series},
  Math. Student \textbf{36} (1968), 222--226 (issued 1969),
  \href{https://users.renyi.hu/~p_erdos/1969-09.pdf}{archive scan}.
\bibitem{erdosgraham1980} P.~Erd\H{o}s and R.~L. Graham, \emph{Old and New Problems and
  Results in Combinatorial Number Theory}, Monographies de l'Enseignement
  Math\'ematique, vol.~28, L'Enseignement Math\'ematique, Geneva, 1980,
  pp.~61--62,
  \href{https://mathweb.ucsd.edu/~ronspubs/80_11_number_theory.pdf}{author-hosted scan}.
\bibitem{erdos1988} P.~Erd\H{o}s, \emph{On the irrationality of certain series: problems and
  results}, in \emph{New Advances in Transcendence Theory}, ed.~A.~Baker,
  Cambridge University Press, 1988, pp.~102--109,
  doi:\href{https://doi.org/10.1017/CBO9780511897184.009}{10.1017/CBO9780511897184.009}.
\bibitem{borwein1992} P.~B. Borwein, \emph{On the irrationality of certain series},
  Math. Proc. Cambridge Philos. Soc. \textbf{112} (1992), no.~1, 141--146,
  doi:\href{https://doi.org/10.1017/S030500410007081X}{10.1017/S030500410007081X}.
\bibitem{apostol} T.~M. Apostol, \emph{Introduction to Analytic Number Theory},
  Springer, 1976,
  doi:\href{https://doi.org/10.1007/978-1-4757-5579-4}{10.1007/978-1-4757-5579-4}.
\bibitem{merca2017} M.~Merca, \emph{The Lambert series factorization theorem},
  Ramanujan J. \textbf{44} (2017), no.~2, 417--435,
  doi:\href{https://doi.org/10.1007/s11139-016-9856-3}{10.1007/s11139-016-9856-3}.
\bibitem{mercaschmidt} M.~Merca and M.~D. Schmidt, \emph{Generating special arithmetic
  functions by Lambert series factorizations}, Contrib. Discrete Math. \textbf{14}
  (2019), no.~1, 31--45,
  doi:\href{https://doi.org/10.55016/ojs/cdm.v14i1.62425}{10.55016/ojs/cdm.v14i1.62425}.
\bibitem{kakeya1914} S.~Kakeya, \emph{On the set of partial sums of an infinite series},
  Proc. Tokyo Math.-Phys. Soc., 2nd ser. \textbf{7} (1914), 250--251,
  doi:\href{https://doi.org/10.11429/ptmps1907.7.14_250}{10.11429/ptmps1907.7.14\_250}.
\bibitem{farey1816} J.~Farey, \emph{On a curious property of vulgar fractions},
  Philos. Mag. \textbf{47} (1816), 385--386,
  doi:\href{https://doi.org/10.1080/14786441608628487}{10.1080/14786441608628487}.
\bibitem{alloucheshallit} J.-P. Allouche and J. Shallit, \emph{Automatic
  Sequences: Theory, Applications, Generalizations}, Cambridge University
  Press, 2003,
  doi:\href{https://doi.org/10.1017/CBO9780511546563}{10.1017/CBO9780511546563}.
\bibitem{coons2010} M.~Coons, \emph{(Non)automaticity of number theoretic
  functions}, J. Th\'eor. Nombres Bordeaux \textbf{22} (2010), no.~2,
  339--352, doi:\href{https://doi.org/10.5802/jtnb.718}{10.5802/jtnb.718}.
\bibitem{nesterenko} Yu.~V. Nesterenko, \emph{Modular functions and transcendence questions},
  Sb. Math. \textbf{187} (1996), no.~9, 1319--1348,
  doi:\href{https://doi.org/10.1070/SM1996v187n09ABEH000158}{10.1070/SM1996v187n09ABEH000158}.
\bibitem{postelmansvanassche} K.~Postelmans and W.~Van Assche,
  \emph{Irrationality of $\zeta_q(1)$ and $\zeta_q(2)$},
  J. Number Theory \textbf{126} (2007), no.~1, 119--154,
  doi:\href{https://doi.org/10.1016/j.jnt.2006.11.011}{10.1016/j.jnt.2006.11.011}.
\bibitem{lucatachiya} F.~Luca and Y.~Tachiya, \emph{Irrationality of Lambert series associated
  with a periodic sequence}, Int. J. Number Theory \textbf{10} (2014), no.~3, 623--636,
  doi:\href{https://doi.org/10.1142/S1793042113501121}{10.1142/S1793042113501121}.
\bibitem{kovactao} V.~Kova\v{c} and T.~Tao, \emph{On several irrationality problems for
  Ahmes series}, Acta Math. Hungar. \textbf{175} (2025), no.~2, 572--608,
  doi:\href{https://doi.org/10.1007/s10474-025-01528-0}{10.1007/s10474-025-01528-0}.
\bibitem{wang2026} H.~Wang, \emph{Positive dyadic density for
  rational weighted binary expansions}, arXiv:2606.24972v2 (2026),
  doi:\href{https://doi.org/10.48550/arXiv.2606.24972}{10.48550/arXiv.2606.24972}.
\bibitem{crandall} R.~Crandall, \emph{The googol-th bit of the Erd\H{o}s--Borwein constant},
  Integers \textbf{12} (2012), no.~5, 811--840,
  doi:\href{https://doi.org/10.1515/integers-2012-0007}{10.1515/integers-2012-0007}.
\bibitem{campbell} J.~M. Campbell, \emph{On the binary digits of the Erd\H{o}s--Borwein
  constant}, arXiv:2605.24160v1 (2026),
  doi:\href{https://doi.org/10.48550/arXiv.2605.24160}{10.48550/arXiv.2605.24160}.
\bibitem{alfordgranvillepomerance} W.~R. Alford, A.~Granville, and C.~Pomerance,
  \emph{There are infinitely many Carmichael numbers}, Ann. of Math. (2)
  \textbf{139} (1994), no.~3, 703--722,
  doi:\href{https://doi.org/10.2307/2118576}{10.2307/2118576}.
\bibitem{taoteravainen} T.~Tao and J.~Ter\"av\"ainen, \emph{Quantitative correlations and
  some problems on prime factors of consecutive integers}, arXiv:2512.01739v2
  (2025; revised 2026),
  doi:\href{https://doi.org/10.48550/arXiv.2512.01739}{10.48550/arXiv.2512.01739}.
\bibitem{lean} L.~de Moura, S.~Kong, J.~Avigad, F.~van Doorn, and J.~von Raumer,
  \emph{The Lean theorem prover (system description)}, \emph{CADE-25}, LNCS 9195
  (2015), 378--388, doi:\href{https://doi.org/10.1007/978-3-319-21401-6_26}{10.1007/978-3-319-21401-6\_26}.
\bibitem{mathlib} The mathlib Community, \emph{The Lean mathematical library},
  in \emph{CPP 2020}, ACM (2020), 367--381, doi:\href{https://doi.org/10.1145/3372885.3373824}{10.1145/3372885.3373824}.
\bibitem{erdosproblems} T.~Bloom (ed.), \emph{Erd\H{o}s Problems},
  \href{https://www.erdosproblems.com/249}{entry 249} and
  \href{https://www.erdosproblems.com/257}{entry 257}, accessed 18 July 2026.

\end{thebibliography}

\end{document}
